### 2017-04-24 manuscript.tex
skeleton document

% Template for PLoS
% Version 3.3 June 2016
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

\usepackage{seqsplit}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

\newcommand{\comment}[1]{{\color{red}(\textsl{#1})}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Quantitative comparison of the potential for viral escape from broad and narrow anti-influenza antibodies} % Please use "title case" (capitalize all terms in the title except conjunctions, prepositions, and articles).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Michael B. Doud\textsuperscript{1,2,3},
Jesse D. Bloom\textsuperscript{1,2*},
\\
\bigskip
\textbf{1} Basic Sciences and Computational Biology Program, Fred Hutchinson Cancer Research Center, Seattle, WA, USA
\\
\textbf{2} Department of Genome Sciences, University of Washington, Seattle, WA,  USA
\\
\textbf{3} Medical Scientist Training Program, University of Washington, Seattle, WA, USA
\\
\bigskip

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* \href{mailto:jbloom@fredhutch.org}{jbloom@fredhutch.org}

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
There is great interest in broad antibody-based immunity against influenza virus.
Even for antibodies recognizing broad swaths of influenza strains, viral genetic variation can influence neutralization sensitivity.
To better understand the potential for viral resistance to antibodies, there is need to evaluate not just whether mutations can influence sensitivity to an antibody, but also to quantify how many mutations do this and by how much, and to make these comparisons across antibodies.
Here we comprehensively map mutations that increase viral resistance to the broadly neutralizing anti-influenza antibody FI6v3, and compare the results to those previously obtained for narrow strain-specific antibodies.
We find that although there are some mutations that enhance resistance to FI6v3, these mutations are both rare and modest in their effect sizes.
We formalize these observations into a framework for quantitative comparison of viral escape potential between different antibodies.
Overall, our work demonstrates that a virus's absolute capacity to escape an antibody can be quantified, and that influenza is in fact extremely limited in its ability to escape some antibodies.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author Summary}
this is the author summary

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}
this is the introduction

\section*{Results}

\subsection*{Results subsection 1}

\subsection*{Results subsection 2}

\section*{Discussion}

\clearpage

\section*{Materials and Methods}

\subsubsection*{methods subsubsection 1}

\subsubsection*{Data availability and source code}
Deep sequencing data has been deposited at the Sequence Read Archive under BioSample accession ().

\section*{Acknowledgments}

\nolinenumbers

\bibliography{references.bib}

\clearpage

\begin{figure}[!h]
\caption{\label{fig:?}
{\bf figure caption.} 
{\bf (A)} 
Libraries ...
{\bf (B)} 
Differential selection ...
}
\end{figure}

\clearpage

\section*{Supporting Information}

% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
\paragraph*{S1 Fig.}
\label{S1_Fig}
{\bf S1 Figure title.} 
Caption text.

\end{document}

### 2017-06-21 manuscript.tex
add some refs and begin methods

@@
 Michael B. Doud\textsuperscript{1,2,3},
+Juhye M. Lee\textsuperscript{1,2,3},
 Jesse D. Bloom\textsuperscript{1,2*},
@@
 Even for antibodies recognizing broad swaths of influenza strains, viral genetic variation can influence neutralization sensitivity.
-To better understand the potential for viral resistance to antibodies, there is need to evaluate not just whether mutations can influence sensitivity to an antibody, but also to quantify how many mutations do this and by how much, and to make these comparisons across antibodies.
-Here we comprehensively map mutations that increase viral resistance to the broadly neutralizing anti-influenza antibody FI6v3, and compare the results to those previously obtained for narrow strain-specific antibodies.
-We find that although there are some mutations that enhance resistance to FI6v3, these mutations are both rare and modest in their effect sizes.
+To better understand the potential for viral resistance to antibodies, there is a need to evaluate not just whether mutations can influence sensitivity to an antibody, but rather to comprehensively quantify how many mutations do this and by how much, and to make these comparisons across antibodies.
+Here we comprehensively map mutations that increase viral resistance to the broadly neutralizing anti-influenza antibodies FI6v3, C179, and S139/1, and compare the results to those previously obtained for narrow strain-specific antibodies.
+We find that although there are some mutations that enhance resistance to the stalk-binding antibody FI6v3 \comment{and C179?}, these mutations are both rare and modest in their effect sizes.
+\comment{..?... In contrast, viral escape from S139/1 ... looks more like the narrow head binders than the broad stalk binders?}
 We formalize these observations into a framework for quantitative comparison of viral escape potential between different antibodies.
-Overall, our work demonstrates that a virus's absolute capacity to escape an antibody can be quantified, and that influenza is in fact extremely limited in its ability to escape some antibodies.
+Overall, our work demonstrates that a virus's absolute capacity to escape an antibody can be quantified, and that influenza is in fact extremely limited in its ability to escape some antibodies \comment{targeting the stalk...
+and also extremely adept at escaping another broadly-reactive antibody targeting the head...
+}.
 
@@
 \section*{Introduction}
-this is the introduction
+
+Broadly neutralizing antibodies against influenza virus are of great interest... \cite{corti2017tackling}.
+
+% blurbs about the specific antibodies we use here:
+C179 isolation and escape mutant selection \cite{okuno1993common}.
+C179 structure with 1957 H2, binding data to many strains, and citations to older studies demonstrating C179 cross-neutralization to H1, H2, H5, H6, and H9:  \cite{dreyfus2013structure}.
+FI6v3 was isolated... \cite{corti2011neutralizing}.
+S139/1 isolation and selection of escape mutants from Aichi H3, Adachi H2, and WSN H1: \cite{yoshida2009cross}.
+S139/1 structure with Victoria75 H3 and binding/neutralization data: \cite{lee2012heterosubtypic}.
+
+Recently we have described new technologies... \cite{doud2017complete, dingens2017comprehensive}.
 
@@
 
-\subsubsection*{methods subsubsection 1}
+\subsubsection*{Antibodies}
+FI6v3 was expressed and purified by the Fred Hutchinson Cancer Research Center protein expression core \comment{i think?}.
+C179 was purchased from Takara Bio Inc (Catalog \# M145).
+S139/1 heavy and light chain variable sequences were obtained from PDB ID 4GMS~\cite{lee2012heterosubtypic} and expressed and purified by the Fred Hutchinson Cancer Research Center protein expression core.
+
+\subsubsection*{Mutant virus selections with antibody}
+Library selections, deep sequencing, and computation of mutation differential selection were performed as previously described~\cite{doud2017complete}. Briefly, ...
+
+\subsubsection*{Computation of the fraction $\phi_{r,a}$ of each mutation that escapes antibody neutralization}
+\comment{depending on how much of paper focus is on this new metric, some of this rationale might be better placed in the results section}
+
+Mutation differential selection values $s_{r,a}$ reflect the enrichment of a mutation in an antibody-selected sample relative to a mock-selected control. 
+The extent of these mutation enrichments are dependent on the stringency of the neutralization: as the concentration of an antibody used to neutralize the mutant virus library is increased, a larger proportion of the virus library is neutralized, and escape mutants become more enriched in the antibody-selected sample relative to the control sample~\cite{doud2017complete}.
+Consequently, due to differences in the concentrations and potencies of various antibodies used in mutational antigenic profiling, quantitative comparisons of the effects of mutations on neutralization escape between various antibodies is confounded by differences in neutralization stringency across experiments.
+
+We defined a new metric that accounts for the neutralization stringency of each experiment so that quantitative comparisons can be made between experiments with varying neutralization stringencies.
+This new measure, $\phi_{r,a}$, is roughly analogous to the fraction of mutant viruses carrying mutation $a$ at site $r$ that escape neutralization by a given antibody.
+The computation of $\phi_{r,a}$ incorporates information about the stringency of neutralization. 
+We define this as  $\gamma$ = (100 - \% neutralization)/100), the fraction of library remaining infectious after antibody neutralization measured by qRT-PCR for each antibody-selection experiment as previously described~\cite{doud2017complete}.
+
+Define $\hat{n}_{r,a}^{sel}$ and $\hat{n}_{r,a}^{mock}$ as the number of error-corrected counts of amino-acid mutant $a$ at site $r$ as defined in~\cite{doud2017complete}, and $N_r^{sel}$ and $N_r^{mock}$ be the total counts at site $r$, as specified for either an antibody selection or matched mock condition.
+Let $P$ be a pseudocount (default to 5 in these analyses) added to amino-acid counts, and let $f_{r}^{sel}$ and $f_{r}^{mock}$ be relative depths of the selected and mock samples at site $r$ for the purposes of scaling the pseudocount by relative read depth as previously described~\cite{doud2017complete}.
+
+Define the pseudocount-adjusted frequency $\rho$ of mutant $a$ at site $r$ in selected and mock samples to be:
+$$\rho_{r,a}^{sel} = \frac{n_{r,a}^{sel}+f_r^{sel}\times P}{N_r^{sel}+f_r^{sel}\times  P}$$
+$$\rho_{r,a}^{mock} = \frac{n_{r,a}^{mock}+f_r^{mock}\times P}{N_r^{mock}+f_r^{mock}\times  P}$$
+
+We define $\phi_{r,a}$ as the net `fraction' of mutant $a$ at site $r$ escaping neutralization (above the average escape fraction $\gamma$), and compute this as:
+$$\phi_{r,a} = \frac{\gamma \times \rho_{r,a}^{sel}}{\rho_{r,a}^{mock}} - \gamma$$
+
+By this definition, mutations with no effect on antibody escape will have $\phi_{r,a} =0$, and mutations completely escaping antibody will have $\phi_{r,a} \approx 1$.
+Care should be taken in interpreting the absolute values of $\phi$, however, since the choice of pseudocount affects the range of $\phi$ values (but not, importantly, the relationships of $\phi$ distributions across antibodies).
+We have found that a value of $P=5$ pseudocounts results in $\phi$ values that are roughly bounded from 0 to 1, and consistently use $P=5$ all analyses.
 
 \subsubsection*{Data availability and source code}
-Deep sequencing data has been deposited at the Sequence Read Archive under BioSample accession ().
+Deep sequencing data has been deposited at the Sequence Read Archive under BioSample accession (\comment{should probably deposit all the new data into the existing biosample SAMN05789126 (Sample name: WSN/1933 HA mutant libraries selected by monoclonal antibodies), which is under bioproject PRJNA309339}).
 

### 2017-07-28 manuscript.tex
upload fig 1 draft and add results subsection headers

@@
 Here we comprehensively map mutations that increase viral resistance to the broadly neutralizing anti-influenza antibodies FI6v3, C179, and S139/1, and compare the results to those previously obtained for narrow strain-specific antibodies.
-We find that although there are some mutations that enhance resistance to the stalk-binding antibody FI6v3 \comment{and C179?}, these mutations are both rare and modest in their effect sizes.
+We find that although there are some mutations that enhance resistance to the stalk-binding antibody FI6v3 and C179, these mutations are both rare and modest in their effect sizes.
 \comment{..?... In contrast, viral escape from S139/1 ... looks more like the narrow head binders than the broad stalk binders?}
@@
 
-\subsection*{Results subsection 1}
+\subsection*{Characterizing escape mutations from a broadly neutralizing receptor site binding antibody}
 
-\subsection*{Results subsection 2}
+\subsection*{Characterizing escape mutations from broadly neutralizing stalk-binding antibodies}
+
+\subsection*{Comparison of escape profiles with mutational tolerance}
+
+\subsection*{Quantitatively comparing escape from different neutralizing antibodies}
 

### 2017-09-29 paper/manuscript.tex
updated paper outline structure to no longer be for _PLoS Pathogens_

--- 2017-06-21 manuscript.tex
@@
-% Template for PLoS
-% Version 3.3 June 2016
-%
-% % % % % % % % % % % % % % % % % % % % % %
-%
-% -- IMPORTANT NOTE
-%
-% This template contains comments intended 
-% to minimize problems and delays during our production 
-% process. Please follow the template instructions
-% whenever possible.
-%
-% % % % % % % % % % % % % % % % % % % % % % % 
-%
-% Once your paper is accepted for publication, 
-% PLEASE REMOVE ALL TRACKED CHANGES in this file 
-% and leave only the final text of your manuscript. 
-% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
-% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
-%
-%
-% There are no restrictions on package use within the LaTeX files except that 
-% no packages listed in the template may be deleted.
-%
-% Please do not include colors or graphics in the text.
-%
-% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
-%
-% % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- FIGURES AND TABLES
-%
-% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
-%
-% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
-% - Figures should be uploaded separately from your manuscript file. 
-% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
-% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
-% For figure citations, please use "Fig" instead of "Figure".
-% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
-%
-% Tables should be cell-based and may not contain:
-% - spacing/line breaks within cells to alter layout or alignment
-% - do not nest tabular environments (no tabular environments within tabular environments)
-% - no graphics or colored text (cell background color/shading OK)
-% See http://journals.plos.org/plosone/s/tables for table guidelines.
-%
-% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
-%
-% IMPORTANT
-% Below are a few tips to help format your equations and other special characters according to our specifications.
-For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
-%
-% For inline equations, please be sure to include all portions of an equation in the math environment.
-For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
-%
-% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
-%
-% Please add line breaks to long display equations when possible in order to fit size of the column. 
-%
-% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
-%
-% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
-%
-% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
-%
-% % % % % % % % % % % % % % % % % % % % % % % % 
-%
-% Please contact latex@plos.org with any questions.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
+% Use only LaTeX2e, calling the article.cls class and 12-point type.
 
-\documentclass[10pt,letterpaper]{article}
-\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
-
-% amsmath and amssymb packages, useful for mathematical formulas and symbols
-\usepackage{amsmath,amssymb}
-
-% Use adjustwidth environment to exceed column width (see example table in text)
-\usepackage{changepage}
-
-% Use Unicode characters when possible
-\usepackage[utf8x]{inputenc}
-
-% textcomp package and marvosym package for additional characters
-\usepackage{textcomp,marvosym}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-% Use nameref to cite supporting information files (see Supporting Information section for more info)
-\usepackage{nameref,hyperref}
-
-% line numbers
-\usepackage[right]{lineno}
-
-% ligatures disabled
-\usepackage{microtype}
-\DisableLigatures[f]{encoding = *, family = * }
-
-% color can be used to apply background shading to table cells only
-\usepackage[table]{xcolor}
-
-% array package and thick rules for tables
-\usepackage{array}
-
-% create "+" rule type for thick vertical lines
-\newcolumntype{+}{!{\vrule width 2pt}}
-
-% create \thickcline for thick horizontal lines of variable length
-\newlength\savedwidth
-\newcommand\thickcline[1]{%
-  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-  \cline{#1}%
-  \noalign{\vskip\arrayrulewidth}%
-  \noalign{\global\arrayrulewidth\savedwidth}%
-}
-
-% \thickhline command for thick horizontal lines that span the table
-\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
-\hline
-\noalign{\global\arrayrulewidth\savedwidth}}
+\documentclass[11pt]{article}
+\usepackage[round,semicolon]{natbib}
+\usepackage[margin=1.3in]{geometry}
+\usepackage{kpfonts}
 
@@
 
-% Remove comment for double spacing
-%\usepackage{setspace} 
-%\doublespacing
+\usepackage{newfloat}
+\usepackage[labelfont=bf]{caption}
+\usepackage{nameref}
+\usepackage{rotating}
+\usepackage{color}
+\usepackage{float}
 
-% Text layout
-\raggedright
-\setlength{\parindent}{0.5cm}
-\textwidth 5.25in 
-\textheight 8.75in
+\setcounter{topnumber}{8}
+\setcounter{bottomnumber}{8}
+\setcounter{totalnumber}{8}
 
-% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
-% Captions will be left justified
-\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
-\renewcommand{\figurename}{Fig}
+\usepackage{newfloat}
+\DeclareFloatingEnvironment[name={Supplementary Figure}]{suppfigure}
 
-% Use the PLoS provided BiBTeX style
-\bibliographystyle{plos2015}
+\definecolor{darkblue}{rgb}{0, 0.0, 0.6}
 
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
+\usepackage{hyperref}
+\hypersetup{colorlinks,citecolor=blue,linkcolor=darkblue,urlcolor=blue}
 
-% Leave date blank
-\date{}
+\usepackage{seqsplit}
 
-% Header and Footer with logo
-\usepackage{lastpage,fancyhdr,graphicx}
-\usepackage{epstopdf}
-\pagestyle{myheadings}
-\pagestyle{fancy}
-\fancyhf{}
-\setlength{\headheight}{27.023pt}
-\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
-\rfoot{\thepage/\pageref{LastPage}}
-\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
-\fancyheadoffset[L]{2.25in}
-\fancyfootoffset[L]{2.25in}
-\lfoot{\sf PLOS}
-
-%% Include all macros below
-
-\newcommand{\lorem}{{\bf LOREM}}
-\newcommand{\ipsum}{{\bf IPSUM}}
+\usepackage{array}
+\newcolumntype{R}[1]{>{\raggedright\arraybackslash}p{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 
@@
 
-%% END MACROS SECTION
+\usepackage{setspace}
+
+\renewcommand{\topfraction}{1}
+\renewcommand{\bottomfraction}{1}
+\renewcommand{\textfraction}{0}
+\renewcommand{\floatpagefraction}{1}
+
+%\renewcommand{\abstractname}{\large SUMMARY}
+
+\title{Contrasting the potential for viral escape from broad and narrow antibodies by complete mapping of the antigenic effects of single amino-acid mutations to influenza hemagglutinin}
+
+\author
+{Michael B. Doud$^{1,2,3,\dagger}$, Juhye M. Lee$^{1,2,3,\dagger}$, and Jesse D. Bloom$^{1,2,*}$\\
+\\
+\scriptsize{$^1$Basic Sciences and Computational Biology Program, Fred Hutchinson Cancer Research Center}\\
+\scriptsize{$^2$Department of Genome Sciences and $^3$Medical Scientist Training Program, University of Washington} \\
+\scriptsize{Seattle, WA, USA} \\
+\scriptsize{$^{\dagger}$These authors contributed equally} \\
+\scriptsize{$^*$Correspondence: \href{jbloom@fredhutch.org}{jbloom@fredhutch.org}}
+}
+
+\date{}
 
 \begin{document}
-\vspace*{0.2in}
 
-% Title must be 250 characters or less.
-\begin{flushleft}
-{\Large
-\textbf\newline{Quantitative comparison of the potential for viral escape from broad and narrow anti-influenza antibodies} % Please use "title case" (capitalize all terms in the title except conjunctions, prepositions, and articles).
-}
-\newline
-% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
-\\
-Michael B. Doud\textsuperscript{1,2,3},
-Juhye M. Lee\textsuperscript{1,2,3},
-Jesse D. Bloom\textsuperscript{1,2*},
-\\
-\bigskip
-\textbf{1} Basic Sciences and Computational Biology Program, Fred Hutchinson Cancer Research Center, Seattle, WA, USA
-\\
-\textbf{2} Department of Genome Sciences, University of Washington, Seattle, WA,  USA
-\\
-\textbf{3} Medical Scientist Training Program, University of Washington, Seattle, WA, USA
-\\
-\bigskip
+\maketitle
+\onehalfspacing
 
-% Use the asterisk to denote corresponding authorship and provide email address in note below.
-* \href{mailto:jbloom@fredhutch.org}{jbloom@fredhutch.org}
+\begin{abstract}
+Broadly neutralizing antibodies are of great interest as therapeutics to treat influenza and are the desired products of future universal influenza vaccines. 
+It is important to understand viral potential for escape from antibodies and to develop ways to compare how easy or difficult it is for viral escape between various antibodies. 
+We previously demonstrated a method to measure the effects of all amino-acid mutations in influenza HA on viral escape from neutralizing antibodies. 
+Here we apply this technique to 'broadly neutralizing' antibodies with reactivity spanning multiple influenza subtypes, including antibodies targeting the HA receptor-binding site and stalk domain.
+We find quantitatively similar distributions of escape mutations to a relatively broad antibody targeting the receptor-binding site (s139) and to several narrow-specificity antibodies targeting nearby antigenic regions of the head domain (H17-L19, others?).
+In contrast to the ease of escape from head-binding antibodies of various breadths, stalk-binding antibodies can only be evaded by a paucity of mutations with relatively small effects.  
+For the antibodies studied here, antibody breadth is not an indicator of difficulty for viral escape.  
+Instead, targeting a mutationally intolerant epitope in the stalk region seems to limit viral escape mutations to a few relatively ineffective options
+\end{abstract}
 
-\end{flushleft}
-% Please keep the abstract below 300 words
-\section*{Abstract}
-There is great interest in broad antibody-based immunity against influenza virus.
-Even for antibodies recognizing broad swaths of influenza strains, viral genetic variation can influence neutralization sensitivity.
-To better understand the potential for viral resistance to antibodies, there is a need to evaluate not just whether mutations can influence sensitivity to an antibody, but rather to comprehensively quantify how many mutations do this and by how much, and to make these comparisons across antibodies.
-Here we comprehensively map mutations that increase viral resistance to the broadly neutralizing anti-influenza antibodies FI6v3, C179, and S139/1, and compare the results to those previously obtained for narrow strain-specific antibodies.
-We find that although there are some mutations that enhance resistance to the stalk-binding antibody FI6v3 \comment{and C179?}, these mutations are both rare and modest in their effect sizes.
-\comment{..?... In contrast, viral escape from S139/1 ... looks more like the narrow head binders than the broad stalk binders?}
-We formalize these observations into a framework for quantitative comparison of viral escape potential between different antibodies.
-Overall, our work demonstrates that a virus's absolute capacity to escape an antibody can be quantified, and that influenza is in fact extremely limited in its ability to escape some antibodies \comment{targeting the stalk...
-and also extremely adept at escaping another broadly-reactive antibody targeting the head...
-}.
-
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLOS ONE authors please skip this step. 
-% Author Summary not valid for PLOS ONE submissions.   
-\section*{Author Summary}
-this is the author summary
-
-\linenumbers
-
-% Use "Eq" instead of "Equation" for equation citations.
-\section*{Introduction}
-
+\section*{INTRODUCTION}
 Broadly neutralizing antibodies against influenza virus are of great interest... \cite{corti2017tackling}.
@@
 
-\section*{Results}
+\section*{RESULTS}
 
-\subsection*{Results subsection 1}
+\subsection*{First subsection.}
+Some text
 
-\subsection*{Results subsection 2}
+\begin{figure}
+\centerline{\includegraphics[width=\textwidth]{figs/antibody_summary_fig/fig.png}}
+\caption{\label{fig:antibody_summary}
+CAPTION}
+\end{figure}
 
-\section*{Discussion}
+\section*{DISCUSSION}
+Some discussion
 
@@
 
-\section*{Materials and Methods}
-
-\subsubsection*{Antibodies}
+\section*{METHODS}
+\subsection*{Antibodies}
 FI6v3 was expressed and purified by the Fred Hutchinson Cancer Research Center protein expression core \comment{i think?}.
@@
 
-\subsubsection*{Mutant virus selections with antibody}
+\subsection*{Mutant virus selections with antibody}
 Library selections, deep sequencing, and computation of mutation differential selection were performed as previously described~\cite{doud2017complete}. Briefly, ...
 
-\subsubsection*{Computation of the fraction $\phi_{r,a}$ of each mutation that escapes antibody neutralization}
+\subsection*{Computation of the fraction $\phi_{r,a}$ of each mutation that escapes antibody neutralization}
 \comment{depending on how much of paper focus is on this new metric, some of this rationale might be better placed in the results section}
@@
 
-\subsubsection*{Data availability and source code}
+\subsection*{Data availability and source code}
 Deep sequencing data has been deposited at the Sequence Read Archive under BioSample accession (\comment{should probably deposit all the new data into the existing biosample SAMN05789126 (Sample name: WSN/1933 HA mutant libraries selected by monoclonal antibodies), which is under bioproject PRJNA309339}).
 
-\section*{Acknowledgments}
+\clearpage
 
-\nolinenumbers
+\small
+\subsection*{ACKNOWLEDGMENTS}
+This work was supported by grants R01GM102198 and R01AI127893 from the NIGMS and NIAID of the NIH.
+MBD was supported in part by training grant T32AI083203 from the NIAID of the NIH.
+JML was supported in part by \comment{CIDID grant}.
+The research of JDB is supported in part by a Faculty Scholar Grant from the Howard Hughes Medical Institute and the Simons Foundation.
 
+\bibliographystyle{mbe}
 \bibliography{references.bib}
@@
 \clearpage
+\normalsize
 
-\begin{figure}[!h]
-\caption{\label{fig:?}
-{\bf figure caption.} 
-{\bf (A)} 
-Libraries ...
-{\bf (B)} 
-Differential selection ...
-}
-\end{figure}
-
-\clearpage
-
-\section*{Supporting Information}
-
-% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
-\paragraph*{S1 Fig.}
-\label{S1_Fig}
-{\bf S1 Figure title.} 
-Caption text.
+\section*{Supplementary Material}
 

### 2017-09-30 analysis_code/pymodules/__init__.py
Created new `Jupyter notebook` downloading sequencing data from SRA

@@ -1,0 +1,1 @@
+"""Makes modules in this directory importable."""

### 2017-09-30 analysis_code/pymodules/utils.py

@@ -1,0 +1,395 @@
+"""Python utilities for iPython notebook analysis.
+
+Juhye Lee."""
+
+import os
+import subprocess
+import glob
+import math
+import time
+import pandas as pd
+import numpy as np
+from scipy.optimize import curve_fit
+from colour import Color
+import functools
+import dms_tools2
+
+import matplotlib
+import matplotlib.lines as mlines
+matplotlib.use("Pdf")
+import pylab as plt
+from IPython.display import Image, display
+
+colorcycle = ['#1f77b4', '#aec7e8', '#ff7f0e', '#ffbb78', '#2ca02c', '#98df8a', '#d62728', '#ff9896', '#9467bd', '#c5b0d5', '#8c564b', '#c49c94', '#e377c2', '#f7b6d2', '#7f7f7f', '#c7c7c7', '#bcbd22', '#dbdb8d', '#17becf', '#9edae5']
+
+orange = '#e69f00'
+skyblue = '#56b4e9'
+bluishgreen = '#009e73'
+yellow = '#f0e442'
+blue = '#0072b2'
+vermillion = '#d55e00'
+reddishpurple = '#cc79a7'
+darkvermillion = '#873b00'
+lightvermillion = '#ff8524'
+
+    
+def MapDiffselColorToSite(diffselfile, scriptfile, script_type='pymol', 
+                          map_type='abs_diffsel', colors = ['#ffff99', '#990000'], sitecolorfile=None,
+                          script_preamble=False, restrict_to_chain=False, prefsfile=None):
+    '''Produces a colormapping based on differential selection and writes a script for `pymol` or `chimera`
+    to color a structure by this colormapping.
+    
+    Uses the data in *diffselfile*, which can be either a mutdiffsel or a sitediffsel file, 
+    depending on the specified *map_type*. 
+    
+    Writes a python script to *scriptfile* for the molecular visualization program specified 
+    in *script_type* (either `pymol` or `chimera`).
+    
+    *colors* is a list of two colors defined by hex or r,g,b codes.
+    
+    How the mapping from differential selection to color is determined is specified by the 
+    following choices for *map_type*:
+    
+        * `abs_diffsel`: plot total absolute differential selection at each site. 
+        Total absolute differential selection will be converted to color by interpolating between 
+        the two *colors*, which will be used to show the minimum and maximum values, respectively.
+        Requries *diffselfile* to be a '*sitediffsel.txt' file.
+        
[…]

### 2017-10-01 analysis_code/pymodules/utils.py
Begin cleaning up `analysis_code.ipynb` and create preliminary figure drafts

@@ -93,7 +93,7 @@ def MapDiffselColorToSite(diffselfile, scriptfile, script_type='pymol',
     column_names = list(df)
     if column_names == ['site', 'wildtype', 'mutation', 'mutdiffsel']:
         filetype = 'mutdiffsel'
-    elif column_names == ['site', 'abs_diffsel', 'positive_diffsel', 'negative_diffsel']:
+    elif column_names == ['site', 'abs_diffsel', 'positive_diffsel', 'negative_diffsel', 'max_diffsel', 'min_diffsel']:
         filetype = 'sitediffsel'
     else:
         raise ValueError('diffsel file does not have appropriate set of column identifiers')

### 2017-10-05 paper/manuscript.tex
added comments about figure adjustments

@@
 
-\newcommand{\comment}[1]{{\color{red}(\textsl{#1})}}
+\newcommand{\comment}[1]{{\color{red}[\textsl{#1}]}}
 
@@
 \begin{figure}
-\centerline{\includegraphics[width=\textwidth]{figs/antibody_summary_fig/fig.png}}
+\centerline{\includegraphics[width=\textwidth]{figs/antibody_summary_fig/antibody_summary_fig.pdf}}
 \caption{\label{fig:antibody_summary}
+CAPTION
+\comment{Can you make the following changes:
+(1) Add H17-L10 and H17-L7.
+These could just be in different shades of red since they are all narrow antibodies like H17-L19.
+(2) Make the structure slightly larger relative to the tree.
+(3) Increase the line widths on the tree and the font sizes everywhere.
+(4) Move the color-based label to the left side of the tree so it falls between the two panels, since it is really labeling both.
+(5) Make the tree branches all black.
+Since color is already being used to represent the different antibodies, I think it is confusing to also use it to distinguish Group 1 and Group 2.}}
+\end{figure}
+
+\begin{figure}
+\centerline{\includegraphics[width=0.8\textwidth]{figs/fracsurvive_example/frac_survive_example_plot.pdf}}
+\caption{\label{fig:fracsurvive_example}
 CAPTION}
+\end{figure}
+
+\begin{figure}
+\centerline{\includegraphics[width=0.8\textwidth]{figs/neutralization_curves/WT_neutralization_curves.pdf}}
+\caption{\label{fig:neutcurves}
+CAPTION
+\comment{Can you modify this plot to bring the formatting more in line with Figure~\ref{fig:fracsurvive_example}. 
+This includes:
+(1) Make the y-axis \emph{fraction infectivity}.
+(2) Make the vertical lines dotted and of the same color as in Figure~\ref{fig:fracsurvive_example}, which is the last color in the \emph{cbbPalette} defined here \url{https://github.com/jbloomlab/HA_antibody_ease_of_escape/blob/master/paper/figs/fracsurvive_example/fracsurvive_example.ipynb}.
+(3) Get rid of the minor tick marks that show up on two of the four plots.
+(4) Add H17-L10 and H17-L9, and make the plot three antibodies per line.
+(5) Get rid of the concentration of S139/1 that we are not using.
+(6) Also, I think it would look better if for each plot the antibody name was just a title above the plot in large letters.
+The x-axis then can then just ``antibody concentration ($\mu$g/ml)''.
+Also, I don't think you need the number labels for the concentrations on either the plot or in the table.
+If your table has the concentration and fraction infectivity, it is obvious which line goes with which concentration just by seeing how they are arranged horizontally.
+(7) Increase the font size in the table (you'll have more space to do this after you get rid of the numbers).
+}
+}
 \end{figure}

### 2017-10-05 paper/manuscript.tex
some work on abstract

@@
 
-\title{Contrasting the potential for viral escape from broad and narrow antibodies by complete mapping of the antigenic effects of single amino-acid mutations to influenza hemagglutinin}
+\title{Quantifying the ease of viral escape from broad and narrow antibodies to influenza virus hemagglutinin}
 
@@
 \begin{abstract}
-Broadly neutralizing antibodies are of great interest as therapeutics to treat influenza and are the desired products of future universal influenza vaccines. 
-It is important to understand viral potential for escape from antibodies and to develop ways to compare how easy or difficult it is for viral escape between various antibodies. 
-We previously demonstrated a method to measure the effects of all amino-acid mutations in influenza HA on viral escape from neutralizing antibodies. 
-Here we apply this technique to 'broadly neutralizing' antibodies with reactivity spanning multiple influenza subtypes, including antibodies targeting the HA receptor-binding site and stalk domain.
-We find quantitatively similar distributions of escape mutations to a relatively broad antibody targeting the receptor-binding site (s139) and to several narrow-specificity antibodies targeting nearby antigenic regions of the head domain (H17-L19, others?).
-In contrast to the ease of escape from head-binding antibodies of various breadths, stalk-binding antibodies can only be evaded by a paucity of mutations with relatively small effects.  
-For the antibodies studied here, antibody breadth is not an indicator of difficulty for viral escape.  
-Instead, targeting a mutationally intolerant epitope in the stalk region seems to limit viral escape mutations to a few relatively ineffective options
+Nearly all viruses exhibit some antigenic variation.
+However, the extent of this variation ranges widely across viruses and antibodies.
+Influenza virus exhibits high antigenic variation, and can completely escape most neutralizing antibodies with single mutations.
+However, some rare antibodies neutralize many influenza virus strains.
+It is unclear how easily the virus might escape such broad antibodies if there was strong selection to do so.
+Here we map all single amino-acid mutations to an H1 influenza hemagglutinin that increase resistance to several broad antibodies.
+Crucially, our approach both identifies antigenic mutations and quantifies their effect size.
+We find mutations that increase resistance to all antibodies tested, but the effect sizes range widely. 
+Single mutations completely escape a broad antibody targeting the hemagglutinin receptor-binding pocket, similar to escape from narrow strain-specific antibodies.   
+In contrast, the few mutations selected by broad antibodies targeting the hemagglutinin stalk have only small effects. 
+Overall, our work demonstrates a method to quantify the ease of viral antibody escape, and shows that antibody breadth is not necessarily an indicator of the difficulty of viral escape.
+%Instead, targeting a mutationally intolerant epitope in the stalk region seems to limit viral escape mutations to a few relatively ineffective options
 \end{abstract}

### 2017-10-05 paper/manuscript.tex
redraft of abstract
revised abstract again

@@
 \begin{abstract}
-Nearly all viruses exhibit some antigenic variation.
-However, the extent of this variation ranges widely across viruses and antibodies.
-Influenza virus exhibits high antigenic variation, and can completely escape most neutralizing antibodies with single mutations.
-However, some rare antibodies neutralize many influenza virus strains.
+Influenza virus can completely escape most neutralizing antibodies with single mutations.
+However, rare antibodies broadly neutralize many viral strains.
 It is unclear how easily the virus might escape such broad antibodies if there was strong selection to do so.
@@
 Overall, our work demonstrates a method to quantify the ease of viral antibody escape, and shows that antibody breadth is not necessarily an indicator of the difficulty of viral escape.
-%Instead, targeting a mutationally intolerant epitope in the stalk region seems to limit viral escape mutations to a few relatively ineffective options
+In addition, our results show that broad antibodies targeting the hemagglutinin stalk are quantifiably harder to escape via point mutations.
 \end{abstract}
@@
 \section*{INTRODUCTION}
+Nearly all viruses exhibit some antigenic variation.
+However, the extent of this variation ranges widely across viruses and antibodies.
 Broadly neutralizing antibodies against influenza virus are of great interest... \cite{corti2017tackling}.

### 2017-10-07 paper/manuscript.tex

@@
 
-\subsection*{First subsection.}
-Some text
+\subsection*{A method to quantify the ease of escape.}
+\comment{Jesse writes a draft of this section.}
+\begin{figure}
+\centerline{\includegraphics[width=0.8\textwidth]{figs/fracsurvive_example/frac_survive_example_plot.pdf}}
+\caption{\label{fig:fracsurvive_example}
+CAPTION}
+\end{figure}
 
+\subsection*{Broad and narrow antibodies that neutralize influenza virus hemagglutinin}
+\comment{Juhye writes a draft of this section.}
 \begin{figure}
@@
 Since color is already being used to represent the different antibodies, I think it is confusing to also use it to distinguish Group 1 and Group 2.}}
-\end{figure}
-
-\begin{figure}
-\centerline{\includegraphics[width=0.8\textwidth]{figs/fracsurvive_example/frac_survive_example_plot.pdf}}
-\caption{\label{fig:fracsurvive_example}
-CAPTION}
 \end{figure}
@@
 \end{figure}
+
+\subsection*{Ease of escape from each antibody}
+This would show the medianavgfracsurvive plot.
+
+\subsection*{For all antibodies, the selected sites of escape occur in the antibody binding footprints}
 

### 2017-10-07 paper/manuscript.tex
analysis notebook thru logo plots

@@
 \begin{figure}
-\centerline{\includegraphics[width=0.8\textwidth]{figs/fracsurvive_example/frac_survive_example_plot.pdf}}
+\centerline{\includegraphics[width=\textwidth]{figs/fracsurvive_example/fracsurvive_fig.pdf}}
 \caption{\label{fig:fracsurvive_example}
@@
 \begin{figure}
-\centerline{\includegraphics[width=\textwidth]{figs/antibody_summary_fig/antibody_summary_fig.pdf}}
+\centerline{\includegraphics[width=\textwidth]{figs/antibody_summary_fig/Ab_summary.pdf}}
 \caption{\label{fig:antibody_summary}
@@
 This would show the medianavgfracsurvive plot.
+
+\begin{figure}
+\centerline{\includegraphics[width=\textwidth]{../analysis_code/results/fracsurvive/summary_medianavgfracsurvive.pdf}}
+\caption{
+CAPTION
+}
+\end{figure}
 

### 2017-10-07 analysis_code/pymodules/__init__.py

@@ -1,1 +1,0 @@
-"""Makes modules in this directory importable."""

### 2017-10-07 analysis_code/pymodules/utils.py

@@ -1,395 +1,0 @@
-"""Python utilities for iPython notebook analysis.
-
-Juhye Lee."""
-
-import os
-import subprocess
-import glob
-import math
-import time
-import pandas as pd
-import numpy as np
-from scipy.optimize import curve_fit
-from colour import Color
-import functools
-import dms_tools2
-
-import matplotlib
-import matplotlib.lines as mlines
-matplotlib.use("Pdf")
-import pylab as plt
-from IPython.display import Image, display
-
-colorcycle = ['#1f77b4', '#aec7e8', '#ff7f0e', '#ffbb78', '#2ca02c', '#98df8a', '#d62728', '#ff9896', '#9467bd', '#c5b0d5', '#8c564b', '#c49c94', '#e377c2', '#f7b6d2', '#7f7f7f', '#c7c7c7', '#bcbd22', '#dbdb8d', '#17becf', '#9edae5']
-
-orange = '#e69f00'
-skyblue = '#56b4e9'
-bluishgreen = '#009e73'
-yellow = '#f0e442'
-blue = '#0072b2'
-vermillion = '#d55e00'
-reddishpurple = '#cc79a7'
-darkvermillion = '#873b00'
-lightvermillion = '#ff8524'
-
-    
-def MapDiffselColorToSite(diffselfile, scriptfile, script_type='pymol', 
-                          map_type='abs_diffsel', colors = ['#ffff99', '#990000'], sitecolorfile=None,
-                          script_preamble=False, restrict_to_chain=False, prefsfile=None):
-    '''Produces a colormapping based on differential selection and writes a script for `pymol` or `chimera`
-    to color a structure by this colormapping.
-    
-    Uses the data in *diffselfile*, which can be either a mutdiffsel or a sitediffsel file, 
-    depending on the specified *map_type*. 
-    
-    Writes a python script to *scriptfile* for the molecular visualization program specified 
-    in *script_type* (either `pymol` or `chimera`).
-    
-    *colors* is a list of two colors defined by hex or r,g,b codes.
-    
-    How the mapping from differential selection to color is determined is specified by the 
-    following choices for *map_type*:
-    
-        * `abs_diffsel`: plot total absolute differential selection at each site. 
-        Total absolute differential selection will be converted to color by interpolating between 
-        the two *colors*, which will be used to show the minimum and maximum values, respectively.
-        Requries *diffselfile* to be a '*sitediffsel.txt' file.
-        
[…]

### 2017-10-08 paper/manuscript.tex
revised abstract

@@
 \usepackage{seqsplit}
+\usepackage{placeins}
 
@@
 \usepackage{newfloat}
-\DeclareFloatingEnvironment[name={Supplementary Figure}]{suppfigure}
+\DeclareFloatingEnvironment[name={Figure}]{suppfigure}
+\renewcommand{\thesuppfigure}{S\arabic{suppfigure}}
 
@@
 
-\title{Quantifying the ease of viral escape from broad and narrow antibodies to influenza virus hemagglutinin}
+\title{Quantifying the ease of viral escape from broad and narrow antibodies to influenza hemagglutinin}
 
@@
 \begin{abstract}
-Influenza virus can completely escape most neutralizing antibodies with single mutations.
+Influenza virus can completely escape most antibodies with single mutations.
 However, rare antibodies broadly neutralize many viral strains.
-It is unclear how easily the virus might escape such broad antibodies if there was strong selection to do so.
-Here we map all single amino-acid mutations to an H1 influenza hemagglutinin that increase resistance to several broad antibodies.
-Crucially, our approach both identifies antigenic mutations and quantifies their effect size.
-We find mutations that increase resistance to all antibodies tested, but the effect sizes range widely. 
-Single mutations completely escape a broad antibody targeting the hemagglutinin receptor-binding pocket, similar to escape from narrow strain-specific antibodies.   
-In contrast, the few mutations selected by broad antibodies targeting the hemagglutinin stalk have only small effects. 
-Overall, our work demonstrates a method to quantify the ease of viral antibody escape, and shows that antibody breadth is not necessarily an indicator of the difficulty of viral escape.
-In addition, our results show that broad antibodies targeting the hemagglutinin stalk are quantifiably harder to escape via point mutations.
+It is unclear how easily influenza virus might escape such broad antibodies if they became widespread due to therapeutic use or vaccination.
+Here we map all single amino-acid mutations that increase resistance to broad antibodies targeting hemagglutinin.
+Crucially, our approach not only identifies antigenic mutations but also quantifies their effect sizes.
+All antibodies select mutations, but the effect sizes vary widely. 
+The virus can escape a broad antibody targeting hemagglutinin's receptor-binding pocket the same way it escapes narrow strain-specific antibodies: via single mutations with huge effects.   
+In contrast, broad antibodies targeting hemagglutinin's stalk only select mutations with small antigenic effects. 
+Therefore, antibody breadth is not necessarily an indicator of the difficulty of viral escape.
+Broad antibodies targeting hemagglutinin's stalk are quantifiably harder to escape than antibodies targeting hemagglutinin's head or receptor-binding pocket.
 \end{abstract}
@@
 \caption{\label{fig:fracsurvive_example}
-CAPTION}
+CAPTION
+\comment{One additional change to this figure: can you increase the spacing between panels, especially the top and bottom set of panels but also a bit between the ones on the same horizontal level.
+Right now it is too crowded.}
+}
 \end{figure}
@@
 CAPTION
-\comment{Can you make the following changes:
-(1) Add H17-L10 and H17-L7.
-These could just be in different shades of red since they are all narrow antibodies like H17-L19.
-(2) Make the structure slightly larger relative to the tree.
-(3) Increase the line widths on the tree and the font sizes everywhere.
-(4) Move the color-based label to the left side of the tree so it falls between the two panels, since it is really labeling both.
-(5) Make the tree branches all black.
-Since color is already being used to represent the different antibodies, I think it is confusing to also use it to distinguish Group 1 and Group 2.}}
+}
 \end{figure}
@@
 \begin{figure}
-\centerline{\includegraphics[width=\textwidth]{../analysis_code/results/fracsurvive/summary_medianavgfracsurvive.pdf}}
+\centerline{\includegraphics[width=\textwidth]{figs/avgfracsurvive.pdf}}
 \caption{
+\label{fig:avgfracsurvive}
 CAPTION
+Figure~\ref{suppfig:maxfracsurvive} shows the fraction surviving for the single most strongly selected amino-acid mutation at each site.
 }
@@
 \section*{Supplementary Material}
+\FloatBarrier
+
+\begin{suppfigure}
+\centerline{\includegraphics[width=\textwidth]{figs/maxfracsurvive.pdf}}
+\caption{\label{suppfig:maxfracsurvive}
+CAPTION}
+\end{suppfigure}
+
+\begin{suppfigure}
+\centerline{\includegraphics[trim=0.1cm 0.02cm 0.1cm 0.03cm,clip=true,width=\textwidth]{figs/logoplots/H17L19_fracsurvive.pdf}}
+\caption{\label{suppfig:H17L19logo}
+CAPTION}
+\end{suppfigure}
+
+\begin{suppfigure}
+\centerline{\includegraphics[trim=0.1cm 0.02cm 0.1cm 0.03cm,clip=true,width=\textwidth]{figs/logoplots/H17L10_fracsurvive.pdf}}
+\caption{\label{suppfig:H17L10logo}
+CAPTION}
+\end{suppfigure}
+
+\begin{suppfigure}
+\centerline{\includegraphics[trim=0.1cm 0.02cm 0.1cm 0.03cm,clip=true,width=\textwidth]{figs/logoplots/H17L7_fracsurvive.pdf}}
+\caption{\label{suppfig:H17L7logo}
+CAPTION}
+\end{suppfigure}
 

### 2017-10-08 paper/manuscript.tex
some progress on Intro

@@
 \usepackage[round,semicolon]{natbib}
-\usepackage[margin=1.3in]{geometry}
+\usepackage[margin=1.4in]{geometry}
 \usepackage{kpfonts}
@@
 \section*{INTRODUCTION}
-Nearly all viruses exhibit some antigenic variation.
-However, the extent of this variation ranges widely across viruses and antibodies.
+Nearly all viruses show some antigenic variation.
+However, the extent of this variation ranges widely.
+For instance, although both measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic,drexler2014robustness} exhibit antigenic variation, the magnitude of this variation is small.
+Therefore, immunity elicited by infection or vaccination provides life-long protection against both these viruses~\citep{?}.
+In contrast, some other viruses exhibit much more extensive antigenic variation.
+An example is human influenza virus.
+Infection with influenza virus provides life-long immunity to the infecting strain~\citep{?}.
+But influenza virus undergoes such rapid antigenic evolution that the effectiveness of this immunity to antigenically drifted viral strains is almost completely eroded within a few years~\citep{?}.
+
+Many explanations have been suggested for the differences in extent of antigenic variation among viruses, including epidemiological factors, the breadth of the immune response, and the virus's inherent evolutionary capacity to escape immunodominant antibody specificities~\citep{lipsitch2007patterns,cobey2014pathogen,fulton2015mutational}.
+Here we will focus on the last of these explanations, as it provides the underlying rationale for recent efforts to develop a ``universal'' influenza vaccine that protects against a wide range of viral strains.
+
+According to this inherent evolutionary capacity explanation, human influenza virus undergoes rapid antigenic drift because most neutralizing antibodies target epitopes on the viral hemagglutinin protein that are highly tolerant of mutational change.
+This explanation is supported by classic experiments showing that it is easy to select viral mutants that completely escape most anti-hemagglutinin antibodies~\citep{}, as well as by the observation that single mutations that cause large changes in antigenicity arise frequently during the natural evolution of human influenza virus at the global scale~\cite{?} and even within individual humans with long-term infections~\cite{?}.
+A corollary of this explanation is that the virus's capacity for antigenic drift would be greatly reduced if most antibodies instead targeted epitopes
+
+This corollary has become of more than academic interest with the discovery of broadly neutralizing antibodies.
+
+Of course, it remains unclear if these epitopes are more conserved because they are inherent intolerant or just aren't under pressure.
+For instance, non-human influenza doesn't undergo much antigenic drift.
+Knowing is important because success of a vaccine would put them under pressure.
+
 Broadly neutralizing antibodies against influenza virus are of great interest... \cite{corti2017tackling}.

### 2017-10-08 paper/manuscript.tex

--- 2017-10-08 paper/manuscript.tex
@@
 \section*{INTRODUCTION}
-Nearly all viruses exhibit some antigenic variation.
-However, the extent of this variation ranges widely across viruses and antibodies.
+Nearly all viruses show some antigenic variation.
+However, the extent of this variation ranges widely.
+For instance, although both measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic,drexler2014robustness} exhibit antigenic variation, the magnitude of this variation is small.
+Therefore, immunity elicited by infection or vaccination provides life-long protection against both these viruses~\citep{?}.
+In contrast, some viruses exhibit much more extensive antigenic variation.
+An example is human influenza virus.
+Infection with influenza virus provides life-long immunity to the infecting strain~\citep{?}.
+But influenza virus undergoes such rapid antigenic evolution that the effectiveness of this immunity to new viral strains is almost completely eroded within a few years~\citep{?}.
+
+One possible explanation for the reason that viruses exhibit differences in the extent of antigenic variation is that they have disparate inherent evolutionary capacities to escape immunodominant antibodies~\citep{lipsitch2007patterns,cobey2014pathogen,fulton2015mutational}.
+According to this explanation, human influenza virus undergoes rapid antigenic drift because most neutralizing antibodies target epitopes on the viral hemagglutinin protein that are highly tolerant of mutational change.
+This explanation is supported by classic experiments showing that it is easy to select viral mutants that escape most anti-hemagglutinin antibodies~\citep{yewdell1979antigenic,webster1980determination}, as well as by the observation that mutations that alter antigenicity arise frequently during influenza's evolution at the global scale~\citep{koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction} and even within individual humans with long-term infections~\citep{xue2017parallel}.
+A corollary of this explanation is that the virus's capacity for antigenic drift would be reduced if most antibodies instead targeted epitopes that were less mutationally tolerant.
+
+Verifying this corollary has become of practical importance with the discovery of broadly neutralizing antibodies against influenza virus.
+These antibodies typically target conserved epitopes in hemagglutinin's stalk~\citep{?} or receptor-binding pocket~\citep{?}, and neutralize a wide range of viral strains.
+Broad antibodies usually make only a modest contribution to natural human immunity, since they are less abundant than antibodies to the antigenically variable head of hemagglutinin~\citep{?}.
+However, major efforts are now underway to elicit broad antibodies by vaccination~\citep{?} or administer them directly as therapeutics or prophylactics~\citep{?}.
+
+If these efforts succeed, the epitopes of broadly neutralizing antibodies will then be under strong antigenic selection in human influenza virus.
+Certainly there is precedent for the idea that changing the immune status of the host population can alter influenza virus evolution: ...
+Of course, it remains unclear if these epitopes are more conserved because they are inherent intolerant or just aren't under pressure.
+For instance, non-human influenza doesn't undergo much antigenic drift.
+Knowing is important because success of a vaccine would put them under pressure.
+
 Broadly neutralizing antibodies against influenza virus are of great interest... \cite{corti2017tackling}.

### 2017-10-09 paper/manuscript.tex
full draft of Intro

--- 2017-10-08 paper/manuscript.tex
@@
 \section*{INTRODUCTION}
-Nearly all viruses exhibit some antigenic variation.
-However, the extent of this variation ranges widely across viruses and antibodies.
+Nearly all viruses show some antigenic variation.
+However, the extent of this variation ranges widely.
+For instance, although both measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic,drexler2014robustness} exhibit antigenic variation, the magnitude of this variation is small.
+Therefore, immunity to both these viruses is lifelong~\citep{?}.
+In contrast, human influenza virus exhibits extensive antigenic variation.
+So although infection with an influenza virus strain provides lifelong immunity to that exact strain~\citep{?}, the virus's rapid antigenic evolution erodes the effectiveness of this immunity to that strain's descendents within just a few years~\citep{?}.
+
+One possible reason that different viruses exhibit different amounts of antigenic variation is that they may have disparate evolutionary capacities to escape the antibodies that are immunodominant during natural immune responses~\citep{lipsitch2007patterns,cobey2014pathogen,fulton2015mutational}.
+According to this explanation, human influenza virus undergoes rapid antigenic drift because most neutralizing antibodies target epitopes on the viral hemagglutinin protein that are highly tolerant of mutational change.
+This explanation is supported by classic experiments showing that it is easy to select viral mutants that escape most antibodies~\citep{yewdell1979antigenic,webster1980determination}, as well as by the observation that mutations that alter antigenicity arise frequently during influenza's evolution globally~\citep{koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction} and within individual humans with long-term infections~\citep{xue2017parallel}.
+A corollary of this explanation is that the virus's capacity for antigenic drift would be reduced if most antibodies instead targeted epitopes that were less mutationally tolerant.
+
+Verifying this corollary has become of practical importance with the discovery of broadly neutralizing antibodies against influenza virus.
+These antibodies typically target conserved epitopes in hemagglutinin's stalk~\citep{?} or receptor-binding pocket~\citep{?}, and neutralize a wide range of viral strains.
+Broad antibodies usually make only a modest contribution to natural human immunity, since they are less abundant than antibodies to antigenically variable epitopes on the head of hemagglutinin~\citep{?}.
+However, major efforts are underway to elicit broad antibodies by vaccination~\citep{?} or administer them directly as therapeutics or prophylactics~\citep{?}.
+
+If these efforts succeed, the epitopes of broad antibodies will be under strong antigenic selection in human influenza virus.
+Would such selection then drive extensive antigenic variation in these epitopes?
+There is precedent for the idea that the immune status of the host population can shape influenza virus antigenic evolution: the virus undergoes faster antigenic drift in long-lived humans that accumulate immune memory than in short-lived swine that are mostly naive~\citep{sheerar1989antigenic,luoh1992hemagglutinin}, and poultry vaccination may accelerate antigenic drift of avian influenza~\citep{lee2004effect}.
+But alternatively, perhaps broad antibodies are broad because the virus has difficulty escaping from them regardless of the selection from host immunity.
+
+So far, there is limited data addressing whether the epitopes of broad antibodies are inherently resistant to viral antigenic evolution.
+Several studies have shown that the globular head domain of hemagglutinin is less mutationally tolerant than the stalk domain where many broad antibodies bind~\citep{thyagarajan2014inherent,wu2014high,heaton2013genome}.
+However, these studies did not select for actual antibody escape, so it is difficult to directly relate their measurements to the virus's evolutionary capacity under immune selection.
+Other work has selected antigenic mutants with a few broad antibodies~\citep{?}, demonstrating that these epitopes are not entirely refractory to change.
+But given that antibodies can select some antigenic variation even in measles virus~\citep{?} and polio virus~\citep{?}, the existence of mutations selectable by broad antibodies does not necessarily imply that influenza virus can escape these antibodies with ease.
+The fundamental problem is that none of these studies provide quantitative data on the ease of viral escape that can be compared across antibodies in an apples-to-apples fashion.
+
+Here we systematically quantify the ease with which influenza virus can escape from a range of broad and narrow antibodies via single amino-acid mutations.
+Specifically, we impose antibody selection on libraries of viruses carrying all amino-acid mutants of hemagglutinin compatible with viral replication, and then measure the fraction of virions with each mutation that survive the selection at several antibody concentrations.
+Critically, this approach quantifies the \emph{magnitude} of the antigenic effect of every mutation in a way that can be directly compared across antibodies.
+We find that even the broadest antibodies can select mutations that increase the virus's resistance to neutralization.
+However, the magnitudes of the antigenic effects vary greatly across antibodies.
+Single mutations can make the virus completely resistant to both narrow strain-specific antibodies and a broad antibody against the receptor-binding pocket.
+But no single mutation does more than modestly increase the virus's resistance to broad antibodies against the hemagglutinin stalk.
+These results suggest that systematically quantifying the antigenic effects of all single amino-acid mutations is a useful complement to simply measuring antibody breadth when trying to assess the potential for antigenic drift.
+
 Broadly neutralizing antibodies against influenza virus are of great interest... \cite{corti2017tackling}.

### 2017-10-09 paper/manuscript.tex

@@
 
-Broadly neutralizing antibodies against influenza virus are of great interest... \cite{corti2017tackling}.
-
-% blurbs about the specific antibodies we use here:
-C179 isolation and escape mutant selection \cite{okuno1993common}.
-C179 structure with 1957 H2, binding data to many strains, and citations to older studies demonstrating C179 cross-neutralization to H1, H2, H5, H6, and H9:  \cite{dreyfus2013structure}.
-FI6v3 was isolated... \cite{corti2011neutralizing}.
-S139/1 isolation and selection of escape mutants from Aichi H3, Adachi H2, and WSN H1: \cite{yoshida2009cross}.
-S139/1 structure with Victoria75 H3 and binding/neutralization data: \cite{lee2012heterosubtypic}.
-
-Recently we have described new technologies... \cite{doud2017complete, dingens2017comprehensive}.
+%Broadly neutralizing antibodies against influenza virus are of great interest... \cite{corti2017tackling}.
 
@@
 CAPTION
-\comment{One additional change to this figure: can you increase the spacing between panels, especially the top and bottom set of panels but also a bit between the ones on the same horizontal level.
-Right now it is too crowded.}
 }
@@
 \comment{Juhye writes a draft of this section.}
+
+% blurbs about the specific antibodies we use here:
+C179 isolation and escape mutant selection \cite{okuno1993common}.
+C179 structure with 1957 H2, binding data to many strains, and citations to older studies demonstrating C179 cross-neutralization to H1, H2, H5, H6, and H9:  \cite{dreyfus2013structure}.
+FI6v3 was isolated... \cite{corti2011neutralizing}.
+S139/1 isolation and selection of escape mutants from Aichi H3, Adachi H2, and WSN H1: \cite{yoshida2009cross}.
+S139/1 structure with Victoria75 H3 and binding/neutralization data: \cite{lee2012heterosubtypic}.
+
 \begin{figure}
@@
 CAPTION
-\comment{Can you modify this plot to bring the formatting more in line with Figure~\ref{fig:fracsurvive_example}. 
-This includes:
-(1) Make the y-axis \emph{fraction infectivity}.
-(2) Make the vertical lines dotted and of the same color as in Figure~\ref{fig:fracsurvive_example}, which is the last color in the \emph{cbbPalette} defined here \url{https://github.com/jbloomlab/HA_antibody_ease_of_escape/blob/master/paper/figs/fracsurvive_example/fracsurvive_example.ipynb}.
-(3) Get rid of the minor tick marks that show up on two of the four plots.
-(4) Add H17-L10 and H17-L9, and make the plot three antibodies per line.
-(5) Get rid of the concentration of S139/1 that we are not using.
-(6) Also, I think it would look better if for each plot the antibody name was just a title above the plot in large letters.
-The x-axis then can then just ``antibody concentration ($\mu$g/ml)''.
-Also, I don't think you need the number labels for the concentrations on either the plot or in the table.
-If your table has the concentration and fraction infectivity, it is obvious which line goes with which concentration just by seeing how they are arranged horizontally.
-(7) Increase the font size in the table (you'll have more space to do this after you get rid of the numbers).
+\comment{
+Get rid of the panel numbers altogether and just have the antibody names.
 }
@@
 \subsection*{Ease of escape from each antibody}
-This would show the medianavgfracsurvive plot.
+\comment{Jesse can work on a rough draft of this section.}
+
+\comment{Juhye, can you create a correlation among replicates supplemental figure.
+In the \texttt{./figs/corrs/} directory are all the plots.
+Is it possible to nicely arrange them on one page, maybe getting rid of the diagonals?}
 

### 2017-10-10 paper/manuscript.tex

@@
 For instance, although both measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic,drexler2014robustness} exhibit antigenic variation, the magnitude of this variation is small.
-Therefore, immunity to both these viruses is lifelong~\citep{?}.
-In contrast, human influenza virus exhibits extensive antigenic variation.
-So although infection with an influenza virus strain provides lifelong immunity to that exact strain~\citep{?}, the virus's rapid antigenic evolution erodes the effectiveness of this immunity to that strain's descendents within just a few years~\citep{?}.
+Therefore, immunity to both these viruses is lifelong~\citep{panum1847iagttagelser,salk1984one}.
+In contrast, human influenza virus exhibits much more antigenic variation.
+So although infection with an influenza virus strain provides long-term immunity to that exact strain~\citep{davies1982christ,yu2008neutralizing}, the virus's rapid antigenic evolution erodes the effectiveness of this immunity to that strain's descendants within four to seven years~\citep{couch1983immunity}.
 
-One possible reason that different viruses exhibit different amounts of antigenic variation is that they may have disparate evolutionary capacities to escape the antibodies that are immunodominant during natural immune responses~\citep{lipsitch2007patterns,cobey2014pathogen,fulton2015mutational}.
+One possible reason that viruses exhibit different amounts of antigenic variation is that they have disparate evolutionary capacities to escape the immunodominant antibodies generated by natural immune responses~\citep{lipsitch2007patterns,cobey2014pathogen,fulton2015mutational}.
 According to this explanation, human influenza virus undergoes rapid antigenic drift because most neutralizing antibodies target epitopes on the viral hemagglutinin protein that are highly tolerant of mutational change.
 This explanation is supported by classic experiments showing that it is easy to select viral mutants that escape most antibodies~\citep{yewdell1979antigenic,webster1980determination}, as well as by the observation that mutations that alter antigenicity arise frequently during influenza's evolution globally~\citep{koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction} and within individual humans with long-term infections~\citep{xue2017parallel}.
-A corollary of this explanation is that the virus's capacity for antigenic drift would be reduced if most antibodies instead targeted epitopes that were less mutationally tolerant.
+A corollary of this explanation is that influenza virus's capacity for antigenic drift would be reduced if most antibodies instead targeted epitopes that were less mutationally tolerant.
 
 Verifying this corollary has become of practical importance with the discovery of broadly neutralizing antibodies against influenza virus.
-These antibodies typically target conserved epitopes in hemagglutinin's stalk~\citep{?} or receptor-binding pocket~\citep{?}, and neutralize a wide range of viral strains.
+These antibodies typically target conserved epitopes in hemagglutinin's stalk~\citep{sui2009structural,ekiert2009antibody,corti2011neutralizing} or receptor-binding pocket~\citep{lee2012heterosubtypic,ekiert2012cross}, and neutralize a wide range of viral strains.
 Broad antibodies usually make only a modest contribution to natural human immunity, since they are less abundant than antibodies to antigenically variable epitopes on the head of hemagglutinin~\citep{?}.
-However, major efforts are underway to elicit broad antibodies by vaccination~\citep{?} or administer them directly as therapeutics or prophylactics~\citep{?}.
+However, major efforts are underway to elicit broad antibodies by vaccination or administer them directly as therapeutics~\citep{krammer2015advances,corti2017tackling}.
 
 If these efforts succeed, the epitopes of broad antibodies will be under strong antigenic selection in human influenza virus.
-Would such selection then drive extensive antigenic variation in these epitopes?
-There is precedent for the idea that the immune status of the host population can shape influenza virus antigenic evolution: the virus undergoes faster antigenic drift in long-lived humans that accumulate immune memory than in short-lived swine that are mostly naive~\citep{sheerar1989antigenic,luoh1992hemagglutinin}, and poultry vaccination may accelerate antigenic drift of avian influenza~\citep{lee2004effect}.
-But alternatively, perhaps broad antibodies are broad because the virus has difficulty escaping from them regardless of the selection from host immunity.
+Might such selection then drive extensive antigenic variation in these epitopes?
+There is precedent for the idea that the immune status of the host population can shape influenza virus evolution: the virus undergoes faster antigenic drift in long-lived humans that accumulate immune memory than in short-lived swine that are mostly naive~\citep{sheerar1989antigenic,luoh1992hemagglutinin}, and poultry vaccination may accelerate antigenic drift of avian influenza~\citep{lee2004effect}.
+But alternatively, perhaps broad antibodies are broad because the virus has difficulty escaping them regardless of the selection from host immunity.
 
-So far, there is limited data addressing whether the epitopes of broad antibodies are inherently resistant to viral antigenic evolution.
+So far, there is limited data to distinguish between these possibilities.
 Several studies have shown that the globular head domain of hemagglutinin is less mutationally tolerant than the stalk domain where many broad antibodies bind~\citep{thyagarajan2014inherent,wu2014high,heaton2013genome}.
-However, these studies did not select for actual antibody escape, so it is difficult to directly relate their measurements to the virus's evolutionary capacity under immune selection.
-Other work has selected antigenic mutants with a few broad antibodies~\citep{?}, demonstrating that these epitopes are not entirely refractory to change.
-But given that antibodies can select some antigenic variation even in measles virus~\citep{?} and polio virus~\citep{?}, the existence of mutations selectable by broad antibodies does not necessarily imply that influenza virus can escape these antibodies with ease.
-The fundamental problem is that none of these studies provide quantitative data on the ease of viral escape that can be compared across antibodies in an apples-to-apples fashion.
+However, these studies did not select for actual antibody escape, so it is difficult to relate their measurements to the virus's evolutionary capacity under immune selection.
+Other work has selected antigenic mutants with a few broad antibodies~\citep{yoshida2009cross,chai2016two}, demonstrating that these epitopes are not entirely refractory to change.
+But given that antibodies can select some antigenic variation even in measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic}, the existence of mutations selectable by broad antibodies does not necessarily imply that influenza virus can escape these antibodies with ease.
+The fundamental problem is that existing studies do not quantify the ease of viral escape in a way that can be compared across antibodies in an apples-to-apples fashion.
 
-Here we systematically quantify the ease with which influenza virus can escape from a range of broad and narrow antibodies via single amino-acid mutations.
-Specifically, we impose antibody selection on libraries of viruses carrying all amino-acid mutants of hemagglutinin compatible with viral replication, and then measure the fraction of virions with each mutation that survive the selection at several antibody concentrations.
+Here we systematically quantify the ease with which influenza virus can escape from several broad and narrow antibodies via single amino-acid mutations.
+Specifically, we impose antibody selection on libraries of viruses carrying all amino-acid mutants of hemagglutinin compatible with viral replication, and then measure the fraction of virions with each mutation that survive at several antibody concentrations.
 Critically, this approach quantifies the \emph{magnitude} of the antigenic effect of every mutation in a way that can be directly compared across antibodies.
-We find that even the broadest antibodies can select mutations that increase the virus's resistance to neutralization.
+We find that even the broadest antibodies select antigenic mutations.
 However, the magnitudes of the antigenic effects vary greatly across antibodies.
-Single mutations can make the virus completely resistant to both narrow strain-specific antibodies and a broad antibody against the receptor-binding pocket.
+Single mutations make the virus completely resistant to both narrow strain-specific antibodies and a broad antibody against the receptor-binding pocket.
 But no single mutation does more than modestly increase the virus's resistance to broad antibodies against the hemagglutinin stalk.
-These results suggest that systematically quantifying the antigenic effects of all single amino-acid mutations is a useful complement to simply measuring antibody breadth when trying to assess the potential for antigenic drift.
+Therefore, broad anti-stalk antibodies are quantifiably more resistant to viral escape via single amino-acid mutations than the other antibodies tested here.
 

### 2017-10-10 paper/manuscript.tex
more tweaks to Intro

@@
 
-\title{Quantifying the ease of viral escape from broad and narrow antibodies to influenza hemagglutinin}
+\title{Quantifying the ease of viral escape from broad and narrow antibodies to an H1 influenza hemagglutinin}
 
@@
 However, rare antibodies broadly neutralize many viral strains.
-It is unclear how easily influenza virus might escape such broad antibodies if they became widespread due to therapeutic use or vaccination.
-Here we map all single amino-acid mutations that increase resistance to broad antibodies targeting hemagglutinin.
+It is unclear how easily influenza virus might escape such antibodies if they became widespread due to therapeutic use or vaccination.
+Here we map all single amino-acid mutations that increase resistance to broad antibodies targeting an H1 hemagglutinin.
 Crucially, our approach not only identifies antigenic mutations but also quantifies their effect sizes.
@@
 The virus can escape a broad antibody targeting hemagglutinin's receptor-binding pocket the same way it escapes narrow strain-specific antibodies: via single mutations with huge effects.   
-In contrast, broad antibodies targeting hemagglutinin's stalk only select mutations with small antigenic effects. 
+In contrast, broad antibodies targeting hemagglutinin's stalk only select mutations with small effects. 
 Therefore, antibody breadth is not necessarily an indicator of the difficulty of viral escape.
@@
 For instance, although both measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic,drexler2014robustness} exhibit antigenic variation, the magnitude of this variation is small.
-Therefore, immunity to both these viruses is lifelong~\citep{panum1847iagttagelser,salk1984one}.
+Therefore, immunity to these viruses is lifelong~\citep{panum1847iagttagelser,salk1984one}.
 In contrast, human influenza virus exhibits much more antigenic variation.
@@
 If these efforts succeed, the epitopes of broad antibodies will be under strong antigenic selection in human influenza virus.
-Might such selection then drive extensive antigenic variation in these epitopes?
-There is precedent for the idea that the immune status of the host population can shape influenza virus evolution: the virus undergoes faster antigenic drift in long-lived humans that accumulate immune memory than in short-lived swine that are mostly naive~\citep{sheerar1989antigenic,luoh1992hemagglutinin}, and poultry vaccination may accelerate antigenic drift of avian influenza~\citep{lee2004effect}.
-But alternatively, perhaps broad antibodies are broad because the virus has difficulty escaping them regardless of the selection from host immunity.
+Might such selection then drive antigenic variation in these epitopes?
+There is precedent for the idea that the immune status of the host population can shape influenza virus evolution: the virus undergoes faster antigenic drift in long-lived humans that accumulate immune memory than in short-lived swine that are mostly naive~\citep{sheerar1989antigenic,luoh1992hemagglutinin}, and poultry vaccination may accelerate antigenic drift of avian influenza~\citep{lee2004effect,cattoli2011antigenic}.
+But alternatively, perhaps broad antibodies are broad because the virus has difficulty escaping them regardless of selection from host immunity.
 
 So far, there is limited data to distinguish between these possibilities.
-Several studies have shown that the globular head domain of hemagglutinin is less mutationally tolerant than the stalk domain where many broad antibodies bind~\citep{thyagarajan2014inherent,wu2014high,heaton2013genome}.
+Several studies have shown that the head domain of hemagglutinin is less mutationally tolerant than the stalk domain where many broad antibodies bind~\citep{thyagarajan2014inherent,wu2014high,heaton2013genome}.
 However, these studies did not select for actual antibody escape, so it is difficult to relate their measurements to the virus's evolutionary capacity under immune selection.
-Other work has selected antigenic mutants with a few broad antibodies~\citep{yoshida2009cross,chai2016two}, demonstrating that these epitopes are not entirely refractory to change.
-But given that antibodies can select some antigenic variation even in measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic}, the existence of mutations selectable by broad antibodies does not necessarily imply that influenza virus can escape these antibodies with ease.
+Other work has shown that it is possible to select antigenic mutants with broad antibodies~\citep{yoshida2009cross,chai2016two}, demonstrating that these epitopes are not entirely refractory to change.
+But given that antibodies can select some antigenic variation even in measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic}, the existence of selectable mutations does not necessarily imply that influenza virus can escape broad antibodies with the same ease that it drifts away from narrow strain-specific ones.
 The fundamental problem is that existing studies do not quantify the ease of viral escape in a way that can be compared across antibodies in an apples-to-apples fashion.
 
-Here we systematically quantify the ease with which influenza virus can escape from several broad and narrow antibodies via single amino-acid mutations.
-Specifically, we impose antibody selection on libraries of viruses carrying all amino-acid mutants of hemagglutinin compatible with viral replication, and then measure the fraction of virions with each mutation that survive at several antibody concentrations.
-Critically, this approach quantifies the \emph{magnitude} of the antigenic effect of every mutation in a way that can be directly compared across antibodies.
+Here we systematically quantify the results of selecting all single amino-acid mutants of an H1 hemagglutinin with several broad and narrow antibodies.
+Critically, our approach quantifies the \emph{magnitude} of the antigenic effect of every mutation in a way that can be directly compared across antibodies.
 We find that even the broadest antibodies select antigenic mutations.

### 2017-10-10 paper/manuscript.tex
bit of progress on first results section

@@
 But given that antibodies can select some antigenic variation even in measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic}, the existence of selectable mutations does not necessarily imply that influenza virus can escape broad antibodies with the same ease that it drifts away from narrow strain-specific ones.
-The fundamental problem is that existing studies do not quantify the ease of viral escape in a way that can be compared across antibodies in an apples-to-apples fashion.
+The fundamental problem is that existing studies have not quantified the ease of viral escape in a way that can be compared across antibodies in an apples-to-apples fashion.
 
@@
 
-\subsection*{A method to quantify the ease of escape.}
+\subsection*{An approach to quantify the fraction of virions with each mutation that escape antibody neutralization}
+A feature of influenza antigenic drift is that single amino-acid mutations often have very large effects on neutralization by monoclonal antibodies targeting hemagglutinin's globular head~\citep{?}.
+We can visualize the outcome of antibody selection on viral populations containing such strong escape mutations in several ways.
+For instance, Figure~\ref{fig:fracsurvive_example} shows hypothetical data for four viral variants (wildtype, and mutants V1K, L2C, and S3A).
+If mutation V1K strongly escapes neutralization, then most variants with this mutation will evade antibody neutralization at a concentration where the other three variants are mostly neutralized (Figure~\ref{fig:fracsurvive_example}A).
+This escape is manifested by a large rightward shift in the neutralization curve for the V1K variant (Figure~\ref{fig:fracsurvive_example}B).
+If we imagine drawing vertical lines through the overlaid neutralization curves for all variants, we can calculate the fraction of variants with each mutation that escape neutralization.
+These fractions can be represented using logo plots as in Figure~\ref{fig:fracsurvive_example}C -- if a mutation is a strong escape mutation, then a large fraction of variants with that mutation will survive at antibody concentrations that eliminate most other variants.
+
 \comment{Jesse writes a draft of this section.}

### 2017-10-11 paper/manuscript.tex
some progress on first results section

--- 2017-10-10 paper/manuscript.tex
@@
 One possible reason that viruses exhibit different amounts of antigenic variation is that they have disparate evolutionary capacities to escape the immunodominant antibodies generated by natural immune responses~\citep{lipsitch2007patterns,cobey2014pathogen,fulton2015mutational}.
-According to this explanation, human influenza virus undergoes rapid antigenic drift because most neutralizing antibodies target epitopes on the viral hemagglutinin protein that are highly tolerant of mutational change.
+According to this explanation, human influenza virus undergoes rapid antigenic drift because most neutralizing antibodies target epitopes on the viral hemagglutinin (HA) protein that are highly tolerant of mutational change.
 This explanation is supported by classic experiments showing that it is easy to select viral mutants that escape most antibodies~\citep{yewdell1979antigenic,webster1980determination}, as well as by the observation that mutations that alter antigenicity arise frequently during influenza's evolution globally~\citep{koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction} and within individual humans with long-term infections~\citep{xue2017parallel}.
@@
 Verifying this corollary has become of practical importance with the discovery of broadly neutralizing antibodies against influenza virus.
-These antibodies typically target conserved epitopes in hemagglutinin's stalk~\citep{sui2009structural,ekiert2009antibody,corti2011neutralizing} or receptor-binding pocket~\citep{lee2012heterosubtypic,ekiert2012cross}, and neutralize a wide range of viral strains.
-Broad antibodies usually make only a modest contribution to natural human immunity, since they are less abundant than antibodies to antigenically variable epitopes on the head of hemagglutinin~\citep{?}.
+These antibodies typically target conserved epitopes in HA's stalk~\citep{sui2009structural,ekiert2009antibody,corti2011neutralizing} or receptor-binding pocket~\citep{lee2012heterosubtypic,ekiert2012cross}, and neutralize a wide range of viral strains.
+Broad antibodies usually make only a modest contribution to natural human immunity, since they are less abundant than antibodies to antigenically variable epitopes on the head of HA~\citep{?}.
 However, major efforts are underway to elicit broad antibodies by vaccination or administer them directly as therapeutics~\citep{krammer2015advances,corti2017tackling}.
@@
 So far, there is limited data to distinguish between these possibilities.
-Several studies have shown that the head domain of hemagglutinin is less mutationally tolerant than the stalk domain where many broad antibodies bind~\citep{thyagarajan2014inherent,wu2014high,heaton2013genome}.
+Several studies have shown that the head domain of HA is less mutationally tolerant than the stalk domain where many broad antibodies bind~\citep{thyagarajan2014inherent,wu2014high,heaton2013genome}.
 However, these studies did not select for actual antibody escape, so it is difficult to relate their measurements to the virus's evolutionary capacity under immune selection.
@@
 But given that antibodies can select some antigenic variation even in measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic}, the existence of selectable mutations does not necessarily imply that influenza virus can escape broad antibodies with the same ease that it drifts away from narrow strain-specific ones.
-The fundamental problem is that existing studies do not quantify the ease of viral escape in a way that can be compared across antibodies in an apples-to-apples fashion.
+The fundamental problem is that existing studies have not quantified the ease of viral escape in a way that can be compared across antibodies in an apples-to-apples fashion.
 
-Here we systematically quantify the results of selecting all single amino-acid mutants of an H1 hemagglutinin with several broad and narrow antibodies.
+Here we systematically quantify the results of selecting all single amino-acid mutants of an H1 HA with several broad and narrow antibodies.
 Critically, our approach quantifies the \emph{magnitude} of the antigenic effect of every mutation in a way that can be directly compared across antibodies.
@@
 Single mutations make the virus completely resistant to both narrow strain-specific antibodies and a broad antibody against the receptor-binding pocket.
-But no single mutation does more than modestly increase the virus's resistance to broad antibodies against the hemagglutinin stalk.
+But no single mutation does more than modestly increase the virus's resistance to broad antibodies against the HA stalk.
 Therefore, broad anti-stalk antibodies are quantifiably more resistant to viral escape via single amino-acid mutations than the other antibodies tested here.
-
-%Broadly neutralizing antibodies against influenza virus are of great interest... \cite{corti2017tackling}.
 
@@
 
-\subsection*{A method to quantify the ease of escape.}
-\comment{Jesse writes a draft of this section.}
+\subsection*{An approach to quantify the fraction of virions with each mutation that escape antibody neutralization}
+We can visualize the outcome of antibody selection on viral populations containing antigenic mutations as in Figure~\ref{fig:fracsurvive_example}.
+Specifically, this figure shows hypothetical data for four viral variants.
+If a mutation strongly escapes neutralization, then all virions with this mutation will survive antibody treatment at a concentration where other virions are mostly neutralized (Figure~\ref{fig:fracsurvive_example}A).
+This escape is manifested by a large rightward shift in the neutralization curve for the mutant (Figure~\ref{fig:fracsurvive_example}B).
+If we draw vertical lines through the overlaid neutralization curves, we can calculate the fraction of virions with each mutation that escape neutralization at each antibody concentration (Figure~\ref{fig:fracsurvive_example}B).
+These fractions can be represented using logo plots, where the height of each letter is proportional to the fraction of virions with that amino acid at a site that survive antibody treatment (Figure~\ref{fig:fracsurvive_example}C).
+In these logo plots, large letters correspond to strong escape mutations.
+
 \begin{figure}
@@
 
-\subsection*{Broad and narrow antibodies that neutralize influenza virus hemagglutinin}
+Now consider the case where a mutation has only a small antigenic effect, and so only slightly increases the fraction of virions that escape neutralization (Figure~\ref{fig:fracsurvive_example}D).
+In this scenario, the neutralization curve is only slightly shifted to the right (Figure~\ref{fig:fracsurvive_example}E).
+In the logo plot representation, the antigenic mutation is only slightly larger than other amino acids (Figure~\ref{fig:fracsurvive_example}F), since possessing the mutation only modestly increases the chance that a virion survives antibody treatment.
+
+Our goal is to determine the fraction of mutant virions that escape antibody neutralization for \emph{all} mutations to influenza HA.
+In other words, we want to generate data like that in Figure~\ref{fig:fracsurvive_example}C,F for every amino acid at every site in HA.
+One way to do this would be to measure individual neutralization curves for the wildtype virus and viruses carrying each of the $19\times565 = 10,735$ amino-acid mutations that can be made to the 565-residue HA protein.
+However, individually creating and assaying that many viral mutants would be extraordinarily time-consuming and expensive.
+
+\subsection*{Broad and narrow antibodies that neutralize influenza virus}
 \comment{Juhye writes a draft of this section.}

### 2017-10-11 paper/manuscript.tex
full draft of first results ection

@@
 We can visualize the outcome of antibody selection on viral populations containing antigenic mutations as in Figure~\ref{fig:fracsurvive_example}.
-Specifically, this figure shows hypothetical data for four viral variants.
+This figure shows hypothetical data for four viral variants.
 If a mutation strongly escapes neutralization, then all virions with this mutation will survive antibody treatment at a concentration where other virions are mostly neutralized (Figure~\ref{fig:fracsurvive_example}A).
@@
 If we draw vertical lines through the overlaid neutralization curves, we can calculate the fraction of virions with each mutation that escape neutralization at each antibody concentration (Figure~\ref{fig:fracsurvive_example}B).
-These fractions can be represented using logo plots, where the height of each letter is proportional to the fraction of virions with that amino acid at a site that survive antibody treatment (Figure~\ref{fig:fracsurvive_example}C).
+These fractions can be represented using logo plots, where the height of each letter is proportional to the fraction of virions with that amino acid at a site that survive the antibody (Figure~\ref{fig:fracsurvive_example}C).
 In these logo plots, large letters correspond to strong escape mutations.
@@
 \caption{\label{fig:fracsurvive_example}
-CAPTION
+{\bf Quantifying the fraction of virions with each mutation that escape antibody neutralization.}
+This figure shows hypothetical data for four viral variants: wildtype and three mutants.
+(A) Virions with the V1K mutation (orange) completely survive an antibody concentration where most other virions are neutralized.
+(B) This resistance is manifested by a large shift in V1K's neutralization curve.
+(C) For each dotted vertical line drawn through the neutralization curves in (B), we calculate the fraction of virions with that mutation that survive the antibody, and indicate this fraction by the height of the letter corresponding to that amino acid at that site.
+(D-F) Similar data to the first three panels, but now V1K has only a small antigenic effect, and so only modestly increases the fraction of virions that survive antibody treatment.
 }
@@
 In the logo plot representation, the antigenic mutation is only slightly larger than other amino acids (Figure~\ref{fig:fracsurvive_example}F), since possessing the mutation only modestly increases the chance that a virion survives antibody treatment.
+These logo plots therefore provide a way to both identify escape mutations and quantify the magnitudes of their antigenic effects in a way that is directly comparable across antibodies.
 
 Our goal is to determine the fraction of mutant virions that escape antibody neutralization for \emph{all} mutations to influenza HA.
-In other words, we want to generate data like that in Figure~\ref{fig:fracsurvive_example}C,F for every amino acid at every site in HA.
-One way to do this would be to measure individual neutralization curves for the wildtype virus and viruses carrying each of the $19\times565 = 10,735$ amino-acid mutations that can be made to the 565-residue HA protein.
-However, individually creating and assaying that many viral mutants would be extraordinarily time-consuming and expensive.
+In other words, we want to generate logo plots like those in Figure~\ref{fig:fracsurvive_example} for every amino acid at every site in HA.
+One way to do this would be to measure individual neutralization curves for each of the $19\times565 = 10,735$ single amino-acid mutants of the 565-residue HA protein.
+However, individually creating and assaying that many viral mutants would be time-consuming and expensive.
+Fortunately, we have recently shown that antibody selection on all viral mutations can be assayed in a single experiment using an approach termed mutational antigenic profiling~\citep{doud2017complete,dingens2017comprehensive}.
+This approach involves generating viral libraries containing all mutations to the protein of interest (in this case, HA), selecting these viruses with or without antibody, and then using an accurate deep-sequencing method to determine the relative frequencies of each mutation.
+
+These frequencies can be analyzed to calculate the fraction of variants with each mutation that survive antibody treatment.
+Specifically, the deep sequencing determines the frequencies of virions carrying amino-acid $a$ at site $r$ in the antibody-selected and mock-selected conditions, which we denote as $\rho_{r,a}^{\rm{selected}}$ and $\rho_{r,a}^{\rm{mock}}$, respectively.
+We can also measure the total fraction of the viral library that survives the antibody treatment, which we denote as $\gamma$.
+The fraction of variants with amino-acid $a$ at site $r$ that survive selection is then simply 
+\begin{equation}
+\label{eq:fracsurvive}
+F_{r,a} = \gamma \frac{\rho_{r,a}^{\rm{selected}}}{\rho_{r,a}^{\rm{mock}}}.
+\end{equation}
+For instance, in Figure~\ref{fig:fracsurvive_example}A, the frequency of virions with the orange mutation is $\rho_{r,a}^{\rm{selected}} = \frac{4}{7}$ in the antibody-selected condition and $\rho_{r,a}^{\rm{mock}} = \frac{4}{16}$ in the mock-selected condition.
+The overall fraction of virions that survive the antibody in Figure~\ref{fig:fracsurvive_example}A is $\gamma = \frac{7}{16}$.
+Therefore, we use Equation~\ref{eq:fracsurvive} to calculate that the fraction of variants with the orange mutation that survive the selection is $F_{r,a} = \frac{7}{16} \times \frac{4/7}{4/16} = 1$.
+Performing the analogous calculation for Figure~\ref{fig:fracsurvive_example}D correctly determines that fraction of virions with the orange mutation that survive the antibody is only 0.5 for the scenario in that figure panel.
+In most of the analyses of real data below, we will plot the fraction surviving \emph{above} the overall library average (this is $\max\left[0, F_{a,r} - \gamma\right]$) since this quantity more clearly accentuates antigenic mutations.
+Details of how the calculations are extended to account for sequencing errors and sampling statistics are in the METHODS.
+Open-source software that performs all steps in the analysis beginning with the deep sequencing data is available at \url{https://jbloomlab.github.io/dms_tools2/}.
 

### 2017-10-11 paper/manuscript.tex
progress on third RESULTS section and part of METHODS

@@
 \usepackage{hyperref}
-\hypersetup{colorlinks,citecolor=blue,linkcolor=darkblue,urlcolor=blue}
+\hypersetup{colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue}
 
@@
 These antibodies typically target conserved epitopes in HA's stalk~\citep{sui2009structural,ekiert2009antibody,corti2011neutralizing} or receptor-binding pocket~\citep{lee2012heterosubtypic,ekiert2012cross}, and neutralize a wide range of viral strains.
-Broad antibodies usually make only a modest contribution to natural human immunity, since they are less abundant than antibodies to antigenically variable epitopes on the head of HA~\citep{?}.
+Broad antibodies are usually less abundant in human serum than antibodies to antigenically variable epitopes on the head of HA~\citep{ellebedy2014induction,andrews2015immune}.
 However, major efforts are underway to elicit broad antibodies by vaccination or administer them directly as therapeutics~\citep{krammer2015advances,corti2017tackling}.
@@
 
-Here we systematically quantify the results of selecting all single amino-acid mutants of an H1 HA with several broad and narrow antibodies.
+Here we systematically quantify the results of selecting all single amino-acid mutations to an H1 HA with several broad and narrow antibodies.
 Critically, our approach quantifies the \emph{magnitude} of the antigenic effect of every mutation in a way that can be directly compared across antibodies.
@@
 \section*{RESULTS}
+\label{sec:results}
 
@@
 We can visualize the outcome of antibody selection on viral populations containing antigenic mutations as in Figure~\ref{fig:fracsurvive_example}.
-This figure shows hypothetical data for four viral variants.
 If a mutation strongly escapes neutralization, then all virions with this mutation will survive antibody treatment at a concentration where other virions are mostly neutralized (Figure~\ref{fig:fracsurvive_example}A).
@@
 
-Now consider the case where a mutation has only a small antigenic effect, and so only slightly increases the fraction of virions that escape neutralization (Figure~\ref{fig:fracsurvive_example}D).
-In this scenario, the neutralization curve is only slightly shifted to the right (Figure~\ref{fig:fracsurvive_example}E).
+Now consider the case where a mutation has just a small antigenic effect, and so only slightly increases the fraction of virions that escape neutralization (Figure~\ref{fig:fracsurvive_example}D).
+In this scenario, the neutralization curve shifts only slightly to the right (Figure~\ref{fig:fracsurvive_example}E).
 In the logo plot representation, the antigenic mutation is only slightly larger than other amino acids (Figure~\ref{fig:fracsurvive_example}F), since possessing the mutation only modestly increases the chance that a virion survives antibody treatment.
@@
 
-These frequencies can be analyzed to calculate the fraction of variants with each mutation that survive antibody treatment.
+These frequencies can be analyzed to calculate the fraction of virions with each mutation that survive antibody treatment.
 Specifically, the deep sequencing determines the frequencies of virions carrying amino-acid $a$ at site $r$ in the antibody-selected and mock-selected conditions, which we denote as $\rho_{r,a}^{\rm{selected}}$ and $\rho_{r,a}^{\rm{mock}}$, respectively.
 We can also measure the total fraction of the viral library that survives the antibody treatment, which we denote as $\gamma$.
-The fraction of variants with amino-acid $a$ at site $r$ that survive selection is then simply 
+The fraction of variants with amino-acid $a$ at site $r$ that survive antibody selection is then simply 
 \begin{equation}
 \label{eq:fracsurvive}
-F_{r,a} = \gamma \frac{\rho_{r,a}^{\rm{selected}}}{\rho_{r,a}^{\rm{mock}}}.
+F_{r,a} = \gamma \times \frac{\rho_{r,a}^{\rm{selected}}}{\rho_{r,a}^{\rm{mock}}}.
 \end{equation}
@@
 Performing the analogous calculation for Figure~\ref{fig:fracsurvive_example}D correctly determines that fraction of virions with the orange mutation that survive the antibody is only 0.5 for the scenario in that figure panel.
-In most of the analyses of real data below, we will plot the fraction surviving \emph{above} the overall library average (this is $\max\left[0, F_{a,r} - \gamma\right]$) since this quantity more clearly accentuates antigenic mutations.
-Details of how the calculations are extended to account for sequencing errors and sampling statistics are in the METHODS.
+In most of the analyses of real data below, we will plot the fraction surviving \emph{above} the overall library average, which is
+\begin{equation}
+\label{eq:fracsurvive_above}
+F_{r,a}^{\rm{aboveavg}} = \max\left(0, F_{r,a} - \gamma\right).
+\end{equation}
+We plot the fraction surviving above the library average rather than the raw fraction surviving since the former quantity more clearly accentuates antigenic mutations against the background of other mutations.
+Details of how the calculations are extended to account for sequencing errors and sampling statistics are in the \nameref{sec:methods}.
 Open-source software that performs all steps in the analysis beginning with the deep sequencing data is available at \url{https://jbloomlab.github.io/dms_tools2/}.
@@
 
-\subsection*{Ease of escape from each antibody}
+\subsection*{Quantifying the antigenic effects of all mutations selected by each antibody}
+We performed mutational antigenic profiling using the three broadly neutralizing antibodies (FI6v3, C179, and S139/1) at the concentrations indicated in Figure~\ref{fig:neutcurves}, and analyzed the data to calculate the fraction of virions carrying each mutation that survived neutralization.
+All experiments were performed in full biological triplicate using three independently generated virus libraries carrying all amino-acid mutants to the HA of the A/WSN/1933 (H1N1) strain~\citep{doud2016accurate}, and in some cases we also included technical replicates with the same virus library.
+We have previously described~\citep{doud2017complete} complete mapping of escape from the three strain-specific antibodies (H17-L19, H17-L10, and H17-L7) at the concentrations indicated in Figure~\ref{fig:neutcurves}, and we also re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived each selection.
+The correlations among replicates are shown in Figure~\ref{suppfig:corr}.
+For the remainder of this paper, we will refer to the median antigenic effect of each mutation taken across all replicates with that antibody at that concentration.
+
 \comment{Jesse can work on a rough draft of this section.}
-
-\comment{Juhye, can you create a correlation among replicates supplemental figure.
-In the \texttt{./figs/corrs/} directory are all the plots.
-Is it possible to nicely arrange them on one page, maybe getting rid of the diagonals?}
 
@@
 \section*{METHODS}
+\label{sec:methods}
 \subsection*{Antibodies}
@@
 
-\subsection*{Computation of the fraction $\phi_{r,a}$ of each mutation that escapes antibody neutralization}
-\comment{depending on how much of paper focus is on this new metric, some of this rationale might be better placed in the results section}
+\subsection*{Calculating the fraction of virions with each mutation that escapes antibody neutralization}
+In prior mutational antigenic profiling work~\citep{doud2017complete,dingens2017comprehensive}, we calculated the differential selection on each mutation as the logarithm of its enrichment relative to wildtype in an antibody-selected sample versus a mock-selected control.
+These \emph{mutation differential selection} values are useful for the analysis of individual experiments.
+However, there is no natural way to compare these values across experiments with different antibodies at different concentrations, since the strength of differential selection depends on details of how the pressure is imposed.
+We therefore developed the new approach in this paper to quantify the antigenic effect of a mutation in units that can be compared across antibodies and concentrations.
 
-Mutation differential selection values $s_{r,a}$ reflect the enrichment of a mutation in an antibody-selected sample relative to a mock-selected control. 
-The extent of these mutation enrichments are dependent on the stringency of the neutralization: as the concentration of an antibody used to neutralize the mutant virus library is increased, a larger proportion of the virus library is neutralized, and escape mutants become more enriched in the antibody-selected sample relative to the control sample~\cite{doud2017complete}.
-Consequently, due to differences in the concentrations and potencies of various antibodies used in mutational antigenic profiling, quantitative comparisons of the effects of mutations on neutralization escape between various antibodies is confounded by differences in neutralization stringency across experiments.
-
-We defined a new metric that accounts for the neutralization stringency of each experiment so that quantitative comparisons can be made between experiments with varying neutralization stringencies.
-This new measure, $\phi_{r,a}$, is roughly analogous to the fraction of mutant viruses carrying mutation $a$ at site $r$ that escape neutralization by a given antibody.
-The computation of $\phi_{r,a}$ incorporates information about the stringency of neutralization. 
-We define this as  $\gamma$ = (100 - \% neutralization)/100), the fraction of library remaining infectious after antibody neutralization measured by qRT-PCR for each antibody-selection experiment as previously described~\cite{doud2017complete}.
-
-Define $\hat{n}_{r,a}^{sel}$ and $\hat{n}_{r,a}^{mock}$ as the number of error-corrected counts of amino-acid mutant $a$ at site $r$ as defined in~\cite{doud2017complete}, and $N_r^{sel}$ and $N_r^{mock}$ be the total counts at site $r$, as specified for either an antibody selection or matched mock condition.
-Let $P$ be a pseudocount (default to 5 in these analyses) added to amino-acid counts, and let $f_{r}^{sel}$ and $f_{r}^{mock}$ be relative depths of the selected and mock samples at site $r$ for the purposes of scaling the pseudocount by relative read depth as previously described~\cite{doud2017complete}.
-
-Define the pseudocount-adjusted frequency $\rho$ of mutant $a$ at site $r$ in selected and mock samples to be:
-$$\rho_{r,a}^{sel} = \frac{n_{r,a}^{sel}+f_r^{sel}\times P}{N_r^{sel}+f_r^{sel}\times  P}$$
-$$\rho_{r,a}^{mock} = \frac{n_{r,a}^{mock}+f_r^{mock}\times P}{N_r^{mock}+f_r^{mock}\times  P}$$
-
-We define $\phi_{r,a}$ as the net `fraction' of mutant $a$ at site $r$ escaping neutralization (above the average escape fraction $\gamma$), and compute this as:
-$$\phi_{r,a} = \frac{\gamma \times \rho_{r,a}^{sel}}{\rho_{r,a}^{mock}} - \gamma$$
-
-By this definition, mutations with no effect on antibody escape will have $\phi_{r,a} =0$, and mutations completely escaping antibody will have $\phi_{r,a} \approx 1$.
-Care should be taken in interpreting the absolute values of $\phi$, however, since the choice of pseudocount affects the range of $\phi$ values (but not, importantly, the relationships of $\phi$ distributions across antibodies).
-We have found that a value of $P=5$ pseudocounts results in $\phi$ values that are roughly bounded from 0 to 1, and consistently use $P=5$ all analyses.
+The general principle of the calculations is illustrated in Figure~\ref{fig:fracsurvive_example} and discussed in the first section of the \nameref{sec:results}.
+Code that performs these calculations is included in the \texttt{dms\_tools2} software package~\citep{bloom2015software} which is available at \url{http://jbloomlab.github.io/dms_tools2}.
 
 \subsection*{Data availability and source code}
-Deep sequencing data has been deposited at the Sequence Read Archive under BioSample accession (\comment{should probably deposit all the new data into the existing biosample SAMN05789126 (Sample name: WSN/1933 HA mutant libraries selected by monoclonal antibodies), which is under bioproject PRJNA309339}).
+Deep sequencing data are available from the Sequence Read Archive under BioSample accession (\comment{add this}).
 
@@
 \FloatBarrier
+
+\begin{suppfigure}
+\centerline{\includegraphics[width=0.6\textwidth]{figs/corrs/site_correlations_bnAbs.pdf}}
+\vspace{0.2in}
+
+\centerline{\includegraphics[width=0.5\textwidth]{figs/corrs/site_correlations_H17LAbs.pdf}}
+\caption{\label{suppfig:corr}
+Correlation of measurements across replicates. 
+Each point represents one site in HA, and gives the fraction of mutant virions surviving selection taken across all amino-acid mutations at that site.
+\comment{Elaborate after writing formula in METHODS.}
+}
+\end{suppfigure}
 

### 2017-10-12 paper/manuscript.tex
added more Methods

@@
 
-\title{Quantifying the ease of viral escape from broad and narrow antibodies to an H1 influenza hemagglutinin}
+\title{Quantifying the ease of viral escape from broad and narrow antibodies to influenza hemagglutinin}
 
@@
 Several studies have shown that the head domain of HA is less mutationally tolerant than the stalk domain where many broad antibodies bind~\citep{thyagarajan2014inherent,wu2014high,heaton2013genome}.
-However, these studies did not select for actual antibody escape, so it is difficult to relate their measurements to the virus's evolutionary capacity under immune selection.
+However, these studies did not select for antibody escape, so it is difficult to relate their measurements to the virus's evolutionary capacity under immune selection.
 Other work has shown that it is possible to select antigenic mutants with broad antibodies~\citep{yoshida2009cross,chai2016two}, demonstrating that these epitopes are not entirely refractory to change.
-But given that antibodies can select some antigenic variation even in measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic}, the existence of selectable mutations does not necessarily imply that influenza virus can escape broad antibodies with the same ease that it drifts away from narrow strain-specific ones.
+But given that antibodies can select some antigenic variation even in measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic}, the existence of selectable mutations does not necessarily imply that influenza virus can escape broad antibodies as easily as it drifts away from narrow strain-specific ones.
 The fundamental problem is that existing studies have not quantified the ease of viral escape in a way that can be compared across antibodies in an apples-to-apples fashion.
@@
 We can visualize the outcome of antibody selection on viral populations containing antigenic mutations as in Figure~\ref{fig:fracsurvive_example}.
-If a mutation strongly escapes neutralization, then all virions with this mutation will survive antibody treatment at a concentration where other virions are mostly neutralized (Figure~\ref{fig:fracsurvive_example}A).
-This escape is manifested by a large rightward shift in the neutralization curve for the mutant (Figure~\ref{fig:fracsurvive_example}B).
+If a mutation strongly escapes neutralization, then all virions with this mutation survive antibody treatment at a concentration where other virions are mostly neutralized (Figure~\ref{fig:fracsurvive_example}A).
+This escape is manifested by a large shift in the neutralization curve for the mutant (Figure~\ref{fig:fracsurvive_example}B).
 If we draw vertical lines through the overlaid neutralization curves, we can calculate the fraction of virions with each mutation that escape neutralization at each antibody concentration (Figure~\ref{fig:fracsurvive_example}B).
@@
 Now consider the case where a mutation has just a small antigenic effect, and so only slightly increases the fraction of virions that escape neutralization (Figure~\ref{fig:fracsurvive_example}D).
-In this scenario, the neutralization curve shifts only slightly to the right (Figure~\ref{fig:fracsurvive_example}E).
+In this scenario, the neutralization curve shifts only slightly (Figure~\ref{fig:fracsurvive_example}E).
 In the logo plot representation, the antigenic mutation is only slightly larger than other amino acids (Figure~\ref{fig:fracsurvive_example}F), since possessing the mutation only modestly increases the chance that a virion survives antibody treatment.
@@
 In other words, we want to generate logo plots like those in Figure~\ref{fig:fracsurvive_example} for every amino acid at every site in HA.
-One way to do this would be to measure individual neutralization curves for each of the $19\times565 = 10,735$ single amino-acid mutants of the 565-residue HA protein.
-However, individually creating and assaying that many viral mutants would be time-consuming and expensive.
-Fortunately, we have recently shown that antibody selection on all viral mutations can be assayed in a single experiment using an approach termed mutational antigenic profiling~\citep{doud2017complete,dingens2017comprehensive}.
+One way to do this would be to measure individual neutralization curves for all $19\times565 = 10,735$ single amino-acid mutants of the 565-residue HA protein.
+However, individually creating and assaying that many viral mutants would be exceedingly time-consuming and expensive.
+Fortunately, we have recently shown that antibody selection on all viral mutations can be assayed in a single experiment using mutational antigenic profiling~\citep{doud2017complete,dingens2017comprehensive}.
 This approach involves generating viral libraries containing all mutations to the protein of interest (in this case, HA), selecting these viruses with or without antibody, and then using an accurate deep-sequencing method to determine the relative frequencies of each mutation.
@@
 Performing the analogous calculation for Figure~\ref{fig:fracsurvive_example}D correctly determines that fraction of virions with the orange mutation that survive the antibody is only 0.5 for the scenario in that figure panel.
-In most of the analyses of real data below, we will plot the fraction surviving \emph{above} the overall library average, which is
+In the analyses of real data below, we will plot the fraction surviving \emph{above} the overall library average, which is
 \begin{equation}
@@
 
-\subsection*{Mutant virus selections with antibody}
-Library selections, deep sequencing, and computation of mutation differential selection were performed as previously described~\cite{doud2017complete}. Briefly, ...
+\subsection*{Neutralization assays}
+\comment{add this}
+
+\subsection*{Mutant virus selections with antibody and deep sequencing}
+Library selections, deep sequencing, and computation of mutation differential selection were performed as previously described~\cite{doud2017complete}.
+
+Describe estimation of $\gamma$ by qPCR.
 
@@
 The general principle of the calculations is illustrated in Figure~\ref{fig:fracsurvive_example} and discussed in the first section of the \nameref{sec:results}.
+Here we provide details on how these calculations are performed.
+The deep sequencing measures the number of times that codon $x$ is observed at site $r$ in both the antibody-selected and mock-selected conditions. 
+Denote these counts as $n_{r,x}^{\rm{selected}}$ and $n_{r,x}^{\rm{mock}}$, respectively.
+We also perform deep sequencing of a control (in this case, plasmid DNA encoding the wildtype HA gene) to estimate the sequencing error rate.
+Denote the counts of codon $x$ at site $r$ in this control as $n_{r,x}^{\rm{err}}$.
+Also denote the total reads at each site $r$ in each sample as
+$N_{r}^{\rm{selected}} = \sum_x n_{r,x}^{\rm{selected}}$,
+$N_{r}^{\rm{mock}} = \sum_x n_{r,x}^{\rm{mock}}$, and
+$N_{r}^{\rm{err}} = \sum_x n_{r,x}^{\rm{err}}$.
+
+We first estimate the rate of sequencing errors at site $r$ as
+\begin{equation}
+\label{eq:epsilonrx}
+\epsilon_{r,x} = \frac{n_{r,x}^{\rm{err}}}{N_{r}^{\rm{err}}}.
+\end{equation}
+For the wildtype identity at site $r$, which we denote as $\operatorname{wt}\left(r\right)$, the value of $\epsilon_{r,\operatorname{wt}\left(r\right)}$ is the fraction of times we correctly observe the wildtype identity $\operatorname{wt}\left(r\right)$ at site $r$ versus observing some spurious mutation.
+For all mutant identities $x \ne \operatorname{wt}\left(r\right)$ at site $r$, $\epsilon_{r,x}$ is the fraction of times we observe the mutation $x$ at site $r$ when the identity is really wildtype.
+We ignore second-order terms where we incorrectly read one mutation as another, as such errors will be very rare as mutations themselves are rare (most codons are wildtype in most sequences).
+
+We next adjust all of the deep sequencing codon counts in the antibody-selected and mock-selected conditions by the error control. 
+Specifically, the error-adjusted counts for the antibody-selected sample are
+\begin{equation}
+\label{eq:erroradjust}
+\hat{n}_{r,x} ^{\rm{selected}}= \begin{cases}
+\max\left[N_r^{\rm{selected}} \times \left(\frac{n_{r,x}^{\rm{selected}}}{N_r^{\rm{selected}}} - \epsilon_{r,x}\right), 0\right] & \mbox{if } x \ne \operatorname{wt}\left(r\right) \\
+n_{r,x} / \epsilon_{r,x} & \mbox{if } x = \operatorname{wt}\left(r\right).
+\end{cases}
+\end{equation}
+An equivalent equation is used to calculate $\hat{n}_{r,x} ^{\rm{mock}}$.
+We then sum the error-adjusted codon counts for each amino acid $a$:
+\begin{eqnarray}
+\hat{n}_{r,a}^{\rm{selected}} &=& \sum\limits_{\left\{x \mid \mathcal{A}\left(x\right) = a\right\}} \hat{n}_{r,x}^{\rm{selected}},
+\end{eqnarray}
+so that $\hat{n}_{r,a}^{\rm{selected}}$ are the error-adjusted counts for the antibody-selected condition summed across all codons $x$ where the encoded amino acid $\mathcal{A}\left(x\right)$ is $a$.
+An equivalent equation is used to calculate $\hat{n}_{r,a} ^{\rm{mock}}$.
+
+Finally, we use these error-adjusted amino-acid counts to estimate the mutation frequencies $\rho_{r,a}^{\rm{selected}}$ and $\rho_{r,a}^{\rm{mock}}$ that are used in Equation~\ref{eq:fracsurvive} to calculate the fraction $F_{r,a}$ of virions with amino acid $a$ at site $r$ that survive the selection.
+When estimating these mutation frequencies, we add a pseudocount of $P = 5$ to the lower-depth sample, and a depth-adjusted pseudocount to the higher depth sample.
+The rationale for adding a pseudocount is to regularize the estimates in the case of low counts.
+Depth adjusted...
+
+The experiments also use some independent method (in this case, qPCR) to measure $\gamma$, the total fraction of the virus library that survives the antibody selection.
+
 Code that performs these calculations is included in the \texttt{dms\_tools2} software package~\citep{bloom2015software} which is available at \url{http://jbloomlab.github.io/dms_tools2}.
+Specifically, the analyses described here were performed with the program \texttt{dms2\_batch\_fracsurvive} distributed with version 2.1.0 of \texttt{dms\_tools2}.
 
@@
 \begin{suppfigure}
-\centerline{\includegraphics[width=0.6\textwidth]{figs/corrs/site_correlations_bnAbs.pdf}}
-\vspace{0.2in}
-
-\centerline{\includegraphics[width=0.5\textwidth]{figs/corrs/site_correlations_H17LAbs.pdf}}
+\centerline{\includegraphics[width=\textwidth]{figs/corrs/site_correlations.pdf}}
 \caption{\label{suppfig:corr}

### 2017-10-12 paper/manuscript.tex
added Methods section on computing fracsurvive

@@
 We then sum the error-adjusted codon counts for each amino acid $a$:
-\begin{eqnarray}
-\hat{n}_{r,a}^{\rm{selected}} &=& \sum\limits_{\left\{x \mid \mathcal{A}\left(x\right) = a\right\}} \hat{n}_{r,x}^{\rm{selected}},
-\end{eqnarray}
+\begin{equation}
+\hat{n}_{r,a}^{\rm{selected}} = \sum\limits_{\left\{x \mid \mathcal{A}\left(x\right) = a\right\}} \hat{n}_{r,x}^{\rm{selected}},
+\end{equation}
 so that $\hat{n}_{r,a}^{\rm{selected}}$ are the error-adjusted counts for the antibody-selected condition summed across all codons $x$ where the encoded amino acid $\mathcal{A}\left(x\right)$ is $a$.
@@
 The rationale for adding a pseudocount is to regularize the estimates in the case of low counts.
-Depth adjusted...
+Specifically, we estimate the mutation frequencies as
+\begin{eqnarray}
+\rho_{r,x}^{\rm{selected}} &=& \frac{n_{r,x}^{\rm{selected}} + f_{r, \rm{selected}} \times P}{N_r^{\rm{selected}} + f_{r, \rm{selected}} \times P \times A} \\
+\rho_{r,x}^{\rm{mock}} &=& \frac{n_{r,x}^{\rm{mock}} + f_{r, \rm{mock}} \times P}{N_r^{\rm{mock}} + f_{r, \rm{mock}} \times P \times A} 
+\end{eqnarray}
+where $A$ is the number of characters (e.g., 20 for amino acids), $f_{r, \rm{selected}}$ and $f_{r, \rm{mock}}$ are the pseudocount adjustment factors defined as:
+\begin{eqnarray}
+f_{r, \rm{selected}} &=& \max\left(1, \frac{N_{r}^{\rm{selected}}}{N_{r}^{\rm{mock}}}\right) \\
+f_{r, \rm{mock}} &=& \max\left(1, \frac{N_{r}^{\rm{mock}}}{N_{r}^{\rm{selected}}}\right).
+\end{eqnarray}
+The pseudocount adjustment factors ensure that $P$ is added to the counts for the lower depth sample, and a proportionally scaled-up pseudocount is added to the higher depth sample.
+The depth scaling is necessary to avoid systematically biasing towards higher mutation frequencies in the lower depth sample.
+It is these estimated mutation frequencies that are used in conjunction with $\gamma$ (the qPCR estimated overall of virions that survive selection) to compute the fraction surviving ($F_{r,a}$) and fraction surviving above average ($F_{r,a}^{\rm{aboveavg}}$) via Equations~\ref{eq:fracsurvive} and \ref{eq:fracsurvive_above}.
 
-The experiments also use some independent method (in this case, qPCR) to measure $\gamma$, the total fraction of the virus library that survives the antibody selection.
+In some cases, we need to summarize the fraction of mutations surviving into a single number for each site, such as for plotting as a function of the site number or displaying on the crystal structure.
+There are 19 different $F_{r,a}^{\rm{aboveavg}}$ values for non-wildtype amino acids for each site. 
+One summary statistic is the fraction surviving above the library average \emph{averaged} over all 19 amino-acid mutations at site $r$:
+\begin{equation}
+\label{eq:avgfracsurvive}
+\mathcal{F}_r^{avg} = \frac{1}{19} \sum\limits_{\left\{a \mid a \ne \operatorname{wt}\left(r\right)\right\}} F_{r,a}^{\rm{aboveavg}}.
+\end{equation}
+Another summary statistic is the \emph{maximum} fraction surviving above average among all 19 amino-acid mutations at site $r$:
+\begin{equation}
+\label{eq:maxfracsurvive}
+\mathcal{F}_r^{max} = \frac{1}{19} \max\limits_{\left\{a \mid a \ne \operatorname{wt}\left(r\right)\right\}}\left( F_{r,a}^{\rm{aboveavg}} \right).
+\end{equation}
 
-Code that performs these calculations is included in the \texttt{dms\_tools2} software package~\citep{bloom2015software} which is available at \url{http://jbloomlab.github.io/dms_tools2}.
-Specifically, the analyses described here were performed with the program \texttt{dms2\_batch\_fracsurvive} distributed with version 2.1.0 of \texttt{dms\_tools2}.
+Code that performs all these calculations is included in the \texttt{dms\_tools2} software package~\citep{bloom2015software} which is available at \url{http://jbloomlab.github.io/dms_tools2}.
+The in this paper were performed using version 2.1.0 of \texttt{dms\_tools2}.
 

### 2017-10-12 paper/manuscript.tex
completed legend for corr figure

@@
 \clearpage
+\small
 
@@
 
-\clearpage
-
-\small
 \subsection*{ACKNOWLEDGMENTS}
@@
 \begin{suppfigure}
-\centerline{\includegraphics[width=\textwidth]{figs/corrs/site_correlations.pdf}}
+\centerline{\includegraphics[width=0.84\textwidth]{figs/corrs/site_correlations.pdf}}
 \caption{\label{suppfig:corr}
-Correlation of measurements across replicates. 
-Each point represents one site in HA, and gives the fraction of mutant virions surviving selection taken across all amino-acid mutations at that site.
-\comment{Elaborate after writing formula in METHODS.}
+Correlations across experimental replicates. 
+Each point represents one site in HA, and gives the fraction surviving above average across all amino-acid mutations at that site, as in Equation~\ref{eq:avgfracsurvive}.
+The replicates are highly correlated for antibodies with strong escape mutations (S139/1, H17-L19, H17-L10, and H17-L7), and reasonably correlated for antibodies with only weak escape mutations (FI6v3 and C179).
 }

### 2017-10-13 paper/manuscript.tex
a bit of progress on Results ection 3

@@
 \subsection*{Quantifying the antigenic effects of all mutations selected by each antibody}
-We performed mutational antigenic profiling using the three broadly neutralizing antibodies (FI6v3, C179, and S139/1) at the concentrations indicated in Figure~\ref{fig:neutcurves}, and analyzed the data to calculate the fraction of virions carrying each mutation that survived neutralization.
+We performed mutational antigenic profiling using the three broad antibodies (FI6v3, C179, and S139/1) at the concentrations indicated in Figure~\ref{fig:neutcurves}, and analyzed the data to calculate the fraction of virions carrying each mutation that survived neutralization.
 All experiments were performed in full biological triplicate using three independently generated virus libraries carrying all amino-acid mutants to the HA of the A/WSN/1933 (H1N1) strain~\citep{doud2016accurate}, and in some cases we also included technical replicates with the same virus library.
-We have previously described~\citep{doud2017complete} complete mapping of escape from the three strain-specific antibodies (H17-L19, H17-L10, and H17-L7) at the concentrations indicated in Figure~\ref{fig:neutcurves}, and we also re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived each selection.
-The correlations among replicates are shown in Figure~\ref{suppfig:corr}.
+We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three strain-specific antibodies (H17-L19, H17-L10, and H17-L7) at the concentrations indicated in Figure~\ref{fig:neutcurves}, and we also re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived each selection.
+The correlations among replicates are in Figure~\ref{suppfig:corr}.
 For the remainder of this paper, we will refer to the median antigenic effect of each mutation taken across all replicates with that antibody at that concentration.
-
-\comment{Jesse can work on a rough draft of this section.}
 
@@
 \label{fig:avgfracsurvive}
-CAPTION
+{\bf The strain-specific and broad anti-receptor-binding-pocket antibodies select mutations with large antigenic effects at multiple sites, but the broad anti-stalk antibodies only select mutations with small effects on neutralization.}
+The fraction surviving above average taken across all amino acids at each site in HA.
+Specifically, the panels in this figure show the $\mathcal{F}_r^{avg}$ values calculated using Equation~\ref{eq:avgfracsurvive}.
+The sites are labeled in H3 numbering.
 Figure~\ref{suppfig:maxfracsurvive} shows the fraction surviving for the single most strongly selected amino-acid mutation at each site.
@@
 \end{figure}
+
+It is immediately obvious that both the three strain-specific antibodies and the antibody targeting HA's receptor-binding pocket (S139/1) select antigenic mutations with large effects at multiple sites.
+Specifically, Figure~\ref{fig:avgfracsurvive} shows that for all four of these antibodies, there are sites where a substantial fraction of virions with any mutation survive antibody concentrations that are far in excess of those needed to nearly complete neutralize the wildtype virus.
 
@@
 Describe estimation of $\gamma$ by qPCR.
+
+\subsection*{H3 sequence numbering}
+\comment{add this}
 

### 2017-10-13 paper/manuscript.tex
Start writing preliminary draft of  2nd results subsection

@@
 % blurbs about the specific antibodies we use here:
-C179 isolation and escape mutant selection \cite{okuno1993common}.
-C179 structure with 1957 H2, binding data to many strains, and citations to older studies demonstrating C179 cross-neutralization to H1, H2, H5, H6, and H9:  \cite{dreyfus2013structure}.
-FI6v3 was isolated... \cite{corti2011neutralizing}.
-S139/1 isolation and selection of escape mutants from Aichi H3, Adachi H2, and WSN H1: \cite{yoshida2009cross}.
-S139/1 structure with Victoria75 H3 and binding/neutralization data: \cite{lee2012heterosubtypic}.
+To quantify escape from broadly-neutralizing antibodies using the approach described above, we profiled an antibody targeting the receptor-binding site (S139/1)~\citep{yoshida2009cross, lee2012heterosubtypic}, and two antibodies targeting the stalk domain (FI6v3 and C179)~\citep{corti2011neutralizing, okuno1993common, dreyfus2013structure}.
+In addition, we included analyses of selections we previously completed with three strain-specific antibodies (H17-L19, H17-L10, and H17-L7)~\citep{doud2017complete}. 
+H17-L19, H17-L10, and H17-L7 target the Ca2, Ca1, and Cb antigenic regions, respectively.
+The crystal structures of the broadly-neutralizing antibodies and the binding footprints of the strain-specific antibodies are shown in Figure~\ref{fig:antibody_summary}A.
+S139/1, C179, and FI6v3 also possess a wide range of breadths (Figure~\ref{fig:antibody_summary}B). 
+For instance, S139/1 and FI6v3 can bind various HA's from group 1 and group 2, while C179 exhibits binding only to certain group 1 HA's.
+We were therefore able to profile the ease of escape from antibodies of different specificities and breadths.
+
+In order to compare the potencies of these antibodies, we completed GFP-based neutralization assays using wild-type A/WSN/1933 (H1N1) virus. 
+Figure~\ref{fig:neutcurves} shows the neutralization curves for each antibody, and all six antibodies show similar potencies.
+
+% (selection of escape mutants from S139/1 and C179 have been reported, but no reported escape mutants from FI6v3)
 

### 2017-10-13 paper/manuscript.tex
labeled all fracsurvive line and logo plots

@@
 
-Our goal is to determine the fraction of mutant virions that escape antibody neutralization for \emph{all} mutations to influenza HA.
-In other words, we want to generate logo plots like those in Figure~\ref{fig:fracsurvive_example} for every amino acid at every site in HA.
-One way to do this would be to measure individual neutralization curves for all $19\times565 = 10,735$ single amino-acid mutants of the 565-residue HA protein.
-However, individually creating and assaying that many viral mutants would be exceedingly time-consuming and expensive.
-Fortunately, we have recently shown that antibody selection on all viral mutations can be assayed in a single experiment using mutational antigenic profiling~\citep{doud2017complete,dingens2017comprehensive}.
-This approach involves generating viral libraries containing all mutations to the protein of interest (in this case, HA), selecting these viruses with or without antibody, and then using an accurate deep-sequencing method to determine the relative frequencies of each mutation.
+Our goal is to determine the fraction of mutant virions that escape antibody neutralization for \emph{all} mutations to HA.
+One way to do this would be to measure individual neutralization curves for each of the $19\times565 = 10,735$ single amino-acid mutants of the 565-residue HA protein.
+However, individually creating and assaying that many mutants would be exceedingly time-consuming and expensive.
+Fortunately, we have shown that antibody selection on all viral mutations can be assayed in a single experiment using mutational antigenic profiling~\citep{doud2017complete,dingens2017comprehensive}.
+This approach involves generating viral libraries containing all mutations to the protein of interest, selecting these viruses with or without antibody, and using an accurate deep-sequencing method to determine the relative frequencies of each mutation.
 
@@
 Specifically, the deep sequencing determines the frequencies of virions carrying amino-acid $a$ at site $r$ in the antibody-selected and mock-selected conditions, which we denote as $\rho_{r,a}^{\rm{selected}}$ and $\rho_{r,a}^{\rm{mock}}$, respectively.
-We can also measure the total fraction of the viral library that survives the antibody treatment, which we denote as $\gamma$.
+We can also measure the total fraction of the viral library that survives the antibody, which we denote as $\gamma$.
 The fraction of variants with amino-acid $a$ at site $r$ that survive antibody selection is then simply 
@@
 \end{equation}
-For instance, in Figure~\ref{fig:fracsurvive_example}A, the frequency of virions with the orange mutation is $\rho_{r,a}^{\rm{selected}} = \frac{4}{7}$ in the antibody-selected condition and $\rho_{r,a}^{\rm{mock}} = \frac{4}{16}$ in the mock-selected condition.
+For instance, in Figure~\ref{fig:fracsurvive_example}A, the frequency of virions with the orange mutation is $\rho_{r,a}^{\rm{selected}} = \frac{4}{7}$ in the antibody selection and $\rho_{r,a}^{\rm{mock}} = \frac{4}{16}$ in the mock selection.
 The overall fraction of virions that survive the antibody in Figure~\ref{fig:fracsurvive_example}A is $\gamma = \frac{7}{16}$.
-Therefore, we use Equation~\ref{eq:fracsurvive} to calculate that the fraction of variants with the orange mutation that survive the selection is $F_{r,a} = \frac{7}{16} \times \frac{4/7}{4/16} = 1$.
+Therefore, we use Equation~\ref{eq:fracsurvive} to calculate that the fraction of variants with the orange mutation that survive is $F_{r,a} = \frac{7}{16} \times \frac{4/7}{4/16} = 1$.
 Performing the analogous calculation for Figure~\ref{fig:fracsurvive_example}D correctly determines that fraction of virions with the orange mutation that survive the antibody is only 0.5 for the scenario in that figure panel.
-In the analyses of real data below, we will plot the fraction surviving \emph{above} the overall library average, which is
+In the analyses of real data below, we will plot the excess fraction surviving \emph{above} the overall library average, which is
 \begin{equation}
-\label{eq:fracsurvive_above}
-F_{r,a}^{\rm{aboveavg}} = \max\left(0, F_{r,a} - \gamma\right).
+\label{eq:fracsurvive_excess}
+F_{r,a}^{\rm{excess}} = \max\left(0, F_{r,a} - \gamma\right).
 \end{equation}
-We plot the fraction surviving above the library average rather than the raw fraction surviving since the former quantity more clearly accentuates antigenic mutations against the background of other mutations.
+We plot the excess fraction surviving above the library average since this quantity more clearly accentuates antigenic mutations against the background of other mutations.
 Details of how the calculations are extended to account for sequencing errors and sampling statistics are in the \nameref{sec:methods}.
@@
 We performed mutational antigenic profiling using the three broad antibodies (FI6v3, C179, and S139/1) at the concentrations indicated in Figure~\ref{fig:neutcurves}, and analyzed the data to calculate the fraction of virions carrying each mutation that survived neutralization.
-All experiments were performed in full biological triplicate using three independently generated virus libraries carrying all amino-acid mutants to the HA of the A/WSN/1933 (H1N1) strain~\citep{doud2016accurate}, and in some cases we also included technical replicates with the same virus library.
+All experiments were performed in full biological triplicate using three independently generated virus libraries carrying all amino-acid mutants to the HA of the A/WSN/1933 (H1N1) strain~\citep{doud2016accurate}.
 We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three strain-specific antibodies (H17-L19, H17-L10, and H17-L7) at the concentrations indicated in Figure~\ref{fig:neutcurves}, and we also re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived each selection.
@@
 \label{fig:avgfracsurvive}
-{\bf The strain-specific and broad anti-receptor-binding-pocket antibodies select mutations with large antigenic effects at multiple sites, but the broad anti-stalk antibodies only select mutations with small effects on neutralization.}
-The fraction surviving above average taken across all amino acids at each site in HA.
-Specifically, the panels in this figure show the $\mathcal{F}_r^{avg}$ values calculated using Equation~\ref{eq:avgfracsurvive}.
-The sites are labeled in H3 numbering.
-Figure~\ref{suppfig:maxfracsurvive} shows the fraction surviving for the single most strongly selected amino-acid mutation at each site.
+{\bf The strain-specific and broad anti-receptor-binding-pocket antibodies select mutations with large antigenic effects, but the broad anti-stalk antibodies only select mutations with small effects.}
+The height of the line at each site indicates the excess fraction of virions with a mutation at that site that survive the antibody selection, averaging across all amino-acid mutations at the site (see Equation~\ref{eq:avgfracsurvive}).
+It is immediately obvious that there are multiple large-effect mutations for H17L19, H17L10, H17L7, and S139/1 -- but not any large effect mutations for FI6v3 and C179.
+Sites are labeled in H3 numbering.
+Figure~\ref{suppfig:maxfracsurvive} shows the excess fraction surviving for the single most strongly selected amino-acid mutation at each site.
+Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo} show data for all mutations for each antibody using logo plots.
 }
@@
 \subsection*{For all antibodies, the selected sites of escape occur in the antibody binding footprints}
+
+\begin{figure}
+\centerline{\includegraphics[width=\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
+\caption{
+\label{fig:structures}
+CAPTION
+}
+\end{figure}
 
@@
 The depth scaling is necessary to avoid systematically biasing towards higher mutation frequencies in the lower depth sample.
-It is these estimated mutation frequencies that are used in conjunction with $\gamma$ (the qPCR estimated overall of virions that survive selection) to compute the fraction surviving ($F_{r,a}$) and fraction surviving above average ($F_{r,a}^{\rm{aboveavg}}$) via Equations~\ref{eq:fracsurvive} and \ref{eq:fracsurvive_above}.
+It is these estimated mutation frequencies that are used in conjunction with $\gamma$ (the qPCR estimated overall of virions that survive selection) to compute the fraction surviving ($F_{r,a}$) and excess fraction surviving above the library average ($F_{r,a}^{\rm{excess}}$) via Equations~\ref{eq:fracsurvive} and \ref{eq:fracsurvive_excess}.
 
-In some cases, we need to summarize the fraction of mutations surviving into a single number for each site, such as for plotting as a function of the site number or displaying on the crystal structure.
-There are 19 different $F_{r,a}^{\rm{aboveavg}}$ values for non-wildtype amino acids for each site. 
+In some cases, we need to summarize the excess fraction of mutations surviving into a single number for each site, such as for plotting as a function of the site number or displaying on the crystal structure.
+There are 19 different $F_{r,a}^{\rm{excess}}$ values for non-wildtype amino acids for each site. 
 One summary statistic is the fraction surviving above the library average \emph{averaged} over all 19 amino-acid mutations at site $r$:
@@
 \label{eq:avgfracsurvive}
-\mathcal{F}_r^{avg} = \frac{1}{19} \sum\limits_{\left\{a \mid a \ne \operatorname{wt}\left(r\right)\right\}} F_{r,a}^{\rm{aboveavg}}.
+\mathcal{F}_r^{avg} = \frac{1}{19} \sum\limits_{\left\{a \mid a \ne \operatorname{wt}\left(r\right)\right\}} F_{r,a}^{\rm{excess}}.
 \end{equation}
@@
 \label{eq:maxfracsurvive}
-\mathcal{F}_r^{max} = \frac{1}{19} \max\limits_{\left\{a \mid a \ne \operatorname{wt}\left(r\right)\right\}}\left( F_{r,a}^{\rm{aboveavg}} \right).
+\mathcal{F}_r^{max} = \frac{1}{19} \max\limits_{\left\{a \mid a \ne \operatorname{wt}\left(r\right)\right\}}\left( F_{r,a}^{\rm{excess}} \right).
 \end{equation}
@@
 \FloatBarrier
+\pagenumbering{arabic}% resets `page` counter to 1
+\renewcommand*{\thepage}{S\arabic{page}}
 
@@
 \caption{\label{suppfig:corr}
-Correlations across experimental replicates. 
-Each point represents one site in HA, and gives the fraction surviving above average across all amino-acid mutations at that site, as in Equation~\ref{eq:avgfracsurvive}.
+{\bf Correlations across experimental replicates.} 
+Each point represents one site in HA, and gives the fraction surviving above average across all amino-acid mutations at that site, as calculated using Equation~\ref{eq:avgfracsurvive}.
 The replicates are highly correlated for antibodies with strong escape mutations (S139/1, H17-L19, H17-L10, and H17-L7), and reasonably correlated for antibodies with only weak escape mutations (FI6v3 and C179).
@@
 \caption{\label{suppfig:maxfracsurvive}
-CAPTION}
+{\bf The excess fraction surviving for the single strongest escape mutation at each site.}
+This plot differs from Figure~\ref{fig:avgfracsurvive} in that the height of the line indicates the excess fraction of virions that survive the antibody selection for the single strongest escape mutation at that site, rather than the average across all amino-acid mutations at that site.
+}
 \end{suppfigure}
@@
 \caption{\label{suppfig:H17L19logo}
-CAPTION}
+{\bf The excess fraction surviving selection with antibody H17L19 for all amino-acid mutations.}
+The excess fraction surviving for each replicate was computed using Equation~\ref{eq:fracsurvive_excess}, then we took the median across all technical and biological replicates for each antibody concentration, and then took the medians of those values across concentrations.
+The height of each letter is proportional to the excess fraction surviving of virions with that mutation.
+The scale bar at the top of the plot relates the letter heights to the actual fractions.
+The sites are labeled using H3 numbering.
+}
 \end{suppfigure}
@@
 \caption{\label{suppfig:H17L10logo}
-CAPTION}
+{\bf The excess fraction surviving selection with antibody H17L10 for all amino-acid mutations.}
+The excess fraction surviving for each replicate was computed using Equation~\ref{eq:fracsurvive_excess}, then we took the median across all technical and biological replicates for each antibody concentration, and then took the medians of those values across concentrations.
+The height of each letter is proportional to the excess fraction surviving of virions with that mutation.
+The scale bar at the top of the plot relates the letter heights to the actual fractions.
+The sites are labeled using H3 numbering.
+}
 \end{suppfigure}
@@
 \caption{\label{suppfig:H17L7logo}
-CAPTION}
+{\bf The excess fraction surviving selection with antibody H17L7 for all amino-acid mutations.}
+The excess fraction surviving for each replicate was computed using Equation~\ref{eq:fracsurvive_excess}, then we took the median across all technical and biological replicates for each antibody concentration, and then took the medians of those values across concentrations.
+The height of each letter is proportional to the excess fraction surviving of virions with that mutation.
+The scale bar at the top of the plot relates the letter heights to the actual fractions.
+The sites are labeled using H3 numbering.
+}
+\end{suppfigure}
+
+\begin{suppfigure}
+\centerline{\includegraphics[trim=0.1cm 0.02cm 0.1cm 0.03cm,clip=true,width=\textwidth]{figs/logoplots/FI6v3_fracsurvive.pdf}}
+\caption{\label{suppfig:FI6v3logo}
+{\bf The excess fraction surviving selection with antibody FI6v3 for all amino-acid mutations.}
+The excess fraction surviving for each replicate was computed using Equation~\ref{eq:fracsurvive_excess}, then we took the median across all technical and biological replicates for each antibody concentration, and then took the medians of those values across concentrations.
+The height of each letter is proportional to the excess fraction surviving of virions with that mutation.
+The scale bar at the top of the plot relates the letter heights to the actual fractions.
+The sites are labeled using H3 numbering.
+}
+\end{suppfigure}
+
+\begin{suppfigure}
+\centerline{\includegraphics[trim=0.1cm 0.02cm 0.1cm 0.03cm,clip=true,width=\textwidth]{figs/logoplots/C179_fracsurvive.pdf}}
+\caption{\label{suppfig:C179logo}
+{\bf The excess fraction surviving selection with antibody C179 for all amino-acid mutations.}
+The excess fraction surviving for each replicate was computed using Equation~\ref{eq:fracsurvive_excess}, then we took the median across all technical and biological replicates for each antibody concentration, and then took the medians of those values across concentrations.
+The height of each letter is proportional to the excess fraction surviving of virions with that mutation.
+The scale bar at the top of the plot relates the letter heights to the actual fractions.
+The sites are labeled using H3 numbering.
+}
+\end{suppfigure}
+
+\begin{suppfigure}
+\centerline{\includegraphics[trim=0.1cm 0.02cm 0.1cm 0.03cm,clip=true,width=\textwidth]{figs/logoplots/S139_fracsurvive.pdf}}
+\caption{\label{suppfig:S139logo}
+{\bf The excess fraction surviving selection with antibody S139/1 for all amino-acid mutations.}
+The excess fraction surviving for each replicate was computed using Equation~\ref{eq:fracsurvive_excess}, then we took the median across all technical and biological replicates for each antibody concentration, and then took the medians of those values across concentrations.
+The height of each letter is proportional to the excess fraction surviving of virions with that mutation.
+The scale bar at the top of the plot relates the letter heights to the actual fractions.
+The sites are labeled using H3 numbering.
+}
 \end{suppfigure}

### 2017-10-13 paper/manuscript.tex
Added a few sentences to various results sections

@@
 Figure~\ref{fig:neutcurves} shows the neutralization curves for each antibody, and all six antibodies show similar potencies.
+We chose to conduct the library selections with concentrations of antibodies that would neutralize nearly all wildtype viruses.
 
@@
 Specifically, Figure~\ref{fig:avgfracsurvive} shows that for all four of these antibodies, there are sites where a substantial fraction of virions with any mutation survive antibody concentrations that are far in excess of those needed to nearly complete neutralize the wildtype virus.
+% Worth mentioning how some escape mutants from H17L19 have reduced fraction escape at the higher concentration, while this is not seen for S139/1?
+
+In contrast, the stalk-targeting antibodies...
 
 \subsection*{For all antibodies, the selected sites of escape occur in the antibody binding footprints}
+We visualized for each mutation the fraction surviving selection as a logoplot, and mapped the mutation with the maximum fraction surviving selection at each site onto the HA structure, shown in Figure~\ref{fig:structures}.
 

### 2017-10-13 paper/manuscript.tex
Very minor adjustments to some sentences in results

@@
 \subsection*{For all antibodies, the selected sites of escape occur in the antibody binding footprints}
-We visualized for each mutation the fraction surviving selection as a logoplot, and mapped the mutation with the maximum fraction surviving selection at each site onto the HA structure, shown in Figure~\ref{fig:structures}.
+We visualized for each mutation the fraction surviving selection as a logoplot, and for each site mapped the mutation with the maximum fraction surviving selection onto the HA structure, shown in Figure~\ref{fig:structures}.
+For all of the antibodies, the sites with mutations exhibiting the greatest fraction escape localize in the antibody-binding footprint.
 

### 2017-10-13 paper/manuscript.tex
Added a few comments as a reminder of what to mention
minor suggestions to second results section

@@
 \subsection*{Broad and narrow antibodies that neutralize influenza virus}
-\comment{Juhye writes a draft of this section.}
+We applied this approach to several broad antibodies that neutralize the A/WSN/1933 (H1N1) influenza virus strain.
+For comparison, we also included several narrow strain-specific antibodies that we have characterized previously~\cite{doud2017complete}.
+Two broad anti-stalk, one all one only group 1.
+One anti receptor
+The narrow ones.
+The last concluding sentence of your paragraph is good.
 
@@
 
-In order to compare the potencies of these antibodies, we completed GFP-based neutralization assays using wild-type A/WSN/1933 (H1N1) virus. 
+To choose appropriate concentrations of these antibodies, we performed neutralization assays using wild-type A/WSN/1933 (H1N1) virus. 
 Figure~\ref{fig:neutcurves} shows the neutralization curves for each antibody, and all six antibodies show similar potencies.
-We chose to conduct the library selections with concentrations of antibodies that would neutralize nearly all wildtype viruses.
-
-% (selection of escape mutants from S139/1 and C179 have been reported, but no reported escape mutants from FI6v3)
+We chose to conduct the library selections with concentrations of antibodies that would neutralize nearly all wildtype viruses (possibly add slightly more detail).
 
@@
 For all of the antibodies, the sites with mutations exhibiting the greatest fraction escape localize in the antibody-binding footprint.
+
+% For S139/1, we see strong escape at three sites.
+These three sites were previously reported as sites of escape in selection experiments with S139/1 (cite Yoshida) G158N mutation introduces a potential glycosylation site
+% The binding footprint of FI6v3 and C179 and the angle of approach of both antibodies is very similar (cite dreyfus). 
+% For these two antibodies, we see selection for a few common escape mutations (S/T at site280, S at site 291, R at 47 / 48 (HA2), but also selection for distinct escape mutations (site 2, site 38, site 323, sites 42, 43)
 

### 2017-10-13 paper/manuscript.tex
a bit more progress on third section of Results

@@
 \subsection*{Quantifying the antigenic effects of all mutations selected by each antibody}
-We performed mutational antigenic profiling using the three broad antibodies (FI6v3, C179, and S139/1) at the concentrations indicated in Figure~\ref{fig:neutcurves}, and analyzed the data to calculate the fraction of virions carrying each mutation that survived neutralization.
+We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}, and analyzed the data to calculate the fraction of virions carrying each mutation that survived neutralization.
 All experiments were performed in full biological triplicate using three independently generated virus libraries carrying all amino-acid mutants to the HA of the A/WSN/1933 (H1N1) strain~\citep{doud2016accurate}.
-We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three strain-specific antibodies (H17-L19, H17-L10, and H17-L7) at the concentrations indicated in Figure~\ref{fig:neutcurves}, and we also re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived each selection.
+We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three narrow strain-specific antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}, and we also re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived neutralization.
 The correlations among replicates are in Figure~\ref{suppfig:corr}.
-For the remainder of this paper, we will refer to the median antigenic effect of each mutation taken across all replicates with that antibody at that concentration.
+For the remainder of this paper, we will refer to the median antigenic effect of each mutation across replicates.
 
@@
 
-It is immediately obvious that both the three strain-specific antibodies and the antibody targeting HA's receptor-binding pocket (S139/1) select antigenic mutations with large effects at multiple sites.
-Specifically, Figure~\ref{fig:avgfracsurvive} shows that for all four of these antibodies, there are sites where a substantial fraction of virions with any mutation survive antibody concentrations that are far in excess of those needed to nearly complete neutralize the wildtype virus.
-% Worth mentioning how some escape mutants from H17L19 have reduced fraction escape at the higher concentration, while this is not seen for S139/1?
+It is immediately obvious that the narrow strain-specific antibodies and the antibody targeting HA's receptor-binding pocket (S139/1) select mutations with large antigenic effects.
+Figure~\ref{fig:avgfracsurvive} shows that for all four of these antibodies, there are multiple sites in HA where mutations enable a substantial fraction of virions to survive high antibody concentrations.
+Specifically, there are mutations that enable over a third of the virions to evade antibody at concentrations where virtually all the wildtype virions are neutralized (Figure~\ref{suppfig:maxfracsurvive}).
+Therefore, the virus can escape these four antibodies with the type of large-effect single amino-acid mutations that are thought to underly to influenza antigenic drift.
 

### 2017-10-15 paper/manuscript.tex
Write draft of results section 2 and figure 2 caption

@@
 \subsection*{Broad and narrow antibodies that neutralize influenza virus}
-We applied this approach to several broad antibodies that neutralize the A/WSN/1933 (H1N1) influenza virus strain.
-For comparison, we also included several narrow strain-specific antibodies that we have characterized previously~\cite{doud2017complete}.
-Two broad anti-stalk, one all one only group 1.
-One anti receptor
-The narrow ones.
-The last concluding sentence of your paragraph is good.
-
-% blurbs about the specific antibodies we use here:
-To quantify escape from broadly-neutralizing antibodies using the approach described above, we profiled an antibody targeting the receptor-binding site (S139/1)~\citep{yoshida2009cross, lee2012heterosubtypic}, and two antibodies targeting the stalk domain (FI6v3 and C179)~\citep{corti2011neutralizing, okuno1993common, dreyfus2013structure}.
-In addition, we included analyses of selections we previously completed with three strain-specific antibodies (H17-L19, H17-L10, and H17-L7)~\citep{doud2017complete}. 
-H17-L19, H17-L10, and H17-L7 target the Ca2, Ca1, and Cb antigenic regions, respectively.
+We applied the approach described above to several broad antibodies that neutralize the A/WSN/1933 (H1N1) influenza virus strain.
+For comparison, we also included three narrow strain-specific antibodies that we previously characterized in~\cite{doud2017complete} and were used in classical antigenic mapping experiments (cite Gerhard, Caton).
 The crystal structures of the broadly-neutralizing antibodies and the binding footprints of the strain-specific antibodies are shown in Figure~\ref{fig:antibody_summary}A.
-S139/1, C179, and FI6v3 also possess a wide range of breadths (Figure~\ref{fig:antibody_summary}B). 
-For instance, S139/1 and FI6v3 can bind various HA's from group 1 and group 2, while C179 exhibits binding only to certain group 1 HA's.
+Two of the broadly-neutralizing antibodies, specifically FI6v3 and C179, target the stalk domain of HA~\citep{corti2011neutralizing, okuno1993common, dreyfus2013structure}. 
+However, FI6v3 exhibits binding activity across group 1 and group 2 HA's, while C179 can bind to only certain group 1 HA's (Figure~\ref{fig:antibody_summary}B).
+The third broad antibody, S139/1, targets the receptor-binding site and can bind HA's from both groups~\citep{yoshida2009cross, lee2012heterosubtypic}.
+The strain-specific antibodies H17-L19, H17-L10, and H17-L7 bind the Ca2, Ca1, and Cb antigenic regions, respectively.
 We were therefore able to profile the ease of escape from antibodies of different specificities and breadths.
 
-To choose appropriate concentrations of these antibodies, we performed neutralization assays using wild-type A/WSN/1933 (H1N1) virus. 
-Figure~\ref{fig:neutcurves} shows the neutralization curves for each antibody, and all six antibodies show similar potencies.
-We chose to conduct the library selections with concentrations of antibodies that would neutralize nearly all wildtype viruses (possibly add slightly more detail).
+To choose appropriate concentrations of these antibodies for selection experiments, we performed neutralization assays using wildtype A/WSN/1933 (H1N1) virus. 
+Figure~\ref{fig:neutcurves} shows the neutralization curves for each antibody.
+All six antibodies exhibit similar potencies against the WSN virus.
+We chose to conduct the library selections with concentrations of antibodies that would allow us to detect genuine escape mutants while neutralizing nearly all wildtype viruses. 
+For the broad antibodies, we included multiple concentrations to impose different stringencies of selection and thereby capture escape mutations of varying effect sizes.
 
@@
 \caption{\label{fig:antibody_summary}
-CAPTION
+{\bf Specificities and breadth of broad and narrow antibodies targeting influenza virus hemagglutinin.}
+(A) We profiled three broadly-neutralizing and three strain-specific antibodies.
+The broad antibodies are denoted by circles and the narrow antibodies are denoted by squares.
+S139/1 (PDB 4GMS,~\cite{lee2012heterosubtypic}) targets the receptor-binding pocket; C179 (PDB 4HLZ,~\cite{dreyfus2013structure}) and FI6v3 (PDB 3ZTN,~\cite{corti2011neutralizing}) target the stalk. 
+The crystal structures of the broad antibodies and the presumed binding footprints of the narrow antibodies~\citep{doud2017complete} are superimposed onto the structure of the HA trimer (PDB 1RVX,~\cite{gamblin2004structure}).
+(B) For each HA in the phylogenetic tree, we denote antibodies with demonstrated binding and/or neutralization activity against that HA. 
+S139/1 not tested for binding against H8, H9, H11; C179 not tested for binding against H8, H11; none of the antibodies tested against H17 and H18.
 }

### 2017-10-15 paper/manuscript.tex
revisions up thru results section 2

--- 2017-10-13 paper/manuscript.tex
@@
 Therefore, antibody breadth is not necessarily an indicator of the difficulty of viral escape.
-Broad antibodies targeting hemagglutinin's stalk are quantifiably harder to escape than antibodies targeting hemagglutinin's head or receptor-binding pocket.
+Broad antibodies targeting hemagglutinin's stalk are quantifiably harder to escape than the other antibodies tested here.
 \end{abstract}
@@
 Verifying this corollary has become of practical importance with the discovery of broadly neutralizing antibodies against influenza virus.
-These antibodies typically target conserved epitopes in HA's stalk~\citep{sui2009structural,ekiert2009antibody,corti2011neutralizing} or receptor-binding pocket~\citep{lee2012heterosubtypic,ekiert2012cross}, and neutralize a wide range of viral strains.
+These antibodies typically target conserved epitopes in HA's stalk~\citep{sui2009structural,ekiert2009antibody,corti2011neutralizing} or receptor-binding pocket~\citep{lee2012heterosubtypic,ekiert2012cross,schmidt2015viral}, and neutralize a wide range of viral strains.
 Broad antibodies are usually less abundant in human serum than antibodies to antigenically variable epitopes on the head of HA~\citep{ellebedy2014induction,andrews2015immune}.
@@
 This escape is manifested by a large shift in the neutralization curve for the mutant (Figure~\ref{fig:fracsurvive_example}B).
-If we draw vertical lines through the overlaid neutralization curves, we can calculate the fraction of virions with each mutation that escape neutralization at each antibody concentration (Figure~\ref{fig:fracsurvive_example}B).
-These fractions can be represented using logo plots, where the height of each letter is proportional to the fraction of virions with that amino acid at a site that survive the antibody (Figure~\ref{fig:fracsurvive_example}C).
-In these logo plots, large letters correspond to strong escape mutations.
+If we draw vertical lines through the overlaid neutralization curves, we can calculate the fraction of virions with each mutation that survive neutralization at each antibody concentration (Figure~\ref{fig:fracsurvive_example}B).
+These fractions can be represented using logo plots, where the height of each letter is proportional to the fraction of virions with that amino acid at a site that survive (Figure~\ref{fig:fracsurvive_example}C).
+Large letters correspond to strong escape mutations.
 
@@
 {\bf Quantifying the fraction of virions with each mutation that escape antibody neutralization.}
-This figure shows hypothetical data for four viral variants: wildtype and three mutants.
+This figure shows hypothetical data for four viral variants.
 (A) Virions with the V1K mutation (orange) completely survive an antibody concentration where most other virions are neutralized.
@@
 
-Now consider the case where a mutation has just a small antigenic effect, and so only slightly increases the fraction of virions that escape neutralization (Figure~\ref{fig:fracsurvive_example}D).
+Now consider the case where a mutation has just a small antigenic effect, and so only slightly increases the fraction of virions that survive neutralization (Figure~\ref{fig:fracsurvive_example}D).
 In this scenario, the neutralization curve shifts only slightly (Figure~\ref{fig:fracsurvive_example}E).
 In the logo plot representation, the antigenic mutation is only slightly larger than other amino acids (Figure~\ref{fig:fracsurvive_example}F), since possessing the mutation only modestly increases the chance that a virion survives antibody treatment.
-These logo plots therefore provide a way to both identify escape mutations and quantify the magnitudes of their antigenic effects in a way that is directly comparable across antibodies.
+These logo plots therefore provide a way to both identify antigenic mutations and quantify the magnitudes of their effects in a way that is directly comparable across antibodies.
 
-Our goal is to determine the fraction of mutant virions that escape antibody neutralization for \emph{all} mutations to HA.
+Our goal is to determine the fraction of mutant virions that survive antibody neutralization for \emph{all} mutations to HA.
 One way to do this would be to measure individual neutralization curves for each of the $19\times565 = 10,735$ single amino-acid mutants of the 565-residue HA protein.
@@
 \end{equation}
-We plot the excess fraction surviving above the library average since this quantity more clearly accentuates antigenic mutations against the background of other mutations.
+%We plot the excess fraction above the library average since this quantity more clearly accentuates antigenic mutations against the background of other mutations.
 Details of how the calculations are extended to account for sequencing errors and sampling statistics are in the \nameref{sec:methods}.
 Open-source software that performs all steps in the analysis beginning with the deep sequencing data is available at \url{https://jbloomlab.github.io/dms_tools2/}.
-
-\subsection*{Broad and narrow antibodies that neutralize influenza virus}
-We applied this approach to several broad antibodies that neutralize the A/WSN/1933 (H1N1) influenza virus strain.
-For comparison, we also included several narrow strain-specific antibodies that we have characterized previously~\cite{doud2017complete}.
-Two broad anti-stalk, one all one only group 1.
-One anti receptor
-The narrow ones.
-The last concluding sentence of your paragraph is good.
-
-% blurbs about the specific antibodies we use here:
-To quantify escape from broadly-neutralizing antibodies using the approach described above, we profiled an antibody targeting the receptor-binding site (S139/1)~\citep{yoshida2009cross, lee2012heterosubtypic}, and two antibodies targeting the stalk domain (FI6v3 and C179)~\citep{corti2011neutralizing, okuno1993common, dreyfus2013structure}.
-In addition, we included analyses of selections we previously completed with three strain-specific antibodies (H17-L19, H17-L10, and H17-L7)~\citep{doud2017complete}. 
-H17-L19, H17-L10, and H17-L7 target the Ca2, Ca1, and Cb antigenic regions, respectively.
-The crystal structures of the broadly-neutralizing antibodies and the binding footprints of the strain-specific antibodies are shown in Figure~\ref{fig:antibody_summary}A.
-S139/1, C179, and FI6v3 also possess a wide range of breadths (Figure~\ref{fig:antibody_summary}B). 
-For instance, S139/1 and FI6v3 can bind various HA's from group 1 and group 2, while C179 exhibits binding only to certain group 1 HA's.
-We were therefore able to profile the ease of escape from antibodies of different specificities and breadths.
-
-To choose appropriate concentrations of these antibodies, we performed neutralization assays using wild-type A/WSN/1933 (H1N1) virus. 
-Figure~\ref{fig:neutcurves} shows the neutralization curves for each antibody, and all six antibodies show similar potencies.
-We chose to conduct the library selections with concentrations of antibodies that would neutralize nearly all wildtype viruses (possibly add slightly more detail).
 
@@
 \caption{\label{fig:antibody_summary}
-CAPTION
+{\bf Epitopes and breadth of broad and narrow antibodies targeting HA.}
+(A) Crystal structures of the broad antibodies and binding footprints of the narrow ones are superimposed on the structure of the HA trimer~\citep[PDB 1RVX;][]{gamblin2004structure}. 
+Broad antibodies are denoted by circles, and narrow antibodies by squares.
+S139/1~\citep[PDB 4GMS;][]{lee2012heterosubtypic} targets the receptor-binding pocket; C179~\citep[PDB 4HLZ;][]{dreyfus2013structure} and FI6v3~\citep[PDB 3ZTN;][]{corti2011neutralizing} target the stalk.
+The footprints for H17-L19, H17-L10, and H17-L7 are those mapped by \citet{doud2017complete}. 
+(B) A phylogenetic tree of HA subtypes, with group 1 at top and group 2 at bottom.
+Circles and squares denote antibodies with reported binding or neutralization activity against that subtype. 
+Not all antibodies have been tested against all subtypes. 
+Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8, and H11; no antibodies have been tested against H17 and H18.
 }
 \end{figure}
+
+\subsection*{Broad and narrow antibodies that neutralize influenza virus}
+We applied our approach to anti-HA antibodies with a range of breadths and epitopes.
+%For comparison, we also included three previously characterized~\citep{doud2017complete} narrow strain-specific antibodies.
+The crystal structures or binding footprints of these antibodies are shown in Figure~\ref{fig:antibody_summary}A.
+We selected two broad antibodies, FI6v3 and C179, that target the stalk of HA~\citep{corti2011neutralizing, okuno1993common, dreyfus2013structure}. 
+FI6v3 is extremely broad, and neutralizes both group 1 and group 2 HAs (Figure~\ref{fig:antibody_summary}B).
+C179 is less broad, and neutralizes only some group 1 HAs (Figure~\ref{fig:antibody_summary}B).
+We also selected a broad antibody, S139/1, that targets HA's receptor-binding site and can neutralize both group 1 and group 2 HAs~\citep{yoshida2009cross, lee2012heterosubtypic}.
+Finally, we re-analyzed deep sequencing data from prior mutational antigenic profiling of three narrow strain-specific antibodies, H17-L19, H17-L10, and H17-L7~\citep{doud2017complete}.
+These narrow antibodies bind the Ca2, Ca1, and Cb antigenic regions on HA's globular head~\citep{caton1982antigenic}, and only neutralize a narrow slice of H1 viruses.
+
+We performed our experiments using the A/WSN/1933 (H1N1) strain of influenza, which is neutralized by all six antibodies with similar potencies (Figure~\ref{fig:neutcurves}).
+We performed the experiments at antibody concentrations where nearly all the wildtype virus was neutralized, allowing us to detect genuine antigenic mutations that increase the fraction of virions that survive beyond the baseline of wildtype (Figure~\ref{fig:neutcurves}).
+We used several concentrations of each broad antibody to increase sensitivity to antigenic mutations of both large and small effect sizes.
 
@@
 \caption{\label{fig:neutcurves}
-CAPTION
-\comment{
-Get rid of the panel numbers altogether and just have the antibody names.
-}
+{\bf Neutralization of wildtype virus by each antibody, and the fraction of virions surviving at each concentration used in our experiments.}
+The curves show neutralization of wildtype A/WSN/1933 virus. 
+Each point represents the mean and standard error of three measurements. 
+\comment{Is the preceding sentence correct?}
+The vertical dotted lines show the concentrations of antibody used in our selections, and the tables give the overall fraction of the mutant virus libraries that survived at each concentration.
 }

### 2017-10-15 paper/manuscript.tex
included all figures in LaTex

@@
 \subsection*{Quantifying the antigenic effects of all mutations selected by each antibody}
-We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}, and analyzed the data to calculate the fraction of virions carrying each mutation that survived neutralization.
-All experiments were performed in full biological triplicate using three independently generated virus libraries carrying all amino-acid mutants to the HA of the A/WSN/1933 (H1N1) strain~\citep{doud2016accurate}.
-We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three narrow strain-specific antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}, and we also re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived neutralization.
+We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}, and calculated the fraction of virions carrying each mutation that survived neutralization.
+All experiments were performed in full biological triplicate using three independently generated virus libraries carrying all single amino-acid mutations to HA~\citep{doud2016accurate}.
+We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three narrow strain-specific antibodies, and we re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived neutralization.
 The correlations among replicates are in Figure~\ref{suppfig:corr}.
@@
 \label{fig:avgfracsurvive}
-{\bf The strain-specific and broad anti-receptor-binding-pocket antibodies select mutations with large antigenic effects, but the broad anti-stalk antibodies only select mutations with small effects.}
-The height of the line at each site indicates the excess fraction of virions with a mutation at that site that survive the antibody selection, averaging across all amino-acid mutations at the site (see Equation~\ref{eq:avgfracsurvive}).
-It is immediately obvious that there are multiple large-effect mutations for H17L19, H17L10, H17L7, and S139/1 -- but not any large effect mutations for FI6v3 and C179.
+{\bf The strain-specific and broad anti-receptor-binding-site antibodies select mutations with large antigenic effects, but the broad anti-stalk antibodies only select small-effect mutations.}
+The height of the lines indicate the excess fraction of virions with a mutation at that site that survive the antibody, averaging across all amino-acid mutations at each site (see Equation~\ref{eq:avgfracsurvive}).
+There are multiple sites of large-effect mutations for H17L19, H17L10, H17L7, and S139/1 -- but not any for FI6v3 and C179.
+Figure~\ref{suppfig:maxfracsurvive} shows the excess fraction surviving for the most strongly selected amino-acid mutation at each site.
+Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo} show data for all mutations for each antibody using logo plots.
 Sites are labeled in H3 numbering.
-Figure~\ref{suppfig:maxfracsurvive} shows the excess fraction surviving for the single most strongly selected amino-acid mutation at each site.
-Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo} show data for all mutations for each antibody using logo plots.
 }
@@
 It is immediately obvious that the narrow strain-specific antibodies and the antibody targeting HA's receptor-binding pocket (S139/1) select mutations with large antigenic effects.
-Figure~\ref{fig:avgfracsurvive} shows that for all four of these antibodies, there are multiple sites in HA where mutations enable a substantial fraction of virions to survive high antibody concentrations.
-Specifically, there are mutations that enable over a third of the virions to evade antibody at concentrations where virtually all the wildtype virions are neutralized (Figure~\ref{suppfig:maxfracsurvive}).
-Therefore, the virus can escape these four antibodies with the type of large-effect single amino-acid mutations that are thought to underly to influenza antigenic drift.
+For all four of these antibodies, there are multiple sites in HA where mutations enable a substantial fraction of virions to survive high antibody concentrations (Figure~\ref{fig:avgfracsurvive}).
+There are mutations that enable over a third of virions to evade these antibodies at concentrations where virtually all wildtype virions are neutralized (Figure~\ref{suppfig:maxfracsurvive}).
+Therefore, the virus can escape these four antibodies with the sort of large-effect single amino-acid mutations that characterize traditional influenza antigenic drift~\citep{yewdell1979antigenic,webster1980determination,koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction}.
 
@@
 \label{fig:structures}
+CAPTION
+}
+\end{figure}
+
+\begin{figure}
+\centerline{\includegraphics[width=\textwidth]{figs/stalkAbs_zoomed/stalkAbs_zoomed.pdf}}
+\caption{
+\label{fig:stalkAbs_zoomed}
+CAPTION
+}
+\end{figure}
+
+\begin{figure}
+\centerline{\includegraphics[width=\textwidth]{figs/FI6v3mutant_neutcurves/FI6v3_mutant_neutcurves.pdf}}
+\caption{
+\label{fig:FI6v3neutcurves}
 CAPTION

### 2017-10-15 paper/manuscript.tex
third section of results

@@
 
-In contrast, the stalk-targeting antibodies...
+In contrast, the stalk-targeting antibodies C179 and FI6v3 select no strong escape mutants. 
+If we look at the results for these antibodies on the same scale as the other antibodies, we see only a few small bumps in the fraction of virions surviving (Figures~\ref{fig:avgfracsurvive} and \ref{suppfig:maxfracsurvive}).
+Only if we zoom in a great deal can we see that there are actually a few sites where mutations slightly increase the fraction of virions surviving C179 and FI6v3 (Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo}).
+But the effect sizes of these antigenic mutations are tiny compared to the other antibodies -- especially for FI6v3.
+Therefore, the HA of A/WSN/1933 influenza virus is far less capable of escaping these anti-stalk antibodies by single mutations than it is of escaping the other four antibodies.
 

### 2017-10-15 paper/manuscript.tex
figure suggestions for 5 and 6

@@
 \setcounter{totalnumber}{8}
+\renewcommand{\topfraction}{1}
+\renewcommand{\bottomfraction}{1}
+\renewcommand{\textfraction}{0}
+\renewcommand{\floatpagefraction}{1}
 
@@
 
+\begin{figure}
+\centerline{\includegraphics[width=\textwidth]{figs/avgfracsurvive.pdf}}
+\caption{
+\label{fig:avgfracsurvive}
+{\bf The strain-specific and anti-receptor-binding-site antibodies select mutations with large antigenic effects, but the anti-stalk antibodies only select small-effect mutations.}
+The excess fraction of virions with a mutation at each site that survive the antibody, averaging across all amino-acid mutations at each site (see Equation~\ref{eq:avgfracsurvive}).
+There are multiple sites of large-effect mutations for H17L19, H17L10, H17L7, and S139/1 -- but none for FI6v3 and C179.
+Figure~\ref{suppfig:maxfracsurvive} shows the excess fraction surviving for the largest-effect mutation at each site.
+Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo} show all mutations using logo plots.
+Sites are labeled in H3 numbering.
+}
+\end{figure}
+
 \subsection*{Quantifying the antigenic effects of all mutations selected by each antibody}
@@
 For the remainder of this paper, we will refer to the median antigenic effect of each mutation across replicates.
-
-\begin{figure}
-\centerline{\includegraphics[width=\textwidth]{figs/avgfracsurvive.pdf}}
-\caption{
-\label{fig:avgfracsurvive}
-{\bf The strain-specific and broad anti-receptor-binding-site antibodies select mutations with large antigenic effects, but the broad anti-stalk antibodies only select small-effect mutations.}
-The height of the lines indicate the excess fraction of virions with a mutation at that site that survive the antibody, averaging across all amino-acid mutations at each site (see Equation~\ref{eq:avgfracsurvive}).
-There are multiple sites of large-effect mutations for H17L19, H17L10, H17L7, and S139/1 -- but not any for FI6v3 and C179.
-Figure~\ref{suppfig:maxfracsurvive} shows the excess fraction surviving for the most strongly selected amino-acid mutation at each site.
-Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo} show data for all mutations for each antibody using logo plots.
-Sites are labeled in H3 numbering.
-}
-\end{figure}
 
@@
 
-\subsection*{For all antibodies, the selected sites of escape occur in the antibody binding footprints}
+\subsection*{The selected mutations are in the binding footprints for all antibodies}
 We visualized for each mutation the fraction surviving selection as a logoplot, and for each site mapped the mutation with the maximum fraction surviving selection onto the HA structure, shown in Figure~\ref{fig:structures}.
@@
 \centerline{\includegraphics[width=\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
+\centerline{\includegraphics[width=0.7\textwidth]{figs/stalkAbs_zoomed/stalkAbs_zoomed.pdf}}
 \caption{
 \label{fig:structures}
-CAPTION
+\comment{Juhye, can you actually merge these two figures into one image.
+Some of the journals (\textit{Cell Host \& Microbe}, \textit{Immunity}) limit papers to 6 figures.
+The first figure can be panel (A).
+Include a healthy amount of vertical space, and make the current (A) and (B) into (B) and (C).
+Also, can you do the following:
+(1) Put the stalk zooms inside a frame box (a thin black or grey line) -- this should make the cutting off look better.
+(2) Clip the zoom boxes a little further on the right so that the zoomed sites are slightly bigger.
+(3) Move the scale bar legend below the zoom panels to separate it from the color labeling of what will now be panel (A).
+(4) Label the scale bar ``fraction surviving''.
+(5) For the tree figure, the logo plot for that site is now shown in panel (A) of the same figure.
+So you can get rid of that, and just include labels that indicate that green means binding / neutralization by C179, and the letters are the amino acid at site 38.
 }
-\end{figure}
-
-\begin{figure}
-\centerline{\includegraphics[width=\textwidth]{figs/stalkAbs_zoomed/stalkAbs_zoomed.pdf}}
-\caption{
-\label{fig:stalkAbs_zoomed}
-CAPTION
 }
@@
 \label{fig:FI6v3neutcurves}
-CAPTION
+\comment{The neutralization curves look good, except can you make the different mutations have different shapes as well as different colors.
+Also, I think it would be good to show the P80D and V135T neutralization curves versus H17-L7 and H17-L19, respectively.
+So your current figure could be panel (A), and then these could be below and be called panel (B) and (C). Each panel could also be labeled with the antibody name.
+You can get these neutralization curves from Mike's paper. 
+I think the data for them is in his paper repository, but if you can't find it then perhaps e-mail him to ask if he knows, or I can help look for it too.
+The rationale here is that it will provide an illustration of the difference between the ``large effect'' mutations for other antibodies and the ``small effect'' ones for these FI6v3. 
+}
 }

### 2017-10-16 paper/manuscript.tex
fix minor typo in intro

@@
 So far, there is limited data to distinguish between these possibilities.
-Several studies have shown that the head domain of HA is less mutationally tolerant than the stalk domain where many broad antibodies bind~\citep{thyagarajan2014inherent,wu2014high,heaton2013genome}.
+Several studies have shown that the head domain of HA is more mutationally tolerant than the stalk domain where many broad antibodies bind~\citep{thyagarajan2014inherent,wu2014high,heaton2013genome}.
 However, these studies did not select for antibody escape, so it is difficult to relate their measurements to the virus's evolutionary capacity under immune selection.

### 2017-10-16 paper/manuscript.tex
Add a few sentences to results section 4

@@
 The curves show neutralization of wildtype A/WSN/1933 virus. 
-Each point represents the mean and standard error of three measurements. 
-\comment{Is the preceding sentence correct?}
+Each point represents the mean and standard deviation of three measurements. 
 The vertical dotted lines show the concentrations of antibody used in our selections, and the tables give the overall fraction of the mutant virus libraries that survived at each concentration.
@@
 \subsection*{The selected mutations are in the binding footprints for all antibodies}
-We visualized for each mutation the fraction surviving selection as a logoplot, and for each site mapped the mutation with the maximum fraction surviving selection onto the HA structure, shown in Figure~\ref{fig:structures}.
-For all of the antibodies, the sites with mutations exhibiting the greatest fraction escape localize in the antibody-binding footprint.
+We visualized for each mutation the fraction surviving selection as a logoplot, and for each site mapped the mutation with the maximum fraction surviving selection onto the HA structure, shown in Figure~\ref{fig:structures}A.
+For all of the antibodies, the sites with mutations exhibiting the greatest fraction escape localize in the antibody epitope.
 
-% For S139/1, we see strong escape at three sites.
-These three sites were previously reported as sites of escape in selection experiments with S139/1 (cite Yoshida) G158N mutation introduces a potential glycosylation site
+For S139/1, we observed strong escape mutations at sites 156, 158, and 193, in H3 numbering. 
+These three sites surround the receptor-binding pocket, and were previously reported as sites of escape in selection experiments with S139/1~\citep{yoshida2009cross}.
+The mutation contributing the greatest fraction of virions surviving selection, G158N, introduces a glycosylation motif (N-X-S/T).
+
 % The binding footprint of FI6v3 and C179 and the angle of approach of both antibodies is very similar (cite dreyfus). 
@@
 \centerline{\includegraphics[width=\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
-\centerline{\includegraphics[width=0.7\textwidth]{figs/stalkAbs_zoomed/stalkAbs_zoomed.pdf}}
 \caption{

### 2017-10-16 paper/manuscript.tex
A bit of progress on results section 4

@@
 We visualized for each mutation the fraction surviving selection as a logoplot, and for each site mapped the mutation with the maximum fraction surviving selection onto the HA structure, shown in Figure~\ref{fig:structures}A.
-For all of the antibodies, the sites with mutations exhibiting the greatest fraction escape localize in the antibody epitope.
+For all of the antibodies, the sites with mutations exhibiting the greatest fraction escape localize in or near the antibody epitope.
+Many amino-acid mutations at most sites can confer escape from the strain-specific antibodies and from S139/1. 
+On the other hand, we see selection for only a handful of amino acids at sites of escape from the broad stalk-targeting antibodies.
 

### 2017-10-17 paper/manuscript.tex
A few more sentences to results section 4, added methods header

@@
 These three sites surround the receptor-binding pocket, and were previously reported as sites of escape in selection experiments with S139/1~\citep{yoshida2009cross}.
-The mutation contributing the greatest fraction of virions surviving selection, G158N, introduces a glycosylation motif (N-X-S/T).
+The mutation contributing the greatest fraction of virions surviving selection, G158N, introduces a glycosylation motif (N-X-S/T) at this site.
 
-% The binding footprint of FI6v3 and C179 and the angle of approach of both antibodies is very similar (cite dreyfus). 
+C179 and FI6v3, which share very similar binding footprints and angles of approach~\citep{dreyfus2013structure}, select for identical escape mutations at several sites.
+Specifically, both antibodies select for mutants K280S and K280T, both of which introduce glycosylation motifs.
+
 % For these two antibodies, we see selection for a few common escape mutations (S/T at site280, S at site 291, R at 47 / 48 (HA2), but also selection for distinct escape mutations (site 2, site 38, site 323, sites 42, 43)
@@
 Describe estimation of $\gamma$ by qPCR.
+
+\subsection*{Construction of HA phylogenetic tree}
+We downloaded HA sequences from the Influenza Research Database, aligned sequences using MUSCLE... \comment{going to add more here}
 

### 2017-10-17 paper/manuscript.tex
Some progress on results section 4 and figure 5 legend

--- 2017-10-16 paper/manuscript.tex
@@
 Circles and squares denote antibodies with reported binding or neutralization activity against that subtype. 
-Not all antibodies have been tested against all subtypes. 
-Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8, and H11; no antibodies have been tested against H17 and H18.
+Note that not all antibodies have been tested against all subtypes. 
 }
@@
 We applied our approach to anti-HA antibodies with a range of breadths and epitopes.
-%For comparison, we also included three previously characterized~\citep{doud2017complete} narrow strain-specific antibodies.
 The crystal structures or binding footprints of these antibodies are shown in Figure~\ref{fig:antibody_summary}A.
@@
 These three sites surround the receptor-binding pocket, and were previously reported as sites of escape in selection experiments with S139/1~\citep{yoshida2009cross}.
-The mutation contributing the greatest fraction of virions surviving selection, G158N, introduces a glycosylation motif (N-X-S/T).
+The mutation contributing the greatest fraction of virions surviving selection, G158N, introduces a glycosylation motif (N-X-S/T) at this site.
+C179 and FI6v3, which share very similar binding footprints and angles of approach~\citep{dreyfus2013structure}, select for identical escape mutations at several sites (Figure~\ref{fig:structures}B). 
+Both antibodies select for the K280S, K280T, and N291S mutations, all of which introduce glycosylation motifs in close proximity to the epitope and all have similar effect sizes in conferring escape from either antibody.
 
-% The binding footprint of FI6v3 and C179 and the angle of approach of both antibodies is very similar (cite dreyfus). 
-% For these two antibodies, we see selection for a few common escape mutations (S/T at site280, S at site 291, R at 47 / 48 (HA2), but also selection for distinct escape mutations (site 2, site 38, site 323, sites 42, 43)
+Despite the similarities in the epitope-targeting of C179 and FI6v3, these two antibodies have very different breadths.
+FI6v3 can neutralize across many group 1 and group 2 HA's, while C179 can neutralize only some group 1 HA's.
+The presence of a glycan on the Asn at site 38 in group 2 HA's but not in group 1 HA's has been posited as a reason for this difference in breadth~\citep{corti2011neutralizing, sui2009structural, ekiert2009antibody}.
+Interestingly, in our experiments, the mutation with the strongest effect size from selection with C179 occurs at site 38, while we do not see observable escape mutations at this site by selection with FI6v3.
+While the H38S mutation does not introduce a glycosylation motif, we found that HA subtypes susceptible to binding or neutralization by C179 tend to possess a histidine at this site, but subtypes not bound or neutralized by C179 possess a glutamine or serine (Figure~\ref{fig:structures}C).
+
+\comment{pdmH1N1 has an NTT glycosylation motif at sites 278-280. Its IC50 for FI6v3 is around 4-fold higher than the IC50 for WSN, according to Corti et al 2011}
+
+\comment{Do you think the mutant neutralization curves should be a separate subsection?}
 
@@
 \label{fig:structures}
-\comment{Juhye, can you actually merge these two figures into one image.
-Some of the journals (\textit{Cell Host \& Microbe}, \textit{Immunity}) limit papers to 6 figures.
-The first figure can be panel (A).
-Include a healthy amount of vertical space, and make the current (A) and (B) into (B) and (C).
-Also, can you do the following:
-(1) Put the stalk zooms inside a frame box (a thin black or grey line) -- this should make the cutting off look better.
-(2) Clip the zoom boxes a little further on the right so that the zoomed sites are slightly bigger.
-(3) Move the scale bar legend below the zoom panels to separate it from the color labeling of what will now be panel (A).
-(4) Label the scale bar ``fraction surviving''.
-(5) For the tree figure, the logo plot for that site is now shown in panel (A) of the same figure.
-So you can get rid of that, and just include labels that indicate that green means binding / neutralization by C179, and the letters are the amino acid at site 38.
-}
+{\bf Mapping of escape mutations from broad and narrow antibodies.}
+(A) Logoplots for a subset of sites with mutations exhibiting some of the largest fraction of virions surviving selection. 
+Sites are labeled in H3 numbering. 
+The height of the scalebar for each antibody corresponds to the excess fraction of virions with a mutation that survive antibody selection. 
+To the right of each logoplot, the excess fraction surviving for the single strongest escape mutation is visualized at each site from white to red on the HA trimer (PDB 1RVX). 
+The strongest shade of red is set to the mutation with the maximum fraction escape for that antibody.
+The crystal structures of each broad antibody are superimposed on HA's structure.
+(B) Zoomed-in view of C179 and FI6v3 epitopes. %Here, the colors are normalized to C179 max mut fracsurvive
+(C) A phylogenetic tree (not to scale) of HA subtypes, with group 1 HA's in black and group 2 HA's in gray.
+The amino-acid identity at site 38 is shown to the right of each subtype.
+Subtypes bound or neutralized by C179 are denoted by green rectangles, while subtypes not bound or neutralized by C179 are denoted by red rectangles.
+Subtypes not tested for binding or neutralization activity are not colored.
+None of the group 2 subtypes are bound/neutralized by C179.
 }
@@
 Describe estimation of $\gamma$ by qPCR.
+
+\subsection*{Construction of HA phylogenetic tree}
+To construct the phylogenetic tree of HA, we downloaded HA sequences from the Influenza Research Database, aligned sequences using MUSCLE... \comment{going to add more here}
+Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8, and H11; no antibodies have been tested against H17 and H18.
 

### 2017-10-17 paper/manuscript.tex
some comments / revisions

@@
 All antibodies select mutations, but the effect sizes vary widely. 
-The virus can escape a broad antibody targeting hemagglutinin's receptor-binding pocket the same way it escapes narrow strain-specific antibodies: via single mutations with huge effects.   
+The virus can escape a broad antibody targeting hemagglutinin's receptor-binding site the same way it escapes narrow strain-specific antibodies: via single mutations with huge effects.   
 In contrast, broad antibodies targeting hemagglutinin's stalk only select mutations with small effects. 
@@
 Verifying this corollary has become of practical importance with the discovery of broadly neutralizing antibodies against influenza virus.
-These antibodies typically target conserved epitopes in HA's stalk~\citep{sui2009structural,ekiert2009antibody,corti2011neutralizing} or receptor-binding pocket~\citep{lee2012heterosubtypic,ekiert2012cross,schmidt2015viral}, and neutralize a wide range of viral strains.
+These antibodies typically target conserved epitopes in HA's stalk~\citep{sui2009structural,ekiert2009antibody,corti2011neutralizing} or receptor-binding site~\citep{lee2012heterosubtypic,ekiert2012cross,schmidt2015viral}, and neutralize a wide range of viral strains.
 Broad antibodies are usually less abundant in human serum than antibodies to antigenically variable epitopes on the head of HA~\citep{ellebedy2014induction,andrews2015immune}.
@@
 However, the magnitudes of the antigenic effects vary greatly across antibodies.
-Single mutations make the virus completely resistant to both narrow strain-specific antibodies and a broad antibody against the receptor-binding pocket.
+Single mutations make the virus completely resistant to both narrow strain-specific antibodies and a broad antibody against the receptor-binding site.
 But no single mutation does more than modestly increase the virus's resistance to broad antibodies against the HA stalk.
@@
 (A) Crystal structures of the broad antibodies and binding footprints of the narrow ones are superimposed on the structure of the HA trimer~\citep[PDB 1RVX;][]{gamblin2004structure}. 
-Broad antibodies are denoted by circles, and narrow antibodies by squares.
-S139/1~\citep[PDB 4GMS;][]{lee2012heterosubtypic} targets the receptor-binding pocket; C179~\citep[PDB 4HLZ;][]{dreyfus2013structure} and FI6v3~\citep[PDB 3ZTN;][]{corti2011neutralizing} target the stalk.
+S139/1~\citep[PDB 4GMS;][]{lee2012heterosubtypic} targets the receptor-binding site; C179~\citep[PDB 4HLZ;][]{dreyfus2013structure} and FI6v3~\citep[PDB 3ZTN;][]{corti2011neutralizing} target the stalk.
 The footprints for H17-L19, H17-L10, and H17-L7 are those mapped by \citet{doud2017complete}. 
 (B) A phylogenetic tree of HA subtypes, with group 1 at top and group 2 at bottom.
-Circles and squares denote antibodies with reported binding or neutralization activity against that subtype. 
-Note that not all antibodies have been tested against all subtypes. 
+Circles (broad antibodies) and squares (narrow antibodies) denote reported antibody binding or neutralization activity against that subtype. 
+Not all antibodies have been tested against all subtypes. 
 }
@@
 We performed our experiments using the A/WSN/1933 (H1N1) strain of influenza, which is neutralized by all six antibodies with similar potencies (Figure~\ref{fig:neutcurves}).
-We performed the experiments at antibody concentrations where nearly all the wildtype virus was neutralized, allowing us to detect genuine antigenic mutations that increase the fraction of virions that survive beyond the baseline of wildtype (Figure~\ref{fig:neutcurves}).
-We used several concentrations of each broad antibody to increase sensitivity to antigenic mutations of both large and small effect sizes.
+We used antibody concentrations where nearly all wildtype virus was neutralized, allowing us to detect antigenic mutations that increase the fraction of virions that survive beyond the baseline of wildtype (Figure~\ref{fig:neutcurves}).
+We used several concentrations of each broad antibody to increase sensitivity to antigenic mutations of both large and small effect.
 
@@
 \label{fig:avgfracsurvive}
-{\bf The strain-specific and anti-receptor-binding-site antibodies select mutations with large antigenic effects, but the anti-stalk antibodies only select small-effect mutations.}
+{\bf Strain-specific and anti-receptor-binding-site antibodies select mutations with large antigenic effects, but anti-stalk antibodies only select small-effect mutations.}
 The excess fraction of virions with a mutation at each site that survive the antibody, averaging across all amino-acid mutations at each site (see Equation~\ref{eq:avgfracsurvive}).
@@
 For all four of these antibodies, there are multiple sites in HA where mutations enable a substantial fraction of virions to survive high antibody concentrations (Figure~\ref{fig:avgfracsurvive}).
-There are mutations that enable over a third of virions to evade these antibodies at concentrations where virtually all wildtype virions are neutralized (Figure~\ref{suppfig:maxfracsurvive}).
+Specifically, for all four antibodies there are mutations that enable over a third of virions to survive at concentrations where virtually all wildtype virions are neutralized (Figure~\ref{suppfig:maxfracsurvive}).
 Therefore, the virus can escape these four antibodies with the sort of large-effect single amino-acid mutations that characterize traditional influenza antigenic drift~\citep{yewdell1979antigenic,webster1980determination,koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction}.
@@
 \begin{figure}
-\centerline{\includegraphics[width=\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
+\centerline{\includegraphics[width=0.8\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
 \caption{
 \label{fig:structures}
-{\bf Mapping of escape mutations from broad and narrow antibodies.}
-(A) Logoplots for a subset of sites with mutations exhibiting some of the largest fraction of virions surviving selection. 
-Sites are labeled in H3 numbering. 
-The height of the scalebar for each antibody corresponds to the excess fraction of virions with a mutation that survive antibody selection. 
-To the right of each logoplot, the excess fraction surviving for the single strongest escape mutation is visualized at each site from white to red on the HA trimer (PDB 1RVX). 
-The strongest shade of red is set to the mutation with the maximum fraction escape for that antibody.
-The crystal structures of each broad antibody are superimposed on HA's structure.
-(B) Zoomed-in view of C179 and FI6v3 epitopes. %Here, the colors are normalized to C179 max mut fracsurvive
-(C) A phylogenetic tree (not to scale) of HA subtypes, with group 1 HA's in black and group 2 HA's in gray.
-The amino-acid identity at site 38 is shown to the right of each subtype.
-Subtypes bound or neutralized by C179 are denoted by green rectangles, while subtypes not bound or neutralized by C179 are denoted by red rectangles.
-Subtypes not tested for binding or neutralization activity are not colored.
-None of the group 2 subtypes are bound/neutralized by C179.
+{\bf Mutations selected by broad and narrow antibodies.}
+(A) For each antibody, the logo plots show the sites where mutations have the largest effect.
+Letter heights are proportional to the excess fraction of virions with that mutation that survive antibody, with the proportionality scaled separately for each antibody as indicated by the scale bars.
+The structures are colored white to red by the excess fraction surviving for the largest-effect mutation at each site, with each antibody again scaled separately.  
+(B) Zoomed-in view of sites of selection from the anti-stalk antibodies C179 and FI6v3, with the same coloring scale for both antibodies.
+(C) A phylogenetic tree (not to scale) of HA subtypes, with group 1 in black and group 2 in gray. 
+The amino acid at site 38 is indicated. 
+Subtypes known to be bound or neutralized by C179 are in green, and subtypes known not to be bound or neutralized are in red.  
+\comment{Change coloring in panel C, as red and green are identical to color-blind people. 
+Maybe use orange and light blue as in Figure 1.
+Also, this figure is too large.
+Here I just scaled it all down, but really we want to just shrink panel B. Can you make it a bit smaller with respect to the rest by scaling the figure part by not the text down.
+Also, the tree can be made shorter (take up less space) by moving the legend to the right side.}
 }

### 2017-10-17 paper/manuscript.tex
updated Figure 5 legend

@@
 \renewcommand{\floatpagefraction}{1}
+
+\usepackage[font=small,labelfont=bf]{caption}
 
@@
 \begin{figure}
-\centerline{\includegraphics[width=0.8\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
+\centerline{\includegraphics[width=0.95\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
 \caption{
@@
 {\bf Mutations selected by broad and narrow antibodies.}
-(A) For each antibody, the logo plots show the sites where mutations have the largest effect.
-Letter heights are proportional to the excess fraction of virions with that mutation that survive antibody, with the proportionality scaled separately for each antibody as indicated by the scale bars.
-The structures are colored white to red by the excess fraction surviving for the largest-effect mutation at each site, with each antibody again scaled separately.  
-(B) Zoomed-in view of sites of selection from the anti-stalk antibodies C179 and FI6v3, with the same coloring scale for both antibodies.
-(C) A phylogenetic tree (not to scale) of HA subtypes, with group 1 in black and group 2 in gray. 
+(A) The logo plots show the sites where mutations have the largest effect.
+Letter heights are proportional to the excess fraction of virions with that mutation that survive antibody, as indicated by the scale bars.
+The structures are colored white to red by the excess fraction surviving for the largest-effect mutation at each site, with each antibody scaled separately.  
+(B) Sites of selection from anti-stalk antibodies, with the same coloring scale for both antibodies.
+(C) Cladogram of HA subtypes, with group 1 in black and group 2 in gray. 
 The amino acid at site 38 is indicated. 
-Subtypes known to be bound or neutralized by C179 are in green, and subtypes known not to be bound or neutralized are in red.  
-\comment{Change coloring in panel C, as red and green are identical to color-blind people. 
-Maybe use orange and light blue as in Figure 1.
-Also, this figure is too large.
-Here I just scaled it all down, but really we want to just shrink panel B. Can you make it a bit smaller with respect to the rest by scaling the figure part by not the text down.
-Also, the tree can be made shorter (take up less space) by moving the legend to the right side.}
+Colors indicate whether a subtype is bound or neutralized by C179 if this has been tested.  
 }

### 2017-10-17 paper/manuscript.tex
legend for Figure 6

@@
 \begin{figure}
-\centerline{\includegraphics[width=\textwidth]{figs/FI6v3mutant_neutcurves/FI6v3_mutant_neutcurves.pdf}}
+\centerline{\includegraphics[width=0.8\textwidth]{figs/FI6v3mutant_neutcurves/FI6v3_mutant_neutcurves.pdf}}
 \caption{
 \label{fig:FI6v3neutcurves}
-\comment{The neutralization curves look good, except can you make the different mutations have different shapes as well as different colors.
-Also, I think it would be good to show the P80D and V135T neutralization curves versus H17-L7 and H17-L19, respectively.
-So your current figure could be panel (A), and then these could be below and be called panel (B) and (C). Each panel could also be labeled with the antibody name.
-You can get these neutralization curves from Mike's paper. 
-I think the data for them is in his paper repository, but if you can't find it then perhaps e-mail him to ask if he knows, or I can help look for it too.
-The rationale here is that it will provide an illustration of the difference between the ``large effect'' mutations for other antibodies and the ``small effect'' ones for these FI6v3. 
-}
+{\bf The mutations selected by FI6v3 increase neutralization resistance, but the effects are small.}
+(A) Neutralization curves of individual viral mutants with FI6v3.
+The mutations K280S, K280T, N291S, G47R (HA2), and K(-8)T are all expected to increase neutralization resistance based on the mutational antigenic profiling (Figure~\ref{fig:structures}A), and they all do so -- but only modestly.
+The mutations K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization based on the mutational antigenic profiling (Figure~\ref{suppfig:FI6v3logo}, and none of them do.
+(B), (C) In contrast to FI6v3, mutations selected by other antibodies have very large effects on neutralization.
+Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L2 taken from \citet{doud2017complete}; that paper also shows additional examples of large-effect escape mutations from these antibodies.
 }

### 2017-10-17 paper/manuscript.tex
Fixed alignment of wildtype sequence in logoplots fig

@@
 
-Despite the similarities in the epitope-targeting of C179 and FI6v3, these two antibodies have very different breadths.
-FI6v3 can neutralize across many group 1 and group 2 HA's, while C179 can neutralize only some group 1 HA's.
-The presence of a glycan on the Asn at site 38 in group 2 HA's but not in group 1 HA's has been posited as a reason for this difference in breadth~\citep{corti2011neutralizing, sui2009structural, ekiert2009antibody}.
-Interestingly, in our experiments, the mutation with the strongest effect size from selection with C179 occurs at site 38, while we do not see observable escape mutations at this site by selection with FI6v3.
-While the H38S mutation does not introduce a glycosylation motif, we found that HA subtypes susceptible to binding or neutralization by C179 tend to possess a histidine at this site, but subtypes not bound or neutralized by C179 possess a glutamine or serine (Figure~\ref{fig:structures}C).
+Despite the similarities in the epitope-targeting of C179 and FI6v3, we also see selection for distinct escape mutations with either antibody.
+The mutation with the strongest effect size from selection by C179 occurs at site 38, while we do not see observable escape mutations at this site by selection with FI6v3.
+Interestingly, we found that HA subtypes susceptible to binding or neutralization by C179 tend to possess a histidine at this site, but subtypes not bound or neutralized by C179 possess a glutamine or serine (Figure~\ref{fig:structures}C).
+This would suggest that susceptibility to neutralization by C179 may in part be determined by the amino-acid identity at site 38.
+Furthermore, the difference in breadth between C179 and FI6v3 has been attributed to FI6v3 being able to accommodate a glycan on the Asn at site 38 which is present in group 2 but not group 1 HA's, while group 1-specific stalk-targeting antibodies clash with this glycan~\citep{corti2011neutralizing, sui2009structural, ekiert2009antibody}.
+The H38S mutation selected by C179 does not introduce a glycosylation motif, which indicates that introduction of a glycan at this site is not the only means of mediating escape from this group 1-specific anti-stalk antibody.
 

### 2017-10-17 paper/manuscript.tex
Some additions to manuscript

@@
 The H38S mutation selected by C179 does not introduce a glycosylation motif, which indicates that introduction of a glycan at this site is not the only means of mediating escape from this group 1-specific anti-stalk antibody.
+The amino-acid identity at site 111 of HA2 can change the orientation of a conserved Trp21 in HA2, also resulting in group 1 and group 2 differences in the binding ability of stalk-targeting antibodies.
+An H111T mutation in the A/South Carolina/1/1918 (H1N1) has been reported to abrogate C179 binding~\citep{dreyfus2013structure}.
+However, we did not observe escape mutations at site 111, but this may be due to strain-specific differences
 

### 2017-10-18 paper/figs/antibody_summary_fig/pymol/create_bnAbHA_fig.py
Added all pymol scripts and PDB structures

@@ -1,65 +1,0 @@
-# Pymol script to create figure of HA with antibodies
-
-# LIST OF PDB FILES
-################
-# 1RVX: 1934 H1 HA
-# 3ZTN: FI6v3 in complex with human H1
-# 4GMS: S139/1 in complex with Vic/75 H3
-# 4HLZ: C179 in complex with H2
-################
-
-import sys
-
-cmd.set('bg_rgb','[1,1,1]') # white
-cmd.set('ray_opaque_background','off')
-cmd.set('specular', '0')
-cmd.set('antialias','2')
-
-cmd.load('1RVX.pdb')
-for c in ['G', 'H', 'I', 'J', 'K', 'L']:
-	cmd.remove('chain {0}'.format(c))
-cmd.select('1RVX_A', '1RVX and chain A')
-cmd.select('1RVX_B', '1RVX and chain B')
-cmd.select('1RVX_C', '1RVX and chain C')
-cmd.color('gray60', '1RVX')
-
-#FI6v3
-cmd.load('3ZTN.pdb')
-cmd.select('3ZTN_B', '3ZTN and chain B')
-cmd.select('3ZTN_HA', '3ZTN and (chain A | chain B)')
-cmd.select('FI6v3', '3ZTN and (chain H | chain L)')
-cmd.super('3ZTN_B', '1RVX_B')
-cmd.color('gray60', '3ZTN_HA')
-
-#C179
-cmd.load('4HLZ.pdb')
-for c in ['C', 'D', 'E', 'F', 'I', 'J', 'K', 'L']:
-	cmd.remove('4HLZ and chain {0}'.format(c))
-cmd.select('4HLZ_B', '4HLZ and chain B')
-cmd.select('4HLZ_HA', '4HLZ and (chain A | chain B)')
-cmd.select('C179', '4HLZ and (chain G | chain H)')
-cmd.super('4HLZ_B', '1RVX_B')
-cmd.color('gray60', '4HLZ_HA')
-
-#S139/1
-cmd.load('4GMS.pdb')
-for c in ['C', 'D', 'E', 'F', 'I', 'J', 'M', 'N']:
-	cmd.remove('4GMS and chain {0}'.format(c))
-cmd.select('4GMS_A', '4GMS and chain A')
-cmd.select('4GMS_HA', '4GMS and (chain A | chain B)')
-cmd.select('S139', '4GMS and (chain H | chain L)')
-cmd.super('4GMS_A', '1RVX_A')
-cmd.color('gray60', '4GMS_HA')
-
[…]

### 2017-10-18 paper/pymol/C179_maxmut_zoomed.py

@@ -1,0 +1,612 @@
+import sys
+
+cmd.set('bg_rgb','[1,1,1]')
+cmd.set('ray_opaque_background', 'off')
+cmd.set('antialias', '2')
+cmd.set('specular', '0')
+
+cmd.load('1RVX_H3_numbering.pdb')
+cmd.hide('everything')
+cmd.show('surface')
+cmd.select('1RVX', '1RVX_H3_numbering')
+cmd.select('HA1', '(chain A | chain C | chain E)')
+cmd.select('HA2', '(chain B | chain D | chain F)')
+
+cmd.alter('HA2 and resi 1', 'b = 0.00011905304247703755')
+cmd.alter('HA2 and resi 10', 'b = 0.0005940000000000003')
+cmd.alter('HA2 and resi 100', 'b = 0.0')
+cmd.alter('HA2 and resi 101', 'b = 0.00012907477623358182')
+cmd.alter('HA2 and resi 102', 'b = 0.0027063733737821122')
+cmd.alter('HA2 and resi 103', 'b = 4.336808689942017e-19')
+cmd.alter('HA2 and resi 104', 'b = 0.000530469904285812')
+cmd.alter('HA2 and resi 105', 'b = 4.336808689942017e-19')
+cmd.alter('HA2 and resi 106', 'b = 0.00030921135204491166')
+cmd.alter('HA2 and resi 107', 'b = 0.00039293193322206')
+cmd.alter('HA2 and resi 108', 'b = 4.336808689942017e-19')
+cmd.alter('HA2 and resi 109', 'b = 4.336808689942017e-19')
+cmd.alter('HA2 and resi 11', 'b = 0.002698615392586744')
+cmd.alter('HA2 and resi 110', 'b = 0.0')
+cmd.alter('HA2 and resi 111', 'b = 0.00155837847230249')
+cmd.alter('HA2 and resi 112', 'b = 0.0')
+cmd.alter('HA2 and resi 113', 'b = 0.00015219465263120908')
+cmd.alter('HA2 and resi 114', 'b = 8.673617379884034e-19')
+cmd.alter('HA2 and resi 115', 'b = 0.00022615271939852803')
+cmd.alter('HA2 and resi 116', 'b = 8.921984603537357e-05')
+cmd.alter('HA2 and resi 117', 'b = 1.7739288228138094e-05')
+cmd.alter('HA2 and resi 118', 'b = 0.0001187251795914218')
[…]

### 2017-10-18 paper/pymol/C179_median_maxmut.py

@@ -1,0 +1,1170 @@
+import sys
+
+cmd.set('bg_rgb','[1,1,1]')
+cmd.set('ray_opaque_background', 'off')
+cmd.set('antialias', '2')
+cmd.set('specular', '0')
+
+cmd.load('1RVX_H3_numbering.pdb')
+cmd.hide('everything')
+cmd.show('surface')
+cmd.select('HA1', 'chain A')
+cmd.select('HA2', 'chain B')
+cmd.color('white', '(chain C | chain D | chain E | chain F)')
+
+cmd.set_color('color1_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color1_HA2', 'resi 1 and HA2')
+cmd.set_color('color10_HA2', '[0.9785530430580994, 0.9783804317346233, 0.9783804317346233]')
+cmd.color('color10_HA2', 'resi 10 and HA2')
+cmd.set_color('color100_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color100_HA2', 'resi 100 and HA2')
+cmd.set_color('color101_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color101_HA2', 'resi 101 and HA2')
+cmd.set_color('color102_HA2', '[0.9713509300813108, 0.9701791889803419, 0.9701791889803419]')
+cmd.color('color102_HA2', 'resi 102 and HA2')
+cmd.set_color('color103_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color103_HA2', 'resi 103 and HA2')
+cmd.set_color('color104_HA2', '[0.9785530430580994, 0.9783804317346233, 0.9783804317346233]')
+cmd.color('color104_HA2', 'resi 104 and HA2')
+cmd.set_color('color105_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color105_HA2', 'resi 105 and HA2')
+cmd.set_color('color106_HA2', '[0.9794706706823998, 0.9793882235767065, 0.9793882235767065]')
+cmd.color('color106_HA2', 'resi 106 and HA2')
+cmd.set_color('color107_HA2', '[0.9794706706823998, 0.9793882235767065, 0.9793882235767065]')
+cmd.color('color107_HA2', 'resi 107 and HA2')
[…]

### 2017-10-18 paper/pymol/FI6v3_maxmutC179scale_zoomed.py

@@ -1,0 +1,604 @@
+import sys
+
+cmd.set('bg_rgb','[1,1,1]')
+cmd.set('ray_opaque_background', 'off')
+cmd.set('antialias', '2')
+cmd.set('specular', '0')
+
+cmd.load('1RVX_H3_numbering.pdb')
+cmd.hide('everything')
+cmd.show('surface')
+cmd.select('1RVX', '1RVX_H3_numbering')
+cmd.select('HA1', '(chain A | chain C | chain E)')
+cmd.select('HA2', '(chain B | chain D | chain F)')
+
+cmd.alter('HA2 and resi 1', 'b = 0.009989797813893538')
+cmd.alter('HA2 and resi 10', 'b = 0.000966924779412968')
+cmd.alter('HA2 and resi 100', 'b = 0.00503774021239961')
+cmd.alter('HA2 and resi 101', 'b = 7.716738155708494e-06')
+cmd.alter('HA2 and resi 102', 'b = 0.0037108812370539187')
+cmd.alter('HA2 and resi 103', 'b = 0.001259991050515719')
+cmd.alter('HA2 and resi 104', 'b = 0.0018299916074075875')
+cmd.alter('HA2 and resi 105', 'b = 0.002270683256287649')
+cmd.alter('HA2 and resi 106', 'b = 0.0039512818924899275')
+cmd.alter('HA2 and resi 107', 'b = 0.0005393462161555283')
+cmd.alter('HA2 and resi 108', 'b = 0.0004422516517032832')
+cmd.alter('HA2 and resi 109', 'b = 0.012250067781034634')
+cmd.alter('HA2 and resi 11', 'b = 0.00043828639116155066')
+cmd.alter('HA2 and resi 110', 'b = 0.0019339761778381745')
+cmd.alter('HA2 and resi 111', 'b = 0.0004810472337673293')
+cmd.alter('HA2 and resi 112', 'b = 0.00573083254314432')
+cmd.alter('HA2 and resi 113', 'b = 0.005803937913066862')
+cmd.alter('HA2 and resi 114', 'b = 0.0017314827237735536')
+cmd.alter('HA2 and resi 115', 'b = 0.0003878709427395052')
+cmd.alter('HA2 and resi 116', 'b = 0.004699022641075086')
+cmd.alter('HA2 and resi 117', 'b = 0.001248816048736724')
+cmd.alter('HA2 and resi 118', 'b = 0.00035060598135770716')
[…]

### 2017-10-18 paper/pymol/FI6v3_median_maxmut.py

@@ -1,0 +1,1162 @@
+import sys
+
+cmd.set('bg_rgb','[1,1,1]')
+cmd.set('ray_opaque_background', 'off')
+cmd.set('antialias', '2')
+cmd.set('specular', '0')
+
+cmd.load('1RVX_H3_numbering.pdb')
+cmd.hide('everything')
+cmd.show('surface')
+cmd.select('HA1', 'chain A')
+cmd.select('HA2', 'chain B')
+cmd.color('white', '(chain C | chain D | chain E | chain F)')
+
+cmd.set_color('color1_HA2', '[0.9134899922836754, 0.8759319889336612, 0.8759319889336612]')
+cmd.color('color1_HA2', 'resi 1 and HA2')
+cmd.set_color('color10_HA2', '[0.9731283069892388, 0.9722526510051813, 0.9722526510051813]')
+cmd.color('color10_HA2', 'resi 10 and HA2')
+cmd.set_color('color100_HA2', '[0.9427452492316157, 0.9313951885066056, 0.9313951885066056]')
+cmd.color('color100_HA2', 'resi 100 and HA2')
+cmd.set_color('color101_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color101_HA2', 'resi 101 and HA2')
+cmd.set_color('color102_HA2', '[0.9520204305030352, 0.9452250408317912, 0.9452250408317912]')
+cmd.color('color102_HA2', 'resi 102 and HA2')
+cmd.set_color('color103_HA2', '[0.9704680294614141, 0.9691366701338548, 0.9691366701338548]')
+cmd.color('color103_HA2', 'resi 103 and HA2')
+cmd.set_color('color104_HA2', '[0.9661114047026034, 0.9638661975607468, 0.9638661975607468]')
+cmd.color('color104_HA2', 'resi 104 and HA2')
+cmd.set_color('color105_HA2', '[0.9626955589043622, 0.9595803654934529, 0.9595803654934529]')
+cmd.color('color105_HA2', 'resi 105 and HA2')
+cmd.set_color('color106_HA2', '[0.95043598139735, 0.942958651004709, 0.942958651004709]')
+cmd.color('color106_HA2', 'resi 106 and HA2')
+cmd.set_color('color107_HA2', '[0.9767293634776331, 0.9763532723823221, 0.9763532723823221]')
+cmd.color('color107_HA2', 'resi 107 and HA2')
[…]

### 2017-10-18 paper/pymol/H17L10_median_maxmut.py

@@ -1,0 +1,1151 @@
+import sys
+
+cmd.set('bg_rgb','[1,1,1]')
+cmd.set('ray_opaque_background', 'off')
+cmd.set('antialias', '2')
+cmd.set('specular', '0')
+
+cmd.load('1RVX_H3_numbering.pdb')
+cmd.hide('everything')
+cmd.show('surface')
+cmd.select('HA1', '(chain A | chain C | chain E)')
+cmd.select('HA2', '(chain B | chain D | chain F)')
+
+cmd.set_color('color1_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color1_HA2', 'resi 1 and HA2')
+cmd.set_color('color10_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color10_HA2', 'resi 10 and HA2')
+cmd.set_color('color100_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color100_HA2', 'resi 100 and HA2')
+cmd.set_color('color101_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color101_HA2', 'resi 101 and HA2')
+cmd.set_color('color102_HA2', '[0.9794706706823998, 0.9793882235767065, 0.9793882235767065]')
+cmd.color('color102_HA2', 'resi 102 and HA2')
+cmd.set_color('color103_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color103_HA2', 'resi 103 and HA2')
+cmd.set_color('color104_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color104_HA2', 'resi 104 and HA2')
+cmd.set_color('color105_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color105_HA2', 'resi 105 and HA2')
+cmd.set_color('color106_HA2', '[0.9794706706823998, 0.9793882235767065, 0.9793882235767065]')
+cmd.color('color106_HA2', 'resi 106 and HA2')
+cmd.set_color('color107_HA2', '[0.9794706706823998, 0.9793882235767065, 0.9793882235767065]')
+cmd.color('color107_HA2', 'resi 107 and HA2')
[…]

### 2017-10-18 paper/pymol/H17L19_median_maxmut.py

@@ -1,0 +1,1152 @@
+import sys
+
+cmd.set('bg_rgb','[1,1,1]')
+cmd.set('ray_opaque_background', 'off')
+cmd.set('antialias', '2')
+cmd.set('specular', '0')
+
+cmd.load('1RVX_H3_numbering.pdb')
+cmd.hide('everything')
+cmd.show('surface')
+cmd.select('HA1', 'chain A')
+cmd.select('HA2', 'chain B')
+cmd.color('white', '(chain C | chain D | chain E | chain F)')
+
+cmd.set_color('color1_HA2', '[0.9785530430580994, 0.9783804317346233, 0.9783804317346233]')
+cmd.color('color1_HA2', 'resi 1 and HA2')
+cmd.set_color('color10_HA2', '[0.9767293634776331, 0.9763532723823221, 0.9763532723823221]')
+cmd.color('color10_HA2', 'resi 10 and HA2')
+cmd.set_color('color100_HA2', '[0.9794706706823998, 0.9793882235767065, 0.9793882235767065]')
+cmd.color('color100_HA2', 'resi 100 and HA2')
+cmd.set_color('color101_HA2', '[0.9767293634776331, 0.9763532723823221, 0.9763532723823221]')
+cmd.color('color101_HA2', 'resi 101 and HA2')
+cmd.set_color('color102_HA2', '[0.9794706706823998, 0.9793882235767065, 0.9793882235767065]')
+cmd.color('color102_HA2', 'resi 102 and HA2')
+cmd.set_color('color103_HA2', '[0.9794706706823998, 0.9793882235767065, 0.9793882235767065]')
+cmd.color('color103_HA2', 'resi 103 and HA2')
+cmd.set_color('color104_HA2', '[0.9794706706823998, 0.9793882235767065, 0.9793882235767065]')
+cmd.color('color104_HA2', 'resi 104 and HA2')
+cmd.set_color('color105_HA2', '[0.9713509300813108, 0.9701791889803419, 0.9701791889803419]')
+cmd.color('color105_HA2', 'resi 105 and HA2')
+cmd.set_color('color106_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color106_HA2', 'resi 106 and HA2')
+cmd.set_color('color107_HA2', '[0.9794706706823998, 0.9793882235767065, 0.9793882235767065]')
+cmd.color('color107_HA2', 'resi 107 and HA2')
[…]

### 2017-10-18 paper/pymol/H17L7_median_maxmut.py

@@ -1,0 +1,1152 @@
+import sys
+
+cmd.set('bg_rgb','[1,1,1]')
+cmd.set('ray_opaque_background', 'off')
+cmd.set('antialias', '2')
+cmd.set('specular', '0')
+
+cmd.load('1RVX_H3_numbering.pdb')
+cmd.hide('everything')
+cmd.show('surface')
+cmd.select('HA1', 'chain C')
+cmd.select('HA2', 'chain D')
+cmd.color('white', '(chain A | chain B | chain E | chain F)')
+
+cmd.set_color('color1_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color1_HA2', 'resi 1 and HA2')
+cmd.set_color('color10_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color10_HA2', 'resi 10 and HA2')
+cmd.set_color('color100_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color100_HA2', 'resi 100 and HA2')
+cmd.set_color('color101_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color101_HA2', 'resi 101 and HA2')
+cmd.set_color('color102_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color102_HA2', 'resi 102 and HA2')
+cmd.set_color('color103_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color103_HA2', 'resi 103 and HA2')
+cmd.set_color('color104_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color104_HA2', 'resi 104 and HA2')
+cmd.set_color('color105_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color105_HA2', 'resi 105 and HA2')
+cmd.set_color('color106_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color106_HA2', 'resi 106 and HA2')
+cmd.set_color('color107_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color107_HA2', 'resi 107 and HA2')
[…]

### 2017-10-18 paper/pymol/S139_median_maxmut.py

@@ -1,0 +1,1171 @@
+import sys
+
+cmd.set('bg_rgb','[1,1,1]')
+cmd.set('ray_opaque_background', 'off')
+cmd.set('antialias', '2')
+cmd.set('specular', '0')
+
+cmd.load('1RVX_H3_numbering.pdb')
+cmd.hide('everything')
+cmd.show('surface')
+cmd.select('HA1', 'chain A')
+cmd.select('HA2', 'chain B')
+cmd.color('white', '(chain C | chain D | chain E | chain F)')
+
+cmd.set_color('color1_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color1_HA2', 'resi 1 and HA2')
+cmd.set_color('color10_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color10_HA2', 'resi 10 and HA2')
+cmd.set_color('color100_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color100_HA2', 'resi 100 and HA2')
+cmd.set_color('color101_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color101_HA2', 'resi 101 and HA2')
+cmd.set_color('color102_HA2', '[0.9749211181213463, 0.9743106788058414, 0.9743106788058414]')
+cmd.color('color102_HA2', 'resi 102 and HA2')
+cmd.set_color('color103_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color103_HA2', 'resi 103 and HA2')
+cmd.set_color('color104_HA2', '[0.9785530430580994, 0.9783804317346233, 0.9783804317346233]')
+cmd.color('color104_HA2', 'resi 104 and HA2')
+cmd.set_color('color105_HA2', '[0.9776392739898438, 0.9773687813364951, 0.9773687813364951]')
+cmd.color('color105_HA2', 'resi 105 and HA2')
+cmd.set_color('color106_HA2', '[0.9803921568627451, 0.9803921568627451, 0.9803921568627451]')
+cmd.color('color106_HA2', 'resi 106 and HA2')
+cmd.set_color('color107_HA2', '[0.9785530430580994, 0.9783804317346233, 0.9783804317346233]')
+cmd.color('color107_HA2', 'resi 107 and HA2')
[…]

### 2017-10-18 paper/pymol/create_bnAbHA_fig.py

@@ -1,0 +1,95 @@
+# Pymol script to create figure of HA with antibodies
+
+# LIST OF PDB FILES
+################
+# 1RVX: 1934 H1 HA
+# 3ZTN: FI6v3 in complex with human H1
+# 4GMS: S139/1 in complex with Vic/75 H3
+# 4HLZ: C179 in complex with H2
+################
+
+import sys
+
+cmd.set('bg_rgb','[1,1,1]') # white
+cmd.set('ray_opaque_background','off')
+cmd.set('specular', '0')
+cmd.set('antialias','2')
+
+cmd.load('1RVX_trimer_sequentialnumbering.pdb')
+cmd.select('1RVX', '1RVX_trimer_sequentialnumbering')
+cmd.select('1RVX_A', '1RVX and chain A')
+cmd.select('1RVX_B', '1RVX and chain B')
+cmd.select('1RVX_C', '1RVX and chain C')
+cmd.color('gray90', '1RVX')
+
+#FI6v3
+cmd.load('3ZTN.pdb')
+cmd.select('3ZTN_B', '3ZTN and chain B')
+cmd.select('3ZTN_HA', '3ZTN and (chain A | chain B)')
+cmd.select('FI6v3', '3ZTN and (chain H | chain L)')
+cmd.super('3ZTN_B', '1RVX_A')
+cmd.remove('3ZTN_HA')
+
+#C179
+cmd.load('4HLZ.pdb')
+for c in ['C', 'D', 'E', 'F', 'I', 'J', 'K', 'L']:
+	cmd.remove('4HLZ and chain {0}'.format(c))
+cmd.select('4HLZ_B', '4HLZ and chain B')
+cmd.select('4HLZ_HA', '4HLZ and (chain A | chain B)')
+cmd.select('C179', '4HLZ and (chain G | chain H)')
+cmd.super('4HLZ_B', '1RVX_A')
+cmd.remove('4HLZ_HA')
+
+#S139/1
+cmd.load('4GMS.pdb')
+for c in ['C', 'D', 'E', 'F', 'I', 'J', 'M', 'N']:
+	cmd.remove('4GMS and chain {0}'.format(c))
+cmd.select('4GMS_A', '4GMS and chain A')
+cmd.select('4GMS_HA', '4GMS and (chain A | chain B)')
+cmd.select('S139', '4GMS and (chain H | chain L)')
+cmd.super('4GMS_A', '1RVX_A')
[…]

### 2017-10-18 paper/manuscript.tex
Some minor additions to methods

@@
 \subsection*{Neutralization assays}
-\comment{add this}
+We performed neutralization assays as previously described in ~\citep{}.
+\comment{going to add more here}
 
 \subsection*{Mutant virus selections with antibody and deep sequencing}
-Library selections, deep sequencing, and computation of mutation differential selection were performed as previously described~\cite{doud2017complete}.
+Library selections, deep sequencing, and computation of mutation differential selection were performed as previously described in ~\cite{doud2017complete}.
+
+To estimate the fraction of the overall library surviving selection, we used a standard curve by making 10-fold serial dilutions of the virus libraries
 

### 2017-10-18 paper/manuscript.tex
Fleshed out methods section

--- 2017-10-17 paper/manuscript.tex
@@
 An H111T mutation in the A/South Carolina/1/1918 (H1N1) has been reported to abrogate C179 binding~\citep{dreyfus2013structure}.
-However, we did not observe escape mutations at site 111, but this may be due to strain-specific differences
+However, we did not observe escape mutations at site 111, but this may be due to strain-specific differences.
 
@@
 
-\comment{Do you think the mutant neutralization curves should be a separate subsection?}
+\subsection*{Validating escape mutation effect sizes with neutralization assays}
 
@@
 \subsection*{Neutralization assays}
-\comment{add this}
+We performed neutralization assays as previously described in~\citep{hooper2013mutant} using WSN GFP reporter viruses.
+Viruses carrying GFP in the PB1 segment were grown using reverse genetics plasmids in complementing 293T-CMV-PB1 and MDCK-SIAT1-CMV-PB1 cells that constitutively express the PB1 protein.
 
 \subsection*{Mutant virus selections with antibody and deep sequencing}
-Library selections, deep sequencing, and computation of mutation differential selection were performed as previously described~\cite{doud2017complete}.
-
-Describe estimation of $\gamma$ by qPCR.
+Library selections and deep sequencing were performed as previously described in~\cite{doud2017complete}. 
+We conducted the antibody selections using the mutant virus libraries from~\cite{doud2016accurate}. 
+We diluted the virus libraries to a concentration of $1 \times 10^{6}$ TCID$_{50}$ per ml and incubated the virus dilutions with an equal volume of antibody at the intended concentration.
+We then incubated the virus-antibody mixture at 37$^\circ$C for 1.5 hours
+Each selection was done in biological triplicate at the antibody concentrations shown in Figure~\ref{fig:neutcurves} concentrations, and technical replicates of some of the concentrations were also performed.
+To estimate the fraction of the overall library surviving selection, denoted as $\gamma$ in this paper, we made duplicate 10-fold serial dilutions of each of the virus libraries to use as a standard curve of infectivity.
+qRT-PCR of the standard curves were performed as previously described in~\cite{doud2017complete} using qPCR primers for WSN influenza nucleoprotein (NP) to quantify viral infectivity and primers for canine GAPDH to correct for RNA amounts.
+Next, we fit a linear regression of the log(infectious dose) versus the mean difference in qPCR Ct values between NP and GAPDH to estimate the total library fraction infectivity after antibody selection.
 
 \subsection*{Construction of HA phylogenetic tree}
-To construct the phylogenetic tree of HA, we downloaded HA sequences from the Influenza Research Database, aligned sequences using MUSCLE... \comment{going to add more here}
-Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8, and H11; no antibodies have been tested against H17 and H18.
+We downloaded one HA sequence per subtype from the NIAID Influenza Research Database (IRD) \comment{cite}, built the phylogenetic tree using RaxML \comment{cite}, and visualized the tree using FigTree \comment{cite}.
+The HA sequences we downloaded were: H1 (A/WSN/1933); pdmH1N1 (A/California/04/2009); H2 (A/Japan/305/1957); H3 (A/Aichi/2/1968); H4 (A/swine/Ontario/01911/2/1999); H5 (A/Vietnam/1203/2004); H6 (A/chicken/Taiwan/0705/1999); H7 (A/Netherlands/219/2003); H8 (A/turkey/Ontario/6118/1968); H9 (A/swine/HongKong/9/1998); H10 (A/mallard/Bavaria/3/2006); H11 (A/duck/England/1/1956); H12 (A/duck/Alberta/60/1976); H13 (A/gull/Maryland/704/1977); H14 (A/mallard/Astrakhan/263/1982); H15 (A/mallard duck/Sweden/139579/2012); H16 (A/black headed gull/Mongolia/1756/2006); H17 (A/little yellow-shouldered bat/Guatemala/060/2010); H18 (A/dark fruit-eating bat/Bolivia/PBV780-781/2011).
+
+We denoted HA's with reported binding or neutralization activity by each antibody, and defined binding activity as \comment{binding to full influenza virus, VLPs, recombinant or purified HA's}.
+Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8 and H11; no antibodies have been tested against H17 and H18.
 
 \subsection*{H3 sequence numbering}
-\comment{add this}
+Unless otherwise indicated, all residues are numbered in H3 numbering in this paper, with the signal peptide in negative numbers, the HA1 subunit as plain numbers, and the HA2 subunit denoted with "(HA2)".
+We converted the sequential residue numbers into H3 numbering by running the Python script available on \url{https://github.com/jbloomlab/HA_numbering}.
+In this script, we used MUSCLE to align the WSN/1933 H1 HA sequence to the 4HMG pdb sequence for the A/Aichi/2/1968 H3 HA, which allowed us to convert from the sequential numbering to the H3 numbering scheme.
 

### 2017-10-18 paper/manuscript.tex
Added a sentence to methods

@@
 
+For the cladogram in Figure~\ref{fig:structures}C, the amino-acid identities at site 38 are from the strains tested against C179 in~\cite{dreyfus2013structure}.
+For subtypes not tested, we used the identity from the corresponding strain listed above.
+
 We denoted HA's with reported binding or neutralization activity by each antibody, and defined binding activity as \comment{binding to full influenza virus, VLPs, recombinant or purified HA's}.
@@
 We converted the sequential residue numbers into H3 numbering by running the Python script available on \url{https://github.com/jbloomlab/HA_numbering}.
-In this script, we used MUSCLE to align the WSN/1933 H1 HA sequence to the 4HMG pdb sequence for the A/Aichi/2/1968 H3 HA, which allowed us to convert from the sequential numbering to the H3 numbering scheme.
+In this script, we used MUSCLE to align the WSN/1933 H1 HA sequence to the 4HMG pdb sequence for the A/Aichi/2/1968 H3 HA \comment{cite Weis1990JMB}, which allowed us to convert from the sequential numbering to the H3 numbering scheme.
 
@@
 \subsection*{Data availability and source code}
-Deep sequencing data are available from the Sequence Read Archive under BioSample accession (\comment{add this}).
+Deep sequencing data are available from the Sequence Read Archive under BioSample accession SAMN05789126.
 

### 2017-10-18 paper/manuscript.tex
text revisions of structure subsection of Results thru paragraph 2

@@
 
-\subsection*{The selected mutations are in the binding footprints for all antibodies}
-We visualized for each mutation the fraction surviving selection as a logoplot, and for each site mapped the mutation with the maximum fraction surviving selection onto the HA structure, shown in Figure~\ref{fig:structures}A.
-For all of the antibodies, the sites with mutations exhibiting the greatest fraction escape localize in or near the antibody epitope.
-Many amino-acid mutations at most sites can confer escape from the strain-specific antibodies and from S139/1. 
-On the other hand, we see selection for only a handful of amino acids at sites of escape from the broad stalk-targeting antibodies.
+\subsection*{The selected mutations are near the binding footprints of the antibodies}
+Antigenic mutations selected by narrow strain-specific antibodies against HA generally occur at residues in or near the physical binding footprint of the antibody~\citep{yewdell1979antigenic,webster1980determination,caton1982antigenic}.
+We examined whether this was also the case for the broad antibodies used in our experiments.
+Figure~\ref{fig:structures}A shows a zoomed-in view of the sites of mutations selected by each antibody, as well as their locations on HA's structure. 
+It is immediately clear that the selected mutations are nearly all in or near the antibody-binding footprint.
 
-For S139/1, we observed strong escape mutations at sites 156, 158, and 193, in H3 numbering. 
-These three sites surround the receptor-binding pocket, and were previously reported as sites of escape in selection experiments with S139/1~\citep{yoshida2009cross}.
+\begin{figure}[h!]
+\centerline{\includegraphics[width=0.91\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
+\caption{
+\label{fig:structures}
+{\bf Mutations selected by broad and narrow antibodies.}
+(A) The logo plots show the sites where mutations have the largest effect.
+Letter heights are proportional to the excess fraction of virions with that mutation that survive antibody, as indicated by the scale bars.
+The structures are colored white to red by the excess fraction surviving for the largest-effect mutation at each site, with each antibody scaled separately.  
+(B) Sites of selection from anti-stalk antibodies, with the same coloring scale for both antibodies.
+(C) Cladogram of HA subtypes, with group 1 in black and group 2 in gray. 
+The amino acid at site 38 is indicated. 
+Colors indicate whether a subtype is bound or neutralized by C179 if this has been tested.  
+}
+\end{figure}
+
+Specifically, for S139/1, we observed strong escape mutations at sites 156, 158, and 193 (Figure~\ref{fig:structures}A; sites are in H3 numbering). 
+These three sites surround HA's receptor-binding pocket, and fall directly in the physical binding footprint of the antibody~\citep{lee2012heterosubtypic}.
+These are also the same three sites where \citet{yoshida2009cross} selected escape mutants in H1, H2, and H3 HAs. 
+Our data show that numerous different amino-acid mutations at each site confer neutralization resistance.
+The mutation with the largest effect, D158N, introduces an N-linked glycosylation site.
+
 The mutation contributing the greatest fraction of virions surviving selection, G158N, introduces a glycosylation motif (N-X-S/T) at this site.
@@
 \subsection*{Validating escape mutation effect sizes with neutralization assays}
-
-\begin{figure}
-\centerline{\includegraphics[width=0.95\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
-\caption{
-\label{fig:structures}
-{\bf Mutations selected by broad and narrow antibodies.}
-(A) The logo plots show the sites where mutations have the largest effect.
-Letter heights are proportional to the excess fraction of virions with that mutation that survive antibody, as indicated by the scale bars.
-The structures are colored white to red by the excess fraction surviving for the largest-effect mutation at each site, with each antibody scaled separately.  
-(B) Sites of selection from anti-stalk antibodies, with the same coloring scale for both antibodies.
-(C) Cladogram of HA subtypes, with group 1 in black and group 2 in gray. 
-The amino acid at site 38 is indicated. 
-Colors indicate whether a subtype is bound or neutralized by C179 if this has been tested.  
-}
-\end{figure}
 

### 2017-10-19 paper/manuscript.tex

@@
 
-Specifically, for S139/1, we observed strong escape mutations at sites 156, 158, and 193 (Figure~\ref{fig:structures}A; sites are in H3 numbering). 
-These three sites surround HA's receptor-binding pocket, and fall directly in the physical binding footprint of the antibody~\citep{lee2012heterosubtypic}.
+Specifically, for broad anti-receptor binding pocket antibody S139/1, we observe strong escape mutations at sites 156, 158, and 193 (Figure~\ref{fig:structures}A; sites are in H3 numbering). 
+These three sites fall directly in the physical binding footprint of the antibody~\citep{lee2012heterosubtypic}.
 These are also the same three sites where \citet{yoshida2009cross} selected escape mutants in H1, H2, and H3 HAs. 
 Our data show that numerous different amino-acid mutations at each site confer neutralization resistance.
-The mutation with the largest effect, D158N, introduces an N-linked glycosylation site.
+The mutation with the largest effect, D158N, introduces an N-linked glycosylation motif.
 
-The mutation contributing the greatest fraction of virions surviving selection, G158N, introduces a glycosylation motif (N-X-S/T) at this site.
-C179 and FI6v3, which share very similar binding footprints and angles of approach~\citep{dreyfus2013structure}, select for identical escape mutations at several sites (Figure~\ref{fig:structures}B). 
-Both antibodies select for the K280S, K280T, and N291S mutations, all of which introduce glycosylation motifs in close proximity to the epitope and all have similar effect sizes in conferring escape from either antibody.
+Although anti-stalk antibodies C179 and FI6v3 only select mutations with small effects, these mutations almost all fall in the physical binding footprints of the antibodies (Figure~\ref{fig:structures}A).
+The two antibodies have similar epitopes and angles of approach~\citep{dreyfus2013structure}, and they select identical mutations at several sites (Figure~\ref{fig:structures}B). 
+The three largest-effect mutations for FI6v3 (K280S, K280T, and N291S) all introduce N-linked glycosylation motifs near the epitope, and all three mutations also have a similar effect size for C179.
 
-Despite the similarities in the epitope-targeting of C179 and FI6v3, we also see selection for distinct escape mutations with either antibody.
-The mutation with the strongest effect size from selection by C179 occurs at site 38, while we do not see observable escape mutations at this site by selection with FI6v3.
-Interestingly, we found that HA subtypes susceptible to binding or neutralization by C179 tend to possess a histidine at this site, but subtypes not bound or neutralized by C179 possess a glutamine or serine (Figure~\ref{fig:structures}C).
-This would suggest that susceptibility to neutralization by C179 may in part be determined by the amino-acid identity at site 38.
-Furthermore, the difference in breadth between C179 and FI6v3 has been attributed to FI6v3 being able to accommodate a glycan on the Asn at site 38 which is present in group 2 but not group 1 HA's, while group 1-specific stalk-targeting antibodies clash with this glycan~\citep{corti2011neutralizing, sui2009structural, ekiert2009antibody}.
-The H38S mutation selected by C179 does not introduce a glycosylation motif, which indicates that introduction of a glycan at this site is not the only means of mediating escape from this group 1-specific anti-stalk antibody.
-The amino-acid identity at site 111 of HA2 can change the orientation of a conserved Trp21 in HA2, also resulting in group 1 and group 2 differences in the binding ability of stalk-targeting antibodies.
+However, C179 also selects several mutations that do not have any apparent effect on recognition by FI6v3 (Figure~\ref{fig:structures}A, Figure~\ref{suppfig:FI6v3logo}).
+The most notable of these C179-specific mutations occurs at site 38.
+The additional breadth of FI6v3 over antibodies such as C179 that neutralize only group 1 HAs has been attributed to FI6v3's ability to accommodate a glycan on the asparagine at site 38 that is present in group 2 but not group 1 HAs~\citep{corti2011neutralizing,sui2009structural,ekiert2009antibody}.
+However, the H38S mutation that has the largest effect on C179 resistance in our experiments does not introduce a glycosylation motif, showing that there are also other ways to escape anti-stalk antibodies at this site.
+Interestingly, group 1 HA subtypes that are susceptible by C179 tend to possess a histidine at site 38, but subtypes that are not bound or neutralized by C179 often possess a serine (Figure~\ref{fig:structures}C).
+%The amino-acid identity at site 111 of HA2 can change the orientation of a conserved Trp21 in HA2, also resulting in group 1 and group 2 differences in the binding ability of stalk-targeting antibodies.
 An H111T mutation in the A/South Carolina/1/1918 (H1N1) has been reported to abrogate C179 binding~\citep{dreyfus2013structure}.
-However, we did not observe escape mutations at site 111, but this may be due to strain-specific differences.
+%However, we did not observe escape mutations at site 111, but this may be due to strain-specific differences.
+
+\comment{still need to add something about signal peptide mutation, maybe refer to H17* antibodies too}
 

### 2017-10-21 paper/manuscript.tex
added British Medical Journal citation

@@
 In contrast, human influenza virus exhibits much more antigenic variation.
-So although infection with an influenza virus strain provides long-term immunity to that exact strain~\citep{davies1982christ,yu2008neutralizing}, the virus's rapid antigenic evolution erodes the effectiveness of this immunity to that strain's descendants within four to seven years~\citep{couch1983immunity}.
+So although infection with an influenza virus strain provides long-term immunity to that exact strain~\citep{fluinboardingschool1978,davies1982christ,yu2008neutralizing}, the virus's rapid antigenic evolution erodes the effectiveness of this immunity to that strain's descendants within four to seven years~\citep{couch1983immunity}.
 

### 2017-10-22 paper/manuscript.tex
small text revisions

@@
 All experiments were performed in full biological triplicate using three independently generated virus libraries carrying all single amino-acid mutations to HA~\citep{doud2016accurate}.
-We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three narrow strain-specific antibodies, and we re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived neutralization.
+%We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three narrow strain-specific antibodies, and we re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived neutralization.
 The correlations among replicates are in Figure~\ref{suppfig:corr}.
@@
 For all four of these antibodies, there are multiple sites in HA where mutations enable a substantial fraction of virions to survive high antibody concentrations (Figure~\ref{fig:avgfracsurvive}).
-Specifically, for all four antibodies there are mutations that enable over a third of virions to survive at concentrations where virtually all wildtype virions are neutralized (Figure~\ref{suppfig:maxfracsurvive}).
+Specifically, there are mutations that enable over a third of virions to survive at concentrations where virtually all wildtype virions are neutralized (Figure~\ref{suppfig:maxfracsurvive}).
 Therefore, the virus can escape these four antibodies with the sort of large-effect single amino-acid mutations that characterize traditional influenza antigenic drift~\citep{yewdell1979antigenic,webster1980determination,koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction}.
@@
 If we look at the results for these antibodies on the same scale as the other antibodies, we see only a few small bumps in the fraction of virions surviving (Figures~\ref{fig:avgfracsurvive} and \ref{suppfig:maxfracsurvive}).
-Only if we zoom in a great deal can we see that there are actually a few sites where mutations slightly increase the fraction of virions surviving C179 and FI6v3 (Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo}).
+Only if we zoom in can we see that there are actually a few sites where mutations slightly increase the fraction of virions surviving C179 and FI6v3 (Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo}).
 But the effect sizes of these antigenic mutations are tiny compared to the other antibodies -- especially for FI6v3.
@@
 
-Specifically, for broad anti-receptor binding pocket antibody S139/1, we observe strong escape mutations at sites 156, 158, and 193 (Figure~\ref{fig:structures}A; sites are in H3 numbering). 
+For the broad anti-receptor binding pocket antibody S139/1, we observe strong escape mutations at sites 156, 158, and 193 (Figure~\ref{fig:structures}A; sites are in H3 numbering). 
 These three sites fall directly in the physical binding footprint of the antibody~\citep{lee2012heterosubtypic}.
@@
 
-However, C179 also selects several mutations that do not have any apparent effect on recognition by FI6v3 (Figure~\ref{fig:structures}A, Figure~\ref{suppfig:FI6v3logo}).
+However, C179 selects several mutations that do not have any apparent effect on FI6v3 (Figure~\ref{fig:structures}A, Figure~\ref{suppfig:FI6v3logo}).
 The most notable of these C179-specific mutations occurs at site 38.
@@
 
+Interestingly, FI6v3 also weakly selects several mutations at residue -8, which is part of HA's signal peptide (Figure~\ref{fig:structures}A). 
+This signal peptide is cleaved from the mature HA protein~\citep{daniels2003n,burke2014recommended}, so presumably these mutations are affecting the density of HA trimers on the virion~\citep{nordholm2017translational} rather than the physical binding of the antibody to HA.
+
 \comment{still need to add something about signal peptide mutation, maybe refer to H17* antibodies too}
 
+\subsection*{Validating escape mutation effect sizes with neutralization assays}
+
 \comment{pdmH1N1 has an NTT glycosylation motif at sites 278-280. Its IC50 for FI6v3 is around 4-fold higher than the IC50 for WSN, according to Corti et al 2011}
-
-\subsection*{Validating escape mutation effect sizes with neutralization assays}
 
@@
 (B), (C) In contrast to FI6v3, mutations selected by other antibodies have very large effects on neutralization.
-Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L2 taken from \citet{doud2017complete}; that paper also shows additional examples of large-effect escape mutations from these antibodies.
+Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 taken from \citet{doud2017complete}; that paper also shows additional examples of large-effect escape mutations from these antibodies.
 }

### 2017-10-22 paper/manuscript.tex
work on neut assay section of results

@@
 \begin{figure}[h!]
-\centerline{\includegraphics[width=0.91\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
+\centerline{\includegraphics[width=0.89\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
 \caption{
@@
 {\bf Mutations selected by broad and narrow antibodies.}
-(A) The logo plots show the sites where mutations have the largest effect.
+(A) Logo plots show sites where mutations have the largest effect.
 Letter heights are proportional to the excess fraction of virions with that mutation that survive antibody, as indicated by the scale bars.
-The structures are colored white to red by the excess fraction surviving for the largest-effect mutation at each site, with each antibody scaled separately.  
+Structures are colored white to red by the excess fraction surviving for the largest-effect mutation at each site, with each antibody scaled separately.  
 (B) Sites of selection from anti-stalk antibodies, with the same coloring scale for both antibodies.
@@
 The amino acid at site 38 is indicated. 
-Colors indicate whether a subtype is bound or neutralized by C179 if this has been tested.  
+Colors indicate whether a subtype is bound or neutralized by C179.  
 }
@@
 For the broad anti-receptor binding pocket antibody S139/1, we observe strong escape mutations at sites 156, 158, and 193 (Figure~\ref{fig:structures}A; sites are in H3 numbering). 
-These three sites fall directly in the physical binding footprint of the antibody~\citep{lee2012heterosubtypic}.
-These are also the same three sites where \citet{yoshida2009cross} selected escape mutants in H1, H2, and H3 HAs. 
+These three sites fall directly in the physical binding footprint of the antibody~\citep{lee2012heterosubtypic}, and are the same three sites where \citet{yoshida2009cross} selected escape mutants in H1, H2, and H3 HAs.
 Our data show that numerous different amino-acid mutations at each site confer neutralization resistance.
@@
 
-Although anti-stalk antibodies C179 and FI6v3 only select mutations with small effects, these mutations almost all fall in the physical binding footprints of the antibodies (Figure~\ref{fig:structures}A).
+Although the anti-stalk antibodies C179 and FI6v3 only select mutations with small effects, these mutations almost all fall in the physical binding footprints of the antibodies (Figure~\ref{fig:structures}A).
 The two antibodies have similar epitopes and angles of approach~\citep{dreyfus2013structure}, and they select identical mutations at several sites (Figure~\ref{fig:structures}B). 
-The three largest-effect mutations for FI6v3 (K280S, K280T, and N291S) all introduce N-linked glycosylation motifs near the epitope, and all three mutations also have a similar effect size for C179.
+The three largest-effect mutations for FI6v3 (K280S, K280T, and N291S) all introduce glycosylation motifs near the epitope, and all three mutations have similar magnitude antigenic effects in both FI6v3 and C179.
 
 However, C179 selects several mutations that do not have any apparent effect on FI6v3 (Figure~\ref{fig:structures}A, Figure~\ref{suppfig:FI6v3logo}).
-The most notable of these C179-specific mutations occurs at site 38.
-The additional breadth of FI6v3 over antibodies such as C179 that neutralize only group 1 HAs has been attributed to FI6v3's ability to accommodate a glycan on the asparagine at site 38 that is present in group 2 but not group 1 HAs~\citep{corti2011neutralizing,sui2009structural,ekiert2009antibody}.
+The most notable of these C179-specific mutations occur at site 38.
+The additional breadth of FI6v3 over antibodies such as C179 that neutralize only group 1 HAs is because FI6v3 can accommodate a glycan on the asparagine at site 38 that is present in group 2 HAs~\citep{corti2011neutralizing,sui2009structural,ekiert2009antibody}.
 However, the H38S mutation that has the largest effect on C179 resistance in our experiments does not introduce a glycosylation motif, showing that there are also other ways to escape anti-stalk antibodies at this site.
@@
 
-Interestingly, FI6v3 also weakly selects several mutations at residue -8, which is part of HA's signal peptide (Figure~\ref{fig:structures}A). 
-This signal peptide is cleaved from the mature HA protein~\citep{daniels2003n,burke2014recommended}, so presumably these mutations are affecting the density of HA trimers on the virion~\citep{nordholm2017translational} rather than the physical binding of the antibody to HA.
+Interestingly, FI6v3 weakly selects several mutations at residue -8, which is part of HA's signal peptide (Figure~\ref{fig:structures}A). 
+This signal peptide is cleaved from the mature HA protein~\citep{daniels2003n,burke2014recommended}, so mutations at this site may affect the density of HA trimers on the virion~\citep{nordholm2017translational} rather than the physical binding of the antibody to HA.
 
-\comment{still need to add something about signal peptide mutation, maybe refer to H17* antibodies too}
-
-\subsection*{Validating escape mutation effect sizes with neutralization assays}
-
-\comment{pdmH1N1 has an NTT glycosylation motif at sites 278-280. Its IC50 for FI6v3 is around 4-fold higher than the IC50 for WSN, according to Corti et al 2011}
+\subsection*{Neutralization assays validate the mutational antigenic profiling}
+Do the mutations identified in our mutational antigenic profiling actually have the expected effect on antibody neutralization?
+We have previously validated many of the large-effect antigenic mutations selected by the narrow antibodies H17-L19, H17-L10, and H17-L7~\citep{doud2017complete}.
+However, the mutations selected by the broad anti-stalk antibodies have much smaller effects -- especially for the broadest antibody, FI6v3.
+We therefore tested some of these FI6v3-selected mutations using neutralization assays on individual viral mutants.
 
@@
 \end{figure}
+
+Figure~\ref{fig:FI6v3neutcurves} shows that the mutational antigenic profiling is highly predictive of the results of the neutralization assays, even for small-effect mutations.
+As discussed in the previous section, the three mutations most strongly selected by FI6v3 introduce glycosylation motifs at sites 278-280 or 289-291 (Figure~\ref{fig:structures}A).
+We created viruses carrying each of these mutations (K280S, K280T, and N291S) and validated that all three mutants were modestly more resistant to FI6v3 (Figure~\ref{fig:FI6v3neutcurves}A).
+As a control, we also validated that a mutation at one of these sites (K280A) that does \emph{not} have an effect in our mutational antigenic profiling does not shift the neutralization curve (Figure~\ref{fig:FI6v3neutcurves}A).
+Interestingly, the HA of pandemic H1N1 has a glycosylation motif at 278-280, and \citet{corti2011neutralizing} have reported that this HA is $\sim$4-fold more resistant to FI6v3 than the A/WSN/1933 HA used in our experiments.
+This reported difference between strains is of similar magnitude to the effect that we measure for introducing the glycosylation motif via the K280S or K280T mutations.
+
+Our mutational antigenic profiling also identified several non-glycosylation-motif mutations that were selected by FI6v3.
+We validated that one of these mutations, G47R in the HA2 chain, indeed increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A) -- although as predicted by the mutational antigenic profiling, the magnitude of the effect was smaller than for the glycosylation-motif mutations.
+The most unexpected mutations identified in the mutational antigenic profiling were at site -8 in the signal peptide.
+We tested one of these mutations, K(-8)T, and found that it indeed caused a modest but detectable increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A).
+As controls, we also tested three more mutations (P80D, V135T, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them affected neutralization resistance.
+
+A remarkable aspect of these validation experiments is the very small effect sizes of the identified mutations.
+Antigenic mutations selected by strain-specific antibodies to HA generally increase the concentration of antibody needed to neutralize the virus by orders of magnitude.
+Neutralization curves for a few examples of such large-effect escape mutants are in Figure~\ref{fig:FI6v3neutcurves}B,C.
+Our work demonstrates that there are no such large-effect single mutations to the A/WSN/1933 HA when it comes to escaping from FI6v3 and C179.
+But the results in Figure~\ref{fig:FI6v3neutcurves}A show that we can still leverage the high sensitivity of mutational antigenic profiling to identify mutations that have much more modest but nonetheless measurable effects on antibody neutralization.
 

### 2017-10-22 paper/manuscript.tex
a few minor revisions to results

@@
 We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}, and calculated the fraction of virions carrying each mutation that survived neutralization.
-All experiments were performed in full biological triplicate using three independently generated virus libraries carrying all single amino-acid mutations to HA~\citep{doud2016accurate}.
+All experiments were performed in full biological triplicate using three independently generated virus libraries carrying single amino-acid mutations to HA~\citep{doud2016accurate}.
 %We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three narrow strain-specific antibodies, and we re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived neutralization.
@@
 
-For the broad anti-receptor binding pocket antibody S139/1, we observe strong escape mutations at sites 156, 158, and 193 (Figure~\ref{fig:structures}A; sites are in H3 numbering). 
+For the broad anti-receptor binding pocket antibody S139/1, there are strong escape mutations at sites 156, 158, and 193 (Figure~\ref{fig:structures}A; sites are in H3 numbering). 
 These three sites fall directly in the physical binding footprint of the antibody~\citep{lee2012heterosubtypic}, and are the same three sites where \citet{yoshida2009cross} selected escape mutants in H1, H2, and H3 HAs.
@@
 However, C179 selects several mutations that do not have any apparent effect on FI6v3 (Figure~\ref{fig:structures}A, Figure~\ref{suppfig:FI6v3logo}).
-The most notable of these C179-specific mutations occur at site 38.
+The most notable of these C179-specific mutations are at site 38.
 The additional breadth of FI6v3 over antibodies such as C179 that neutralize only group 1 HAs is because FI6v3 can accommodate a glycan on the asparagine at site 38 that is present in group 2 HAs~\citep{corti2011neutralizing,sui2009structural,ekiert2009antibody}.
 However, the H38S mutation that has the largest effect on C179 resistance in our experiments does not introduce a glycosylation motif, showing that there are also other ways to escape anti-stalk antibodies at this site.
-Interestingly, group 1 HA subtypes that are susceptible by C179 tend to possess a histidine at site 38, but subtypes that are not bound or neutralized by C179 often possess a serine (Figure~\ref{fig:structures}C).
+Interestingly, group 1 HA subtypes that are susceptible to C179 tend to possess a histidine at site 38, but subtypes that are not bound or neutralized by C179 often possess a serine (Figure~\ref{fig:structures}C).
 %The amino-acid identity at site 111 of HA2 can change the orientation of a conserved Trp21 in HA2, also resulting in group 1 and group 2 differences in the binding ability of stalk-targeting antibodies.
@@
 The mutations K280S, K280T, N291S, G47R (HA2), and K(-8)T are all expected to increase neutralization resistance based on the mutational antigenic profiling (Figure~\ref{fig:structures}A), and they all do so -- but only modestly.
-The mutations K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization based on the mutational antigenic profiling (Figure~\ref{suppfig:FI6v3logo}, and none of them do.
+The mutations K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization based on the mutational antigenic profiling (Figure~\ref{suppfig:FI6v3logo}), and none of them do.
 (B), (C) In contrast to FI6v3, mutations selected by other antibodies have very large effects on neutralization.
-Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 taken from \citet{doud2017complete}; that paper also shows additional examples of large-effect escape mutations from these antibodies.
+Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 taken from \citet{doud2017complete}.
 }
@@
 As discussed in the previous section, the three mutations most strongly selected by FI6v3 introduce glycosylation motifs at sites 278-280 or 289-291 (Figure~\ref{fig:structures}A).
-We created viruses carrying each of these mutations (K280S, K280T, and N291S) and validated that all three mutants were modestly more resistant to FI6v3 (Figure~\ref{fig:FI6v3neutcurves}A).
+We created viruses carrying each of these mutations (K280S, K280T, and N291S) and validated that all three modestly increased resistance to FI6v3 (Figure~\ref{fig:FI6v3neutcurves}A).
 As a control, we also validated that a mutation at one of these sites (K280A) that does \emph{not} have an effect in our mutational antigenic profiling does not shift the neutralization curve (Figure~\ref{fig:FI6v3neutcurves}A).
@@
 Our mutational antigenic profiling also identified several non-glycosylation-motif mutations that were selected by FI6v3.
-We validated that one of these mutations, G47R in the HA2 chain, indeed increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A) -- although as predicted by the mutational antigenic profiling, the magnitude of the effect was smaller than for the glycosylation-motif mutations.
-The most unexpected mutations identified in the mutational antigenic profiling were at site -8 in the signal peptide.
+We validated that one of these mutations, G47R in the HA2 chain, increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A) -- although as predicted by the mutational antigenic profiling, the magnitude of the effect was small.
+The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.
 We tested one of these mutations, K(-8)T, and found that it indeed caused a modest but detectable increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A).
@@
 
-A remarkable aspect of these validation experiments is the very small effect sizes of the identified mutations.
+A notable aspect of these validation experiments is the very small effect sizes of the identified mutations.
 Antigenic mutations selected by strain-specific antibodies to HA generally increase the concentration of antibody needed to neutralize the virus by orders of magnitude.
-Neutralization curves for a few examples of such large-effect escape mutants are in Figure~\ref{fig:FI6v3neutcurves}B,C.
-Our work demonstrates that there are no such large-effect single mutations to the A/WSN/1933 HA when it comes to escaping from FI6v3 and C179.
-But the results in Figure~\ref{fig:FI6v3neutcurves}A show that we can still leverage the high sensitivity of mutational antigenic profiling to identify mutations that have much more modest but nonetheless measurable effects on antibody neutralization.
+Neutralization curves for examples of such large-effect escape mutants are in Figure~\ref{fig:FI6v3neutcurves}B,C.
+Although there are no such large-effect single mutations that escape FI6v3 and C179, the results in Figure~\ref{fig:FI6v3neutcurves}A show that we can still use mutational antigenic profiling to identify mutations that have small but measurable effects on resistance to these antibodies.
 

### 2017-10-22 paper/manuscript.tex
work on last section of Results

@@
 
-A notable aspect of these validation experiments is the very small effect sizes of the identified mutations.
+A notable aspect of these validation experiments is the very small effect sizes of the identified mutations on neutralization by FI6v3.
 Antigenic mutations selected by strain-specific antibodies to HA generally increase the concentration of antibody needed to neutralize the virus by orders of magnitude.
@@
 Although there are no such large-effect single mutations that escape FI6v3 and C179, the results in Figure~\ref{fig:FI6v3neutcurves}A show that we can still use mutational antigenic profiling to identify mutations that have small but measurable effects on resistance to these antibodies.
+
+\subsection*{Limited HA mutational tolerance does not fully explain the lack of strong escape mutations from the anti-stalk antibodies}
+Why are there no large-effect escape mutations from the anti-stalk antibodies?
+One possibility is that all the HA sites in the antibody-binding interfaces are intolerant of mutations, meaning that viruses with mutations at these sites cannot replicate and so are not present in our mutant virus libraries.
+Another possibility is that mutations are tolerated at some HA sites in the antibody-binding interface, but that the binding energetics are distributed across sites in such a way that none of these tolerated mutations strongly affect neutralization.
+
+\begin{figure}
+\centerline{\includegraphics[width=\textwidth]{figs/prefs_fracsurvive/prefs_fracsurvive_logoplots.pdf}}
+\caption{
+\label{fig:muttolerance}
+CAPTION
+}
+\end{figure}
+
+We can examine these possibilities using deep mutational scanning data that measures the tolerance of HA for each possible amino-acid mutation.
+Specifically, we have previously selected our A/WSN/1933 virus HA mutant libraries for variants that can replicate in cell culture, and then used deep sequencing to estimate the preference of each site in HA for each possible amino acid~\citep{doud2016accurate}.
+Figure~\ref{fig:muttolerance} shows these amino-acid preferences for all sites in HA within 4 angstroms of each broad antibody, with the antigenic effects of the mutations overlaid.
+Although some sites in the HA-antibody interfaces strongly prefer a single amino acid, for all antibodies there are also interface sites that tolerate a fairly wide range of amino acids.
+In most cases the mutations selected by the antibodies occur at these mutationally tolerant sites.
+However, there are exceptions -- for instance, the H38S mutation selected by C179 is rather disfavored with respect to viral growth, but has a large enough antigenic effect to still be detected in our mutational antigenic profiling.
+
+The data in Figure~\ref{fig:muttolerance} show that the lack of large-effect escape mutants from FI6v3 and C179 is not due to complete mutational intolerance of the antibody-HA interface.
+Rather, only a fraction of mutations to HA sites in the interface actually reduce antibody neutralization.
+This finding is reminiscent of prior work showing that the binding energetics at protein-protein interfaces can be asymmetrically distributed across sites~\citep{jin1992high,cunningham1993comparison,dall1998mutational}.
+The broad anti-stalk antibodies appear to distribute their binding energetics in such a way that altering the mutationally tolerant HA sites has relatively little effect on neutralization.
 

### 2017-10-22 paper/manuscript.tex
added legend to mutational tolerance figure

@@
 \label{fig:muttolerance}
-CAPTION
+{\bf Mutational tolerance of HA sites in the antibody-binding interfaces.}
+These plots show all HA sites within 4 angstroms of the antibody in the crystal structure, plus any additional sites (marked with a *) where we identified antigenic mutations. 
+The logo plots at bottom show the preference of each HA site for each amino acid under selection for viral growth as measured by \citet{doud2016accurate}.
+For instance, site 153 in HA only tolerates tryptophan, so W occupies the entire height of the preference logo stack.
+In contrast, site 156 tolerates many amino acids, all of which contribute to the height of the preference logo stack. 
+Above the preference logo stacks are logo plots showing the excess fraction surviving antibody treatment as measured in the current study.
+Note that scale for these antigenic effects is 10$\times$ smaller for the FI6v3 and C179 than for S139/1.
+Overall, it is clear that many mutations that are tolerated at interface sites that do not have a detectable antigenic effect.
 }

### 2017-10-23 paper/manuscript.tex
a few more revisions to Results

@@
 {\bf Epitopes and breadth of broad and narrow antibodies targeting HA.}
-(A) Crystal structures of the broad antibodies and binding footprints of the narrow ones are superimposed on the structure of the HA trimer~\citep[PDB 1RVX;][]{gamblin2004structure}. 
+(A) Crystal structures of the broad antibodies and binding footprints of the narrow ones superimposed on the structure of the HA trimer~\citep[PDB 1RVX;][]{gamblin2004structure}. 
 S139/1~\citep[PDB 4GMS;][]{lee2012heterosubtypic} targets the receptor-binding site; C179~\citep[PDB 4HLZ;][]{dreyfus2013structure} and FI6v3~\citep[PDB 3ZTN;][]{corti2011neutralizing} target the stalk.
@@
 \subsection*{Broad and narrow antibodies that neutralize influenza virus}
-We applied our approach to anti-HA antibodies with a range of breadths and epitopes.
+We applied this approach to anti-HA antibodies with a range of breadths and epitopes.
 The crystal structures or binding footprints of these antibodies are shown in Figure~\ref{fig:antibody_summary}A.
@@
 
-Interestingly, FI6v3 weakly selects several mutations at residue -8, which is part of HA's signal peptide (Figure~\ref{fig:structures}A). 
-This signal peptide is cleaved from the mature HA protein~\citep{daniels2003n,burke2014recommended}, so mutations at this site may affect the density of HA trimers on the virion~\citep{nordholm2017translational} rather than the physical binding of the antibody to HA.
+The FI6v3 antibody also weakly selects several mutations at residue -8, which is part of HA's signal peptide (Figure~\ref{fig:structures}A). 
+This signal peptide is cleaved from the mature HA protein~\citep{daniels2003n,burke2014recommended}, so mutations at this site likely affect HA's expression~\citep{nordholm2017translational} and subsequently its density on virions rather than affecting the physical binding of the antibody to HA.
 
@@
 Antigenic mutations selected by strain-specific antibodies to HA generally increase the concentration of antibody needed to neutralize the virus by orders of magnitude.
-Neutralization curves for examples of such large-effect escape mutants are in Figure~\ref{fig:FI6v3neutcurves}B,C.
+Neutralization curves for such large-effect escape mutants are in Figure~\ref{fig:FI6v3neutcurves}B,C.
 Although there are no such large-effect single mutations that escape FI6v3 and C179, the results in Figure~\ref{fig:FI6v3neutcurves}A show that we can still use mutational antigenic profiling to identify mutations that have small but measurable effects on resistance to these antibodies.
@@
 Why are there no large-effect escape mutations from the anti-stalk antibodies?
-One possibility is that all the HA sites in the antibody-binding interfaces are intolerant of mutations, meaning that viruses with mutations at these sites cannot replicate and so are not present in our mutant virus libraries.
-Another possibility is that mutations are tolerated at some HA sites in the antibody-binding interface, but that the binding energetics are distributed across sites in such a way that none of these tolerated mutations strongly affect neutralization.
+One possibility is that all HA sites in the antibody-binding footprint are intolerant of mutations, meaning that viruses with mutations at these sites cannot replicate and so are not present in our mutant virus libraries.
+Another possibility is that mutations are tolerated at some HA sites in the antibody footprint, but that the binding energetics are distributed across sites in such a way that none of these tolerated mutations strongly affect neutralization.
 
@@
 \label{fig:muttolerance}
-{\bf Mutational tolerance of HA sites in the antibody-binding interfaces.}
+{\bf Mutational tolerance of HA sites in the antibody-binding footprints.}
 These plots show all HA sites within 4 angstroms of the antibody in the crystal structure, plus any additional sites (marked with a *) where we identified antigenic mutations. 
@@
 Note that scale for these antigenic effects is 10$\times$ smaller for the FI6v3 and C179 than for S139/1.
-Overall, it is clear that many mutations that are tolerated at interface sites that do not have a detectable antigenic effect.
+Overall, it is clear that many mutations that are tolerated at footprint sites that do not have a detectable antigenic effect.
 }
@@
 Figure~\ref{fig:muttolerance} shows these amino-acid preferences for all sites in HA within 4 angstroms of each broad antibody, with the antigenic effects of the mutations overlaid.
-Although some sites in the HA-antibody interfaces strongly prefer a single amino acid, for all antibodies there are also interface sites that tolerate a fairly wide range of amino acids.
+Although some HA sites in the antibody footprints strongly prefer a single amino acid, for all antibodies there are also footprint sites that tolerate a fairly wide range of amino acids.
 In most cases the mutations selected by the antibodies occur at these mutationally tolerant sites.
@@
 
-The data in Figure~\ref{fig:muttolerance} show that the lack of large-effect escape mutants from FI6v3 and C179 is not due to complete mutational intolerance of the antibody-HA interface.
-Rather, only a fraction of mutations to HA sites in the interface actually reduce antibody neutralization.
+The data in Figure~\ref{fig:muttolerance} show that the lack of large-effect escape mutants from FI6v3 and C179 is not due to complete mutational intolerance of HA sites in the antibody-binding footprints.
+Rather, only a fraction of mutations to HA sites in the footprint actually reduce antibody neutralization.
 This finding is reminiscent of prior work showing that the binding energetics at protein-protein interfaces can be asymmetrically distributed across sites~\citep{jin1992high,cunningham1993comparison,dall1998mutational}.

### 2017-10-23 paper/manuscript.tex
very rough outline of Discussion

@@
 The logo plots at bottom show the preference of each HA site for each amino acid under selection for viral growth as measured by \citet{doud2016accurate}.
-For instance, site 153 in HA only tolerates tryptophan, so W occupies the entire height of the preference logo stack.
+For instance, site 153 only tolerates tryptophan, so W occupies the entire height of the preference logo stack.
 In contrast, site 156 tolerates many amino acids, all of which contribute to the height of the preference logo stack. 
@@
 Note that scale for these antigenic effects is 10$\times$ smaller for the FI6v3 and C179 than for S139/1.
-Overall, it is clear that many mutations that are tolerated at footprint sites that do not have a detectable antigenic effect.
+Overall, it is clear that many mutations are tolerated at footprint sites but do not have a detectable antigenic effect.
 }
@@
 \section*{DISCUSSION}
-Some discussion
+We have quantified the ease of escape.
+There are clear differences among antibodies
+Stalk antibodies are harder to escape.
+
+However, despite the lack of strong escape for stalk, there are small-effect antigenic mutations.
+Points to an advantage over the classical approach which requires large effect.
+Also, it's nice that when we don't find anything we can make statements about what exists.
+
+Interesting that breadth does not directly associate with ease of escape.
+
+Important caveat is that we are looking at the effects of single mutations in the background of A/WSN/1933.
+Things could be different in other backgrounds and/or with combinations~\citep{gong2013stability,harms2014historical,starr2017alternative}.
+\citet{wu2017diversity}.
+Nonetheless, given that HA antigenic drift mostly involves point mutations...
+
+Overall, we have described a new hybrid experimental-analytical approach to quantify antigenic effects.
+Useful both for characterizing the epitopes and escape mutants for single antibodies, but more importantly for comparing across antibodies.
+As antibodies are increasingly considered for anti-viral therapeutics, the ability to comprehensively quantify the potential for viral escape will be useful
 

### 2017-10-23 paper/manuscript.tex
some progress on Discussion

@@
 \section*{DISCUSSION}
-We have quantified the ease of escape.
-There are clear differences among antibodies
-Stalk antibodies are harder to escape.
+We have quantified the effects of all single amino-acid mutations to HA on influenza virus neutralization by a collection of broad and narrow antibodies.
+Our results show that the virus's inherent evolutionary capacity for escape via point mutations differs across antibodies. 
+Interestingly, antibody breadth is not always an indicator of the difficulty of viral escape. 
+As expected, single amino-acid mutations can make the virus completely resistant to narrow strain-specific antibodies against HA's globular head.
+However, such mutations can also enable the virus to escape a broad antibody targeting HA's receptor-binding pocket, despite the fact that this antibody broadly neutralizes multiple HA subtypes.
+But at least for the virus strain used here, no single mutations have a comparably large effect on neutralization by two broad antibodies targeting HA's stalk.
+Therefore, these stalk antibodies are quantifiably more difficult for the virus to escape.
+
+Although there are no large-effect escape mutations from the broad anti-stalk antibodies, there are mutations that more modestly affect neutralization.
+This finding emphasizes the importance of identifying antigenic mutations in a way that accounts for effect sizes.
+The classic approach for selecting escape mutations involves treating a virus stock with antibody at a saturating concentration that completely neutralizes wildtype, and looking for viral mutants that survive this treatment~\citep{yewdell1979antigenic,webster1980determination}.
+There are no such single mutations for the HA and broad anti-stalk antibodies tested here, since no mutations shift the neutralization curve enough to enable a mutant to survive antibody concentrations that fully neutralize wildtype.
+However, the comprehensive and sensitive nature of our approach enable us to identify a variety of mutations that have modest ($<$10-fold) effects on neutralization by even the broadest antibody.
 
@@
 
-Interesting that breadth does not directly associate with ease of escape.
-
 Important caveat is that we are looking at the effects of single mutations in the background of A/WSN/1933.
 Things could be different in other backgrounds and/or with combinations~\citep{gong2013stability,harms2014historical,starr2017alternative}.
+\citet{das2013defining}
 \citet{wu2017diversity}.
 Nonetheless, given that HA antigenic drift mostly involves point mutations...
+
+Virus's potential for escape is ultimately combination of how easy plus how strong selection pressure...
 

### 2017-10-23 paper/manuscript.tex
added epistasis paragraph to Discussion

--- 2017-10-23 paper/manuscript.tex
@@
 \section*{DISCUSSION}
-We have quantified the ease of escape.
-There are clear differences among antibodies
-Stalk antibodies are harder to escape.
+We have quantified how all single amino-acid mutations to influenza virus HA affect neutralization by a collection of broad and narrow antibodies.
+Our results show that the virus's inherent evolutionary capacity for escape via point mutations differs across antibodies. 
+Interestingly, antibody breadth is not always an indicator of the difficulty of viral escape. 
+As expected, single amino-acid mutations can make the virus completely resistant to narrow strain-specific antibodies against HA's globular head.
+However, such mutations can also enable the virus to escape the broad S139/1 antibody targeting HA's receptor-binding pocket, despite the fact that this antibody neutralizes multiple HA subtypes.
+But no single mutations have a comparably large effect on neutralization by two broad antibodies targeting HA's stalk, FI6v3 and C179.
+Therefore, these stalk antibodies are quantifiably more difficult for the virus to escape.
 
-However, despite the lack of strong escape for stalk, there are small-effect antigenic mutations.
-Points to an advantage over the classical approach which requires large effect.
-Also, it's nice that when we don't find anything we can make statements about what exists.
+Although there are no large-effect escape mutations from the broad anti-stalk antibodies, there are mutations that more modestly affect neutralization.
+This finding emphasizes the importance of identifying antigenic mutations in a way that accounts for effect sizes.
+The classic approach for selecting escape mutations involves treating a virus stock with antibody at a concentration that completely neutralizes wildtype, and looking for viral mutants that survive this treatment~\citep{yewdell1979antigenic,webster1980determination}.
+There are no such single mutations for the HA and broad anti-stalk antibodies tested here, since no mutations shift the neutralization curve enough to enable survival at antibody concentrations that fully neutralize wildtype.
+However, our approach shows that there are mutations that have more modest ($<$10-fold) effects on neutralization by even the broadest antibody.
 
-Interesting that breadth does not directly associate with ease of escape.
-
-Important caveat is that we are looking at the effects of single mutations in the background of A/WSN/1933.
-Things could be different in other backgrounds and/or with combinations~\citep{gong2013stability,harms2014historical,starr2017alternative}.
-\citet{wu2017diversity}.
-Nonetheless, given that HA antigenic drift mostly involves point mutations...
+It is important to note that our experiments examine \emph{single} amino-acid mutations to the HA from one influenza virus strain.
+The protein evolution literature is full of examples of epistatic interactions that enable multiple mutations to access phenotypes not accessible by single mutations~\citep{gong2013stability,harms2014historical,starr2017alternative}.
+Such epistasis is relevant to HA's evolution.  
+For instance, work by \citet{das2013defining} suggests that the sequential accumulation of mutations can shift the spectrum of available antibody-escape mutations.
+\citet{wu2017diversity} have used deep mutational scanning to directly demonstrate that rampant epistasis enables HA's receptor-binding pocket to accommodate combinations of individually deleterious mutations, some of which affect sensitivity to antibodies.
+Therefore, our work does not imply any absolute limits on the possibilities for antibody escape when evolution is given sufficient time to explore combinations of mutations.
+However, single mutations are the most accessible form of genetic variation, and much of influenza virus's natural antigenic drift involves individual mutations that reduce sensitivity to immunodominant antibody specificities~\citep{yewdell1979antigenic,webster1980determination,koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction}.
+Quantifying the antigenic effects of all such mutations therefore provides a relevant measure of ease of viral antibody escape.
 

### 2017-10-23 paper/manuscript.tex
full draft of Discussion

@@
 
-Overall, we have described a new hybrid experimental-analytical approach to quantify antigenic effects.
-Useful both for characterizing the epitopes and escape mutants for single antibodies, but more importantly for comparing across antibodies.
-As antibodies are increasingly considered for anti-viral therapeutics, the ability to comprehensively quantify the potential for viral escape will be useful
+A major rationale for studying broadly neutralizing antibodies is that they are hoped to be more resistant to viral evolutionary escape than the antibodies that dominate natural immune responses to influenza virus~\citep{krammer2015advances,corti2017tackling}.
+We have used a new approach to quantify the extent to which this is actually true, and shown that anti-viral activity of broad anti-stalk antibodies is indeed more -- although certainly not completely -- resistant to erosion by viral point mutations.
+Going forward, we suggest that quantifying the ease of viral escape will be a useful complement to more traditional techniques that characterize the structural epitopes and strain breadth of anti-viral antibodies.
 

### 2017-10-23 paper/manuscript.tex
a few very small edits to main text

@@
 The mutations K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization based on the mutational antigenic profiling (Figure~\ref{suppfig:FI6v3logo}), and none of them do.
-(B), (C) In contrast to FI6v3, mutations selected by other antibodies have very large effects on neutralization.
+(B), (C) In contrast to FI6v3, mutations selected by narrow antibodies have very large effects on neutralization.
 Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 taken from \citet{doud2017complete}.
@@
 We tested one of these mutations, K(-8)T, and found that it indeed caused a modest but detectable increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A).
-As controls, we also tested three more mutations (P80D, V135T, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them affected neutralization resistance.
+As controls, we also tested three mutations (P80D, V135T, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them affected neutralization resistance.
 
@@
 These plots show all HA sites within 4 angstroms of the antibody in the crystal structure, plus any additional sites (marked with a *) where we identified antigenic mutations. 
-The logo plots at bottom show the preference of each HA site for each amino acid under selection for viral growth as measured by \citet{doud2016accurate}.
+The logo plots at bottom show the preference of each HA site for each amino acid under selection for viral replication as measured by \citet{doud2016accurate}.
 For instance, site 153 only tolerates tryptophan, so W occupies the entire height of the preference logo stack.
@@
 As expected, single amino-acid mutations can make the virus completely resistant to narrow strain-specific antibodies against HA's globular head.
-However, such mutations can also enable the virus to escape the broad S139/1 antibody targeting HA's receptor-binding pocket, despite the fact that this antibody neutralizes multiple HA subtypes.
+However, such mutations can also enable the virus to escape the broad S139/1 antibody targeting HA's receptor-binding pocket, despite the fact that this antibody neutralizes multiple subtypes.
 But no single mutations have a comparably large effect on neutralization by two broad antibodies targeting HA's stalk, FI6v3 and C179.
-Therefore, these stalk antibodies are quantifiably more difficult for the virus to escape.
+Therefore, these anti-stalk antibodies are quantifiably more difficult for the virus to escape.
 

### 2017-10-23 paper/manuscript.tex
work on Methods up to neut assays

@@
 \subsection*{Antibodies}
-FI6v3 was expressed and purified by the Fred Hutchinson Cancer Research Center protein expression core \comment{i think?}.
-C179 was purchased from Takara Bio Inc (Catalog \# M145).
-S139/1 heavy and light chain variable sequences were obtained from PDB ID 4GMS~\cite{lee2012heterosubtypic} and expressed and purified by the Fred Hutchinson Cancer Research Center protein expression core.
+C179 IgG was purchased from Takara Bio Inc (Catalog \#M145).
+FI6v3 was purified from 293F cells transduced with a lentiviral vector encoding a commercially synthesized gene for the IgG form of the antibody, with the heavy and light chains reverse-translated from the protein sequence in the PDB structure 3ZTN~\citep{corti2011neutralizing} as described in \citet{balazs2013broad}.
+Genes encoding S139/1 in IgG form were were reverse-translated from the protein sequence in PDB structure 4GMS~\citep{lee2012heterosubtypic}, and used to express and purify protein by the Fred Hutchinson Cancer Research Center protein expression core.
 
 \subsection*{Neutralization assays}
-We performed neutralization assays as previously described in~\citep{hooper2013mutant} using WSN GFP reporter viruses.
-Viruses carrying GFP in the PB1 segment were grown using reverse genetics plasmids in complementing 293T-CMV-PB1 and MDCK-SIAT1-CMV-PB1 cells that constitutively express the PB1 protein.
+We performed neutralization assays using influenza viruses that carried GFP in the PB1 segment.
+These PB1flank-eGFP were generated in co-cultures of 293T-CMV-PB1 and MDCK-SIAT1-CMV-PB1 cells as described in \citet{bloom2010permissive}, using the standard bi-directional pHW181-PB2, ..., pHW188-NS reverse-genetics plasmids~\citep{hoffmann2000dna} for all genes \emph{except} PB1, plus the pHH-PB1flank-eGFP plasmid~\citep{bloom2010permissive}.
+Each mutant was generated repeating this process using a version of the pHW184-HA plasmid that had been engineered by site-directed mutagenesis to carry the indicated mutation.
+The neutralization assays themselves were performed by using a plate reader to quantify the GFP signal produced by MDCK-SIAT1-CMV-PB1 cells infected by PB1flank-eGFP virus that had been incubated with the indicated antibody concentration as described in \citet{hooper2013mutant}.
+All neutralization curves represent the mean and standard error of three measurements.
 
@@
 \subsection*{ACKNOWLEDGMENTS}
-This work was supported by grants R01GM102198 and R01AI127893 from the NIGMS and NIAID of the NIH.
+We thank Adam Dingens, Lauren Gentles, and Jeremy Roop for helpful comments on the project and manuscript.
+We thank the Fred Hutchinson Cancer Research Center genomics core for performing the Illumina deep sequencing, and the protein expression core for expressing and purifying the C179 antibody.
+This work was supported by grant R01AI127893 from the NIAID of the NIH.
 MBD was supported in part by training grant T32AI083203 from the NIAID of the NIH.
-JML was supported in part by \comment{CIDID grant}.
+JML was supported in part by the Center for Inference and Dynamics of Infectious Diseases (CIDID), which is funded by grant U54GM111274 from the NIGMS of the NIH.
 The research of JDB is supported in part by a Faculty Scholar Grant from the Howard Hughes Medical Institute and the Simons Foundation.
+The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
 

### 2017-10-23 paper/manuscript.tex
added fracsurvive supp table

@@
 \renewcommand{\thesuppfigure}{S\arabic{suppfigure}}
+\DeclareFloatingEnvironment[name={Table}]{supptable}
+\renewcommand{\thesupptable}{S\arabic{supptable}}
 
@@
 
-\subsection*{Mutant virus selections with antibody and deep sequencing}
-Library selections and deep sequencing were performed as previously described in~\cite{doud2017complete}. 
-We conducted the antibody selections using the mutant virus libraries from~\cite{doud2016accurate}. 
-We diluted the virus libraries to a concentration of $1 \times 10^{6}$ TCID$_{50}$ per ml and incubated the virus dilutions with an equal volume of antibody at the intended concentration.
-We then incubated the virus-antibody mixture at 37$^\circ$C for 1.5 hours
-Each selection was done in biological triplicate at the antibody concentrations shown in Figure~\ref{fig:neutcurves} concentrations, and technical replicates of some of the concentrations were also performed.
-To estimate the fraction of the overall library surviving selection, denoted as $\gamma$ in this paper, we made duplicate 10-fold serial dilutions of each of the virus libraries to use as a standard curve of infectivity.
+\subsection*{Mutant virus libraries}
+The mutant virus libraries are those described in \citet{doud2016accurate}, and were produced in full biological triplicate.
+Briefly, these libraries were generated by using codon mutagenesis~\citep{bloom2014experimentally} to introduce random codon mutations into plasmid-encoded HA, and then generating the virus libraries using a helper-virus strategy that avoids the bottlenecks associated with standard influenza reverse genetics.
+Full details of the library generation and sequencing statistics that quantify how completely each of the triplicate libraries covers the possible amino-acid mutations are in \citet{doud2016accurate}.
+
+\subsection*{Mutational antigenic profiling}
+The mutational antigenic profiling was performed as described in \citet{doud2017complete}. 
+Briefly, we diluted each virus libraries to a concentration of $10^{6}$ TCID$_{50}$ per ml and incubated the virus dilutions with an equal volume of antibody at the intended concentration at 37$^\circ$C for 1.5 hours.
+The final antibody concentrations in these mixtures are shown in Figure~\ref{fig:neutcurves} concentrations.
+We performed three fully independent replicates of each selection using the three replicate mutant virus libraries.
+In addition, we performed technical replicates (independent neutralization experiments on the \emph{same} virus library) in some cases as indicated in Figure~\ref{suppfig:corr}.
+The virus-antibody mixtures were used to infect cells, and viral RNA was extracted, reverse-transcribed, and PCR amplified as described in \citet{doud2017complete}.
+In order to obtain high accuracy in the Illumina deep sequencing, we used the barcoded-subamplicon sequencing strategy described in \citet{doud2017complete}, which is a slight modification of the strategy of \citet{wu2014high}.
+
+We also estimated the overall fraction of virions surviving each antibody selection.
+These fractions are denoted by $\gamma$ in this paper.
+The average of these fractions across libraries are reported in Figure~\ref{fig:neutcurves}, and the values for each individual replicate are in Table~\ref{supptab:fracsurvive}.
+\comment{I AM HERE}
+we made duplicate 10-fold serial dilutions of each of the virus libraries to use as a standard curve of infectivity. 
 qRT-PCR of the standard curves were performed as previously described in~\cite{doud2017complete} using qPCR primers for WSN influenza nucleoprotein (NP) to quantify viral infectivity and primers for canine GAPDH to correct for RNA amounts.
@@
 
+\begin{supptable}
+\centering
+\begin{tabular}{cccc}
+antibody &  concentration ($\mu$g/ml) &  replicate &  fraction surviving \\
+\hline
+FI6v3 &            0.1 &  1a &             0.01662 \\
+ FI6v3 &            0.1 &  1b &             0.01390 \\
+FI6v3 &            0.2 &  1a &             0.00465 \\
+FI6v3 &            0.2 &  1b &             0.00345 \\
+FI6v3 &            0.1 &  2 &             0.02322 \\
+FI6v3 &            0.2 &  2 &             0.00278 \\
+FI6v3 &            0.1 &  3 &             0.00903 \\
+FI6v3 &            0.2 &  3 &             0.00144 \\
+S139/1 &          100.0 &  1 &             0.02490 \\
+S139/1 &          200.0 &  1 &             0.01470 \\
+S139/1 &          300.0 &  1 &             0.01270 \\
+S139/1 &          100.0 &  2 &             0.02190 \\
+S139/1 &          200.0 &  2 &             0.01720 \\
+S139/1 &          300.0 &  2 &             0.00854 \\
+S139/1 &          100.0 &  3 &             0.05180 \\
+S139/1 &          200.0 &  3 &             0.04060 \\
+S139/1 &          300.0 &  3 &             0.03750 \\
+C179 &            1.0 &    1a &             0.00941 \\
+C179 &            1.0 &    1b &             0.00890 \\
+C179 &            1.0 &    1c &             0.00960 \\
+C179 &            2.5 &    1 &             0.00450 \\
+C179 &            1.0 &    2 &             0.00554 \\
+C179 &            2.5 &    2 &             0.00198 \\
+C179 &            1.0 &    3 &             0.00256 \\
+C179 &            2.5 &    3 &             0.00100 \\
+\end{tabular}
+\caption{\label{supptab:fracsurvive}
+The total fraction of virions surviving each antibody treatment at each concentration as estimated by qPCR.
+These are the quantities referred to as $\gamma$.
+This table shows the values for the broad antibodies; values for the narrow H17-L17, H17-L10, and H17-L7 antibodies are reported in \citet{doud2017complete}.
+}
+\end{supptable}
+
 \end{document}

### 2017-10-23 paper/manuscript.tex
H3 numbering subsection of Methods

@@
 \renewcommand{\thesupptable}{S\arabic{supptable}}
+\DeclareFloatingEnvironment[name={File}]{suppfile}
+\renewcommand{\thesuppfile}{S\arabic{suppfile}}
 
@@
 
+\subsection*{Construction of HA phylogenetic tree}
+We downloaded one HA sequence per subtype from the NIAID Influenza Research Database (IRD) \comment{cite}, built the phylogenetic tree using RaxML \comment{cite}, and visualized the tree using FigTree \comment{cite}.
+The HA sequences we downloaded were: H1 (A/WSN/1933); pdmH1N1 (A/California/04/2009); H2 (A/Japan/305/1957); H3 (A/Aichi/2/1968); H4 (A/swine/Ontario/01911/2/1999); H5 (A/Vietnam/1203/2004); H6 (A/chicken/Taiwan/0705/1999); H7 (A/Netherlands/219/2003); H8 (A/turkey/Ontario/6118/1968); H9 (A/swine/HongKong/9/1998); H10 (A/mallard/Bavaria/3/2006); H11 (A/duck/England/1/1956); H12 (A/duck/Alberta/60/1976); H13 (A/gull/Maryland/704/1977); H14 (A/mallard/Astrakhan/263/1982); H15 (A/mallard duck/Sweden/139579/2012); H16 (A/black headed gull/Mongolia/1756/2006); H17 (A/little yellow-shouldered bat/Guatemala/060/2010); H18 (A/dark fruit-eating bat/Bolivia/PBV780-781/2011).
+
+For the cladogram in Figure~\ref{fig:structures}C, the amino-acid identities at site 38 are from the strains tested against C179 in~\cite{dreyfus2013structure}.
+For subtypes not tested, we used the identity from the corresponding strain listed above.
+
+We denoted HA's with reported binding or neutralization activity by each antibody, and defined binding activity as \comment{binding to full influenza virus, VLPs, recombinant or purified HA's}.
+Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8 and H11; no antibodies have been tested against H17 and H18.
+
+\subsection*{H3 sequence numbering}
+Unless otherwise indicated, all residues are numbered in the H3 numbering scheme, with the signal peptide in negative numbers, the HA1 subunit as plain numbers, and the HA2 subunit denoted with ``(HA2)''.
+The conversion between sequential numbering of the A/WSN/1933 HA and the H3 numbering scheme was performed using the Python script available at \url{https://github.com/jbloomlab/HA_numbering}.
+File~\ref{suppfile:HAnumbering} gives the numbering conversion.
+
 \subsection*{Mutant virus libraries}
@@
 The average of these fractions across libraries are reported in Figure~\ref{fig:neutcurves}, and the values for each individual replicate are in Table~\ref{supptab:fracsurvive}.
-\comment{I AM HERE}
-we made duplicate 10-fold serial dilutions of each of the virus libraries to use as a standard curve of infectivity. 
-qRT-PCR of the standard curves were performed as previously described in~\cite{doud2017complete} using qPCR primers for WSN influenza nucleoprotein (NP) to quantify viral infectivity and primers for canine GAPDH to correct for RNA amounts.
-Next, we fit a linear regression of the log(infectious dose) versus the mean difference in qPCR Ct values between NP and GAPDH to estimate the total library fraction infectivity after antibody selection.
-
-\subsection*{Construction of HA phylogenetic tree}
-We downloaded one HA sequence per subtype from the NIAID Influenza Research Database (IRD) \comment{cite}, built the phylogenetic tree using RaxML \comment{cite}, and visualized the tree using FigTree \comment{cite}.
-The HA sequences we downloaded were: H1 (A/WSN/1933); pdmH1N1 (A/California/04/2009); H2 (A/Japan/305/1957); H3 (A/Aichi/2/1968); H4 (A/swine/Ontario/01911/2/1999); H5 (A/Vietnam/1203/2004); H6 (A/chicken/Taiwan/0705/1999); H7 (A/Netherlands/219/2003); H8 (A/turkey/Ontario/6118/1968); H9 (A/swine/HongKong/9/1998); H10 (A/mallard/Bavaria/3/2006); H11 (A/duck/England/1/1956); H12 (A/duck/Alberta/60/1976); H13 (A/gull/Maryland/704/1977); H14 (A/mallard/Astrakhan/263/1982); H15 (A/mallard duck/Sweden/139579/2012); H16 (A/black headed gull/Mongolia/1756/2006); H17 (A/little yellow-shouldered bat/Guatemala/060/2010); H18 (A/dark fruit-eating bat/Bolivia/PBV780-781/2011).
-
-For the cladogram in Figure~\ref{fig:structures}C, the amino-acid identities at site 38 are from the strains tested against C179 in~\cite{dreyfus2013structure}.
-For subtypes not tested, we used the identity from the corresponding strain listed above.
-
-We denoted HA's with reported binding or neutralization activity by each antibody, and defined binding activity as \comment{binding to full influenza virus, VLPs, recombinant or purified HA's}.
-Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8 and H11; no antibodies have been tested against H17 and H18.
-
-\subsection*{H3 sequence numbering}
-Unless otherwise indicated, all residues are numbered in H3 numbering in this paper, with the signal peptide in negative numbers, the HA1 subunit as plain numbers, and the HA2 subunit denoted with "(HA2)".
-We converted the sequential residue numbers into H3 numbering by running the Python script available on \url{https://github.com/jbloomlab/HA_numbering}.
-In this script, we used MUSCLE to align the WSN/1933 H1 HA sequence to the 4HMG pdb sequence for the A/Aichi/2/1968 H3 HA \comment{cite Weis1990JMB}, which allowed us to convert from the sequential numbering to the H3 numbering scheme.
+The fractions were estimated using qPCR against the viral NP and canine GAPDH as described in \citet{doud2017complete}.
+Briefly, we made duplicate 10-fold serial dilutions of each of the virus libraries to use as a standard curve.
+We also performed qPCR on the cells infected with the virus-antibody mix.
+To estimate the fractions, we used linear regression to fit a line relating logarithm of the viral infectious dose in the standard curve to the difference in Ct values between NP and GAPDH, and then interpolated the fraction surviving for each selection from this regression.
 
@@
 \caption{\label{supptab:fracsurvive}
-The total fraction of virions surviving each antibody treatment at each concentration as estimated by qPCR.
+{\bf The total fraction of virions surviving each antibody treatment at each concentration as estimated by qPCR.}
 These are the quantities referred to as $\gamma$.
@@
 \end{supptable}
+\clearpage
+
+\begin{suppfile}
+\caption{\label{suppfile:HAnumbering}
+{\bf Conversion from sequential numbering of the A/WSN/1933 HA to H3 numbering.}
+In this CSV file, the \emph{original} column gives the residue number in sequential (1, 2, ...) numbering of the A/WSN/1933 HA, and the \emph{new} column gives the residue number in H3 numbering.
+}
+\end{suppfile}
 

### 2017-10-23 paper/manuscript.tex
updated methods about trees

@@
 
-\subsection*{Construction of HA phylogenetic tree}
-We downloaded one HA sequence per subtype from the NIAID Influenza Research Database (IRD) \comment{cite}, built the phylogenetic tree using RaxML \comment{cite}, and visualized the tree using FigTree \comment{cite}.
-The HA sequences we downloaded were: H1 (A/WSN/1933); pdmH1N1 (A/California/04/2009); H2 (A/Japan/305/1957); H3 (A/Aichi/2/1968); H4 (A/swine/Ontario/01911/2/1999); H5 (A/Vietnam/1203/2004); H6 (A/chicken/Taiwan/0705/1999); H7 (A/Netherlands/219/2003); H8 (A/turkey/Ontario/6118/1968); H9 (A/swine/HongKong/9/1998); H10 (A/mallard/Bavaria/3/2006); H11 (A/duck/England/1/1956); H12 (A/duck/Alberta/60/1976); H13 (A/gull/Maryland/704/1977); H14 (A/mallard/Astrakhan/263/1982); H15 (A/mallard duck/Sweden/139579/2012); H16 (A/black headed gull/Mongolia/1756/2006); H17 (A/little yellow-shouldered bat/Guatemala/060/2010); H18 (A/dark fruit-eating bat/Bolivia/PBV780-781/2011).
+\subsection*{Inference of HA phylogenetic tree}
+To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database~\citep{?}, inferred the phylogenetic tree using RaxML~\citep{?} with a \comment{what model of substitution?}, and visualized the tree using FigTree (\comment{url}).
+The HA sequences used were: H1 (A/WSN/1933); pdmH1N1 (A/California/04/2009); H2 (A/Japan/305/1957); H3 (A/Aichi/2/1968); H4 (A/swine/Ontario/01911/2/1999); H5 (A/Vietnam/1203/2004); H6 (A/chicken/Taiwan/0705/1999); H7 (A/Netherlands/219/2003); H8 (A/turkey/Ontario/6118/1968); H9 (A/swine/HongKong/9/1998); H10 (A/mallard/Bavaria/3/2006); H11 (A/duck/England/1/1956); H12 (A/duck/Alberta/60/1976); H13 (A/gull/Maryland/704/1977); H14 (A/mallard/Astrakhan/263/1982); H15 (A/mallard duck/Sweden/139579/2012); H16 (A/black headed gull/Mongolia/1756/2006); H17 (A/little yellow-shouldered bat/Guatemala/060/2010); H18 (A/dark fruit-eating bat/Bolivia/PBV780-781/2011).
+In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize. 
+Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8 and H11; and no antibodies have been tested against H17 and H18.
+The narrow H17-L19, H17-L10, and H17-L7 antibodies have not been tested against any other subtypes -- however, since these antibodies have a very limited range even among H1 HAs~\citep{caton1982antigenic}, we assume that they do not bind other subtypes.
 
-For the cladogram in Figure~\ref{fig:structures}C, the amino-acid identities at site 38 are from the strains tested against C179 in~\cite{dreyfus2013structure}.
-For subtypes not tested, we used the identity from the corresponding strain listed above.
-
-We denoted HA's with reported binding or neutralization activity by each antibody, and defined binding activity as \comment{binding to full influenza virus, VLPs, recombinant or purified HA's}.
-Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8 and H11; no antibodies have been tested against H17 and H18.
+For the cladogram in Figure~\ref{fig:structures}C, the amino-acid identities at site 38 are from the strains tested against C179 by \citet{dreyfus2013structure}. 
+For subtypes not tested, the amino-acid identity reported is that in the strain listed above.
 

### 2017-10-24 paper/manuscript.tex
Completed revisions to Methods

@@
 All neutralization curves represent the mean and standard error of three measurements. 
+\subsection*{H3 sequence numbering}
+Unless otherwise indicated, all residues are numbered in the H3 numbering scheme, with the signal peptide in negative numbers, the HA1 subunit as plain numbers, and the HA2 subunit denoted with ``(HA2)''.
+The conversion between sequential numbering of the A/WSN/1933 HA and the H3 numbering scheme was performed using the Python script available at \url{https://github.com/jbloomlab/HA_numbering}.
+File~\ref{suppfile:HAnumbering} gives the numbering conversion.
 
 \subsection*{Inference of HA phylogenetic tree}
-To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database~\citep{?}, inferred the phylogenetic tree using RaxML~\citep{?} with a \comment{what model of substitution?}, and visualized the tree using FigTree (\comment{url}).
-The HA sequences used were: H1 (A/WSN/1933); pdmH1N1 (A/California/04/2009); H2 (A/Japan/305/1957); H3 (A/Aichi/2/1968); H4 (A/swine/Ontario/01911/2/1999); H5 (A/Vietnam/1203/2004); H6 (A/chicken/Taiwan/0705/1999); H7 (A/Netherlands/219/2003); H8 (A/turkey/Ontario/6118/1968); H9 (A/swine/HongKong/9/1998); H10 (A/mallard/Bavaria/3/2006); H11 (A/duck/England/1/1956); H12 (A/duck/Alberta/60/1976); H13 (A/gull/Maryland/704/1977); H14 (A/mallard/Astrakhan/263/1982); H15 (A/mallard duck/Sweden/139579/2012); H16 (A/black headed gull/Mongolia/1756/2006); H17 (A/little yellow-shouldered bat/Guatemala/060/2010); H18 (A/dark fruit-eating bat/Bolivia/PBV780-781/2011).
+To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database~\citep{zhang2017influenza}, inferred the phylogenetic tree using RaxML~\citep{stamatakis2014raxml} with a \comment{what model of substitution?}, and visualized the tree using FigTree (\url{http://tree.bio.ed.ac.uk/software/figtree/}).
+The HA sequences used are in File~\ref{suppfile:HAsubtypes}. 
 In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize. 
@@
 For the cladogram in Figure~\ref{fig:structures}C, the amino-acid identities at site 38 are from the strains tested against C179 by \citet{dreyfus2013structure}. 
-For subtypes not tested, the amino-acid identity reported is that in the strain listed above.
-
-\subsection*{H3 sequence numbering}
-Unless otherwise indicated, all residues are numbered in the H3 numbering scheme, with the signal peptide in negative numbers, the HA1 subunit as plain numbers, and the HA2 subunit denoted with ``(HA2)''.
-The conversion between sequential numbering of the A/WSN/1933 HA and the H3 numbering scheme was performed using the Python script available at \url{https://github.com/jbloomlab/HA_numbering}.
-File~\ref{suppfile:HAnumbering} gives the numbering conversion.
+For subtypes not tested, the amino-acid identity reported is that in the strain for that subtype in File~\ref{suppfile:HAsubtypes}.
 
@@
 To estimate the fractions, we used linear regression to fit a line relating logarithm of the viral infectious dose in the standard curve to the difference in Ct values between NP and GAPDH, and then interpolated the fraction surviving for each selection from this regression.
+
+\subsection*{Analysis of deep sequencing data}
+The deep sequencing data were analyzed using version 2.1.0 of the \texttt{dms\_tools2} software package~\citep{bloom2015software}, which is available at \url{http://jbloomlab.github.io/dms_tools2}.
+File~\ref{suppfile:analysis_code} contains a Jupyter notebook that performs all steps of the analysis beginning with downloading the FASTQ files from the Sequence Read archive.
+Detailed statistics about the sequencing depth and error rates are shown in this Jupyter notebook.
 
@@
 
-Code that performs all these calculations is included in the \texttt{dms\_tools2} software package~\citep{bloom2015software} which is available at \url{http://jbloomlab.github.io/dms_tools2}.
-The in this paper were performed using version 2.1.0 of \texttt{dms\_tools2}.
+In this paper, Figures~\ref{fig:avgfracsurvive} and \ref{suppfig:maxfracsurvive} show the median of excess fraction surviving taken across all biological and technical replicates at a given antibody concentration.
+The subsequent logo plots show the medians of these values taken across all concentrations for each antibody.
+The numerical values plotted in these logo plots are in File~\ref{suppfile:fracsurvive}.
+
+Code that performs these fraction surviving analyses has been added to version 2.1.0 of the \texttt{dms\_tools2} software package~\citep{bloom2015software} which is available at \url{http://jbloomlab.github.io/dms_tools2}.
 
@@
 Deep sequencing data are available from the Sequence Read Archive under BioSample accession SAMN05789126.
+Computer code that analyzes these data to generate the results described in this paper is in File~\ref{suppfile:analysis_code}.
 
@@
 
+\begin{suppfile}
+\caption{\label{suppfile:HAsubtypes}
+{\bf Sequences used to infer the tree for all HA subtypes.}
+This FASTA file gives the HA sequences used to infer the tree of subtypes in Figure~\ref{fig:antibody_summary}.
+\comment{Juhye, can you add this file into the \texttt{./paper/figs/} directory.}
+}
+\end{suppfile}
+
+\begin{suppfile}
+\caption{\label{suppfile:analysis_code}
+{\bf Computer code and data for the analysis of the mutational antigenic profiling data.}
+The code in this ZIP file performs the entire computational analysis beginning with downloading the FASTQ files from the Sequence Read Archive.
+The ZIP file contains a \texttt{README} file that explains the contents in detail.
+The actual analysis is performed by the Jupyter notebook \texttt{analysis\_notebook.ipynb}, which includes embedded plots summarizing key statistics and results.
+An HTML version of this notebook is also included as \texttt{analysis\_notebook.html}.
+}
+\end{suppfile}
+
+\begin{suppfile}
+\caption{\label{suppfile:fracsurvive}
+{\bf The excess fraction surviving for each mutation for each antibody.}
+This file is a ZIP of CSV files giving the numerical values plotted in the logo plots.
+These are median excess fraction surviving taken first across replicates and then across antibody concentrations.
+}
+\end{suppfile}
+
 \end{document}

### 2017-10-24 paper/manuscript.tex
Added some details to Methods

@@
 \subsection*{Inference of HA phylogenetic tree}
-To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database~\citep{zhang2017influenza}, inferred the phylogenetic tree using RaxML~\citep{stamatakis2014raxml} with a \comment{what model of substitution?}, and visualized the tree using FigTree (\url{http://tree.bio.ed.ac.uk/software/figtree/}).
+To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database~\citep{zhang2017influenza}, inferred the phylogenetic tree using RaxML~\citep{stamatakis2014raxml} with a GTR model, and visualized the tree using FigTree (\url{http://tree.bio.ed.ac.uk/software/figtree/}).
 The HA sequences used are in File~\ref{suppfile:HAsubtypes}. 
-In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize. 
+In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize~\citep{yoshida2009cross, lee2012heterosubtypic, okuno1993common, dreyfus2013structure, corti2011neutralizing}.
 Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8 and H11; and no antibodies have been tested against H17 and H18.
@@
 The fractions were estimated using qPCR against the viral NP and canine GAPDH as described in \citet{doud2017complete}.
-Briefly, we made duplicate 10-fold serial dilutions of each of the virus libraries to use as a standard curve.
+Briefly, we made duplicate 10-fold serial dilutions of each of the virus libraries to use as a standard curve of infectivity.
 We also performed qPCR on the cells infected with the virus-antibody mix.

### 2017-10-24 paper/manuscript.tex
small tweaks to align with _Immunity_ formatting
Finish edits to manuscript

@@
 Figure~\ref{fig:structures}A shows a zoomed-in view of the sites of mutations selected by each antibody, as well as their locations on HA's structure. 
-It is immediately clear that the selected mutations are nearly all in or near the antibody-binding footprint.
+It is immediately clear that the selected mutations are nearly all in or close to the antibody-binding footprint.
 
@@
 
-Although the anti-stalk antibodies C179 and FI6v3 only select mutations with small effects, these mutations almost all fall in the physical binding footprints of the antibodies (Figure~\ref{fig:structures}A).
+Although the anti-stalk antibodies C179 and FI6v3 only select mutations with small effects, these mutations almost all fall in or near the physical binding footprints of the antibodies (Figure~\ref{fig:structures}A).
 The two antibodies have similar epitopes and angles of approach~\citep{dreyfus2013structure}, and they select identical mutations at several sites (Figure~\ref{fig:structures}B). 
@@
 As a control, we also validated that a mutation at one of these sites (K280A) that does \emph{not} have an effect in our mutational antigenic profiling does not shift the neutralization curve (Figure~\ref{fig:FI6v3neutcurves}A).
-Interestingly, the HA of pandemic H1N1 has a glycosylation motif at 278-280, and \citet{corti2011neutralizing} have reported that this HA is $\sim$4-fold more resistant to FI6v3 than the A/WSN/1933 HA used in our experiments.
-This reported difference between strains is of similar magnitude to the effect that we measure for introducing the glycosylation motif via the K280S or K280T mutations.
 
@@
 The neutralization assays themselves were performed by using a plate reader to quantify the GFP signal produced by MDCK-SIAT1-CMV-PB1 cells infected by PB1flank-eGFP virus that had been incubated with the indicated antibody concentration as described in \citet{hooper2013mutant}.
-All neutralization curves represent the mean and standard error of three measurements. 
+All neutralization curves represent the mean and standard deviation of three measurements. 
 \subsection*{H3 sequence numbering}
@@
 The mutant virus libraries are those described in \citet{doud2016accurate}, and were produced in full biological triplicate.
-Briefly, these libraries were generated by using codon mutagenesis~\citep{bloom2014experimentally} to introduce random codon mutations into plasmid-encoded HA, and then generating the virus libraries using a helper-virus strategy that avoids the bottlenecks associated with standard influenza reverse genetics.
+Briefly, these libraries were generated by using codon mutagenesis~\citep{bloom2014experimentally} to introduce random codon mutations into plasmid-encoded HA, and then using a helper-virus strategy that avoids the bottlenecks associated with standard influenza reverse genetics to create the virus libraries.
 Full details of the library generation and sequencing statistics that quantify how completely each of the triplicate libraries covers the possible amino-acid mutations are in \citet{doud2016accurate}.
@@
 The mutational antigenic profiling was performed as described in \citet{doud2017complete}. 
-Briefly, we diluted each virus libraries to a concentration of $10^{6}$ TCID$_{50}$ per ml and incubated the virus dilutions with an equal volume of antibody at the intended concentration at 37$^\circ$C for 1.5 hours.
-The final antibody concentrations in these mixtures are shown in Figure~\ref{fig:neutcurves} concentrations.
+Briefly, we diluted each of the virus libraries to a concentration of $10^{6}$ TCID$_{50}$ per ml and incubated the virus dilutions with an equal volume of antibody at the intended concentration at 37$^\circ$C for 1.5 hours.
+The final antibody concentrations in these mixtures are shown in Figure~\ref{fig:neutcurves}.
 We performed three fully independent replicates of each selection using the three replicate mutant virus libraries.
@@
 The deep sequencing data were analyzed using version 2.1.0 of the \texttt{dms\_tools2} software package~\citep{bloom2015software}, which is available at \url{http://jbloomlab.github.io/dms_tools2}.
-File~\ref{suppfile:analysis_code} contains a Jupyter notebook that performs all steps of the analysis beginning with downloading the FASTQ files from the Sequence Read archive.
+File~\ref{suppfile:analysis_code} contains a Jupyter notebook that performs all steps of the analysis beginning with downloading the FASTQ files from the Sequence Read Archive.
 Detailed statistics about the sequencing depth and error rates are shown in this Jupyter notebook.
@@
 This FASTA file gives the HA sequences used to infer the tree of subtypes in Figure~\ref{fig:antibody_summary}.
-\comment{Juhye, can you add this file into the \texttt{./paper/figs/} directory.}
 }

### 2017-10-24 paper/manuscript.tex
a few minor text revisions

@@
 
-%\renewcommand{\abstractname}{\large SUMMARY}
+\renewcommand{\abstractname}{\large SUMMARY}
 
@@
 This escape is manifested by a large shift in the neutralization curve for the mutant (Figure~\ref{fig:fracsurvive_example}B).
-If we draw vertical lines through the overlaid neutralization curves, we can calculate the fraction of virions with each mutation that survive neutralization at each antibody concentration (Figure~\ref{fig:fracsurvive_example}B).
+If we draw vertical lines through the overlaid neutralization curves, we can calculate the fraction of virions with each mutation that survive neutralization at each antibody concentration.
 These fractions can be represented using logo plots, where the height of each letter is proportional to the fraction of virions with that amino acid at a site that survive (Figure~\ref{fig:fracsurvive_example}C).
@@
 (B) Sites of selection from anti-stalk antibodies, with the same coloring scale for both antibodies.
-(C) Cladogram of HA subtypes, with group 1 in black and group 2 in gray. 
+(C) Cladogram of group 1 HA subtypes.
+\comment{Juhye: can you update panel C to only show group 1} 
 The amino acid at site 38 is indicated. 
@@
 The FI6v3 antibody also weakly selects several mutations at residue -8, which is part of HA's signal peptide (Figure~\ref{fig:structures}A). 
-This signal peptide is cleaved from the mature HA protein~\citep{daniels2003n,burke2014recommended}, so mutations at this site likely affect HA's expression~\citep{nordholm2017translational} and subsequently its density on virions rather than affecting the physical binding of the antibody to HA.
+This signal peptide is cleaved from the mature HA protein~\citep{daniels2003n,burke2014recommended}, although mutations at this site can affect HA's expression level~\citep{nordholm2017translational}.
 
@@
 (A) Neutralization curves of individual viral mutants with FI6v3.
-The mutations K280S, K280T, N291S, G47R (HA2), and K(-8)T are all expected to increase neutralization resistance based on the mutational antigenic profiling (Figure~\ref{fig:structures}A), and they all do so -- but only modestly.
-The mutations K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization based on the mutational antigenic profiling (Figure~\ref{suppfig:FI6v3logo}), and none of them do.
+The mutations K280S, K280T, N291S, G47R (HA2), and K(-8)T are all expected to increase neutralization resistance based on the mutational antigenic profiling (Figure~\ref{fig:structures}A). 
+The mutations K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization based on the mutational antigenic profiling (Figure~\ref{suppfig:FI6v3logo}).
 (B), (C) In contrast to FI6v3, mutations selected by narrow antibodies have very large effects on neutralization.
@@
 The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.
-We tested one of these mutations, K(-8)T, and found that it indeed caused a modest but detectable increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A).
+We tested one of these mutations, K(-8)T, and found that it did appear to lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A).
 As controls, we also tested three mutations (P80D, V135T, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them affected neutralization resistance.
@@
 Note that scale for these antigenic effects is 10$\times$ smaller for the FI6v3 and C179 than for S139/1.
-Overall, it is clear that many mutations are tolerated at footprint sites but do not have a detectable antigenic effect.
 }
@@
 The data in Figure~\ref{fig:muttolerance} show that the lack of large-effect escape mutants from FI6v3 and C179 is not due to complete mutational intolerance of HA sites in the antibody-binding footprints.
-Rather, only a fraction of mutations to HA sites in the footprint actually reduce antibody neutralization.
+Some HA sites in each antibody footprint are fairly mutationally tolerant, and contain a range of mutations in the viral libraries used in our antibody selections.
+However, our mutational antigenic profiling shows that only a fraction of mutations at a fraction of these sites actually affect antibody neutralization.
 This finding is reminiscent of prior work showing that the binding energetics at protein-protein interfaces can be asymmetrically distributed across sites~\citep{jin1992high,cunningham1993comparison,dall1998mutational}.
@@
 A major rationale for studying broadly neutralizing antibodies is that they are hoped to be more resistant to viral evolutionary escape than the antibodies that dominate natural immune responses to influenza virus~\citep{krammer2015advances,corti2017tackling}.
-We have used a new approach to quantify the extent to which this is actually true, and shown that anti-viral activity of broad anti-stalk antibodies is indeed more -- although certainly not completely -- resistant to erosion by viral point mutations.
+We have used a new approach to quantify the extent to which this is actually true, and shown that the anti-viral activity of broad anti-stalk antibodies is indeed more -- although certainly not completely -- resistant to erosion by viral point mutations.
 Going forward, we suggest that quantifying the ease of viral escape will be a useful complement to more traditional techniques that characterize the structural epitopes and strain breadth of anti-viral antibodies.
@@
 These PB1flank-eGFP were generated in co-cultures of 293T-CMV-PB1 and MDCK-SIAT1-CMV-PB1 cells as described in \citet{bloom2010permissive}, using the standard bi-directional pHW181-PB2, ..., pHW188-NS reverse-genetics plasmids~\citep{hoffmann2000dna} for all genes \emph{except} PB1, plus the pHH-PB1flank-eGFP plasmid~\citep{bloom2010permissive}.
-Each mutant was generated repeating this process using a version of the pHW184-HA plasmid that had been engineered by site-directed mutagenesis to carry the indicated mutation.
+Each mutant was generated by repeating this process using a version of the pHW184-HA plasmid that had been engineered by site-directed mutagenesis to carry the indicated mutation.
 The neutralization assays themselves were performed by using a plate reader to quantify the GFP signal produced by MDCK-SIAT1-CMV-PB1 cells infected by PB1flank-eGFP virus that had been incubated with the indicated antibody concentration as described in \citet{hooper2013mutant}.
@@
 The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+
+\subsection*{AUTHOR CONTRIBUTIONS}
+MBD and JML performed the experiments.
+All three authors designed the project, contributed to the computer code, analyzed the data, and wrote the paper.
 

### 2017-10-27 paper/manuscript.tex
mike text and fig5 edits

@@
 Single mutations make the virus completely resistant to both narrow strain-specific antibodies and a broad antibody against the receptor-binding site.
-But no single mutation does more than modestly increase the virus's resistance to broad antibodies against the HA stalk.
+But no single mutation does more than modestly increase the virus's resistance to two broad antibodies against the HA stalk.
 Therefore, broad anti-stalk antibodies are quantifiably more resistant to viral escape via single amino-acid mutations than the other antibodies tested here.
@@
 {\bf Epitopes and breadth of broad and narrow antibodies targeting HA.}
-(A) Crystal structures of the broad antibodies and binding footprints of the narrow ones superimposed on the structure of the HA trimer~\citep[PDB 1RVX;][]{gamblin2004structure}. 
+(A) Crystal structures of the broad antibodies and sites of escape mutations selected by the narrow ones superimposed on the structure of the HA trimer~\citep[PDB 1RVX;][]{gamblin2004structure}. 
 S139/1~\citep[PDB 4GMS;][]{lee2012heterosubtypic} targets the receptor-binding site; C179~\citep[PDB 4HLZ;][]{dreyfus2013structure} and FI6v3~\citep[PDB 3ZTN;][]{corti2011neutralizing} target the stalk.
-The footprints for H17-L19, H17-L10, and H17-L7 are those mapped by \citet{doud2017complete}. 
-(B) A phylogenetic tree of HA subtypes, with group 1 at top and group 2 at bottom.
+The sites of escape mutations for H17-L19, H17-L10, and H17-L7 are those mapped by \citet{doud2017complete}. 
+(B) A phylogenetic tree of HA subtypes.
 Circles (broad antibodies) and squares (narrow antibodies) denote reported antibody binding or neutralization activity against that subtype. 
@@
 We applied this approach to anti-HA antibodies with a range of breadths and epitopes.
-The crystal structures or binding footprints of these antibodies are shown in Figure~\ref{fig:antibody_summary}A.
-We selected two broad antibodies, FI6v3 and C179, that target the stalk of HA~\citep{corti2011neutralizing, okuno1993common, dreyfus2013structure}. 
+The crystal structures or sites of escape mutations selected by these antibodies are shown in Figure~\ref{fig:antibody_summary}A.
+We chose two broad antibodies, FI6v3 and C179, that target the stalk of HA~\citep{corti2011neutralizing, okuno1993common, dreyfus2013structure}. 
 FI6v3 is extremely broad, and neutralizes both group 1 and group 2 HAs (Figure~\ref{fig:antibody_summary}B).
@@
 
-We performed our experiments using the A/WSN/1933 (H1N1) strain of influenza, which is neutralized by all six antibodies with similar potencies (Figure~\ref{fig:neutcurves}).
-We used antibody concentrations where nearly all wildtype virus was neutralized, allowing us to detect antigenic mutations that increase the fraction of virions that survive beyond the baseline of wildtype (Figure~\ref{fig:neutcurves}).
+We performed our experiments using the A/WSN/1933 (H1N1) strain of influenza, which is neutralized by all six antibodies (Figure~\ref{fig:neutcurves}).
+We used antibody concentrations sufficient for nearly complete neutralization of wildtype virus, allowing us to detect antigenic mutations that increase the fraction of mutant virions surviving beyond the baseline of wildtype (Figure~\ref{fig:neutcurves}).
 We used several concentrations of each broad antibody to increase sensitivity to antigenic mutations of both large and small effect.
@@
 \caption{\label{fig:neutcurves}
-{\bf Neutralization of wildtype virus by each antibody, and the fraction of virions surviving at each concentration used in our experiments.}
-The curves show neutralization of wildtype A/WSN/1933 virus. 
+{\bf Neutralization of wildtype virus by each antibody, and the fraction of mutant library virions surviving at each concentration used in our experiments.}
+The curves show neutralization of the wildtype A/WSN/1933 virus. 
 Each point represents the mean and standard deviation of three measurements. 
-The vertical dotted lines show the concentrations of antibody used in our selections, and the tables give the overall fraction of the mutant virus libraries that survived at each concentration.
+The vertical dotted lines show the concentrations of antibody that were then used in the mutant virus library selections, and the tables give the overall fraction of the mutant virus libraries that survived at each concentration, determined by qRT-PCR.
 }
@@
 \subsection*{The selected mutations are near the binding footprints of the antibodies}
-Antigenic mutations selected by narrow strain-specific antibodies against HA generally occur at residues in or near the physical binding footprint of the antibody~\citep{yewdell1979antigenic,webster1980determination,caton1982antigenic}.
-We examined whether this was also the case for the broad antibodies used in our experiments.
+Antigenic mutations selected by narrow strain-specific antibodies against HA are thought to occur at residues in or near the physical binding footprint of the antibody~\citep{yewdell1979antigenic,webster1980determination,caton1982antigenic}.
+We examined whether this was the case for the broad antibodies used in our experiments.
 Figure~\ref{fig:structures}A shows a zoomed-in view of the sites of mutations selected by each antibody, as well as their locations on HA's structure. 
@@
 (B) Sites of selection from anti-stalk antibodies, with the same coloring scale for both antibodies.
+Selection for serine or threonine at sites 280 and 291 introduces glycosylation sites at 278 and 289, respectively.
 (C) Cladogram of group 1 HA subtypes.
-\comment{Juhye: can you update panel C to only show group 1} 
 The amino acid at site 38 is indicated. 
@@
 Our data show that numerous different amino-acid mutations at each site confer neutralization resistance.
-The mutation with the largest effect, D158N, introduces an N-linked glycosylation motif.
+The mutation with the largest effect, G158N, introduces an N-linked glycosylation motif.
 
@@
 Figure~\ref{fig:FI6v3neutcurves} shows that the mutational antigenic profiling is highly predictive of the results of the neutralization assays, even for small-effect mutations.
-As discussed in the previous section, the three mutations most strongly selected by FI6v3 introduce glycosylation motifs at sites 278-280 or 289-291 (Figure~\ref{fig:structures}A).
+As discussed in the previous section, the three mutations most strongly selected by FI6v3 introduce glycosylation motifs at sites 278-280 or 289-291 (Figure~\ref{fig:structures}A,B).
 We created viruses carrying each of these mutations (K280S, K280T, and N291S) and validated that all three modestly increased resistance to FI6v3 (Figure~\ref{fig:FI6v3neutcurves}A).
@@
 We tested one of these mutations, K(-8)T, and found that it did appear to lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A).
-As controls, we also tested three mutations (P80D, V135T, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them affected neutralization resistance.
+As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them affected neutralization resistance.
 
@@
 Neutralization curves for such large-effect escape mutants are in Figure~\ref{fig:FI6v3neutcurves}B,C.
-Although there are no such large-effect single mutations that escape FI6v3 and C179, the results in Figure~\ref{fig:FI6v3neutcurves}A show that we can still use mutational antigenic profiling to identify mutations that have small but measurable effects on resistance to these antibodies.
+Although there are no such large-effect single mutations that escape FI6v3 or C179, the results in Figure~\ref{fig:FI6v3neutcurves}A show that we can still use mutational antigenic profiling to identify mutations that have small but measurable effects on resistance to these antibodies.
 
@@
 Above the preference logo stacks are logo plots showing the excess fraction surviving antibody treatment as measured in the current study.
-Note that scale for these antigenic effects is 10$\times$ smaller for the FI6v3 and C179 than for S139/1.
+Note that scale for these antigenic effects is 10$\times$ smaller for FI6v3 and C179 than for S139/1.
 }
@@
 However, such mutations can also enable the virus to escape the broad S139/1 antibody targeting HA's receptor-binding pocket, despite the fact that this antibody neutralizes multiple subtypes.
-But no single mutations have a comparably large effect on neutralization by two broad antibodies targeting HA's stalk, FI6v3 and C179.
+But no single mutation has a comparably large effect on neutralization by two broad antibodies targeting HA's stalk, FI6v3 and C179.
 Therefore, these anti-stalk antibodies are quantifiably more difficult for the virus to escape.
@@
 The average of these fractions across libraries are reported in Figure~\ref{fig:neutcurves}, and the values for each individual replicate are in Table~\ref{supptab:fracsurvive}.
-The fractions were estimated using qPCR against the viral NP and canine GAPDH as described in \citet{doud2017complete}.
+The fractions were estimated using qRT-PCR against the viral NP and canine GAPDH as described in \citet{doud2017complete}.
 Briefly, we made duplicate 10-fold serial dilutions of each of the virus libraries to use as a standard curve of infectivity.
@@
 \subsection*{ACKNOWLEDGMENTS}
-We thank Adam Dingens, Lauren Gentles, and Jeremy Roop for helpful comments on the project and manuscript.
+We thank Adam Dingens, Sarah Hilton, Katherine Xue, Lauren Gentles, and Jeremy Roop for helpful comments on the project and manuscript.
 We thank the Fred Hutchinson Cancer Research Center genomics core for performing the Illumina deep sequencing, and the protein expression core for expressing and purifying the C179 antibody.

### 2017-10-27 paper/manuscript.tex
a few more tiny manuscript revisions

@@
 %We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three narrow strain-specific antibodies, and we re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived neutralization.
-The correlations among replicates are in Figure~\ref{suppfig:corr}.
+The correlations among replicates are shown in Figure~\ref{suppfig:corr}.
 For the remainder of this paper, we will refer to the median antigenic effect of each mutation across replicates.
@@
 \begin{figure}[h!]
-\centerline{\includegraphics[width=0.89\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
+\centerline{\includegraphics[width=0.9\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
 \caption{
@@
 The amino acid at site 38 is indicated. 
-Colors indicate whether a subtype is bound or neutralized by C179.  
+Colors indicate whether a subtype has been reported in the literature to be bound or neutralized by C179.  
 }

### 2017-11-27 paper/Immunity_reviews.tex
added _Immunity_ reviews

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

\usepackage{color}

\usepackage[parfill]{parskip}

%SetFonts

%SetFonts

\title{Reviewers' comments from \textit{Immunity}}
\author{}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle

\section*{Editor Comments}

Dear Dr. Bloom,

Thank you for sending Immunity your manuscript titled "Quantifying the ease of viral escape from broad and narrow antibodies to influenza hemagglutinin" (IMMUNITY-D-17-01026). We have received the reports of the reviewers and we enclose the comments they have provided for transmission to the authors. Unfortunately, the recommendation of the referees is against publication in Immunity.

Jesse- I'm sorry for the length of the review process. Thank you for your patience. Unfortunately, it was felt that the manuscript does not provide a sufficient level of conceptual advance over work published by your lab and others, and there were also concerns with the physiological relevance of the findings and whether the conclusions are broadly relevant and generalizable. Unfortunately, the level of enthusiasm for the work was not sufficient for us to invite a revision, and we feel that it would be more efficient to seek another venue for the work. I know that this manuscript represents a substantial amount of careful work, and I?m sorry that I can?t provide you with a better outcome on this occasion. The manuscript may be a better fit for another Cell Press journal. I have provided links below that can be used should you wish to transfer the manuscript. Please let me know if you have any questions about transfers or if I can be of any further assistance. -Kavitha

\section*{Reviewer \#1: Summary}

Doud et al. used mutational antigenic profiling to systematically map the influenza HA resistant mutations against three broadly neutralizing antibodies - one targets the HA receptor-bind site (S139/1), and two target the HA stalk region (C179 and FI6v3). The authors then analyzed the data generated in this study along with the data generated in their previous study (Doud et al., PLoS Pathog 2017), which describes the mapping of influenza HA resistant mutations against three strain-specific neutralizing antibodies - H17-L7, H17-L10, and H17-L19. Here, the authors made a slight improvement in the analytical method that was originally described in their previous study (Doud et al., PLoS Pathog 2017). The improvement is based on the introduction of the experimentally determined parameter $\gamma$, which describes the fraction of viral library that survives the antibody selection, and is measured using RT-qPCR. Using this improved analytical method, the authors observed strong resistance mutations to S139/1 (broad), H17-L7 (strain-specific), H17-L10 (strain-specific), and H17-L19 (strain-specific), but not to C179 (broad) and FI6v3 (broad). Only weak resistant mutations, including previously unknown ones, were identified for C179 and FI6v3. Based on this result, Doud et al. claimed that antibody breadth is not necessarily an indicator of the difficulty of viral escape and stalk-binding antibodies are more difficult to escape. By pulling the data from their previous study (Doud and Bloom, Viruses 2016), the authors further showed that the lack of strong resistant mutations to stalk-binding antibodies (C179 and FI6v3) is not due to the lack of mutational tolerance.

While this paper is technically sound, the novelty and biological insights are limited. Despite the slight improvement, which allows results from mutational antigenic profiling to be normalized and compared across different antibodies, the methodology that Doud et al. employed is largely the same as they described in their previous study (Doud et al., PLoS Pathog 2017). Also, the mutant libraries used in this study have already been described in authors' previous studies (Doud and Bloom, Viruses 2016; Doud et al., PLoS Pathog 2017). The claim that antibody breadth is not correlated to the ease of viral escape is potentially interesting, but largely relies on the observation of resistant mutations to only one broadly neutralizing antibody S139/1. Therefore, the generality of this claim is unclear.

{\color{red} This is a good and fair summary.
We should make sure that we are not over-stating the generality of the claims.}

Specific comments:

1. Based on Figure 4 and 5, the number of residues that can give rise to resistant mutations to S139/1 seem to be less than to two of the strain-specific antibodies (H17-L10 and H17-L19), and maybe also to H17-L7. Does that suggest S139/1 is slightly more difficult to escape as compared to strain-specific antibodies?
{\color{red} We could comment briefly on the number of sites of escape for the different residues.}

2. Related to the concern above, although the methodology in this study normalized the antigenic profiling results from different antibodies, a quantitative comparison of the ease of escape among different antibodies was not performed. Is it possible to rank the ease of escape among different antibodies?
{\color{red} We could easily rank the antibodies by something like the median effect of the top 5 escape mutations. Mike had a figure that looked a bit like that.}

3. In Figure 6A, the fraction infectivity of K(-8)T at low concentration of FI6v3 is consistent higher that that of WT. Is there any explanation? On a related point the resistant mutation to FI6v3 at residue -8 is an interesting observation. Can the authors suggest possible mechanism?
{\color{red} I think the effect is so small that we aren't even sure that it is true, which is why we didn't really discuss this.. We could mention something about possibly modulating expression level...}

4. In Figure 6A, V135T seems to show some resistance at FI6v3 at level that is similar to K(-8)T, although the authors claim that V135T does not affect the neutralization activity of FI6v3. Can the authors quantify the inhibition effect (e.g. IC50) in this validation experiment to facilitate interpretation of the results?
{\color{red} I'm not sure if we gain anything by trying to do statistics on this, since the data is all shown and the effect is clearly tiny. 
Might be best just to remain very circumspect about K(-8)T?}

5. In the Figure 2B and in the methods section, the authors state that S139/1 has not been tested against H9. However, S139/1 has been shown to binding to H9 by ELISA but without any neutralization activity (Yoshida et al., PLoS Pathog 2009).
{\color{red} We should add this fact.}

\section*{Reviewer \#2} 
This is a manuscript examining the escape of influenza to broad and narrow antibodies in order to quantify escape. It is a well-written paper and the authors are to be commended for their encyclopedic referencing of classic literature. I appreciate the desire of the authors to perform an "apples-to-apples" comparison, and there is much to like about this paper, but unfortunately, this analysis has flaws that preclude me recommending it for publication.

The manuscript focuses on neutralization as the primary mechanism of pressure on influenza and uses neutralization measurements, antibody selection of viruses, and a model to interpret the results. In particular, the authors compare an antibody that binds near the receptor binding site (S139/1), two stem-directed antibodies (FI6v3 and C179), and a series of strain-specific antibodies (H17 series).

Specific concerns

1. Although mAb S139/1 is called a receptor binding site (RBS) antibody, the epitope is actually adjacent to the RBS. The abstract of the Yoshida paper called it "a novel conformational epitope adjacent to the receptor-binding domain of HA." It is not clear what a true RBS antibody (eg, 5J8, CH65) would do in this model.
{\color{red} We can certainly say that it binds \emph{near} the RBS rather than in the RBS.
I don't think it is worth adding additional antibodies at this time.}

2. Stem epitopes and the RBS have different degrees of exposure on properly formed, infectious influenza virions. Although a number of publications have shown that stem antibodies can access the epitope on infectious virions, the data shows that 1-3 orders of magnitude more antibody is needed to neutralize infectious virions compared with pseudovirions (where HA packing is less dense) [see Joyce MG et al. Cell 166(3): 609-623 (2016) and Corti D et al. Science 333(6044): 850-856 (2011)]. This raises the question of whether a model dependent on neutralization of infectious virions is relevant to assess in vivo pressure on the HA stem epitope. Looking at the Doud and Bloom 2016 paper, it is not clear whether the virions produced in the helper virus system would be packed tightly (as infectious influenza is) or loosely (like in the pseudovirus assay).
{\color{red}
This is one that we could address if given a chance.
We are using full-replication competent viruses that should have physiological HA density. 
We can add a line that mentions this.
}

3. In vivo protection afforded by stem-directed antibodies has been convincingly shown to be Fc-FcR mediated [see DiLillo DJ et al. J Clin Invest 126(2): 605-610 (2016) and DiLillo DJ et al. Nat Med 20(2): 143-151 (2014)], making the relevance of a finding based on neutralization (whether tightly or loosely packed) unclear.
{\color{red}
We can mention something about FcR.
However, I would guess that mutations that escape neutralization probably also escape any in vivo activity.
}

4. The amount of antibody used is wildly different among the antibodies tested and it is not clear why some antibodies (eg, S139/1) needed very high concentrations in the assay while others (eg, C179) required two orders of magnitude less. This is particularly relevant given the comment above about epitopic exposure on virions.
{\color{red}
We should emphasize that we are choosing concentrations where about the same \emph{fraction} (between 0.01 and 0.001) of the library escapes the antibody.
The neutralization curves are for wildtype library, and these fractions are computed on the mutant libraries.
}

5. The neutralization curves (Fig. 3) that show all antibodies had complete or near-complete neutralization at 1 $\mu$g/mL. Since the point is to have an ``apples-to-apples'' comparison, it would seem that comparing the ability of antibodies to select viruses at similar concentrations, and especially at concentrations likely to be present in a physiologic condition, would be the relevant assay.
{\color{red}
See answer above.
}

\section*{Reviewer \#3} 
Boud et al., used a previously characterized HA single AA mutant library and a previously developed method for quantifying viral survival to quantify the viral escape from three broad neutralizing antibodies. While the work is interesting, there are several problems.

1. Their previously published HA sequencing method uses a molecular barcode, which helps to correct the sequencing and PCR errors. However, they did not explain why they still need to use a ratio between antibody selected and mock treated to observe the true mutant.
{\color{red} 
This is a dumb comment. 
I'm not sure if it is even worth trying to clarify in the manuscript since I think it is obvious to anyone who thinks about it.
}

2. Some of the mutants may have a competitive edge over other mutants regardless of antibodies, while others, the inhibition of other competitive mutants by antibodies may give them some advantage. As their data in Figure 7 revealed, sometime the selected mutation is rather disfavored with respect to viral growth, which would affect the growth of these survivors if they are in the original viral pool without antibody selection. Thus, the use of ratio makes it impossible to distinguish these two scenarios. Also, would these selected mutants that are disfavor the viral growth develop additional mutation to favor the viral growth in long-term?
{\color{red}
This also seems like a dumb comment.
I think we make fairly clear that the goal is to identify mutations that mediate escape, not to quantify their effect on viral growth.
I guess maybe the term ``ease of escape'' is confusing in this sense?
We are pretty clear in the Discussion that we are only looking at single mutations, and that there could be compensatory changes -- we spend a whole paragraph on this.
}

3. Also, how did authors make sure there is no drop out in the library at the beginning of the assay?
{\color{red}
These plots are in the notebook. 
I suppose we could move some of them to be supporting figures -- or just refer to the notebook more clearly?
}

4. Full HA sequences diversity vs abundance curve need to be shown in order for readers to fully understand the dynamic changes of viral species before and after antibody addition.
{\color{red}
See comment above.
}

5. Part of the data, the narrow antibody selection, are from their previous publication with a similar conclusion, although the mutation patterns have some differences. The new analysis applied a background correction so that they can compare data obtained from different experiments. So the question is how reproducible is their method? Do they have some kind of metrics for false negative and false positive based on their criterial?
{\color{red}
This also sort of a dumb comment.
It's actually a good sign that we get essentially the same results using two slightly different methods.
The point of data analyses is that the conclusions should be robust to different reasonable analysis methods, not that the results should be identical.
I don't think we can really develop a metric for false positive / negative rates.
}

6. Why they choose to use P=5 for this study while 10 in the previous one? How that affects the results?
{\color{red}
The pseudocount choice is inherently arbitrary.
See the comment above.
}

7. Figure 6 lacks statistic tests.
{\color{red}
I still think it is better to show the actual data than run some semi-bogus statistical test.
But we could add some sort of test statistic I suppose...}

8. For the last result section, I do not think they have enough experimental evidence to favor one hypothesis over the other without testing dependent mutations.
{\color{red}
I don't think we draw any overly strong conclusions here, but I suppose we could be a bit more circumspect.
}

\end{document}

### 2017-11-27 paper/Immunity_reviews.tex
Added note to Reviewer 2 comment

@@
 The abstract of the Yoshida paper called it "a novel conformational epitope adjacent to the receptor-binding domain of HA." It is not clear what a true RBS antibody (eg, 5J8, CH65) would do in this model.
-{\color{red} We can certainly say that it binds \emph{near} the RBS rather than in the RBS.
-I don't think it is worth adding additional antibodies at this time.}
+{\color{red} Although the three sites of escape found by Yoshida et al were indeed found to be near the RBS, they didn't crystallize the antibody in complex with HA.
+However, [Lee et al 2012 PNAS] published the crystal structure, and as can be seen in figure 1, the HCDR2 loop of S139/1 inserts into the receptor binding site.}
 

### 2017-11-28 paper/manuscript.tex
tweaks for _Immunity_ submission
re-formatted citations / sections for _Nature Microbiology_

@@
 \documentclass[11pt]{article}
-\usepackage[round,semicolon]{natbib}
+\usepackage[super]{natbib}
 \usepackage[margin=1.4in]{geometry}
@@
 
-\renewcommand{\abstractname}{\large SUMMARY}
+\renewcommand{\abstractname}{\large ABSTRACT}
 
@@
 Therefore, antibody breadth is not necessarily an indicator of the difficulty of viral escape.
-Broad antibodies targeting hemagglutinin's stalk are quantifiably harder to escape than the other antibodies tested here.
+Broadly neutralizing antibodies targeting hemagglutinin's stalk are quantifiably harder to escape than the other antibodies tested here.
 \end{abstract}
@@
 However, the extent of this variation ranges widely.
-For instance, although both measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic,drexler2014robustness} exhibit antigenic variation, the magnitude of this variation is small.
-Therefore, immunity to these viruses is lifelong~\citep{panum1847iagttagelser,salk1984one}.
+For instance, although both measles virus\cite{birrer1981antigenic,ter1981antigenic} and polio virus\cite{crainic1983natural,diamond1985antigenic,drexler2014robustness} exhibit antigenic variation, the magnitude of this variation is small.
+Therefore, immunity to these viruses is lifelong\cite{panum1847iagttagelser,salk1984one}.
 In contrast, human influenza virus exhibits much more antigenic variation.
-So although infection with an influenza virus strain provides long-term immunity to that exact strain~\citep{fluinboardingschool1978,davies1982christ,yu2008neutralizing}, the virus's rapid antigenic evolution erodes the effectiveness of this immunity to that strain's descendants within four to seven years~\citep{couch1983immunity}.
+So although infection with an influenza virus strain provides long-term immunity to that exact strain\cite{fluinboardingschool1978,davies1982christ,yu2008neutralizing}, the virus's rapid antigenic evolution erodes the effectiveness of this immunity to that strain's descendants within four to seven years\cite{couch1983immunity}.
 
-One possible reason that viruses exhibit different amounts of antigenic variation is that they have disparate evolutionary capacities to escape the immunodominant antibodies generated by natural immune responses~\citep{lipsitch2007patterns,cobey2014pathogen,fulton2015mutational}.
+One possible reason that viruses exhibit different amounts of antigenic variation is that they have disparate evolutionary capacities to escape the immunodominant antibodies generated by natural immune responses\cite{lipsitch2007patterns,cobey2014pathogen,fulton2015mutational}.
 According to this explanation, human influenza virus undergoes rapid antigenic drift because most neutralizing antibodies target epitopes on the viral hemagglutinin (HA) protein that are highly tolerant of mutational change.
-This explanation is supported by classic experiments showing that it is easy to select viral mutants that escape most antibodies~\citep{yewdell1979antigenic,webster1980determination}, as well as by the observation that mutations that alter antigenicity arise frequently during influenza's evolution globally~\citep{koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction} and within individual humans with long-term infections~\citep{xue2017parallel}.
+This explanation is supported by classic experiments showing that it is easy to select viral mutants that escape most antibodies\cite{yewdell1979antigenic,webster1980determination}, as well as by the observation that mutations that alter antigenicity arise frequently during influenza's evolution globally\cite{koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction} and within individual humans with long-term infections\cite{xue2017parallel}.
 A corollary of this explanation is that influenza virus's capacity for antigenic drift would be reduced if most antibodies instead targeted epitopes that were less mutationally tolerant.
@@
 Verifying this corollary has become of practical importance with the discovery of broadly neutralizing antibodies against influenza virus.
-These antibodies typically target conserved epitopes in HA's stalk~\citep{sui2009structural,ekiert2009antibody,corti2011neutralizing} or receptor-binding site~\citep{lee2012heterosubtypic,ekiert2012cross,schmidt2015viral}, and neutralize a wide range of viral strains.
-Broad antibodies are usually less abundant in human serum than antibodies to antigenically variable epitopes on the head of HA~\citep{ellebedy2014induction,andrews2015immune}.
-However, major efforts are underway to elicit broad antibodies by vaccination or administer them directly as therapeutics~\citep{krammer2015advances,corti2017tackling}.
+These antibodies typically target conserved epitopes in HA's stalk\cite{sui2009structural,ekiert2009antibody,corti2011neutralizing} or receptor-binding site\cite{lee2012heterosubtypic,ekiert2012cross,schmidt2015viral}, and neutralize a wide range of viral strains.
+Broad antibodies are usually less abundant in human serum than antibodies to antigenically variable epitopes on the head of HA\cite{ellebedy2014induction,andrews2015immune}.
+However, major efforts are underway to elicit broad antibodies by vaccination or administer them directly as therapeutics\cite{krammer2015advances,corti2017tackling}.
 
@@
 Might such selection then drive antigenic variation in these epitopes?
-There is precedent for the idea that the immune status of the host population can shape influenza virus evolution: the virus undergoes faster antigenic drift in long-lived humans that accumulate immune memory than in short-lived swine that are mostly naive~\citep{sheerar1989antigenic,luoh1992hemagglutinin}, and poultry vaccination may accelerate antigenic drift of avian influenza~\citep{lee2004effect,cattoli2011antigenic}.
+There is precedent for the idea that the immune status of the host population can shape influenza virus evolution: the virus undergoes faster antigenic drift in long-lived humans that accumulate immune memory than in short-lived swine that are mostly naive\cite{sheerar1989antigenic,luoh1992hemagglutinin}, and poultry vaccination may accelerate antigenic drift of avian influenza\cite{lee2004effect,cattoli2011antigenic}.
 But alternatively, perhaps broad antibodies are broad because the virus has difficulty escaping them regardless of selection from host immunity.
@@
 So far, there is limited data to distinguish between these possibilities.
-Several studies have shown that the head domain of HA is more mutationally tolerant than the stalk domain where many broad antibodies bind~\citep{thyagarajan2014inherent,wu2014high,heaton2013genome}.
+Several studies have shown that the head domain of HA is more mutationally tolerant than the stalk domain where many broad antibodies bind\cite{thyagarajan2014inherent,wu2014high,heaton2013genome}.
 However, these studies did not select for antibody escape, so it is difficult to relate their measurements to the virus's evolutionary capacity under immune selection.
-Other work has shown that it is possible to select antigenic mutants with broad antibodies~\citep{yoshida2009cross,chai2016two}, demonstrating that these epitopes are not entirely refractory to change.
-But given that antibodies can select some antigenic variation even in measles virus~\citep{birrer1981antigenic,ter1981antigenic} and polio virus~\citep{crainic1983natural,diamond1985antigenic}, the existence of selectable mutations does not necessarily imply that influenza virus can escape broad antibodies as easily as it drifts away from narrow strain-specific ones.
+Other work has shown that it is possible to select antigenic mutants with broad antibodies\cite{yoshida2009cross,chai2016two}, demonstrating that these epitopes are not entirely refractory to change.
+But given that antibodies can select some antigenic variation even in measles virus\cite{birrer1981antigenic,ter1981antigenic} and polio virus\cite{crainic1983natural,diamond1985antigenic}, the existence of selectable mutations does not necessarily imply that influenza virus can escape broad antibodies as easily as it drifts away from narrow strain-specific ones.
 The fundamental problem is that existing studies have not quantified the ease of viral escape in a way that can be compared across antibodies in an apples-to-apples fashion.
@@
 However, individually creating and assaying that many mutants would be exceedingly time-consuming and expensive.
-Fortunately, we have shown that antibody selection on all viral mutations can be assayed in a single experiment using mutational antigenic profiling~\citep{doud2017complete,dingens2017comprehensive}.
+Fortunately, we have shown that antibody selection on all viral mutations can be assayed in a single experiment using mutational antigenic profiling\cite{doud2017complete,dingens2017comprehensive}.
 This approach involves generating viral libraries containing all mutations to the protein of interest, selecting these viruses with or without antibody, and using an accurate deep-sequencing method to determine the relative frequencies of each mutation.
@@
 {\bf Epitopes and breadth of broad and narrow antibodies targeting HA.}
-(A) Crystal structures of the broad antibodies and sites of escape mutations selected by the narrow ones superimposed on the structure of the HA trimer~\citep[PDB 1RVX;][]{gamblin2004structure}. 
-S139/1~\citep[PDB 4GMS;][]{lee2012heterosubtypic} targets the receptor-binding site; C179~\citep[PDB 4HLZ;][]{dreyfus2013structure} and FI6v3~\citep[PDB 3ZTN;][]{corti2011neutralizing} target the stalk.
-The sites of escape mutations for H17-L19, H17-L10, and H17-L7 are those mapped by \citet{doud2017complete}. 
+(A) Crystal structures of the broad antibodies and sites of escape mutations selected by the narrow ones superimposed on the structure of the HA trimer (PDB 1RVX\cite{gamblin2004structure}). 
+S139/1 (PDB 4GMS\cite{lee2012heterosubtypic}) targets the receptor-binding site; C179 (PDB 4HLZ)\cite{dreyfus2013structure}) and FI6v3 (PDB 3ZTN\cite{corti2011neutralizing}) target the stalk. 
+The sites of escape mutations for H17-L19, H17-L10, and H17-L7 are those mapped by Doud et al\cite{doud2017complete}. 
 (B) A phylogenetic tree of HA subtypes.
@@
 The crystal structures or sites of escape mutations selected by these antibodies are shown in Figure~\ref{fig:antibody_summary}A.
-We chose two broad antibodies, FI6v3 and C179, that target the stalk of HA~\citep{corti2011neutralizing, okuno1993common, dreyfus2013structure}. 
+We chose two broad antibodies, FI6v3 and C179, that target the stalk of HA\cite{corti2011neutralizing, okuno1993common, dreyfus2013structure}. 
 FI6v3 is extremely broad, and neutralizes both group 1 and group 2 HAs (Figure~\ref{fig:antibody_summary}B).
 C179 is less broad, and neutralizes only some group 1 HAs (Figure~\ref{fig:antibody_summary}B).
-We also selected a broad antibody, S139/1, that targets HA's receptor-binding site and can neutralize both group 1 and group 2 HAs~\citep{yoshida2009cross, lee2012heterosubtypic}.
-Finally, we re-analyzed deep sequencing data from prior mutational antigenic profiling of three narrow strain-specific antibodies, H17-L19, H17-L10, and H17-L7~\citep{doud2017complete}.
-These narrow antibodies bind the Ca2, Ca1, and Cb antigenic regions on HA's globular head~\citep{caton1982antigenic}, and only neutralize a narrow slice of H1 viruses.
+We also selected a broad antibody, S139/1, that targets HA's receptor-binding site and can neutralize both group 1 and group 2 HAs\cite{yoshida2009cross, lee2012heterosubtypic}.
+Finally, we re-analyzed deep sequencing data from prior mutational antigenic profiling of three narrow strain-specific antibodies, H17-L19, H17-L10, and H17-L7\cite{doud2017complete}.
+These narrow antibodies bind the Ca2, Ca1, and Cb antigenic regions on HA's globular head\cite{caton1982antigenic}, and only neutralize a narrow slice of H1 viruses.
 
@@
 We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}, and calculated the fraction of virions carrying each mutation that survived neutralization.
-All experiments were performed in full biological triplicate using three independently generated virus libraries carrying single amino-acid mutations to HA~\citep{doud2016accurate}.
-%We have previously described~\citep{doud2017complete} mutational antigenic profiling of the three narrow strain-specific antibodies, and we re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived neutralization.
+All experiments were performed in full biological triplicate using three independently generated virus libraries carrying single amino-acid mutations to HA\cite{doud2016accurate}.
+%We have previously described\cite{doud2017complete} mutational antigenic profiling of the three narrow strain-specific antibodies, and we re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived neutralization.
 The correlations among replicates are shown in Figure~\ref{suppfig:corr}.
@@
 Specifically, there are mutations that enable over a third of virions to survive at concentrations where virtually all wildtype virions are neutralized (Figure~\ref{suppfig:maxfracsurvive}).
-Therefore, the virus can escape these four antibodies with the sort of large-effect single amino-acid mutations that characterize traditional influenza antigenic drift~\citep{yewdell1979antigenic,webster1980determination,koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction}.
+Therefore, the virus can escape these four antibodies with the sort of large-effect single amino-acid mutations that characterize traditional influenza antigenic drift\cite{yewdell1979antigenic,webster1980determination,koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction}.
 
@@
 \subsection*{The selected mutations are near the binding footprints of the antibodies}
-Antigenic mutations selected by narrow strain-specific antibodies against HA are thought to occur at residues in or near the physical binding footprint of the antibody~\citep{yewdell1979antigenic,webster1980determination,caton1982antigenic}.
+Antigenic mutations selected by narrow strain-specific antibodies against HA are thought to occur at residues in or near the physical binding footprint of the antibody\cite{yewdell1979antigenic,webster1980determination,caton1982antigenic}.
 We examined whether this was the case for the broad antibodies used in our experiments.
@@
 For the broad anti-receptor binding pocket antibody S139/1, there are strong escape mutations at sites 156, 158, and 193 (Figure~\ref{fig:structures}A; sites are in H3 numbering). 
-These three sites fall directly in the physical binding footprint of the antibody~\citep{lee2012heterosubtypic}, and are the same three sites where \citet{yoshida2009cross} selected escape mutants in H1, H2, and H3 HAs.
+These three sites fall directly in the physical binding footprint of the antibody\cite{lee2012heterosubtypic}, and are the same three sites where previous work has selected escape mutants in H1, H2, and H3 HAs\cite{yoshida2009cross}.
 Our data show that numerous different amino-acid mutations at each site confer neutralization resistance.
@@
 Although the anti-stalk antibodies C179 and FI6v3 only select mutations with small effects, these mutations almost all fall in or near the physical binding footprints of the antibodies (Figure~\ref{fig:structures}A).
-The two antibodies have similar epitopes and angles of approach~\citep{dreyfus2013structure}, and they select identical mutations at several sites (Figure~\ref{fig:structures}B). 
+The two antibodies have similar epitopes and angles of approach\cite{dreyfus2013structure}, and they select identical mutations at several sites (Figure~\ref{fig:structures}B). 
 The three largest-effect mutations for FI6v3 (K280S, K280T, and N291S) all introduce glycosylation motifs near the epitope, and all three mutations have similar magnitude antigenic effects in both FI6v3 and C179.
@@
 The most notable of these C179-specific mutations are at site 38.
-The additional breadth of FI6v3 over antibodies such as C179 that neutralize only group 1 HAs is because FI6v3 can accommodate a glycan on the asparagine at site 38 that is present in group 2 HAs~\citep{corti2011neutralizing,sui2009structural,ekiert2009antibody}.
+The additional breadth of FI6v3 over antibodies such as C179 that neutralize only group 1 HAs is because FI6v3 can accommodate a glycan on the asparagine at site 38 that is present in group 2 HAs\cite{corti2011neutralizing,sui2009structural,ekiert2009antibody}.
 However, the H38S mutation that has the largest effect on C179 resistance in our experiments does not introduce a glycosylation motif, showing that there are also other ways to escape anti-stalk antibodies at this site.
@@
 %The amino-acid identity at site 111 of HA2 can change the orientation of a conserved Trp21 in HA2, also resulting in group 1 and group 2 differences in the binding ability of stalk-targeting antibodies.
-An H111T mutation in the A/South Carolina/1/1918 (H1N1) has been reported to abrogate C179 binding~\citep{dreyfus2013structure}.
+An H111T mutation in the A/South Carolina/1/1918 (H1N1) has been reported to abrogate C179 binding\cite{dreyfus2013structure}.
 %However, we did not observe escape mutations at site 111, but this may be due to strain-specific differences.
@@
 The FI6v3 antibody also weakly selects several mutations at residue -8, which is part of HA's signal peptide (Figure~\ref{fig:structures}A). 
-This signal peptide is cleaved from the mature HA protein~\citep{daniels2003n,burke2014recommended}, although mutations at this site can affect HA's expression level~\citep{nordholm2017translational}.
+This signal peptide is cleaved from the mature HA protein\cite{daniels2003n,burke2014recommended}, although mutations at this site can affect HA's expression level\cite{nordholm2017translational}.
 
@@
 Do the mutations identified in our mutational antigenic profiling actually have the expected effect on antibody neutralization?
-We have previously validated many of the large-effect antigenic mutations selected by the narrow antibodies H17-L19, H17-L10, and H17-L7~\citep{doud2017complete}.
+We have previously validated many of the large-effect antigenic mutations selected by the narrow antibodies H17-L19, H17-L10, and H17-L7\cite{doud2017complete}.
 However, the mutations selected by the broad anti-stalk antibodies have much smaller effects -- especially for the broadest antibody, FI6v3.
@@
 (B), (C) In contrast to FI6v3, mutations selected by narrow antibodies have very large effects on neutralization.
-Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 taken from \citet{doud2017complete}.
+Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 taken from Doud et al\cite{doud2017complete}.
 }
@@
 These plots show all HA sites within 4 angstroms of the antibody in the crystal structure, plus any additional sites (marked with a *) where we identified antigenic mutations. 
-The logo plots at bottom show the preference of each HA site for each amino acid under selection for viral replication as measured by \citet{doud2016accurate}.
+The logo plots at bottom show the preference of each HA site for each amino acid under selection for viral replication as measured by Doud and Bloom\cite{doud2016accurate}.
 For instance, site 153 only tolerates tryptophan, so W occupies the entire height of the preference logo stack.
@@
 We can examine these possibilities using deep mutational scanning data that measures the tolerance of HA for each possible amino-acid mutation.
-Specifically, we have previously selected our A/WSN/1933 virus HA mutant libraries for variants that can replicate in cell culture, and then used deep sequencing to estimate the preference of each site in HA for each possible amino acid~\citep{doud2016accurate}.
+Specifically, we have previously selected our A/WSN/1933 virus HA mutant libraries for variants that can replicate in cell culture, and then used deep sequencing to estimate the preference of each site in HA for each possible amino acid\cite{doud2016accurate}.
 Figure~\ref{fig:muttolerance} shows these amino-acid preferences for all sites in HA within 4 angstroms of each broad antibody, with the antigenic effects of the mutations overlaid.
@@
 However, our mutational antigenic profiling shows that only a fraction of mutations at a fraction of these sites actually affect antibody neutralization.
-This finding is reminiscent of prior work showing that the binding energetics at protein-protein interfaces can be asymmetrically distributed across sites~\citep{jin1992high,cunningham1993comparison,dall1998mutational}.
+This finding is reminiscent of prior work showing that the binding energetics at protein-protein interfaces can be asymmetrically distributed across sites\cite{jin1992high,cunningham1993comparison,dall1998mutational}.
 The broad anti-stalk antibodies appear to distribute their binding energetics in such a way that altering the mutationally tolerant HA sites has relatively little effect on neutralization.
@@
 This finding emphasizes the importance of identifying antigenic mutations in a way that accounts for effect sizes.
-The classic approach for selecting escape mutations involves treating a virus stock with antibody at a concentration that completely neutralizes wildtype, and looking for viral mutants that survive this treatment~\citep{yewdell1979antigenic,webster1980determination}.
+The classic approach for selecting escape mutations involves treating a virus stock with antibody at a concentration that completely neutralizes wildtype, and looking for viral mutants that survive this treatment\cite{yewdell1979antigenic,webster1980determination}.
 There are no such single mutations for the HA and broad anti-stalk antibodies tested here, since no mutations shift the neutralization curve enough to enable survival at antibody concentrations that fully neutralize wildtype.
@@
 It is important to note that our experiments examine \emph{single} amino-acid mutations to the HA from one influenza virus strain.
-The protein evolution literature is full of examples of epistatic interactions that enable multiple mutations to access phenotypes not accessible by single mutations~\citep{gong2013stability,harms2014historical,starr2017alternative}.
+The protein evolution literature is full of examples of epistatic interactions that enable multiple mutations to access phenotypes not accessible by single mutations\cite{gong2013stability,harms2014historical,starr2017alternative}.
 Such epistasis is relevant to HA's evolution.  
-For instance, work by \citet{das2013defining} suggests that the sequential accumulation of mutations can shift the spectrum of available antibody-escape mutations.
-\citet{wu2017diversity} have used deep mutational scanning to directly demonstrate that rampant epistasis enables HA's receptor-binding pocket to accommodate combinations of individually deleterious mutations, some of which affect sensitivity to antibodies.
+For instance, work by Das et al\cite{das2013defining} suggests that the sequential accumulation of mutations can shift the spectrum of available antibody-escape mutations.
+Wu et al\cite{wu2017diversity} have used deep mutational scanning to directly demonstrate that rampant epistasis enables HA's receptor-binding pocket to accommodate combinations of individually deleterious mutations, some of which affect sensitivity to antibodies.
 Therefore, our work does not imply any absolute limits on the possibilities for antibody escape when evolution is given sufficient time to explore combinations of mutations.
-However, single mutations are the most accessible form of genetic variation, and much of influenza virus's natural antigenic drift involves individual mutations that reduce sensitivity to immunodominant antibody specificities~\citep{yewdell1979antigenic,webster1980determination,koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction}.
+However, single mutations are the most accessible form of genetic variation, and much of influenza virus's natural antigenic drift involves individual mutations that reduce sensitivity to immunodominant antibody specificities\cite{yewdell1979antigenic,webster1980determination,koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction}.
 Quantifying the antigenic effects of all such mutations therefore provides a relevant measure of ease of viral antibody escape.
 
-A major rationale for studying broadly neutralizing antibodies is that they are hoped to be more resistant to viral evolutionary escape than the antibodies that dominate natural immune responses to influenza virus~\citep{krammer2015advances,corti2017tackling}.
+A major rationale for studying broadly neutralizing antibodies is that they are hoped to be more resistant to viral evolutionary escape than the antibodies that dominate natural immune responses to influenza virus\cite{krammer2015advances,corti2017tackling}.
 We have used a new approach to quantify the extent to which this is actually true, and shown that the anti-viral activity of broad anti-stalk antibodies is indeed more -- although certainly not completely -- resistant to erosion by viral point mutations.
@@
 C179 IgG was purchased from Takara Bio Inc (Catalog \#M145).
-FI6v3 was purified from 293F cells transduced with a lentiviral vector encoding a commercially synthesized gene for the IgG form of the antibody, with the heavy and light chains reverse-translated from the protein sequence in the PDB structure 3ZTN~\citep{corti2011neutralizing} as described in \citet{balazs2013broad}.
-Genes encoding S139/1 in IgG form were were reverse-translated from the protein sequence in PDB structure 4GMS~\citep{lee2012heterosubtypic}, and used to express and purify protein by the Fred Hutchinson Cancer Research Center protein expression core.
+FI6v3 was purified from 293F cells transduced with a lentiviral vector encoding a commercially synthesized gene for the IgG form of the antibody, with the heavy and light chains reverse-translated from the protein sequence in the PDB structure 3ZTN\cite{corti2011neutralizing} as described previously\cite{balazs2013broad}.
+Genes encoding S139/1 in IgG form were were reverse-translated from the protein sequence in PDB structure 4GMS\cite{lee2012heterosubtypic}, and used to express and purify protein by the Fred Hutchinson Cancer Research Center protein expression core.
 
@@
 We performed neutralization assays using influenza viruses that carried GFP in the PB1 segment.
-These PB1flank-eGFP were generated in co-cultures of 293T-CMV-PB1 and MDCK-SIAT1-CMV-PB1 cells as described in \citet{bloom2010permissive}, using the standard bi-directional pHW181-PB2, ..., pHW188-NS reverse-genetics plasmids~\citep{hoffmann2000dna} for all genes \emph{except} PB1, plus the pHH-PB1flank-eGFP plasmid~\citep{bloom2010permissive}.
+These PB1flank-eGFP were generated in co-cultures of 293T-CMV-PB1 and MDCK-SIAT1-CMV-PB1 cells as described previously\cite{bloom2010permissive}, using the standard bi-directional pHW181-PB2, ..., pHW188-NS reverse-genetics plasmids\cite{hoffmann2000dna} for all genes \emph{except} PB1, plus the pHH-PB1flank-eGFP plasmid\cite{bloom2010permissive}.
 Each mutant was generated by repeating this process using a version of the pHW184-HA plasmid that had been engineered by site-directed mutagenesis to carry the indicated mutation.
-The neutralization assays themselves were performed by using a plate reader to quantify the GFP signal produced by MDCK-SIAT1-CMV-PB1 cells infected by PB1flank-eGFP virus that had been incubated with the indicated antibody concentration as described in \citet{hooper2013mutant}.
+The neutralization assays themselves were performed by using a plate reader to quantify the GFP signal produced by MDCK-SIAT1-CMV-PB1 cells infected by PB1flank-eGFP virus that had been incubated with the indicated antibody concentration as described previously\cite{hooper2013mutant}.
 All neutralization curves represent the mean and standard deviation of three measurements. 
@@
 \subsection*{Inference of HA phylogenetic tree}
-To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database~\citep{zhang2017influenza}, inferred the phylogenetic tree using RaxML~\citep{stamatakis2014raxml} with a GTR model, and visualized the tree using FigTree (\url{http://tree.bio.ed.ac.uk/software/figtree/}).
+To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database\cite{zhang2017influenza}, inferred the phylogenetic tree using RaxML\cite{stamatakis2014raxml} with a GTR model, and visualized the tree using FigTree (\url{http://tree.bio.ed.ac.uk/software/figtree/}).
 The HA sequences used are in File~\ref{suppfile:HAsubtypes}. 
-In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize~\citep{yoshida2009cross, lee2012heterosubtypic, okuno1993common, dreyfus2013structure, corti2011neutralizing}.
+In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize\cite{yoshida2009cross, lee2012heterosubtypic, okuno1993common, dreyfus2013structure, corti2011neutralizing}.
 Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8 and H11; and no antibodies have been tested against H17 and H18.
-The narrow H17-L19, H17-L10, and H17-L7 antibodies have not been tested against any other subtypes -- however, since these antibodies have a very limited range even among H1 HAs~\citep{caton1982antigenic}, we assume that they do not bind other subtypes.
+The narrow H17-L19, H17-L10, and H17-L7 antibodies have not been tested against any other subtypes -- however, since these antibodies have a very limited range even among H1 HAs\cite{caton1982antigenic}, we assume that they do not bind other subtypes.
 
-For the cladogram in Figure~\ref{fig:structures}C, the amino-acid identities at site 38 are from the strains tested against C179 by \citet{dreyfus2013structure}. 
+For the cladogram in Figure~\ref{fig:structures}C, the amino-acid identities at site 38 are from the strains tested against C179 by by Dreyfus et al\cite{dreyfus2013structure}. 
 For subtypes not tested, the amino-acid identity reported is that in the strain for that subtype in File~\ref{suppfile:HAsubtypes}.
@@
 \subsection*{Mutant virus libraries}
-The mutant virus libraries are those described in \citet{doud2016accurate}, and were produced in full biological triplicate.
-Briefly, these libraries were generated by using codon mutagenesis~\citep{bloom2014experimentally} to introduce random codon mutations into plasmid-encoded HA, and then using a helper-virus strategy that avoids the bottlenecks associated with standard influenza reverse genetics to create the virus libraries.
-Full details of the library generation and sequencing statistics that quantify how completely each of the triplicate libraries covers the possible amino-acid mutations are in \citet{doud2016accurate}.
+The mutant virus libraries are those described in Doud and Bloom\cite{doud2016accurate}, and were produced in full biological triplicate.
+Briefly, these libraries were generated by using codon mutagenesis\cite{bloom2014experimentally} to introduce random codon mutations into plasmid-encoded HA, and then using a helper-virus strategy that avoids the bottlenecks associated with standard influenza reverse genetics to create the virus libraries.
+Full details of the library generation and sequencing statistics that quantify how completely each of the triplicate libraries covers the possible amino-acid mutations have been described previously\cite{doud2016accurate}.
 
 \subsection*{Mutational antigenic profiling}
-The mutational antigenic profiling was performed as described in \citet{doud2017complete}. 
+The mutational antigenic profiling was performed as described previously\cite{doud2017complete}. 
 Briefly, we diluted each of the virus libraries to a concentration of $10^{6}$ TCID$_{50}$ per ml and incubated the virus dilutions with an equal volume of antibody at the intended concentration at 37$^\circ$C for 1.5 hours.
@@
 In addition, we performed technical replicates (independent neutralization experiments on the \emph{same} virus library) in some cases as indicated in Figure~\ref{suppfig:corr}.
-The virus-antibody mixtures were used to infect cells, and viral RNA was extracted, reverse-transcribed, and PCR amplified as described in \citet{doud2017complete}.
-In order to obtain high accuracy in the Illumina deep sequencing, we used the barcoded-subamplicon sequencing strategy described in \citet{doud2017complete}, which is a slight modification of the strategy of \citet{wu2014high}.
+The virus-antibody mixtures were used to infect cells, and viral RNA was extracted, reverse-transcribed, and PCR amplified as described previously\cite{doud2017complete}.
+In order to obtain high accuracy in the Illumina deep sequencing, we used the barcoded-subamplicon sequencing strategy described by Doud and Bloom\cite{doud2016accurate}, which is a slight modification of the strategy of Wu et al\cite{wu2014high}.
 
@@
 The average of these fractions across libraries are reported in Figure~\ref{fig:neutcurves}, and the values for each individual replicate are in Table~\ref{supptab:fracsurvive}.
-The fractions were estimated using qRT-PCR against the viral NP and canine GAPDH as described in \citet{doud2017complete}.
+The fractions were estimated using qRT-PCR against the viral NP and canine GAPDH as described previously\cite{doud2017complete}.
 Briefly, we made duplicate 10-fold serial dilutions of each of the virus libraries to use as a standard curve of infectivity.
@@
 \subsection*{Analysis of deep sequencing data}
-The deep sequencing data were analyzed using version 2.1.0 of the \texttt{dms\_tools2} software package~\citep{bloom2015software}, which is available at \url{http://jbloomlab.github.io/dms_tools2}.
+The deep sequencing data were analyzed using version 2.1.0 of the \texttt{dms\_tools2} software package\cite{bloom2015software}, which is available at \url{http://jbloomlab.github.io/dms_tools2}.
 File~\ref{suppfile:analysis_code} contains a Jupyter notebook that performs all steps of the analysis beginning with downloading the FASTQ files from the Sequence Read Archive.
@@
 \subsection*{Calculating the fraction of virions with each mutation that escapes antibody neutralization}
-In prior mutational antigenic profiling work~\citep{doud2017complete,dingens2017comprehensive}, we calculated the differential selection on each mutation as the logarithm of its enrichment relative to wildtype in an antibody-selected sample versus a mock-selected control.
+In prior mutational antigenic profiling work\cite{doud2017complete,dingens2017comprehensive}, we calculated the differential selection on each mutation as the logarithm of its enrichment relative to wildtype in an antibody-selected sample versus a mock-selected control.
 These \emph{mutation differential selection} values are useful for the analysis of individual experiments.
@@
 
-Code that performs these fraction surviving analyses has been added to version 2.1.0 of the \texttt{dms\_tools2} software package~\citep{bloom2015software} which is available at \url{http://jbloomlab.github.io/dms_tools2}.
+Code that performs these fraction surviving analyses has been added to version 2.1.0 of the \texttt{dms\_tools2} software package\cite{bloom2015software} which is available at \url{http://jbloomlab.github.io/dms_tools2}.
 
@@
 Deep sequencing data are available from the Sequence Read Archive under BioSample accession SAMN05789126.
-Computer code that analyzes these data to generate the results described in this paper is in File~\ref{suppfile:analysis_code}.
+Computer code that analyzes these data to generate all the results described in this paper is in File~\ref{suppfile:analysis_code}.
 
@@
 
-\bibliographystyle{mbe}
+\bibliographystyle{naturemag}
 \bibliography{references.bib}
@@
 These are the quantities referred to as $\gamma$.
-This table shows the values for the broad antibodies; values for the narrow H17-L17, H17-L10, and H17-L7 antibodies are reported in \citet{doud2017complete}.
+This table shows the values for the broad antibodies; values for the narrow H17-L17, H17-L10, and H17-L7 antibodies have been reported previously\cite{doud2017complete}.
 }

### 2017-11-28 paper/manuscript.tex
revisions up to neutralization assay subsection of results

@@
 \documentclass[11pt]{article}
-\usepackage[super]{natbib}
+\usepackage[super,comma]{natbib}
 \usepackage[margin=1.4in]{geometry}
@@
 
-\title{Quantifying the ease of viral escape from broad and narrow antibodies to influenza hemagglutinin}
+\title{Quantifying the effects of single mutations on viral escape from broad and narrow antibodies to an H1 influenza hemagglutinin}
 
@@
 However, rare antibodies broadly neutralize many viral strains.
-It is unclear how easily influenza virus might escape such antibodies if they became widespread due to therapeutic use or vaccination.
+It is unclear how easily influenza virus might escape such antibodies if it was under strong pressure to do so.
 Here we map all single amino-acid mutations that increase resistance to broad antibodies targeting an H1 hemagglutinin.
@@
 All antibodies select mutations, but the effect sizes vary widely. 
-The virus can escape a broad antibody targeting hemagglutinin's receptor-binding site the same way it escapes narrow strain-specific antibodies: via single mutations with huge effects.   
+The virus can escape a broad antibody that targets residues in hemagglutinin's receptor-binding site the same way it escapes narrow strain-specific antibodies: via single mutations with huge effects.
 In contrast, broad antibodies targeting hemagglutinin's stalk only select mutations with small effects. 
-Therefore, antibody breadth is not necessarily an indicator of the difficulty of viral escape.
-Broadly neutralizing antibodies targeting hemagglutinin's stalk are quantifiably harder to escape than the other antibodies tested here.
+Therefore, among the antibodies we have examined, breadth is an imperfect indicator of the potential for viral escape via single mutations.
+Broadly neutralizing antibodies targeting the H1 hemagglutinin stalk are quantifiably harder to escape than the other antibodies tested here.
 \end{abstract}
@@
 However, these studies did not select for antibody escape, so it is difficult to relate their measurements to the virus's evolutionary capacity under immune selection.
-Other work has shown that it is possible to select antigenic mutants with broad antibodies\cite{yoshida2009cross,chai2016two}, demonstrating that these epitopes are not entirely refractory to change.
+Other work has shown that it is possible to select antigenic mutants with broad antibodies\cite{yoshida2009cross,chai2016two,anderson2017natural}, demonstrating that these epitopes are not entirely refractory to change.
 But given that antibodies can select some antigenic variation even in measles virus\cite{birrer1981antigenic,ter1981antigenic} and polio virus\cite{crainic1983natural,diamond1985antigenic}, the existence of selectable mutations does not necessarily imply that influenza virus can escape broad antibodies as easily as it drifts away from narrow strain-specific ones.
@@
 However, the magnitudes of the antigenic effects vary greatly across antibodies.
-Single mutations make the virus completely resistant to both narrow strain-specific antibodies and a broad antibody against the receptor-binding site.
+Single mutations make the virus completely resistant to both narrow strain-specific antibodies and a broad antibody that targets residues in HA's receptor-binding site.
 But no single mutation does more than modestly increase the virus's resistance to two broad antibodies against the HA stalk.
@@
 \end{equation}
-%We plot the excess fraction above the library average since this quantity more clearly accentuates antigenic mutations against the background of other mutations.
 Details of how the calculations are extended to account for sequencing errors and sampling statistics are in the \nameref{sec:methods}.
@@
 (A) Crystal structures of the broad antibodies and sites of escape mutations selected by the narrow ones superimposed on the structure of the HA trimer (PDB 1RVX\cite{gamblin2004structure}). 
-S139/1 (PDB 4GMS\cite{lee2012heterosubtypic}) targets the receptor-binding site; C179 (PDB 4HLZ)\cite{dreyfus2013structure}) and FI6v3 (PDB 3ZTN\cite{corti2011neutralizing}) target the stalk. 
+S139/1 (PDB 4GMS\cite{lee2012heterosubtypic}) targets residues in the receptor-binding pocket; C179 (PDB 4HLZ\cite{dreyfus2013structure}) and FI6v3 (PDB 3ZTN\cite{corti2011neutralizing}) target the stalk.
 The sites of escape mutations for H17-L19, H17-L10, and H17-L7 are those mapped by Doud et al\cite{doud2017complete}. 
@@
 C179 is less broad, and neutralizes only some group 1 HAs (Figure~\ref{fig:antibody_summary}B).
-We also selected a broad antibody, S139/1, that targets HA's receptor-binding site and can neutralize both group 1 and group 2 HAs\cite{yoshida2009cross, lee2012heterosubtypic}.
+We also selected a broad antibody, S139/1, that crystallographic studies have shown binds to residues in HA's receptor-binding pocket\cite{lee2012heterosubtypic}, and which can neutralize both group 1 and group 2 HAs\cite{yoshida2009cross, lee2012heterosubtypic}.
 Finally, we re-analyzed deep sequencing data from prior mutational antigenic profiling of three narrow strain-specific antibodies, H17-L19, H17-L10, and H17-L7\cite{doud2017complete}.
@@
 We performed our experiments using the A/WSN/1933 (H1N1) strain of influenza, which is neutralized by all six antibodies (Figure~\ref{fig:neutcurves}).
-We used antibody concentrations sufficient for nearly complete neutralization of wildtype virus, allowing us to detect antigenic mutations that increase the fraction of mutant virions surviving beyond the baseline of wildtype (Figure~\ref{fig:neutcurves}).
-We used several concentrations of each broad antibody to increase sensitivity to antigenic mutations of both large and small effect.
+Our experiments utilize fully infectious influenza virus rather than pseudovirus, which is important since the accessibility of some epitopes can vary with HA density, which differs between fully infectious virus and pseudovirus\cite{corti2011neutralizing,joyce2016vaccine}.
 
@@
 The vertical dotted lines show the concentrations of antibody that were then used in the mutant virus library selections, and the tables give the overall fraction of the mutant virus libraries that survived at each concentration, determined by qRT-PCR.
+As described in the text, the antibody concentrations were chosen to give similar fractions of the mutant virus libraries that survive, rather than to fall at uniform positions on the neutralization curves of the wildtype virus.
 }
 \end{figure}
+
+The wildtype virus is neutralized with similar potency by all the antibodies (Figure~\ref{fig:neutcurves}).
+However, our selections are performed on mutant virus libraries, not wildtype virus.
+Because these libraries have different capacities to escape each antibody, the fraction of each library that survives high antibody concentrations will vary among antibodies.
+For instance, at concentrations that neutralize 99\% of the wildtype virus, we expect a larger fraction of a library to survive an antibody with many escape mutations than an antibody with few escape mutations.
+Therefore, rather than using the same concentration for all antibodies, we selected concentrations for each antibody where between 2\% and 0.1\% of the libraries survived in order to strongly select for escape mutations (Figure~\ref{fig:neutcurves}).
+Slight differences among antibodies in the fraction surviving within this range should not strongly affect our results, since Equations~\ref{eq:fracsurvive} and \ref{eq:fracsurvive_excess} account for such differences via the $\gamma$ term.
+However, to confirm the robustness of our results, we used several concentrations of each broad antibody (Figure~\ref{fig:neutcurves}).
 
@@
 \subsection*{Quantifying the antigenic effects of all mutations selected by each antibody}
-We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}, and calculated the fraction of virions carrying each mutation that survived neutralization.
+We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}. 
 All experiments were performed in full biological triplicate using three independently generated virus libraries carrying single amino-acid mutations to HA\cite{doud2016accurate}.
-%We have previously described\cite{doud2017complete} mutational antigenic profiling of the three narrow strain-specific antibodies, and we re-analyzed those deep sequencing data to calculate the fraction of virions with each mutation that survived neutralization.
 The correlations among replicates are shown in Figure~\ref{suppfig:corr}.
@@
 
-It is immediately obvious that the narrow strain-specific antibodies and the antibody targeting HA's receptor-binding pocket (S139/1) select mutations with large antigenic effects.
+It is immediately obvious that the narrow strain-specific antibodies and the antibody targeting residues in HA's receptor-binding pocket (S139/1) select mutations with large antigenic effects.
 For all four of these antibodies, there are multiple sites in HA where mutations enable a substantial fraction of virions to survive high antibody concentrations (Figure~\ref{fig:avgfracsurvive}).
@@
 
-For the broad anti-receptor binding pocket antibody S139/1, there are strong escape mutations at sites 156, 158, and 193 (Figure~\ref{fig:structures}A; sites are in H3 numbering). 
+For the S139/1 antibody that targets residues in the HA receptor-binding pocket, there are strong escape mutations at sites 156, 158, and 193 (Figure~\ref{fig:structures}A; sites are in H3 numbering).
 These three sites fall directly in the physical binding footprint of the antibody\cite{lee2012heterosubtypic}, and are the same three sites where previous work has selected escape mutants in H1, H2, and H3 HAs\cite{yoshida2009cross}.
@@
 The FI6v3 antibody also weakly selects several mutations at residue -8, which is part of HA's signal peptide (Figure~\ref{fig:structures}A). 
-This signal peptide is cleaved from the mature HA protein\cite{daniels2003n,burke2014recommended}, although mutations at this site can affect HA's expression level\cite{nordholm2017translational}.
+This signal peptide is cleaved from the mature HA protein\cite{daniels2003n,burke2014recommended}, although mutations at this site can affect HA's expression level\cite{nordholm2017translational}, which might conceivably affect HA density on virions and subsequently antibody neutralization\cite{corti2011neutralizing,joyce2016vaccine}.
 
@@
 Briefly, these libraries were generated by using codon mutagenesis\cite{bloom2014experimentally} to introduce random codon mutations into plasmid-encoded HA, and then using a helper-virus strategy that avoids the bottlenecks associated with standard influenza reverse genetics to create the virus libraries.
+Although a helper virus is used to generate the libraries from plasmids, the viruses in the resulting library carry the full complement of genes and are fully infectious and replication-competent\cite{doud2016accurate}.
+This fact is important, since the accessibility of HA epitopes can depend on virion HA density, which is often lower in pseudovirus than in fully infectious virus\cite{corti2011neutralizing,joyce2016vaccine}.
 Full details of the library generation and sequencing statistics that quantify how completely each of the triplicate libraries covers the possible amino-acid mutations have been described previously\cite{doud2016accurate}.

### 2017-11-29 paper/Immunity_reviews.tex
mike thoughts in blue

@@
 We should make sure that we are not over-stating the generality of the claims.}
+{\color{blue} I agree with this too, specifically about the limited number of antibodies that we used and how this should limit generalization of the finding.}
 
@@
 {\color{red} We could comment briefly on the number of sites of escape for the different residues.}
+{\color{blue} While also mentioning that both number of sites and effect size of mutations at those sites are relevant (and I guess also the number of mutations within each site as well...), with the emphasis being the difference in effect sizes between the antibodies, but I agree with the reviewer that we should also draw attention to the number of sites of escape.}
 
@@
 {\color{red} We could easily rank the antibodies by something like the median effect of the top 5 escape mutations. Mike had a figure that looked a bit like that.}
+{\color{blue} I do still like my plots of rank-ordered mutations...
+perhaps there is a place in the figures we could squeeze that in.
+I do think it is a nice way of visualizing both number of escape mutations and effect sizes.
+Could be useful as we comment on the # of sites issue raised above, although the plots as I was making them were by mutation not by site, but we could also show sites ranked by maximum effect size or something along those lines.}
 
@@
 We could mention something about possibly modulating expression level...}
+{\color{blue} This is also why I was pushing a little to deemphasize any interpretations about this mutant, or specifically the validation data for it.
+The plateau in the neutralization curve is funky (more likely technical issue than biological), and normally I wouldn't be too concerned about something like that if there was a large and obvious IC50 shift, but in this case when the curve shift is so small to begin with, a funky plateau kind of throws everything into question.
+I DO still think the original result from the library experiments is potentially super interesting, and like to think about expression level modulation and stuff, but want to keep it clear that it's speculation, and not hang any hats on the neutralization assay.}
 
@@
 Might be best just to remain very circumspect about K(-8)T?}
+{\color{blue} I agree best to remove as much as willing about K(-8)T, or just say the effect size of that one (and M17L) is so small that it is within the error the neutralization assay.
+Could drop the validation data for K(-8)T and M17L completely (or to supplement), and consider dropping panels B and C which are essentially reprints, and use the extra space to show the rank-order mutation effect size plots comparing all the antibodies.
+Then the figure is showing that we validated small effect sizes for the strongest FI6v3 mutations, and that the mutations are few in number and small in effect size for the stalk abs relative to the head abs with the "mike's plot" mentioned above.
+I also agree not worth doing statistics. I commend the reviewer for very politely calling us out on this.}
 
@@
 It is a well-written paper and the authors are to be commended for their encyclopedic referencing of classic literature.
-I appreciate the desire of the authors to perform an "apples-to-apples" comparison, and there is much to like about this paper, but unfortunately, this analysis has flaws that preclude me recommending it for publication.
+{\color{blue} (Mad Props, Jesse)} I appreciate the desire of the authors to perform an "apples-to-apples" comparison, and there is much to like about this paper, but unfortunately, this analysis has flaws that preclude me recommending it for publication.
 
@@
 However, [Lee et al 2012 PNAS] published the crystal structure, and as can be seen in figure 1, the HCDR2 loop of S139/1 inserts into the receptor binding site.}
+{\color{blue} So clearly Reviewer \#2 isn't Ian Wilson.}
 
@@
 }
+{\color{blue} I guess to really satisfy this reviewer we would have to do some sort of experiment measuring protein levels in virions or EM or something.
+It seems nitpicky, and I would hedge that infectious virus grown in lab has physiological density of surface proteins, but do we really know that?
+I could imagine weird density in the helper-virus-rescued P0 virus, remember we used to speculate that funky things could be going on at that P0 stage secondary to the HAs being expressed off of plasmids and not incoming viral RNPs and all that --- but I think that doesn't matter since we are using P1 library in the experiments here.}
 
@@
 }
+{\color{blue} I too would guess some correlation between neutralization escape effect and FcR-based clearance escape effect.
+But I think in light of the good points the reviewer makes about surface density impacting amount of antibody needed to neutralize, and that the relevant density of HA on the cell surface during antibody Fc - FcR engagement  might be different than the density on virions during neutralization, etc, it might be good to lay out a caveat or just mention FcR-based viral clearance somewhere.
+I would have to think about this a little more.}
 
@@
 }
+{\color{blue} I would say 'measured' instead of 'computed' on the mutant libraries and just make it clear that we measured the library neutralization with qrtpcr.
+This is already stated clearly in the figure 3 caption.}
 
@@
 }
+{\color{blue} I agree, to me, apples to apples is better achieved by roughly matching neutralization stringency.
+"concentrations likely to be present in a physiologic condition" seems really hard to define and translate from physiologic infection to in vitro neutralization.}
 
 \section*{Reviewer \#3} 
-Boud et al., used a previously characterized HA single AA mutant library and a previously developed method for quantifying viral survival to quantify the viral escape from three broad neutralizing antibodies.
+Boud et al.{\color{blue} (no comment)}, used a previously characterized HA single AA mutant library and a previously developed method for quantifying viral survival to quantify the viral escape from three broad neutralizing antibodies.
 While the work is interesting, there are several problems.
@@
 }
+{\color{blue} Agreed, no need to address}
 
@@
 }
+{\color{blue} agree, i think we are clear. }
 
@@
 But we could add some sort of test statistic I suppose...}
+{\color{blue} I'm getting tired now and agree with all comments above and feel less inclined to really change anything with respect to this third review, which is not very useful.
+I wanted to say here that I agree with not doing a statistical test for this.}
 
@@
 }
+{\color{blue} I think the paragraph in discussion about single mutations is sufficient.}
 

### 2017-11-30 paper/manuscript.tex
full pass on editing paper

@@
 We have previously validated many of the large-effect antigenic mutations selected by the narrow antibodies H17-L19, H17-L10, and H17-L7\cite{doud2017complete}.
-However, the mutations selected by the broad anti-stalk antibodies have much smaller effects -- especially for the broadest antibody, FI6v3.
+However, the mutations selected by the broad anti-stalk antibodies have much smaller effects in our mutational antigenic profiling -- especially for the broadest antibody, FI6v3.
 We therefore tested some of these FI6v3-selected mutations using neutralization assays on individual viral mutants.
@@
 The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.
-We tested one of these mutations, K(-8)T, and found that it did appear to lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A).
+We tested one of these mutations, K(-8)T, and it did appear to lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A).
 As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them affected neutralization resistance.
@@
 
-The data in Figure~\ref{fig:muttolerance} show that the lack of large-effect escape mutants from FI6v3 and C179 is not due to complete mutational intolerance of HA sites in the antibody-binding footprints.
+The data in Figure~\ref{fig:muttolerance} show that the lack of large-effect escape mutants from FI6v3 and C179 is not entirely due to the mutational intolerance of HA sites in the antibody-binding footprints.
 Some HA sites in each antibody footprint are fairly mutationally tolerant, and contain a range of mutations in the viral libraries used in our antibody selections.
@@
 This finding is reminiscent of prior work showing that the binding energetics at protein-protein interfaces can be asymmetrically distributed across sites\cite{jin1992high,cunningham1993comparison,dall1998mutational}.
-The broad anti-stalk antibodies appear to distribute their binding energetics in such a way that altering the mutationally tolerant HA sites has relatively little effect on neutralization.
+The broad anti-stalk antibodies therefore appear to both mostly target mutationally intolerant sites and distribute their binding energetics in such a way that altering the mutationally tolerant HA sites has relatively little effect on neutralization.
 
@@
 As expected, single amino-acid mutations can make the virus completely resistant to narrow strain-specific antibodies against HA's globular head.
-However, such mutations can also enable the virus to escape the broad S139/1 antibody targeting HA's receptor-binding pocket, despite the fact that this antibody neutralizes multiple subtypes.
+However, such mutations can also enable the virus to escape the broad S139/1 antibody targeting residues in HA's receptor-binding pocket, despite the fact that this antibody neutralizes multiple subtypes.
 But no single mutation has a comparably large effect on neutralization by two broad antibodies targeting HA's stalk, FI6v3 and C179.
@@
 However, our approach shows that there are mutations that have more modest ($<$10-fold) effects on neutralization by even the broadest antibody. 
+When interpreting the magnitude of these effects, is important to note that our experiments only measure how mutations affect neutralization, and not how they effect Fc-mediated effector functions that are responsible for much of the \textit{in vivo} protection afforded by anti-stalk antibodies derive from Fc-mediated effector functions~\cite{dilillo2014broadly,dilillo2016broadly}.
 
-It is important to note that our experiments examine \emph{single} amino-acid mutations to the HA from one influenza virus strain.
+Another important caveat that our experiments examine \emph{single} amino-acid mutations to the HA from one influenza virus strain.
 The protein evolution literature is full of examples of epistatic interactions that enable multiple mutations to access phenotypes not accessible by single mutations\cite{gong2013stability,harms2014historical,starr2017alternative}.
@@
 A major rationale for studying broadly neutralizing antibodies is that they are hoped to be more resistant to viral evolutionary escape than the antibodies that dominate natural immune responses to influenza virus\cite{krammer2015advances,corti2017tackling}.
-We have used a new approach to quantify the extent to which this is actually true, and shown that the anti-viral activity of broad anti-stalk antibodies is indeed more -- although certainly not completely -- resistant to erosion by viral point mutations.
-Going forward, we suggest that quantifying the ease of viral escape will be a useful complement to more traditional techniques that characterize the structural epitopes and strain breadth of anti-viral antibodies.
+We have used a new approach to quantify the extent to which this is actually true, and shown that virus neutralization by broad anti-stalk antibodies is indeed more -- although certainly not completely -- resistant to erosion by viral point mutations.
+Going forward, we suggest that completely mapping viral escape mutations will be a useful complement to more traditional techniques that simply characterize the breadth of anti-viral antibodies against circulating strains.
 
@@
 File~\ref{suppfile:analysis_code} contains a Jupyter notebook that performs all steps of the analysis beginning with downloading the FASTQ files from the Sequence Read Archive.
-Detailed statistics about the sequencing depth and error rates are shown in this Jupyter notebook.
+Detailed statistics about the sequencing depth and error rates are shown in this Jupyter notebook and its HTML rendering in File~\ref{suppfile:analysis_html}.
 
@@
 Deep sequencing data are available from the Sequence Read Archive under BioSample accession SAMN05789126.
-Computer code that analyzes these data to generate all the results described in this paper is in File~\ref{suppfile:analysis_code}.
+Computer code that analyzes these data to generate all the results described in this paper is in File~\ref{suppfile:analysis_code}, and an HTML version of the analysis notebook is in File~\ref{suppfile:analysis_html}.
 
@@
 The actual analysis is performed by the Jupyter notebook \texttt{analysis\_notebook.ipynb}, which includes embedded plots summarizing key statistics and results.
-An HTML version of this notebook is also included as \texttt{analysis\_notebook.html}.
+An HTML version of this notebook is also included as File~\ref{suppfile:analysis_code}.
+}
+\end{suppfile}
+
+\begin{suppfile}
+\caption{\label{suppfile:analysis_html}
+{\bf HTML version of the analysis notebook.}
+This file is an HTML rendering of the Jupyter notebook in File~\ref{suppfile:analysis_code}.
+It contains detailed plots for all aspects of the deep sequencing data and its analysis.
 }

### 2017-11-30 paper/manuscript.tex
text mention that S139 is tested against H9

@@
 In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize\cite{yoshida2009cross, lee2012heterosubtypic, okuno1993common, dreyfus2013structure, corti2011neutralizing}.
-Among broad antibodies, S139/1 has not been tested against H8, H9, and H11; C179 has not been tested against H8 and H11; and no antibodies have been tested against H17 and H18.
+Among broad antibodies, S139/1 has not been tested against H8 and H11; C179 has not been tested against H8 and H11; and no antibodies have been tested against H17 and H18.
 The narrow H17-L19, H17-L10, and H17-L7 antibodies have not been tested against any other subtypes -- however, since these antibodies have a very limited range even among H1 HAs\cite{caton1982antigenic}, we assume that they do not bind other subtypes.

### 2017-11-30 paper/manuscript.tex
text fix

@@
 However, our approach shows that there are mutations that have more modest ($<$10-fold) effects on neutralization by even the broadest antibody. 
-When interpreting the magnitude of these effects, is important to note that our experiments only measure how mutations affect neutralization, and not how they effect Fc-mediated effector functions that are responsible for much of the \textit{in vivo} protection afforded by anti-stalk antibodies derive from Fc-mediated effector functions~\cite{dilillo2014broadly,dilillo2016broadly}.
+When interpreting the magnitude of these effects, is important to note that our experiments only measure how mutations affect neutralization, and not how they affect Fc-mediated functions that are responsible for much of the \textit{in vivo} protection afforded by anti-stalk antibodies\cite{dilillo2014broadly,dilillo2016broadly}.
 

### 2017-11-30 paper/manuscript.tex
a few more tweaks

@@
 The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.
-We tested one of these mutations, K(-8)T, and it did appear to lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A).
-As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them affected neutralization resistance.
+We tested one of these mutations, K(-8)T, and it did appear to lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A), although we are not confident that this effect exceeds the noise in our neutralization assays.
+As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them affected neutralization resistance within the sensitivity of our assays.
 
@@
 A major rationale for studying broadly neutralizing antibodies is that they are hoped to be more resistant to viral evolutionary escape than the antibodies that dominate natural immune responses to influenza virus\cite{krammer2015advances,corti2017tackling}.
-We have used a new approach to quantify the extent to which this is actually true, and shown that virus neutralization by broad anti-stalk antibodies is indeed more -- although certainly not completely -- resistant to erosion by viral point mutations.
+We have used a new approach to quantify the extent to which this is actually true, and shown that neutralization of an H1 virus by broad anti-stalk antibodies is indeed more -- although certainly not completely -- resistant to erosion by viral point mutations.
 Going forward, we suggest that completely mapping viral escape mutations will be a useful complement to more traditional techniques that simply characterize the breadth of anti-viral antibodies against circulating strains.

### 2017-11-30 paper/manuscript.tex
fix per Mike comment about IC50s not really being very similar across antibodies
Fixed minor typo in Discussion

@@
 
-The wildtype virus is neutralized with similar potency by all the antibodies (Figure~\ref{fig:neutcurves}).
+The wildtype virus is neutralized by all the antibodies, with IC50s between 0.01 and 1 $\mu$g/ml (Figure~\ref{fig:neutcurves}).
 However, our selections are performed on mutant virus libraries, not wildtype virus.
@@
 However, our approach shows that there are mutations that have more modest ($<$10-fold) effects on neutralization by even the broadest antibody. 
-When interpreting the magnitude of these effects, is important to note that our experiments only measure how mutations affect neutralization, and not how they affect Fc-mediated functions that are responsible for much of the \textit{in vivo} protection afforded by anti-stalk antibodies\cite{dilillo2014broadly,dilillo2016broadly}.
+When interpreting the magnitude of these effects, it is important to note that our experiments only measure how mutations affect neutralization, and not how they affect Fc-mediated functions that are responsible for much of the \textit{in vivo} protection afforded by anti-stalk antibodies\cite{dilillo2014broadly,dilillo2016broadly}.
 

### 2018-01-16 paper/NatureMicrobiologyReviews.tex
added *Nature Microbiology* reviews

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format

\usepackage[margin=1.2in]{geometry}

\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
\usepackage{color}
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

\title{Response to reviews of ``Quantifying the effects of single mutations on viral escape from broad and narrow antibodies to an H1 influenza hemagglutinin'' from \textit{Nature Microbiology}}
\author{Michael B. Doud, Juhye M. Lee, and Jesse D. Bloom}

\begin{document}
\maketitle

\subsection*{Overview of reviewer response and changes to the manuscript.}

We received four reviews for the original submission to \textit{Nature Microbiology}.
All four reviewers found the work to be scientifically sound.
Two reviewers (\#1 and \#4) were also highly enthusiastic about the impact.
The other two reviewers (\#2 and \#3) concurred about scientific soundness, but were less enthusiastic about impact.
The most critical was Reviewer \#3, who still wrote that ``[t]he study is performed to an extremely high technical standard'' --- a positive comment to receive from the most negative reviewer!

Given that all reviewers found the work to be scientifically sound and at least half found it to be high impact, we request that the paper be considered for \textit{Nature Communications} with the current reviews.
Impact is subjective, and the fact that half the reviewers found the paper of sufficient impact for \textit{Nature Microbiology} indicates that it will interest many in the field. 
In addition, we emphasize the following impactful aspects of our study:
\begin{itemize}\parskip1pt
\item It introduces a rigorous method to compare the potential for immune-escape by single amino-acid mutations across antibodies.
\item It is the first paper to completely map how all single amino-acid mutations affect neutralization by broadly neutralizing anti-influenza antibodies.
\item It is the first direct head-to-head comparison of immune-escape from anti-head versus anti-stalk antibodies to influenza hemagglutinin.
\end{itemize}

Below we provide a detailed point-by-point response to the reviews.
As you will see, we have been able to clarify and address all the questions / concerns.

In conclusion, our paper has been rigorously evaluated by four expert reviewers, who all found it to be scientifically sound, raising only minor points that we have clarified in our response / revision.
The only question was whether our work was of sufficient impact for \textit{Nature Microbiology}, with the reviewers splitting about equally on this question.
The \textit{Nature Communications} policies read: ``While \textit{Nature} and Nature research journals strive to publish the highest impact research within these disciplines, \textit{Nature Communications} is committed to publishing important advances of significance to specialists within each field.''
Our study clearly meets these criteria, and so is suitable for \textit{Nature Communications}.

\subsection*{Response to reviews}
Below find our responses to the reviews received from \textit{Nature Microbiology}.
The original comments {\color{blue} are in blue}, and our responses are in black.

\color{blue}

\subsubsection*{Reviewer Expertise:}

\begin{itemize}\parskip1pt
\item Referee \#1: influenza antibody neutralization and escape
\item Referee \#2: viral evolution, drift, deep-sequencing, computation
\item Referee \#3: influenza antibody neutralization and escape
\item Referee \#4: computational analysis of viral evolution
\end{itemize}

\subsubsection*{Reviewer \#1 (Remarks to the Author):}

In their manuscript Doud and colleagues quantify the effects of single mutations on viral escape from both narrow and broadly binding anti-hemagglutinin antibodies. The methodology is based on a mutant library of an H1N1 strain which are analyzed by deep sequencing after a single infection cycle in the presence or absence of mAb. The data presented in the manuscript is of high significance for the field and the work is technically sound. There are some issues that the authors need to address but no major weaknesses were identified.

Specific points

1) Multicylce replication might have changed the data somewhat because it would have allowed the virus to acquire additional mutations.

2) The virus used, WSN, is highly artificial and does not reflect actual human influenza viruses. It is trypsin independent (human viruses are not) and it is actually neurotropic (which is highly unusual). Also, it is a mouse adapted virus. While this is not necessarily a problem in the context of the presented data this caveat should be extensively discussed and avoided in the future.

3) Page 2, beginning of second last paragraph: It is unclear if such a vaccine would really enhance the pressure dramatically. This would depend on vaccine coverage and potency. Currently, the drift of influenza viruses is driven by immunity induced by natural infection, not by vaccines (global vaccination rates are extremely low, even in high income countries in Europe the coverage is low).

4) Data deposition at github: How long will data/software stored on this platform be available? If versions are update, is this tracked/recorded?

5) Figure 6: It might be a good idea to quantify the IC50 shifts and - since the neutralization assay was hopefully done at least in triplicates - check what is significant and what not. E.g. the effect of the -8 mutation for FI6 doesn't look real.

6) It would be important to quantify differences in mAb binding to the different escape mutants shown in Figure 6, e.g. by SPR/BLI. Right now it is unclear which mutant produces a complete loss of binding, if affinity changes result in the loss of neutralization etc. This needs to be explored.

7) Page 15/first page of discussion: The authors did not find a single-mutant virus with full loss of neutralization by anti-stalk mAbs. It has been noted that it is relatively difficult to make escape mutants against anti-stalk mAbs and the process usually involves serial passaging of virus in the presence of sub-neutralizing mAb concentrations - but usually full escape mutants are then recovered. In most cases these escape mutants have a) more than one mutation and b) have no completely loss of binding (but complete loss of neutralization). But there are also exceptions to this rule. Two papers have so far described escape from stalk-based mAbs in detail (https://www.ncbi.nlm.nih.gov/pubmed/27351973 and https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4362269/). It is strongly suggested that the authors consult this papers and include their findings in the discussion.

8) It would be good to have a negative control experiment with a non-binding human or mouse mAb. This is missing but should actually be the 'real' negative control.

\subsubsection*{Reviewer \#2 (Remarks to the Author):}

The authors use a previously described method of mutational antigenic profiling, a very clever and elegant approach exploited in previous papers in PLOS Pathogens and CHM. This approach involves generating viral libraries containing all mutations in the protein of interest, selecting these viruses with or without antibody, and deep-sequencing to determine the relative frequencies of each mutation. The full methodology (except for summary statistics described in Equations 10 and 11), as well as all the results for the narrow antibodies, were described in their previous paper (Doud et al., 2017, PLOS Pathogens). Figure 7 seems to be imported from another paper (Doud et al., 2016, Viruses).

In this work, they compare three broad and three narrow antibodies. The broad: FI6v3, binds stalk, neutralizes both group 1 and 2 HAs; C179 binds stalk, neutralizes only some group 1 HAs; S139/1, binds receptor-binding pocket, neutralizes both group 1 and group 2 HAs. The authors compare these in a re-analysis of data from narrow, strain-specific antibodies: H17-L19, H17-L10, and H17-L7, bind the Ca2, Ca1, and Cb antigenic regions on HA’s globular head.

Using these three broad antibodies and WSN/33 Flu, they selected for escape mutations, and compared them to previous narrow antibody data. They find that escape from narrow or RBS-targeting Ab is relatively easy by single point mutation; while those targeting the stalk are difficult to escape by single point mutation. So breadth of antibody activity per se, is not a predictor of likelihood of escape.

They go on to show that the selected mutations are nearly all in or close to the antibody-binding footprint. While this may not be surprising, it is a nice set of data. Another nice aspect, is that the authors show that numerous different amino acid mutations at each site confer neutralization resistance - most studies tend to identify just one or two, rather than all of the possible aa.

They further show that the mutational antigenic profiling is highly predictive of the
results of the neutralization assays, even for small-effect mutations, as was previously shown for large-effect mutations from narrow antibody escape.

They compare the sites where mutations arise from broad antibody, to previous mutational scanning data that reveals whether these regions are tolerant or not to mutation. They find that this is only partially true, but cannot account for all lack of escape mutations; however, they do not uncover what the other reasons are.

Some questions or comments that come to mind:

1. If the rationale is to ask whether current circulating strains could escape immunity based on broad neutralising antibody elicited by vaccination, could the authors not use circulating strains for their experiments, or at least more lab strains that only WSN?

2. What difference in terms of viral fitness is there between a mutation present in large proportion (for the narrow antibodies) vs present at low frequency (for the stalk-targeting antibodies). Figure 7 partly answers this question; however, it would be interesting to confirm the fitness of those mutants using more conventional growth curves, binding assays, etc.

3. Are all changes in amino-acid obtained after one nucleotide change? Are the authors able to detect multiple nucleotide or amino acid changes?

4. The major caveat of this study is that it only takes into account one mutation. Authors could check a combination of the mutations they obtain with their experiments (especially the ones escaping the stalk-targeting antibodies).

5. Figure 6 confirms some of the findings using neutralisation assays. As the authors state, it is unclear whether some of the observations actually exceed the noise of the assay. In particular, K(-8)T looks as far from WT as does K280A. It could be that (1) all these differences are due to noise or that (2) there is actually a biological difference between WT and K280A. Whether (1) occurs is unclear: “three biological replicates” should perhaps be more detailed in the Materials and Methods section as to what this means: are they three rows of the same plate performed on the same day, on different days, etc.? According to the authors’ previous paper (Doug et al., 2017, PLOS Pathogens), they performed “three replicate dilution columns”. The neutralisation curves seem therefore to have been performed at different days (besides, the WT curves are the same or similar to that of Figure 3, and the neutralisation curves from the narrow antibodies were imported from the
authors’ previous paper). This can induce a batch effect, which may not be linear, thus the normalisation performed (with no antibody and background noise) may not regress this out. From the standpoint of experimental design, these experiments should be repeated, with all viruses for a given antibody performed in the same batch; the triplicates will then give a more accurate measurement of noise. In general, all neutralisation experiments should be repeated, so as to make sure the findings are reproducible. If this has been done, it was not clear from the figure legend, and could be added for clarity. To address (2), authors should provide details of Fr,a for K280A. If I understood correctly, because Fr,a-excess (defined in Equation 2) hides mutations with lower fitness than average, it is not possible for the reader to assess whether this mutation is “neutral” or it has lower fitness with the antibody under study.

6. Could the authors detail, in the legend, how the curves of Figure 6 were fitted?

\subsubsection*{Reviewer \#3 (Remarks to the Author):}

This paper identifies escape mutations against anti-influenza HA antibodies and quantifies their effects. The authors employed a high-throughput approach known as mutational antigenic profiling, which was developed in their previous study (1). The authors made a minor technical improvement in this study to enable a fairer comparison across profiling results from different antibodies. Two broadly neutralizing antibodies that target the HA stem (mouse C179 and human FI6v3), and one that targets the HA receptor-binding site (S139/1) were examined in this study. Additionally, the authors reanalyzed the results from their previous study (1) on three strain-specific antibodies (H17-L7, H17-L10, and H17-L19). The authors observed strong escape mutations against the strain-specific antibodies and broadly neutralizing antibody S139/1. In contrast, only weak escape mutations were identified against the two stem antibodies. The authors conclude that “breadth is an imperfect indicator of
the potential for viral escape”. The authors also conclude that broadly neutralizing antibodies targeting the H1 hemagglutinin stalk are “quantifiably harder to escape than the other antibodies tested here”. The last two conclusions are not unexpected and are not particularly novel insights.

The study is, however, performed to an extremely high technical standard. Nevertheless, the impact is limited. The improvement in methodology over their previous study (1) is minor. Also, it is known that strong escape mutations against S139/1, one of the broadly neutralizing antibodies tested in this study (although not nearly as broad as FI6v3), can be generated rapidly through passaging the virus in the presence of the antibody (2). In fact, strong escape mutations against some of the stalk antibodies that were not tested in this study, such as CR8020 (3), CR8043 (4), CR6261 (5) and 39.29 (6), have also been generated through single point mutations. Overall, the impact and novelty of the study is not sufficient for publication in Nature Microbiology.

Specific comments:

1.  Strong escape mutations to C179 have been identified in the original paper that discovered the antibody (7). But it seems like those mutations were not identified here. Can the authors provide some explanations for such a discrepancy?

2.  How is the curve fitting performed for the neutralization assay results? I do not seem to see any description for the curve-fitting method.

References

1.  Doud MB, Hensley SE, Bloom JD. Complete mapping of viral escape from neutralizing antibodies. PLoS Pathog. 2017;13(3):e1006271.

2.  Yoshida R, Igarashi M, Ozaki H, Kishida N, Tomabechi D, Kida H, et al. Cross-protective potential of a novel monoclonal antibody directed against antigenic site B of the hemagglutinin of influenza A viruses. PLoS Pathog. 2009;5(3):e1000350.

3.  Ekiert DC, Friesen RH, Bhabha G, Kwaks T, Jongeneelen M, Yu W, et al. A highly conserved neutralizing epitope on group 2 influenza A viruses. Science. 2011;333(6044):843-50.

4.  Friesen RH, Lee PS, Stoop EJ, Hoffman RM, Ekiert DC, Bhabha G, et al. A common solution to group 2 influenza virus neutralization. Proc Natl Acad Sci USA. 2014;111(1):445-50.

5.  Throsby M, van den Brink E, Jongeneelen M, Poon LL, Alard P, Cornelissen L, et al. Heterosubtypic neutralizing monoclonal antibodies cross-protective against H5N1 and H1N1 recovered from human IgM+ memory B cells. PLoS One. 2008;3(12):e3942.

6.  Chai N, Swem LR, Reichelt M, Chen-Harris H, Luis E, Park S, et al. Two escape mechanisms of influenza A virus to a broadly neutralizing stalk-binding antibody. PLoS Pathog. 2016;12(6):e1005702.

7.  Okuno Y, Isegawa Y, Sasao F, Ueda S. A common neutralizing epitope conserved between the hemagglutinins of influenza A virus H1 and H2 strains. J Virol. 1993;67(5):2552-8.

\subsubsection*{Reviewer \#4 (Remarks to the Author):}

This paper tackles a really nice question: Are some antibodies broad because they target sites that face little immune pressure, or are they broad because the epitopes they target cannot easily mutate to escape them? Put differently, are conserved epitopes that are targeted by some antibodies conserved because they've not had to change, or are they conserved because no beneficial changes are accessible? This question is obviously relevant for "universal" vaccine development, but it's also important for understanding strain structure and evolution more generally.

To investigate this question, the authors analyzed the ability of a H1 flu strain to escape six well characterized antibodies via single mutations. They showed that for most antibodies, including a broadly neutralizing antibody, single mutations allowed escape from antibody detection. In the other two antibodies, no mutations really rescued the virus. By comparing the mutations conferring antibody escape to those that are generally tolerated during replication, the authors further showed that mutational tolerance does not fully explain failure to escape: despite substitutions, some antibodies are harder to escape than others.

These are important observations. My only real criticism of the work is that I found the description of mutational effects confusing as early as the abstract. Does the measured frequency of antibody escape mutants reflect only antibody selection, or does it also include selection for replicative capacity? The distinction between antibody selection and general fitness is not made until p. 13, but I think readers (especially those familiar with other work from the lab) might benefit if these distinctions were made in the introduction and review of the methods. My take is that the fraction surviving antibody escape includes both, in that non-viable mutants were rare in the initial library (again implied on p. 13), but this could be clarified. It seems the contribution of each might be quantifiable, or at the relative mutational tolerances of the different epitopes compared.

Minor points:

1. How do the concentrations for selecting escape mutants in vitro compare to likely concentrations in vivo? Concentrations may be too high or low in vivo to promote adaptation.

2. Outside these experiments, how much data is there on the antibodies' neutralization capacities within HA subtypes? i.e., how broad or narrow do we really know them to be? Obviously such data might be complementary.

3. Additional suggested references:

"the viruses's rapid antigenic evolution erodes the effectiveness of this immunity to that strain's descendants within four to seven years..." -- Kucharski et al. PLOS Biol 2015 (although numbers might need revising)

on immune status driving the rate of antigenic evolution -- Wen et al. Proc B 2016

4. In "...we expect a larger fraction of a library to survive an antibody with many escape mutations..." (and similar phrases), the "many escape mutations" read as though they modify "antibody."

\end{document}

### 2018-01-16 paper/manuscript_NatureMicrobiology_submission.tex
created copy of *Nature Microbiology* submission

--- 2017-11-30 paper/manuscript.tex
@@
 
-Another important caveat that our experiments examine \emph{single} amino-acid mutations to the HA from one influenza virus strain.
+Another important caveat is that our experiments examine \emph{single} amino-acid mutations to the HA from one influenza virus strain.
 The protein evolution literature is full of examples of epistatic interactions that enable multiple mutations to access phenotypes not accessible by single mutations\cite{gong2013stability,harms2014historical,starr2017alternative}.
@@
 We thank Adam Dingens, Sarah Hilton, Katherine Xue, Lauren Gentles, and Jeremy Roop for helpful comments on the project and manuscript.
-We thank the Fred Hutchinson Cancer Research Center genomics core for performing the Illumina deep sequencing, and the protein expression core for expressing and purifying the C179 antibody.
+We thank the Fred Hutchinson Cancer Research Center genomics core for performing the Illumina deep sequencing, and the protein expression core for expressing and purifying the S139/1 antibody.
 This work was supported by grant R01AI127893 from the NIAID of the NIH.

### 2018-01-16 paper/NatureMicrobiologyReviews.tex
intro to reviewer response

@@
 Below we provide a detailed point-by-point response to the reviews.
-As you will see, we have been able to clarify and address all the questions / concerns.
+As you will see, we have addressed all the questions / concerns.
 
-In conclusion, our paper has been rigorously evaluated by four expert reviewers, who all found it to be scientifically sound, raising only minor points that we have clarified in our response / revision.
-The only question was whether our work was of sufficient impact for \textit{Nature Microbiology}, with the reviewers splitting about equally on this question.
+In conclusion, our paper has been evaluated by four expert reviewers, who all found it to be scientifically sound, raising only minor points that we have clarified in our response / revision.
+The only question was whether our work was of sufficient impact for \textit{Nature Microbiology}, with the reviewers splitting on this question.
 The \textit{Nature Communications} policies read: ``While \textit{Nature} and Nature research journals strive to publish the highest impact research within these disciplines, \textit{Nature Communications} is committed to publishing important advances of significance to specialists within each field.''
-Our study clearly meets these criteria, and so is suitable for \textit{Nature Communications}.
+Our study clearly meets these criteria from the evaluations of all reviewers, and so is suitable for \textit{Nature Communications}.
 
@@
 
+{\color{black}
+Thanks for selecting reviewers who cover the full spectrum of relevant expertise.}
+
 \subsubsection*{Reviewer \#1 (Remarks to the Author):}
@@
 There are some issues that the authors need to address but no major weaknesses were identified.
+
+{\color{black}
+We appreciate the fact that the reviewer find our work to be both technically sound and of high impact.}
 

### 2018-01-17 neutralization_assays/neutcurve.py
started code for neut curve fitting

@@ -1,0 +1,36 @@
+"""
+=============
+neutcurve
+=============
+Module for fitting and analyzing neutralization curves.
+"""
+
+
+def fourParamLogistic(c, ic50, slope, top, bottom):
+    """4-parameter logistic neutralization curve.
+
+    Evaluates :math:`f(c) = b + \\frac{t - b}{1 + (c/i)^s}`
+
+    In the argument definitions below, we interpret the
+    meaning of this curve in the case where :math:`s > 0`,
+    in which case :math:`f(c)` can be thought of as
+    the fraction surviving at concentration `c`.
+
+    Args:
+        `c` (float)
+            Concentration of antibody
+        `ic50` (float)
+            Inhibitory concentration 50%, :math:`i` in equation.
+        `slope` (float)
+            Hill slope of curve, :math:`s` in equation.
+        `bottom` (float)
+            Bottom of curve (value as :math:`c` gets large),
+            :math:`b` in equation.
+        `top` (float)
+            Top of curve (value as :math:`c` get small),
+            :math:`t` in equation.
+
+    Returns:
+        A float giving :math:`f(c)`, which is interpreted
+        as the fraction surviving at concentration `c`.
+    """

### 2018-01-19 neutralization_assays/neutcurve.py
partial work on neutcurve plotter

@@ -6,31 +6,134 @@ Module for fitting and analyzing neutralization curves.
 """
 
 
-def fourParamLogistic(c, ic50, slope, top, bottom):
+import functools
+import scipy
+import scipy.optimize
+
+
+class fourParamLogistic:
     """4-parameter logistic neutralization curve.
 
     Evaluates :math:`f(c) = b + \\frac{t - b}{1 + (c/i)^s}`
 
-    In the argument definitions below, we interpret the
-    meaning of this curve in the case where :math:`s > 0`,
-    in which case :math:`f(c)` can be thought of as
-    the fraction surviving at concentration `c`.
+    When :math:`s > 0`, then :math:`f(c)`, then :math:`f(c)`
+    is the fraction surviving at concentration :math:`c`.
 
-    Args:
-        `c` (float)
-            Concentration of antibody
+    Each object has the following attributed:
         `ic50` (float)
-            Inhibitory concentration 50%, :math:`i` in equation.
+            Inhibitory concentration 50%, :math:`i` in equation
+            above
         `slope` (float)
-            Hill slope of curve, :math:`s` in equation.
+            Hill slope of curve, :math:`s` in equation above.
         `bottom` (float)
             Bottom of curve (value as :math:`c` gets large),
-            :math:`b` in equation.
+            :math:`b` in equation above.
         `top` (float)
             Top of curve (value as :math:`c` get small),
-            :math:`t` in equation.
+            :math:`t` in equation above.
 
-    Returns:
-        A float giving :math:`f(c)`, which is interpreted
-        as the fraction surviving at concentration `c`.
+    See `__init__` method for how to initialize an object.
+
+    To use an object to evaluate neutralization values:
+
+    >>> top = 1
+    >>> bottom = 0
+    >>> slope = 5
+    >>> ic50 = 0.03
+    >>> neut = fourParamLogistic(ic50, slope, top, bottom)
+    >>> scipy.allclose(0.5, neut.fracsurvive(ic50))
+    True
+    >>> scipy.allclose(bottom, neut.fracsurvive(1e3))
+    True
+    >>> scipy.allclose(top, neut.fracsurvive(1e-3))
+    True
     """
+
+
+    def __init__(self, ic50=0, slope=1, top=1, bottom=0):
+        """Initialize a new object."""
+        self.ic50 = ic50
+        self.slope = slope
+        self.top = top
+        self.bottom = bottom
+
+
+    def fracsurvive(self, c):
[…]

### 2018-01-21 neutralization_assays/neutcurve.py
analyze replicate neutralization curves for FI6v3

@@ -1,139 +1,0 @@
-"""
-=============
-neutcurve
-=============
-Module for fitting and analyzing neutralization curves.
-"""
-
-
-import functools
-import scipy
-import scipy.optimize
-
-
-class fourParamLogistic:
-    """4-parameter logistic neutralization curve.
-
-    Evaluates :math:`f(c) = b + \\frac{t - b}{1 + (c/i)^s}`
-
-    When :math:`s > 0`, then :math:`f(c)`, then :math:`f(c)`
-    is the fraction surviving at concentration :math:`c`.
-
-    Each object has the following attributed:
-        `ic50` (float)
-            Inhibitory concentration 50%, :math:`i` in equation
-            above
-        `slope` (float)
-            Hill slope of curve, :math:`s` in equation above.
-        `bottom` (float)
-            Bottom of curve (value as :math:`c` gets large),
-            :math:`b` in equation above.
-        `top` (float)
-            Top of curve (value as :math:`c` get small),
-            :math:`t` in equation above.
-
-    See `__init__` method for how to initialize an object.
-
-    To use an object to evaluate neutralization values:
-
-    >>> top = 1
-    >>> bottom = 0
-    >>> slope = 5
-    >>> ic50 = 0.03
-    >>> neut = fourParamLogistic(ic50, slope, top, bottom)
-    >>> scipy.allclose(0.5, neut.fracsurvive(ic50))
-    True
-    >>> scipy.allclose(bottom, neut.fracsurvive(1e3))
-    True
-    >>> scipy.allclose(top, neut.fracsurvive(1e-3))
-    True
-    """
-
-
-    def __init__(self, ic50=0, slope=1, top=1, bottom=0):
-        """Initialize a new object."""
-        self.ic50 = ic50
-        self.slope = slope
-        self.top = top
-        self.bottom = bottom
-
-
-    def fracsurvive(self, c):
-        """Evaluates fraction surviving at concentration `c`."""
-        return self._fracsurvive_allargs(c, self.ic50,
-                self.slope, self.top, self.bottom)
-
-
-    def _fracsurvive_allargs(self, c, ic50, slope, top, bottom):
-        """Evaluates fraction surviving with called params.
-
-        Differs from `fracsurvive` in that it does **not** use
-        the current object values for parameters.
-        """
-        return bottom + (top - bottom) / (1 + (c / ic50)**slope)
-
-
-    def fitcurve(self, cs, fs, fixtop=False, fixbottom=False):
[…]

### 2018-01-21 neutralization_assays/platereaderdata/merge_FI6v3_CSVs.py

@@ -1,0 +1,12 @@
+import glob
+import pandas
+
+dfs = []
+for f in glob.glob('FI6v3_*.csv'):
+    df = pandas.read_csv(f)
+    sample = f.split('_', 1)[1].split('.')[0].replace(
+            '_HA2', '-HA2').replace('-8T', '(-8T)')
+    df = df.rename(columns=dict(
+            [(str(i), '{0}-{1}'.format(sample, i)) for i in [1, 2, 3]]))
+    dfs.append(df.set_index('concentration'))
+pandas.concat(dfs, axis=1).to_csv('FI6v3_neutcurves.csv')

### 2018-01-21 paper/NatureMicrobiologyReviews.tex
revision and response about neutralization curves: show replicates, fit IC50s, test significance

@@
 \usepackage[margin=1.2in]{geometry}
+
+\usepackage{hyperref}
+%\hypersetup{colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue}
 
@@
 
+{\color{red}
+Need to add response.}
+
 2) The virus used, WSN, is highly artificial and does not reflect actual human influenza viruses.
@@
 
+{\color{red}
+Need to add response.}
+
 3) Page 2, beginning of second last paragraph: It is unclear if such a vaccine would really enhance the pressure dramatically.
@@
 
+{\color{red}
+Need to add response.}
+
 4) Data deposition at github: How long will data/software stored on this platform be available? If versions are update, is this tracked/recorded?
+
+{\color{red}
+Need to add response.}
 
@@
 
+{\color{black}
+This is a good comment that was also made by several other reviewers.
+We therefore now do as suggested by showing full data for triplicate measurements with significance testing.
+
+Specifically, all the neutralization assays with FI6v3 were performed in triplicate on the same day to eliminate batch effects, with each replicate involving independent serial dilution of the antibody in a separate column of a 96-well plate (this is now more fully explained in the Methods).
+In the original manuscript, we only showed the mean and standard deviation in Figure~6A.
+We have now added a supplementary figure (Figure~S9) that shows all the replicates.
+In addition, as explained in the legend to the new Figure~S9, we fit IC50 values to each replicate, and then use the three IC50's to test the null hypothesis that each mutant has an IC50 indistinguishable from wildtype.
+
+The statistical testing results in Figure~S9B support the conclusions in the text. 
+The four mutations that are \emph{not} expected to differ from wildtype with respect to FI6v3 neutralization (K280A, V135T, P80D, and M17L(HA2)) do \emph{not} have IC50s that are significantly different than wildtype.
+The four mutations expected to have the largest increases in neutralization resistance (K280S, G47R(HA2), K280T, and N291S) are all significantly more resistant to FI6v3 than wildtype, with corrected $P$-values ranging from 0.00072 to 0.023.
+The mutation found by our mutational antigenic profiling but that elicited the most skepticism both from us and the reviewers is K(-8)T.
+However, the statistical testing suggests that it really is slightly but significantly more resistant to FI6v3, with a corrected $P$-value of 0.04. 
+But we agree that the effect is very small, and in the revised text we clearly emphasize this fact and advise caution in over-interpreting the result:
+
+\begin{quote}
+\textsl{The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.
+We tested one of these mutations, K(-8)T, and it did lead to a very slight increase in neutralization resistance (Figure~6A, Figure~S9)---although despite the significance testing in Figure~S9, we remain circumspect about the magnitude of this effect relative to the noise in our neutralization assays.
+}
+\end{quote}
+
+These revisions address this important comment made by several reviewers.
+}
+
 6) It would be important to quantify differences in mAb binding to the different escape mutants shown in Figure 6, e.g.
@@
 This needs to be explored.
+
+{\color{red}
+Need to add response.}
 
@@
 
+{\color{red}
+Need to add response.}
+
 8) It would be good to have a negative control experiment with a non-binding human or mouse mAb. This is missing but should actually be the 'real' negative control.
+
+{\color{red}
+Need to add response.}
 
@@
 The broad: FI6v3, binds stalk, neutralizes both group 1 and 2 HAs; C179 binds stalk, neutralizes only some group 1 HAs; S139/1, binds receptor-binding pocket, neutralizes both group 1 and group 2 HAs.
-The authors compare these in a re-analysis of data from narrow, strain-specific antibodies: H17-L19, H17-L10, and H17-L7, bind the Ca2, Ca1, and Cb antigenic regions on HA’s globular head.
+The authors compare these in a re-analysis of data from narrow, strain-specific antibodies: H17-L19, H17-L10, and H17-L7, bind the Ca2, Ca1, and Cb antigenic regions on HA's globular head.
 
@@
 
+{\color{red}
+Need to add response.}
+
 2.
@@
 
+{\color{red}
+Need to add response.}
+
 3. Are all changes in amino-acid obtained after one nucleotide change? Are the authors able to detect multiple nucleotide or amino acid changes?
+
+{\color{red}
+Need to add response.}
 
@@
 Authors could check a combination of the mutations they obtain with their experiments (especially the ones escaping the stalk-targeting antibodies).
+
+{\color{red}
+Need to add response.}
 
@@
 It could be that (1) all these differences are due to noise or that (2) there is actually a biological difference between WT and K280A.
-Whether (1) occurs is unclear: “three biological replicates” should perhaps be more detailed in the Materials and Methods section as to what this means: are they three rows of the same plate performed on the same day, on different days, etc.?
-According to the authors’ previous paper (Doug et al., 2017, PLOS Pathogens), they performed “three replicate dilution columns”.
-The neutralisation curves seem therefore to have been performed at different days (besides, the WT curves are the same or similar to that of Figure 3, and the neutralisation curves from the narrow antibodies were imported from the
-authors’ previous paper).
+Whether (1) occurs is unclear: ``three biological replicates'' should perhaps be more detailed in the Materials and Methods section as to what this means: are they three rows of the same plate performed on the same day, on different days, etc.?
+According to the authors' previous paper (Doud et al., 2017, PLOS Pathogens), they performed ``three replicate dilution columns''.
+The neutralisation curves seem therefore to have been performed at different days (besides, the WT curves are the same or similar to that of Figure 3, and the neutralisation curves from the narrow antibodies were imported from the authors' previous paper).
 This can induce a batch effect, which may not be linear, thus the normalisation performed (with no antibody and background noise) may not regress this out.
@@
 If this has been done, it was not clear from the figure legend, and could be added for clarity.
-To address (2), authors should provide details of Fr,a for K280A.
-If I understood correctly, because Fr,a-excess (defined in Equation 2) hides mutations with lower fitness than average, it is not possible for the reader to assess whether this mutation is “neutral” or it has lower fitness with the antibody under study.
+To address (2), authors should provide details of $F_{r,a}$ for K280A.
+If I understood correctly, because $F_{r,a}^{\rm{excess}}$ (defined in Equation 2) hides mutations with lower fitness than average, it is not possible for the reader to assess whether this mutation is ``neutral'' or it has lower fitness with the antibody under study.
+
+{\color{black}
+This is a good point, and was also made by several other reviewers.
+
+We have provided a detailed response in answer to question 5 by Reviewer \#1.
+Our revisions include showing the neutralization curves for all replicates and performing statistical testing.
+All the neutralization curves for antibody FI6v3 were performed on the same day.
+This is important because as the reviewer notes, there can be batch effects.
+It was not feasible to perform all assays for all mutants on the same day, but all assays that are being compared are on the same day.
+So for instance, if you look at Figure~6, you will see that the wildtype curves in panels A, B, and C are slightly different since each was performed a different day.
+But each panel is for a single antibody, and all curves with that antibody were performed on the same day---that is, we repeated the wildtype in each batch of assays to avoid any confounding batch effects.
+Therefore, we have already done what the reviewer suggests, we just hadn't clearly shown all the data before. 
+That is remedied by the new Figure~S9.
+
+As described in the response to question 5 by Reviewer \#1, the new statistical analysis finds that the effect for K(-8)T is significant, whereas that for K280A is not.
+We agree that the effect for K(-8)T is somewhat perplexing, and we wish that it would go away since it is not central to any of our claims, and has elicited skepticism from both us and the reviewers.
+In the original manuscript, we conveyed this skepticism by saying, \textsl{``we are not confident that this effect exceeds the noise in our neutralization assays.''}.
+However, the new statistical analysis described in response to question 5 by Reviewer \#1 finds the effect is significant, but we continue to add strong caveats (see the quoted new text in the response to question 5 by Reviewer \#1).
+So rest assured that we are not trying to exaggerate the effect of K(-8)T; we sincerely wish it would go away as it would simplify the reviewer responses!
+However, this mutation was picked up by our mutational antigenic profiling (which was performed in triplicate), and then again by a totally independent approach (neutralization assays, which were again performed in triplicate).
+So in good conscience, we feel that we have to present the result with K(-8)T, however perplexing.
+
+The reviewer also points out that it might be interesting to look at the $F_{r,a}$ value for K280A, since it might have reduced neutralization resistance.
+First, we note that the statistical analysis in the new Figure~S9 shows that the difference between K280A and wildtype is \emph{not} statistically significant, so we do not comment on possible increased neutralization sensitivity or the $F_{r,a}$ for this mutation in the paper.
+However, we have added a new supplementary file (File~S6) that gives all the $F_{r,a}$ values in case anyone wants to examine them.
+The value of $F_{r,a}$ for this mutation is 0.00382, which is slightly below the library average of 0.00955.
+}
 
 6. Could the authors detail, in the legend, how the curves of Figure 6 were fitted?
+
+{\color{black}
+For the quantitative analyses of IC50 values, we fit four-parameter logistic curves, although we fixed the top value to one, so there are really only three parameters.
+This is now explained in the legend to the new Figure~S9, which reports the actual IC50 values and shows the curves for individual replicates.
+The exact computer code used to fit the curves is at \url{https://jbloomlab.github.io/dms_tools2/dms_tools2.neutcurve.html}.
+}
 
@@
 In contrast, only weak escape mutations were identified against the two stem antibodies.
-The authors conclude that “breadth is an imperfect indicator of
-the potential for viral escape”.
-The authors also conclude that broadly neutralizing antibodies targeting the H1 hemagglutinin stalk are “quantifiably harder to escape than the other antibodies tested here”.
+The authors conclude that ``breadth is an imperfect indicator of the potential for viral escape''.
+The authors also conclude that broadly neutralizing antibodies targeting the H1 hemagglutinin stalk are ``quantifiably harder to escape than the other antibodies tested here''.
 The last two conclusions are not unexpected and are not particularly novel insights.
@@
 
+{\color{red}
+Need to add response.}
+
 Specific comments:
@@
 
+{\color{red}
+Need to add response.}
+
 2.  How is the curve fitting performed for the neutralization assay results? I do not seem to see any description for the curve-fitting method.
+
+{
+\color{black}
+This excellent question was also asked by other reviewers, and we apologize for not providing more details in the original submission.
+We have made extensive revisions to better describe the neutralization curves, show the curves for all individual replicates, calculate IC50s and the statistical significance of their differences from wildtype, and describe the curve fitting procedures.
+These revisions are detailed in response to question 5 of Reviewer \#1 and questions 5 and 6 of Reviewer \#2.}
 
@@
 Obviously such data might be complementary.
+
+{\color{red}
+Need to add response.}
 
@@
 
+{\color{red}
+Need to add response.}
+
 4. In "...we expect a larger fraction of a library to survive an antibody with many escape mutations..." (and similar phrases), the "many escape mutations" read as though they modify "antibody."
 
+{\color{red}
+Need to add response.}
+
 \end{document}

### 2018-01-21 paper/manuscript.tex
Fixed typo in Discussion
revision and response about neutralization curves: show replicates, fit IC50s, test significance

@@
 \usepackage{placeins}
+\usepackage{booktabs}
 
@@
 The excess fraction of virions with a mutation at each site that survive the antibody, averaging across all amino-acid mutations at each site (see Equation~\ref{eq:avgfracsurvive}).
-There are multiple sites of large-effect mutations for H17L19, H17L10, H17L7, and S139/1 -- but none for FI6v3 and C179.
+There are multiple sites of large-effect mutations for H17L19, H17L10, H17L7, and S139/1---but none for FI6v3 and C179.
 Figure~\ref{suppfig:maxfracsurvive} shows the excess fraction surviving for the largest-effect mutation at each site.
@@
 Only if we zoom in can we see that there are actually a few sites where mutations slightly increase the fraction of virions surviving C179 and FI6v3 (Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo}).
-But the effect sizes of these antigenic mutations are tiny compared to the other antibodies -- especially for FI6v3.
+But the effect sizes of these antigenic mutations are tiny compared to the other antibodies---especially for FI6v3.
 Therefore, the HA of A/WSN/1933 influenza virus is far less capable of escaping these anti-stalk antibodies by single mutations than it is of escaping the other four antibodies.
@@
 We have previously validated many of the large-effect antigenic mutations selected by the narrow antibodies H17-L19, H17-L10, and H17-L7\cite{doud2017complete}.
-However, the mutations selected by the broad anti-stalk antibodies have much smaller effects in our mutational antigenic profiling -- especially for the broadest antibody, FI6v3.
+However, the mutations selected by the broad anti-stalk antibodies have much smaller effects in our mutational antigenic profiling---especially for the broadest antibody, FI6v3.
 We therefore tested some of these FI6v3-selected mutations using neutralization assays on individual viral mutants.
@@
 (A) Neutralization curves of individual viral mutants with FI6v3.
-The mutations K280S, K280T, N291S, G47R (HA2), and K(-8)T are all expected to increase neutralization resistance based on the mutational antigenic profiling (Figure~\ref{fig:structures}A). 
-The mutations K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization based on the mutational antigenic profiling (Figure~\ref{suppfig:FI6v3logo}).
+The mutations K280S, K280T, N291S, G47R (HA2), and K(-8)T are all expected to increase neutralization resistance based on the mutational antigenic profiling (Figure~\ref{fig:structures}A), whereas K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization (Figure~\ref{suppfig:FI6v3logo}).
+All neutralization curves in this panel were performed in triplicate on the same day.
+This panel shows the average of the replicates; Figure~\ref{suppfig:FI6v3replicates} shows the curves for each replicate individually and performs statistical testing of whether the IC50s for mutants are significantly different than for wildtype.
 (B), (C) In contrast to FI6v3, mutations selected by narrow antibodies have very large effects on neutralization.
@@
 As discussed in the previous section, the three mutations most strongly selected by FI6v3 introduce glycosylation motifs at sites 278-280 or 289-291 (Figure~\ref{fig:structures}A,B).
-We created viruses carrying each of these mutations (K280S, K280T, and N291S) and validated that all three modestly increased resistance to FI6v3 (Figure~\ref{fig:FI6v3neutcurves}A).
-As a control, we also validated that a mutation at one of these sites (K280A) that does \emph{not} have an effect in our mutational antigenic profiling does not shift the neutralization curve (Figure~\ref{fig:FI6v3neutcurves}A).
+We created viruses carrying each of these mutations (K280S, K280T, and N291S) and validated that all three modestly but significantly increased resistance to FI6v3 (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates}).
+As a control, we also validated that a mutation at one of these sites (K280A) that does \emph{not} have an effect in our mutational antigenic profiling does not significantly shift the neutralization curve (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates}).
 
 Our mutational antigenic profiling also identified several non-glycosylation-motif mutations that were selected by FI6v3.
-We validated that one of these mutations, G47R in the HA2 chain, increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A) -- although as predicted by the mutational antigenic profiling, the magnitude of the effect was small.
+We validated that one of these mutations, G47R in the HA2 chain, significantly increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates})---although as predicted by the mutational antigenic profiling, the magnitude of the effect was small.
 The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.
-We tested one of these mutations, K(-8)T, and it did appear to lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A), although we are not confident that this effect exceeds the noise in our neutralization assays.
-As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them affected neutralization resistance within the sensitivity of our assays.
+We tested one of these mutations, K(-8)T, and it did lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates})---although despite the significance testing in Figure~\ref{suppfig:FI6v3replicates}, we remain circumspect about the magnitude of this effect relative to the noise in our neutralization assays.
+As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them significantly affected neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates}).
 
@@
 In most cases the mutations selected by the antibodies occur at these mutationally tolerant sites.
-However, there are exceptions -- for instance, the H38S mutation selected by C179 is rather disfavored with respect to viral growth, but has a large enough antigenic effect to still be detected in our mutational antigenic profiling.
+However, there are exceptions---for instance, the H38S mutation selected by C179 is rather disfavored with respect to viral growth, but has a large enough antigenic effect to still be detected in our mutational antigenic profiling.
 
@@
 A major rationale for studying broadly neutralizing antibodies is that they are hoped to be more resistant to viral evolutionary escape than the antibodies that dominate natural immune responses to influenza virus\cite{krammer2015advances,corti2017tackling}.
-We have used a new approach to quantify the extent to which this is actually true, and shown that neutralization of an H1 virus by broad anti-stalk antibodies is indeed more -- although certainly not completely -- resistant to erosion by viral point mutations.
+We have used a new approach to quantify the extent to which this is actually true, and shown that neutralization of an H1 virus by broad anti-stalk antibodies is indeed more---although certainly not completely---resistant to erosion by viral point mutations.
 Going forward, we suggest that completely mapping viral escape mutations will be a useful complement to more traditional techniques that simply characterize the breadth of anti-viral antibodies against circulating strains.
@@
 The neutralization assays themselves were performed by using a plate reader to quantify the GFP signal produced by MDCK-SIAT1-CMV-PB1 cells infected by PB1flank-eGFP virus that had been incubated with the indicated antibody concentration as described previously\cite{hooper2013mutant}.
-All neutralization curves represent the mean and standard deviation of three measurements. 
+All neutralization curves in Figure~\ref{fig:FI6v3neutcurves}A represent the mean and standard deviation of three measurements, with the individual replicates shown in Figure~\ref{suppfig:FI6v3replicates}.
+All the neutralization assays for FI6v3 were performed on the same day to eliminate batch effects, with each replicate involving independent serial dilution of the antibody in a separate column of a 96-well plate.
+
 \subsection*{H3 sequence numbering}
@@
 Among broad antibodies, S139/1 has not been tested against H8 and H11; C179 has not been tested against H8 and H11; and no antibodies have been tested against H17 and H18.
-The narrow H17-L19, H17-L10, and H17-L7 antibodies have not been tested against any other subtypes -- however, since these antibodies have a very limited range even among H1 HAs\cite{caton1982antigenic}, we assume that they do not bind other subtypes.
+The narrow H17-L19, H17-L10, and H17-L7 antibodies have not been tested against any other subtypes---however, since these antibodies have a very limited range even among H1 HAs\cite{caton1982antigenic}, we assume that they do not bind other subtypes.
 
@@
 \subsection*{Analysis of deep sequencing data}
-The deep sequencing data were analyzed using version 2.1.0 of the \texttt{dms\_tools2} software package\cite{bloom2015software}, which is available at \url{http://jbloomlab.github.io/dms_tools2}.
+The deep sequencing data were analyzed using version 2.2.1 of the \texttt{dms\_tools2} software package\cite{bloom2015software}, which is available at \url{http://jbloomlab.github.io/dms_tools2}.
 File~\ref{suppfile:analysis_code} contains a Jupyter notebook that performs all steps of the analysis beginning with downloading the FASTQ files from the Sequence Read Archive.
@@
 
-In this paper, Figures~\ref{fig:avgfracsurvive} and \ref{suppfig:maxfracsurvive} show the median of excess fraction surviving taken across all biological and technical replicates at a given antibody concentration.
+In this paper, Figures~\ref{fig:avgfracsurvive} and \ref{suppfig:maxfracsurvive} show the median of excess fraction surviving taken across all biological and technical replicates at a given antibody concentration (Equation~\ref{eq:fracsurvive_excess}).
 The subsequent logo plots show the medians of these values taken across all concentrations for each antibody.
-The numerical values plotted in these logo plots are in File~\ref{suppfile:fracsurvive}.
+The numerical values plotted in these logo plots are in File~\ref{suppfile:fracsurvive_excess}.
+The fraction surviving values \emph{not} adjusted to be in excess of the library average (Equation~\ref{eq:fracsurvive}) are in File~\ref{suppfile:fracsurvive}.
 
@@
 
+\begin{suppfigure}
+{\bf \LARGE A} \\
+\centerline{\includegraphics[width=\textwidth]{figs/FI6v3_replicate_neutcurves.pdf}} \\
+\vspace{0.1in}
+{\bf \LARGE B}
+\begin{center}
+\input{figs/FI6v3_replicate_IC50s}
+\end{center}
+\caption{\label{suppfig:FI6v3replicates}
+{\bf Replicates of the FI6v3 neutralization curves in Figure~\ref{fig:FI6v3neutcurves}A.}
+The neutralization assays were performed in triplicate for all nine mutants and wildtype.
+Figure~\ref{fig:FI6v3neutcurves}A shows the \emph{average} of those replicates.
+(A) The neutralization data for each replicate shown individually, with IC50 values fit using a four-parameter logistic curve with the top value constrained to one (see \url{https://jbloomlab.github.io/dms_tools2/dms_tools2.neutcurve.html} for the code used for the fitting.)
+(B) Table of the IC50 values for each replicate.
+We used an unpaired Student's t-test with unequal variances to test the null hypothesis that each mutant had an IC50 indistinguishable from wildtype.
+We then used Bonferroni's method to correct the $P$-values for multiple testing, and report these corrected values.
+}
+\end{suppfigure}
+
 \begin{supptable}
@@
 \begin{suppfile}
-\caption{\label{suppfile:fracsurvive}
+\caption{\label{suppfile:fracsurvive_excess}
 {\bf The excess fraction surviving for each mutation for each antibody.}
@@
 These are median excess fraction surviving taken first across replicates and then across antibody concentrations.
+See Equation~\ref{eq:fracsurvive_excess}.
+}
+\end{suppfile}
+
+\begin{suppfile}
+\caption{\label{suppfile:fracsurvive}
+{\bf The fraction surviving for each mutation for each antibody.}
+This file differs from File~\ref{suppfile:fracsurvive_excess} only in that the values are \emph{not} adjusted to be in excess of the library average (e.g., they are from Equation~\ref{eq:fracsurvive} rather than Equation~\ref{eq:fracsurvive_excess}).
 }

### 2018-01-21 paper/manuscript_NatureMicrobiology_submission.tex
revision and response about neutralization curves: show replicates, fit IC50s, test significance

@@
 Our mutational antigenic profiling also identified several non-glycosylation-motif mutations that were selected by FI6v3.
-We validated that one of these mutations, G47R in the HA2 chain, increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A) -- although as predicted by the mutational antigenic profiling, the magnitude of the effect was small.
+We validated that one of these mutations, G47R in the HA2 chain, increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A)---although as predicted by the mutational antigenic profiling, the magnitude of the effect was small.
 The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.

### 2018-01-21 paper/NatureMicrobiologyReviews.tex
Responded to some Reviewer 1 comments: no multi-cycle growth, WSN is lab adapted, GitHub is reliable, no irrelevant antibody control

@@
 
-{\color{red}
-Need to add response.}
+{\color{black}
+This is true.
+Our approach systematically quantifies how all \emph{single} mutations affect viral neutralization.
+This contrasts with an alternative approach of using multiple rounds of growth to select for multi-mutants.
+There are strengths and weaknesses to both approaches.
+Examining all single mutations makes the results complete and quantitative---but we can't make claims about the effects of multiple mutations.
+In contrast, multi-cycle growth sometimes finds combinations of mutations, but is not complete and quantitative, since there are too many multiple mutants to examine them all, and it is random which mutations arise in any given multi-cycle growth experiment.
+In particular, multi-cycle growth assays do not generate data that can be quantitatively compared across antibodies using the methods here.
+Therefore, the strength of our approach (examining all single mutations in a way that is comparable across antibodies) comes with the limitation that we do not find multiple mutants.
+
+We have taken the following pains to emphasize this limitation. 
+First, the title of the paper reads ``Quantifying the effects of \emph{single mutations} on viral escape from broad and narrow antibodies to an H1 influenza hemagglutinin'' (emphasis added). 
+Stating this fact in the title makes the limitation immediately and directly clear.
+
+In addition, we have included the following paragraph in the Discussion:
+
+\begin{quote}
+\textsl{Another important caveat is that our experiments examine {\bf single} amino-acid mutations to the HA from one influenza virus strain.
+The protein evolution literature is full of examples of epistatic interactions that enable multiple mutations to access phenotypes not accessible by single mutations.
+Such epistasis is relevant to HA's evolution.  
+For instance, work by Das et al suggests that the sequential accumulation of mutations can shift the spectrum of available antibody-escape mutations.
+Wu et al have used deep mutational scanning to directly demonstrate that rampant epistasis enables HA's receptor-binding pocket to accommodate combinations of individually deleterious mutations, some of which affect sensitivity to antibodies.
+Therefore, our work does not imply any absolute limits on the possibilities for antibody escape when evolution is given sufficient time to explore combinations of mutations.
+However, single mutations are the most accessible form of genetic variation, and much of influenza virus's natural antigenic drift involves individual mutations that reduce sensitivity to immunodominant antibody specificities.
+Quantifying the antigenic effects of all such mutations therefore provides a relevant measure of ease of viral antibody escape. }
+\end{quote}
+}
 
@@
 
-{\color{red}
-Need to add response.}
+{\color{black}
+The reviewer is correct that A/WSN/1933 is a lab-adapted virus.
+As the reviewer notes, in his/her comment, this is not necessarily a problem for our work since we are simply examining antibody neutralization rather than pathogenesis, fitness, or transmission. 
+Nonetheless, the reviewer is correct that it is important to clearly mention this caveat. 
+We have now added the following text when we first discuss this strain:
+
+\begin{quote}
+\textsl{We performed our experiments using the lab-adapted A/WSN/1933 (H1N1) strain of influenza.
+This strain is derived from an early seasonal H1N1 that was extensively passaged in the lab, where it adapted to become neurotropic and trypsin independent.
+But despite these unusual properties, the virus is neutralized by most broad antibodies that target other H1 viruses, including those used in this study (Figure~3).}
+\end{quote}
+}
 
@@
 
-4) Data deposition at github: How long will data/software stored on this platform be available? If versions are update, is this tracked/recorded?
+4) Data deposition at GitHub: How long will data/software stored on this platform be available? If versions are updated, is this tracked/recorded?
 
-{\color{red}
-Need to add response.}
+{\color{black}
+GitHub is the leading platform for archiving and tracking computer code.
+It is used widely in computational biology, and also by Google and Microsoft, among others. For instance, see: \url{https://www.wired.com/2015/03/github-conquered-google-microsoft-everyone-else/}
+
+GitHub uses version tracking to record all updates to computer code, and make it possible to access any prior version of the software.
+For instance, while \url{https://github.com/jbloomlab/dms_tools2} provides the latest version of the code, the version used in this study (version 2.2.1) will always be available \url{https://github.com/jbloomlab/dms_tools2/tree/2.2.1}, and it is easy to examine modifications made to this version in any new updates.
+
+While it is impossible to guarantee that GitHub will exist forever, the same could be said for \textit{Nature Press} (which could go out of business) or PubMed (which could be de-funded by the NIH or the Trump administration).
+The best that we can say is that GitHub is considered the most stable and reliable way to archive and track software, as evidenced by its use not only in computational biology but also by major corporations such as Google and Microsoft.j
+
+In addition, all the Jupyter notebooks and major results are provided in supplementary files, which will be available as long as \textit{Nature Press} continues to maintain supporting information.
+}
 
@@
 
-8) It would be good to have a negative control experiment with a non-binding human or mouse mAb. This is missing but should actually be the 'real' negative control.
+8) It would be good to have a negative control experiment with a non-binding human or mouse mAb. This is missing but should actually be the ``real'' negative control.
 
-{\color{red}
-Need to add response.}
+{\color{black}
+We see the rationale behind this suggestion, but do we not believe it is essential for the current work. 
+Our claim is that we specifically identify mutations that increase resistance to each of the individual monoclonal antibodies.
+With the three broad antibodies tested in this paper (FI6v3, C179, S139/1) and the four narrow antibodies tested by Doud et al (2016), we have now examined seven different antibodies.
+The only overlap between antibodies in the identified resistance mutations comes when those antibodies target similar epitopes (e.g., FI6v3 and C179)---there is no overlap in the mutations identified by antibodies targeting different epitopes.
+This fact indicates that we are identifying bona fide resistance mutations to the specific antibodies used, rather than mutations with some general effect on all IgGs.
+
+Figure~6 provides further data indicating that these are genuine antibody-specific resistance mutations rather than mutations generally selected by any IgG.
+For instance, the mutations that increase resistance to H17-L19 (V135T) or H17-L7 (P80D) do not affect neutralization by FI6v3, demonstrating that they are specific to the antibodies that selected them.
+Finally, we note that there are numerous published papers where the same monoclonal antibody is tested against a wide range of viral variants for the purposes of comparing IC50s across viruses without also including data showing that all viral variants are unaffected by an irrelevant antibody (e.g., Table~1 of \textit{Nature}, 489:526-532; Figure~3 of \textit{Nature Microbiology},; 2:16199).
+}
 

### 2018-01-21 paper/manuscript.tex
Responded to some Reviewer 1 comments: no multi-cycle growth, WSN is lab adapted, GitHub is reliable, no irrelevant antibody control

@@
 
-We performed our experiments using the A/WSN/1933 (H1N1) strain of influenza, which is neutralized by all six antibodies (Figure~\ref{fig:neutcurves}).
+We performed our experiments using the lab-adapted A/WSN/1933 (H1N1) strain of influenza.
+This strain is derived from an early seasonal H1N1 that was extensively passaged in the lab, where it adapted to become neurotropic and trypsin independent~\cite{sun2010modifications}.
+But despite these unusual properties, the virus is neutralized by most broad antibodies that target other H1 viruses, including those used in this study (Figure~\ref{fig:neutcurves}).
 Our experiments utilize fully infectious influenza virus rather than pseudovirus, which is important since the accessibility of some epitopes can vary with HA density, which differs between fully infectious virus and pseudovirus\cite{corti2011neutralizing,joyce2016vaccine}.

### 2018-01-21 paper/NatureMicrobiologyReviews.tex
completed responses to Reviewer 1

@@
 
-{\color{red}
-Need to add response.}
+{\color{black}
+This is a good point. We have changed ``will be under strong antigenic selection'' to ``could come under stronger antigenic selection.''
+}
 
@@
 While it is impossible to guarantee that GitHub will exist forever, the same could be said for \textit{Nature Press} (which could go out of business) or PubMed (which could be de-funded by the NIH or the Trump administration).
-The best that we can say is that GitHub is considered the most stable and reliable way to archive and track software, as evidenced by its use not only in computational biology but also by major corporations such as Google and Microsoft.j
+The best that we can say is that GitHub is considered the most stable and reliable way to archive and track software, as evidenced by its use not only in computational biology but also by major corporations such as Google and Microsoft.
 
-In addition, all the Jupyter notebooks and major results are provided in supplementary files, which will be available as long as \textit{Nature Press} continues to maintain supporting information.
+In addition, all the Jupyter notebooks and major results are provided in supplementary files, which will be available on the \textit{Nature Press} website.
 }
@@
 
-{\color{red}
-Need to add response.}
+{\color{black}
+The reviewer is correct that we do not include biophysical measurements of why the mutants have increased neutralization resistance.
+However, biophysically characterizeing antibody binding to the mutants is beyond the scope of the current study.
+We are careful to avoid making any claims about the mechanisms of the mutations---we simply say that they reduce neutralization.
+This is in line with many other studies in the field.
+While biophysically oriented studies certainly include such measurements, there are numerous papers on viral evolution and antibody breadth that characterize neutralization as the phenotype of interest.
+Our study falls into this latter class.
+}
 
@@
 But there are also exceptions to this rule.
-Two papers have so far described escape from stalk-based mAbs in detail (https://www.ncbi.nlm.nih.gov/pubmed/27351973 and https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4362269/).
-It is strongly suggested that the authors consult this papers and include their findings in the discussion.
+Two papers have so far described escape from stalk-based mAbs in detail (\url{https://www.ncbi.nlm.nih.gov/pubmed/27351973} and \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4362269/}).
+It is strongly suggested that the authors consult these papers and include their findings in the discussion.
 
-{\color{red}
-Need to add response.}
+{\color{black}
+Thank you for pointing out these two important references, which we now discuss.
+We also think it is interesting that both of these studies selected their escape mutants in group 2 HAs rather than a group 1 HA like the one in our study.
+Anecdotally, it seems to us that there are more described broad anti-stalk antibodies against group 1 than group 2 HAs.
+We obviously have no way to quantify this rigorously, but in the added text that mentions the two studies suggested by the reviewer, we also point out the difference in HA groups:
+
+\begin{quote}
+\textsl{Interestingly, most previous studies (Chai, 2016; Dunand, 2016) that have reported selecting single mutations with large effects ($\gg$10-fold) on neutralization by anti-stalk antibodies have used group 2 (e.g., H3 or H7) HAs rather than group 1 HAs like the one used in our work.}
+\end{quote}
+}
 

### 2018-01-21 paper/manuscript.tex
completed responses to Reviewer 1

@@
 
-If these efforts succeed, the epitopes of broad antibodies will be under strong antigenic selection in human influenza virus.
+If these efforts succeed, the epitopes of broad antibodies could come under stronger antigenic selection in human influenza virus.
 Might such selection then drive antigenic variation in these epitopes?
@@
 \section*{DISCUSSION}
-We have quantified how all single amino-acid mutations to influenza virus HA affect neutralization by a collection of broad and narrow antibodies.
+We have quantified how all single amino-acid mutations to an H1 influenza virus HA affect neutralization by a collection of broad and narrow antibodies.
 Our results show that the virus's inherent evolutionary capacity for escape via point mutations differs across antibodies. 
@@
 The classic approach for selecting escape mutations involves treating a virus stock with antibody at a concentration that completely neutralizes wildtype, and looking for viral mutants that survive this treatment\cite{yewdell1979antigenic,webster1980determination}.
-There are no such single mutations for the HA and broad anti-stalk antibodies tested here, since no mutations shift the neutralization curve enough to enable survival at antibody concentrations that fully neutralize wildtype.
+There are no such single mutations for the H1 HA and broad anti-stalk antibodies tested here, since no mutations shift the neutralization curve enough to enable survival at antibody concentrations that fully neutralize wildtype.
 However, our approach shows that there are mutations that have more modest ($<$10-fold) effects on neutralization by even the broadest antibody. 
+Interestingly, most previous studies~\cite{chai2016two,dunand2015preexisting} that have reported selecting single mutations with large effects ($\gg$10-fold) on neutralization by anti-stalk antibodies have used group 2 (e.g., H3 or H7) HAs rather than group 1 HAs like the one used in our work.
 When interpreting the magnitude of these effects, it is important to note that our experiments only measure how mutations affect neutralization, and not how they affect Fc-mediated functions that are responsible for much of the \textit{in vivo} protection afforded by anti-stalk antibodies\cite{dilillo2014broadly,dilillo2016broadly}.

### 2018-01-21 paper/NatureMicrobiologyReviews.tex
finished replying to Reviewer 2 comments

@@
 
+{\color{black}
+All of the results for the broadly neutralizing antibodies are new to this study.
+The reviewer is correct that the data for the narrow antibodies in Figure 6B,C is from Doud et al (2017), and the mutational tolerance data in Figure 7 is from Doud and Bloom (2016).
+This fact is clearly stated; for instance, the figure legends say ``Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 taken from Doud et al (2017)'' and ``The logo plots at bottom show the preference of each HA site of each amino acid under selection for viral replication as measured by Doud and Bloom (2016).''
+There is nothing wrong with including previous data for comparative purposes.
+For instance, new reports of protein crystal structures often compare the results to prior structures, and often include these prior structures in figures for points of comparison.
+In fact, a major strength of our approach is that it puts how mutations affect neutralization into units that can be meaningfully compared across studies.
+
+The experimental methodology is the same as in Doud et al (2017), although the entire analysis framework (all Equations in the paper, not just Equations 10 and 11) are new.
+The strength of this analysis framework is that it enables the results to be quantitatively compared across antibodies, which was not possible with the approach in Doud et al (2017). 
+}
+
 In this work, they compare three broad and three narrow antibodies.
@@
 
+{\color{black}
+The foregoing paragraphs are an accurate summary.
+We think it is valuable to know that the mutational antigenic profiling also works well for small-effect mutations.
+}
+
 They compare the sites where mutations arise from broad antibody, to previous mutational scanning data that reveals whether these regions are tolerant or not to mutation.
@@
 
-{\color{red}
-Need to add response.}
+{\color{black}
+We certainly agree that it would be interesting to examine additional strains.
+However, prior to our work, no study had ever completely mapped the effect of all single mutations to any viral strain with respect to neutralization by any broadly neutralizing influenza antibody, so we believe that just doing this with one strain is a sufficient step forward in knowledge to be valuable.
+As mentioned in the response to question 2 by Reviewer \#1, we now more clearly explain that WSN is a lab-adapted strain.
+}
 
@@
 
-{\color{red}
-Need to add response.}
+{\color{black}
+The neutralization assays in Figure~6 show that the mutations for the stalk-targeting antibodies only have weak effects on neutralization, while Figure~7 shows that many of them don't have much effect on viral growth.
+In addition, we were able to generate all of the stalk-antibody mutations for the neutralization assays, so they are not too badly attenuated for viral fitness in cell culture.
+However, we avoid making strong statements about the effects of mutations on viral growth beyond the data in Figure~7 because it is beyond the scope of our study to do growth curves for such a large number of mutations.}
 
@@
 
-{\color{red}
-Need to add response.}
+{\color{black}
+We examined all amino-acid changes, including those that involved multiple-nucleotide changes.
+The reason is that our libraries contained all codon mutations (e.g., the 63 non-wildtype codons at each site) rather than just all single nucleotide mutations.
+We have added text that clarifies this point:
+\begin{quote}
+\textsl{Importantly, as described previously (Doud and Bloom 2016), these virus libraries were generated by mutagenizing HA at the \emph{codon} level rather than at the nucleotide level.
+Performing codon mutagenesis is important, because single-nucleotide mutations access only about a third of the possible amino-acid mutations from a given codon, whereas codon mutations access all possible amino-acid mutations.}
+\end{quote}
+}
 
@@
 
-{\color{red}
-Need to add response.}
+{\color{black}
+We agree that this is an important caveat, and it was also mentioned by other reviewers.
+See our response to question 1 by Reviewer \#1 for how we have addressed this caveat.}
 
@@
 
-{\color{red}
-Need to add response.}
+{\color{black}
+We are disappointed that the reviewer does not view the work of sufficient impact and novelty for \textit{Nature Microbiology}, but we are appreciative of the statement that the study is ``performed to an extremely high technical standard.''
+Personally, we feel that direct quantitative comparison of the escape from narrow and broad antibodies is valuable, since other comparisons have been more anecdotal.
+In any case, we are now re-submitting to \textit{Nature Communications}, which places less emphasis on impact.
+We also note that at least two of the four reviewers found the work to be of high impact---and even Reviewer \#3 praises it on a technical level.
+Therefore, we think it is appropriate for \textit{Nature Communications}.}
 

### 2018-01-21 paper/manuscript.tex
finished replying to Reviewer 2 comments

@@
 C179 is less broad, and neutralizes only some group 1 HAs (Figure~\ref{fig:antibody_summary}B).
-We also selected a broad antibody, S139/1, that crystallographic studies have shown binds to residues in HA's receptor-binding pocket\cite{lee2012heterosubtypic}, and which can neutralize both group 1 and group 2 HAs\cite{yoshida2009cross, lee2012heterosubtypic}.
+We also chose a broad antibody, S139/1, that crystallographic studies have shown binds to residues in HA's receptor-binding pocket\cite{lee2012heterosubtypic}, and which can neutralize both group 1 and group 2 HAs\cite{yoshida2009cross, lee2012heterosubtypic}.
 Finally, we re-analyzed deep sequencing data from prior mutational antigenic profiling of three narrow strain-specific antibodies, H17-L19, H17-L10, and H17-L7\cite{doud2017complete}.
@@
 All experiments were performed in full biological triplicate using three independently generated virus libraries carrying single amino-acid mutations to HA\cite{doud2016accurate}.
-The correlations among replicates are shown in Figure~\ref{suppfig:corr}.
+Importantly, as described previously\cite{doud2016accurate}, these virus libraries were generated by mutagenizing HA at the \emph{codon} level rather than at the nucleotide level.
+Performing codon mutagenesis is important, because single-nucleotide mutations access only about a third of the possible amino-acid mutations from a given codon, whereas codon mutations access all possible amino-acid mutations.
+
+The correlations among replicates of the mutational antigenic profiling in terms of the measured fraction-surviving above average for each possible amino-acid mutations are shown in Figure~\ref{suppfig:corr}.
 For the remainder of this paper, we will refer to the median antigenic effect of each mutation across replicates.

### 2018-01-22 paper/NatureMicrobiologyReviews.tex
finished draft of reviewer responses

@@
 
-\usepackage[margin=1.2in]{geometry}
+\usepackage[margin=1.1in]{geometry}
 
@@
 All four reviewers found the work to be scientifically sound.
-Two reviewers (\#1 and \#4) were also highly enthusiastic about the impact.
-The other two reviewers (\#2 and \#3) concurred about scientific soundness, but were less enthusiastic about impact.
-The most critical was Reviewer \#3, who still wrote that ``[t]he study is performed to an extremely high technical standard'' --- a positive comment to receive from the most negative reviewer!
+Two reviewers (\#1 and \#4) were also enthusiastic about impact.
+The other two (\#2 and \#3) concurred about scientific soundness, but were less enthusiastic about impact.
+The most critical was Reviewer \#3, who still wrote that ``[t]he study is performed to an extremely high technical standard''---a positive comment to receive from the most negative reviewer!
 
-Given that all reviewers found the work to be scientifically sound and at least half found it to be high impact, we request that the paper be considered for \textit{Nature Communications} with the current reviews.
+Because all reviewers found the work to be scientifically sound and at least half found it to be high impact, we request that the paper be considered for \textit{Nature Communications} with the current reviews.
 Impact is subjective, and the fact that half the reviewers found the paper of sufficient impact for \textit{Nature Microbiology} indicates that it will interest many in the field. 
-In addition, we emphasize the following impactful aspects of our study:
+We emphasize the following impactful aspects of our study:
 \begin{itemize}\parskip1pt
@@
 
-Below we provide a detailed point-by-point response to the reviews.
-As you will see, we have addressed all the questions / concerns.
+Below we provide a detailed response to the reviews.
+We have addressed all the questions / concerns.
+The only major request was that we show all triplicate neutralization assays and perform statistical analyses, which we now do.
+The rest of the response is rather long because the reviews were so thorough, but you will see that all points are addressed.
 
-In conclusion, our paper has been evaluated by four expert reviewers, who all found it to be scientifically sound, raising only minor points that we have clarified in our response / revision.
+In conclusion, our paper has been evaluated by four reviewers who all found it to be sound, raising only minor points that we have clarified in our revision.
 The only question was whether our work was of sufficient impact for \textit{Nature Microbiology}, with the reviewers splitting on this question.
 The \textit{Nature Communications} policies read: ``While \textit{Nature} and Nature research journals strive to publish the highest impact research within these disciplines, \textit{Nature Communications} is committed to publishing important advances of significance to specialists within each field.''
-Our study clearly meets these criteria from the evaluations of all reviewers, and so is suitable for \textit{Nature Communications}.
+Our study clearly meets these criteria according to the evaluations of all reviewers.
 
@@
 Thank you for pointing out these two important references, which we now discuss.
-We also think it is interesting that both of these studies selected their escape mutants in group 2 HAs rather than a group 1 HA like the one in our study.
+Two additional references were pointed out by Reviewer \#3.
+We also think it is interesting that all of these studies selected their escape mutants in group 2 HAs rather than a group 1 HA like the one in our study.
 Anecdotally, it seems to us that there are more described broad anti-stalk antibodies against group 1 than group 2 HAs.
@@
 \begin{quote}
-\textsl{Interestingly, most previous studies (Chai, 2016; Dunand, 2016) that have reported selecting single mutations with large effects ($\gg$10-fold) on neutralization by anti-stalk antibodies have used group 2 (e.g., H3 or H7) HAs rather than group 1 HAs like the one used in our work.}
+\textsl{Interestingly, most previous studies (Chai, 2016; Ekiert, 2011; Friesen, 2014; Dunand, 2016) that have reported selecting single mutations with large effects ($\gg$10-fold) on neutralization by anti-stalk antibodies have used group 2 (e.g., H3 or H7) HAs rather than group 1 HAs like the one used in our work.}
 \end{quote}
@@
 This is important because as the reviewer notes, there can be batch effects.
-It was not feasible to perform all assays for all mutants on the same day, but all assays that are being compared are on the same day.
-So for instance, if you look at Figure~6, you will see that the wildtype curves in panels A, B, and C are slightly different since each was performed a different day.
-But each panel is for a single antibody, and all curves with that antibody were performed on the same day---that is, we repeated the wildtype in each batch of assays to avoid any confounding batch effects.
-Therefore, we have already done what the reviewer suggests, we just hadn't clearly shown all the data before. 
-That is remedied by the new Figure~S9.
+If you look at Figure~6, you will see that the wildtype curves in panels A, B, and C are slightly different since each was performed a different day.
+But each panel is for a single antibody, and all curves with that antibody were performed on the same day---we repeated the wildtype in each batch of assays to avoid any confounding batch effects.
+Therefore, we have already done what the reviewer suggests, and the new Figure~S9 now clearly explains this.
 
 As described in the response to question 5 by Reviewer \#1, the new statistical analysis finds that the effect for K(-8)T is significant, whereas that for K280A is not.
-We agree that the effect for K(-8)T is somewhat perplexing, and we wish that it would go away since it is not central to any of our claims, and has elicited skepticism from both us and the reviewers.
-In the original manuscript, we conveyed this skepticism by saying, \textsl{``we are not confident that this effect exceeds the noise in our neutralization assays.''}.
-However, the new statistical analysis described in response to question 5 by Reviewer \#1 finds the effect is significant, but we continue to add strong caveats (see the quoted new text in the response to question 5 by Reviewer \#1).
+We agree that the effect for K(-8)T is somewhat perplexing.
+In the original manuscript, we said, \textsl{``we are not confident that this effect exceeds the noise in our neutralization assays.''}.
+However, the new statistical analysis described in response to question 5 by Reviewer \#1 finds the effect is significant, but we continue to add strong caveats (see the response to question 5 by Reviewer \#1).
 So rest assured that we are not trying to exaggerate the effect of K(-8)T; we sincerely wish it would go away as it would simplify the reviewer responses!
-However, this mutation was picked up by our mutational antigenic profiling (which was performed in triplicate), and then again by a totally independent approach (neutralization assays, which were again performed in triplicate).
+However, this mutation was identified by our mutational antigenic profiling (which was performed in triplicate), and then again by a totally independent approach (neutralization assays, again performed in triplicate).
 So in good conscience, we feel that we have to present the result with K(-8)T, however perplexing.
@@
 {\color{black}
-We are disappointed that the reviewer does not view the work of sufficient impact and novelty for \textit{Nature Microbiology}, but we are appreciative of the statement that the study is ``performed to an extremely high technical standard.''
-Personally, we feel that direct quantitative comparison of the escape from narrow and broad antibodies is valuable, since other comparisons have been more anecdotal.
-In any case, we are now re-submitting to \textit{Nature Communications}, which places less emphasis on impact.
-We also note that at least two of the four reviewers found the work to be of high impact---and even Reviewer \#3 praises it on a technical level.
-Therefore, we think it is appropriate for \textit{Nature Communications}.}
+We are disappointed that the reviewer does not view the work of sufficient impact for \textit{Nature Microbiology}, but are appreciative of the statement that the study is ``performed to an extremely high technical standard.''
+We feel that direct quantitative comparison of the escape from narrow and broad antibodies is valuable, since other comparisons have been more anecdotal.
+We also note that at least two of the four reviewers found the work to be of high impact.
+Therefore, we think it is appropriate for \textit{Nature Communications}.
+
+Thank you for the references mentioned above. 
+Interestingly, references 3, 4, and 6 all describe selecting escape from \emph{group 2} HAs (e.g., H3 or H7), not a group 1 HA like the one we have used.
+We now cite these references and mention this point as described in the response to question 7 by Reviewer \#1.
+It is true that reference 5 mentioned by the reviewer does mention an escape mutation from a group 1 HA from CR6261, but it never actually shows any neutralization data for that mutation, so we are unable to comment on it in detail.
+}
 
@@
 
-{\color{red}
-Need to add response.}
+{\color{black}
+The original paper on C179 selected a strong escape mutant virus in an H1 background.
+That virus had three mutations, and although the authors argue in the text that only one of them is important, they never actually show that the others are not required as secondary changes.
+We cite this paper, and use it to inform the discussion of multiple mutants that is described in response to question 1 from Reviewer \#1.
+}
 
@@
 \color{black}
-This excellent question was also asked by other reviewers, and we apologize for not providing more details in the original submission.
-We have made extensive revisions to better describe the neutralization curves, show the curves for all individual replicates, calculate IC50s and the statistical significance of their differences from wildtype, and describe the curve fitting procedures.
+This question was also asked by other reviewers.
+We now better describe the neutralization curves, show curves for all replicates, calculate IC50s and the statistical significance of their differences from wildtype, and describe the curve fitting procedures.
 These revisions are detailed in response to question 5 of Reviewer \#1 and questions 5 and 6 of Reviewer \#2.}
@@
 
+{\color{black}
+Thanks for these references, they are now all cited.
+}
+
 \subsubsection*{Reviewer \#4 (Remarks to the Author):}
@@
 
+{\color{black}
+Thanks for the kind comments on the importance of the paper and the questions it addresses.
+
+The point about clearly distinguishing between how mutations affect antigenicity and viral growth is well taken.
+We have added text at several places to clarify this.
+For instance, as soon as we define our measures of antigenic effects, we say:
+\begin{quote}
+\textsl{Importantly, Equations~1 and 2 correct for effects on viral growth due to normalization by the mock-selected control, and so measure only antigenicity and not viral growth provided that the virus at least grows well enough to be present in the library.}
+\end{quote}}
+
 Minor points:
@@
 
-{\color{red}
-Need to add response.}
+{\color{black}
+We are unable to answer this question, because the concentrations of any given antibody specificity can vary widely among people.
+In addition, it is difficult to relate \textit{in vitro} neutralization to \textit{in vivo} activity due to Fc-mediated effector functions.
+This is a limitation not just of our study, but of virtually all work on antibody neutralization.
+We mention this caveat in the Discussion:
+\begin{quote}
+\textsl{When interpreting the magnitude of these effects, it is important to note that our experiments only measure how mutations affect neutralization, and not how they affect Fc-mediated functions that are responsible for much of the in vivo protection afforded by anti-stalk antibodies.}
+\end{quote}
+}
 
@@
 
-{\color{red}
-Need to add response.}
+{\color{black}
+This is an important question.
+Unfortunately, most of the antibodies have been characterized against only a very small number of the very large number of strains within HA subtypes.
+We summarize the subtype-level information in Figure 2B, and cite all studies that look at breadth within subtypes (references 24, 25, 29, 45, 46).
+However, because different studies use different strains, the data in these studies do not enable useful comparisons across antibodies.
+Our study only looks at one strain---but it least it is complete for the mutations to this HA, and so can be compared across antibodies.}
 
@@
 
+{\color{black}
+This is a good reference, and we now cite it at the relevant point.
+Interestingly, the distribution of infection frequencies they derive from epidemiology is similar to the 4-7 years we had previously cited from a human challenge study.
+However, we have updated the text to say ``$\sim$5 years'' to better reflect the ambiguity.}
+
 on immune status driving the rate of antigenic evolution -- Wen et al. Proc B 2016
 
-{\color{red}
-Need to add response.}
+{\color{black}
+This is a good reference, and we now cite it at the relevant point.}
 
@@
 
-{\color{red}
-Need to add response.}
+{\color{black}
+This is a good point, we have clarified the wording in these phrases.}
 

### 2018-01-22 paper/manuscript.tex
finished draft of reviewer responses

@@
 In contrast, human influenza virus exhibits much more antigenic variation.
-So although infection with an influenza virus strain provides long-term immunity to that exact strain\cite{fluinboardingschool1978,davies1982christ,yu2008neutralizing}, the virus's rapid antigenic evolution erodes the effectiveness of this immunity to that strain's descendants within four to seven years\cite{couch1983immunity}.
+So although infection with an influenza virus strain provides long-term immunity to that exact strain\cite{fluinboardingschool1978,davies1982christ,yu2008neutralizing}, the virus's rapid antigenic evolution erodes the effectiveness of this immunity to that strain's descendants within $\sim$5 years\cite{couch1983immunity,kucharski2015estimating}.
 
@@
 According to this explanation, human influenza virus undergoes rapid antigenic drift because most neutralizing antibodies target epitopes on the viral hemagglutinin (HA) protein that are highly tolerant of mutational change.
-This explanation is supported by classic experiments showing that it is easy to select viral mutants that escape most antibodies\cite{yewdell1979antigenic,webster1980determination}, as well as by the observation that mutations that alter antigenicity arise frequently during influenza's evolution globally\cite{koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction} and within individual humans with long-term infections\cite{xue2017parallel}.
+This explanation is supported by classic experiments showing that it is easy to select viral mutants that escape most antibodies\cite{yewdell1979antigenic,webster1980determination}, as well as by the observation that mutations that alter antigenicity arise frequently during influenza's evolution globally\cite{koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction,wen2016explaining} and within individual humans with long-term infections\cite{xue2017parallel}.
 A corollary of this explanation is that influenza virus's capacity for antigenic drift would be reduced if most antibodies instead targeted epitopes that were less mutationally tolerant.
@@
 However, these studies did not select for antibody escape, so it is difficult to relate their measurements to the virus's evolutionary capacity under immune selection.
-Other work has shown that it is possible to select antigenic mutants with broad antibodies\cite{yoshida2009cross,chai2016two,anderson2017natural}, demonstrating that these epitopes are not entirely refractory to change.
+Other work has shown that it is possible to select antigenic mutants with broad antibodies\cite{yoshida2009cross,chai2016two,ekiert2011highly,friesen2014common,dunand2015preexisting,anderson2017natural}, demonstrating that these epitopes are not entirely refractory to change.
 But given that antibodies can select some antigenic variation even in measles virus\cite{birrer1981antigenic,ter1981antigenic} and polio virus\cite{crainic1983natural,diamond1985antigenic}, the existence of selectable mutations does not necessarily imply that influenza virus can escape broad antibodies as easily as it drifts away from narrow strain-specific ones.
@@
 \end{equation}
+Importantly, Equations~\ref{eq:fracsurvive} and \ref{eq:fracsurvive_excess} correct for effects on viral growth due to normalization by the mock-selected control, and so measure only antigenicity and not viral growth provided that the virus at least grows well enough to be present in the library.
 Details of how the calculations are extended to account for sequencing errors and sampling statistics are in the \nameref{sec:methods}.
@@
 Because these libraries have different capacities to escape each antibody, the fraction of each library that survives high antibody concentrations will vary among antibodies.
-For instance, at concentrations that neutralize 99\% of the wildtype virus, we expect a larger fraction of a library to survive an antibody with many escape mutations than an antibody with few escape mutations.
+For instance, at concentrations that neutralize 99\% of the wildtype virus, we expect a larger fraction of a library to survive an antibody for which there are many HA escape mutations than an antibody with few HA escape mutations.
 Therefore, rather than using the same concentration for all antibodies, we selected concentrations for each antibody where between 2\% and 0.1\% of the libraries survived in order to strongly select for escape mutations (Figure~\ref{fig:neutcurves}).
@@
 However, our approach shows that there are mutations that have more modest ($<$10-fold) effects on neutralization by even the broadest antibody. 
-Interestingly, most previous studies~\cite{chai2016two,dunand2015preexisting} that have reported selecting single mutations with large effects ($\gg$10-fold) on neutralization by anti-stalk antibodies have used group 2 (e.g., H3 or H7) HAs rather than group 1 HAs like the one used in our work.
+Interestingly, most previous studies~\cite{chai2016two,ekiert2011highly,friesen2014common,dunand2015preexisting} that have reported selecting single mutations with large effects ($\gg$10-fold) on neutralization by anti-stalk antibodies have used group 2 (e.g., H3 or H7) HAs rather than group 1 HAs like the one used in our work.
 When interpreting the magnitude of these effects, it is important to note that our experiments only measure how mutations affect neutralization, and not how they affect Fc-mediated functions that are responsible for much of the \textit{in vivo} protection afforded by anti-stalk antibodies\cite{dilillo2014broadly,dilillo2016broadly}.

### 2018-01-22 paper/NatureMicrobiologyReviews.tex
some very minor fixes

@@
 Our claim is that we specifically identify mutations that increase resistance to each of the individual monoclonal antibodies.
-With the three broad antibodies tested in this paper (FI6v3, C179, S139/1) and the four narrow antibodies tested by Doud et al (2016), we have now examined seven different antibodies.
+With the three broad antibodies tested in this paper (FI6v3, C179, S139/1) and the four narrow antibodies tested by Doud et al (2017), we have now examined seven different antibodies.
 The only overlap between antibodies in the identified resistance mutations comes when those antibodies target similar epitopes (e.g., FI6v3 and C179)---there is no overlap in the mutations identified by antibodies targeting different epitopes.
@@
 Put differently, are conserved epitopes that are targeted by some antibodies conserved because they've not had to change, or are they conserved because no beneficial changes are accessible?
-This question is obviously relevant for "universal" vaccine development, but it's also important for understanding strain structure and evolution more generally.
+This question is obviously relevant for ``universal'' vaccine development, but it's also important for understanding strain structure and evolution more generally.
 
@@
 However, because different studies use different strains, the data in these studies do not enable useful comparisons across antibodies.
-Our study only looks at one strain---but it least it is complete for the mutations to this HA, and so can be compared across antibodies.}
+Our study only looks at one strain---but at least it is complete for the mutations to this HA, and so can be compared across antibodies.}
 
@@
 
-"the viruses's rapid antigenic evolution erodes the effectiveness of this immunity to that strain's descendants within four to seven years..." -- Kucharski et al.
+``the viruses's rapid antigenic evolution erodes the effectiveness of this immunity to that strain's descendants within four to seven years...'' -- Kucharski et al.
 PLOS Biol 2015 (although numbers might need revising)
@@
 
-4. In "...we expect a larger fraction of a library to survive an antibody with many escape mutations..." (and similar phrases), the "many escape mutations" read as though they modify "antibody."
+4. In ``...we expect a larger fraction of a library to survive an antibody with many escape mutations...'' (and similar phrases), the ``many escape mutations'' read as though they modify ``antibody.''
 

### 2018-01-22 paper/NatureMicrobiologyReviews.tex
shortened reviewer response a bit

@@
 Examining all single mutations makes the results complete and quantitative---but we can't make claims about the effects of multiple mutations.
-In contrast, multi-cycle growth sometimes finds combinations of mutations, but is not complete and quantitative, since there are too many multiple mutants to examine them all, and it is random which mutations arise in any given multi-cycle growth experiment.
-In particular, multi-cycle growth assays do not generate data that can be quantitatively compared across antibodies using the methods here.
+In contrast, multi-cycle growth sometimes finds combinations of mutations, but is not complete and quantitative, since it is random which mutations arise in any given multi-cycle growth experiment.
 Therefore, the strength of our approach (examining all single mutations in a way that is comparable across antibodies) comes with the limitation that we do not find multiple mutants.
@@
 First, the title of the paper reads ``Quantifying the effects of \emph{single mutations} on viral escape from broad and narrow antibodies to an H1 influenza hemagglutinin'' (emphasis added).
-Stating this fact in the title makes the limitation immediately and directly clear.
 
@@
 The reviewer is correct that A/WSN/1933 is a lab-adapted virus.
-As the reviewer notes, in his/her comment, this is not necessarily a problem for our work since we are simply examining antibody neutralization rather than pathogenesis, fitness, or transmission. 
+As the reviewer also notes, this is not actually a problem for our work since we are examining antibody neutralization rather than pathogenesis, fitness, or transmission. 
 Nonetheless, the reviewer is correct that it is important to clearly mention this caveat. 
-We have now added the following text when we first discuss this strain:
+We have added the following text:
 
@@
 This is a good comment that was also made by several other reviewers.
-We therefore now do as suggested by showing full data for triplicate measurements with significance testing.
+We therefore now show full data for triplicate measurements with significance testing.
 
-Specifically, all the neutralization assays with FI6v3 were performed in triplicate on the same day to eliminate batch effects, with each replicate involving independent serial dilution of the antibody in a separate column of a 96-well plate (this is now more fully explained in the Methods).
+Specifically, all neutralization assays with FI6v3 were performed in triplicate on the same day to eliminate batch effects, with each replicate involving independent serial dilution of the antibody in a separate column of a 96-well plate (this is now explained in the Methods).
 In the original manuscript, we only showed the mean and standard deviation in Figure~6A.
-We have now added a supplementary figure (Figure~S9) that shows all the replicates.
-In addition, as explained in the legend to the new Figure~S9, we fit IC50 values to each replicate, and then use the three IC50's to test the null hypothesis that each mutant has an IC50 indistinguishable from wildtype.
+We have added Figure~S9, which shows all the replicates.
+In addition, as explained in Figure~S9, we fit IC50 values to each replicate, and used the three IC50's to test the null hypothesis that each mutant has an IC50 indistinguishable from wildtype.
 
-The statistical testing results in Figure~S9B support the conclusions in the text. 
+The statistical testing in Figure~S9B supports the conclusions in the text. 
 The four mutations that are \emph{not} expected to differ from wildtype with respect to FI6v3 neutralization (K280A, V135T, P80D, and M17L(HA2)) do \emph{not} have IC50s that are significantly different than wildtype.
 The four mutations expected to have the largest increases in neutralization resistance (K280S, G47R(HA2), K280T, and N291S) are all significantly more resistant to FI6v3 than wildtype, with corrected $P$-values ranging from 0.00072 to 0.023.
-The mutation found by our mutational antigenic profiling but that elicited the most skepticism both from us and the reviewers is K(-8)T.
+The mutation found by our mutational antigenic profiling that elicited the most skepticism both from us and the reviewers is K(-8)T.
 However, the statistical testing suggests that it really is slightly but significantly more resistant to FI6v3, with a corrected $P$-value of 0.04. 
-But we agree that the effect is very small, and in the revised text we clearly emphasize this fact and advise caution in over-interpreting the result:
+However the effect is very small, and in the revised text we clearly emphasize this fact and advise caution in over-interpreting the result:
 
@@
 The reviewer is correct that we do not include biophysical measurements of why the mutants have increased neutralization resistance.
-However, biophysically characterizeing antibody binding to the mutants is beyond the scope of the current study.
-We are careful to avoid making any claims about the mechanisms of the mutations---we simply say that they reduce neutralization.
+However, biophysically characterizeing antibody binding is beyond the scope of our study.
+We do not make any claims about the mechanisms of the mutations---we simply say that they reduce neutralization.
 This is in line with many other studies in the field.
@@
 {\color{black}
-Thank you for pointing out these two important references, which we now discuss.
-Two additional references were pointed out by Reviewer \#3.
-We also think it is interesting that all of these studies selected their escape mutants in group 2 HAs rather than a group 1 HA like the one in our study.
-Anecdotally, it seems to us that there are more described broad anti-stalk antibodies against group 1 than group 2 HAs.
-We obviously have no way to quantify this rigorously, but in the added text that mentions the two studies suggested by the reviewer, we also point out the difference in HA groups:
+Thank you for pointing out these important references, which we now discuss.
+Two more references were pointed out by Reviewer \#3.
+Interestingly, all these studies selected their escape mutants in group 2 HAs rather than a group 1 HA like the one in our study.
+Anecdotally, it seems that there are more described broad anti-stalk antibodies against group 1 than group 2 HAs.
+We obviously have no way to quantify this rigorously, but in the added text that mentions the studies suggested by the reviewer, we also point out the difference in HA groups:
 
@@
 We see the rationale behind this suggestion, but do we not believe it is essential for the current work. 
-Our claim is that we specifically identify mutations that increase resistance to each of the individual monoclonal antibodies.
-With the three broad antibodies tested in this paper (FI6v3, C179, S139/1) and the four narrow antibodies tested by Doud et al (2017), we have now examined seven different antibodies.
-The only overlap between antibodies in the identified resistance mutations comes when those antibodies target similar epitopes (e.g., FI6v3 and C179)---there is no overlap in the mutations identified by antibodies targeting different epitopes.
-This fact indicates that we are identifying bona fide resistance mutations to the specific antibodies used, rather than mutations with some general effect on all IgGs.
+Our claim is that we specifically identify mutations that increase resistance to each of the individual antibodies.
+With the three broad antibodies tested in this paper (FI6v3, C179, S139/1) and the four narrow antibodies tested by Doud et al (2017), we have examined seven different antibodies.
+The only overlap in the identified resistance mutations comes when the antibodies target similar epitopes (e.g., FI6v3 and C179)---there is no overlap in the mutations identified by antibodies targeting different epitopes.
+This fact indicates that we are identifying bona fide resistance mutations, rather than mutations with some general effect on all IgGs.
 
-Figure~6 provides further data indicating that these are genuine antibody-specific resistance mutations rather than mutations generally selected by any IgG.
-For instance, the mutations that increase resistance to H17-L19 (V135T) or H17-L7 (P80D) do not affect neutralization by FI6v3, demonstrating that they are specific to the antibodies that selected them.
-Finally, we note that there are numerous published papers where the same monoclonal antibody is tested against a wide range of viral variants for the purposes of comparing IC50s across viruses without also including data showing that all viral variants are unaffected by an irrelevant antibody (e.g., Table~1 of \textit{Nature}, 489:526-532; Figure~3 of \textit{Nature Microbiology},; 2:16199).
+Figure~6 provides further data indicating that we find antibody-specific resistance mutations rather than mutations generally selected by any IgG.
+For instance, the mutations that increase resistance to H17-L19 (V135T) or H17-L7 (P80D) do \emph{not} affect neutralization by FI6v3, showing they are specific to the antibodies that selected them.
+Finally, note that there are numerous papers where a monoclonal antibody is tested against a range of viral variants to compare IC50s without including data showing that all viral variants are unaffected by an irrelevant antibody (e.g., Table~1 of \textit{Nature}, 489:526-532; Figure~3 of \textit{Nature Microbiology},; 2:16199).
 }
@@
 {\color{black}
-All of the results for the broadly neutralizing antibodies are new to this study.
-The reviewer is correct that the data for the narrow antibodies in Figure 6B,C is from Doud et al (2017), and the mutational tolerance data in Figure 7 is from Doud and Bloom (2016).
-This fact is clearly stated; for instance, the figure legends say ``Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 taken from Doud et al (2017)'' and ``The logo plots at bottom show the preference of each HA site of each amino acid under selection for viral replication as measured by Doud and Bloom (2016).''
+All the results for the broadly neutralizing antibodies are new to this study.
+The data for the narrow antibodies in Figure 6B,C is from Doud et al (2017), and the mutational tolerance data in Figure 7 is from Doud and Bloom (2016).
+This fact is clearly stated; for instance, the figure legends say ``Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 \emph{taken from Doud et al (2017)}'' and ``The logo plots at bottom show the preference of each HA site of each amino acid under selection for viral replication \emph{as measured by Doud and Bloom (2016)}.''
 There is nothing wrong with including previous data for comparative purposes.
-For instance, new reports of protein crystal structures often compare the results to prior structures, and often include these prior structures in figures for points of comparison.
-In fact, a major strength of our approach is that it puts how mutations affect neutralization into units that can be meaningfully compared across studies.
+For instance, papers on crystal structures often compare the results to prior structures on related molecules.
+In fact, a major strength of our approach is that it puts the results into units that can be meaningfully compared across studies.
 
-The experimental methodology is the same as in Doud et al (2017), although the entire analysis framework (all Equations in the paper, not just Equations 10 and 11) are new.
-The strength of this analysis framework is that it enables the results to be quantitatively compared across antibodies, which was not possible with the approach in Doud et al (2017). 
+The experimental methodology is the same as in Doud et al (2017), although the entire analysis framework (all equations, not just Equations 10 and 11) are new.
+The strength of this framework is that it enables the results to be compared across antibodies, which was not possible with the approach in Doud et al (2017). 
 }
@@
 {\color{black}
-We certainly agree that it would be interesting to examine additional strains.
-However, prior to our work, no study had ever completely mapped the effect of all single mutations to any viral strain with respect to neutralization by any broadly neutralizing influenza antibody, so we believe that just doing this with one strain is a sufficient step forward in knowledge to be valuable.
+We agree that it would be interesting to examine additional strains.
+However, prior to our work, no study had completely mapped the effect of all single mutations to any viral strain with respect to neutralization by any broad antibodies---so doing this with one strain is a substantial step forward.
 As mentioned in the response to question 2 by Reviewer \#1, we now more clearly explain that WSN is a lab-adapted strain.
@@
 {\color{black}
-The neutralization assays in Figure~6 show that the mutations for the stalk-targeting antibodies only have weak effects on neutralization, while Figure~7 shows that many of them don't have much effect on viral growth.
-In addition, we were able to generate all of the stalk-antibody mutations for the neutralization assays, so they are not too badly attenuated for viral fitness in cell culture.
+The neutralization assays in Figure~6 show that mutations for the stalk-targeting antibodies only have weak effects on neutralization, while Figure~7 shows that many of them don't have much effect on viral growth.
+In addition, we were able to generate all the stalk-antibody escape mutants for the neutralization assays, so they are not too badly attenuated in cell culture.
 However, we avoid making strong statements about the effects of mutations on viral growth beyond the data in Figure~7 because it is beyond the scope of our study to do growth curves for such a large number of mutations.}
@@
 {\color{black}
-We examined all amino-acid changes, including those that involved multiple-nucleotide changes.
+Our study tested all amino-acid changes, including those that involved multiple-nucleotide changes.
 The reason is that our libraries contained all codon mutations (e.g., the 63 non-wildtype codons at each site) rather than just all single nucleotide mutations.
@@
 {\color{black}
-We agree that this is an important caveat, and it was also mentioned by other reviewers.
-See our response to question 1 by Reviewer \#1 for how we have addressed this caveat.}
+We agree that this is an important caveat.
+See our response to question 1 by Reviewer \#1.}
 
@@
 {\color{black}
-This is a good point, and was also made by several other reviewers.
+A similar point was made by several other reviewers.
+We have provided a detailed response in answer to question 5 by Reviewer \#1.
 
-We have provided a detailed response in answer to question 5 by Reviewer \#1.
-Our revisions include showing the neutralization curves for all replicates and performing statistical testing.
-All the neutralization curves for antibody FI6v3 were performed on the same day.
+Our revisions showithe neutralization curves for all replicates and performing statistical testing.
+All neutralization curves for antibody FI6v3 were performed on the same day.
 This is important because as the reviewer notes, there can be batch effects.
-If you look at Figure~6, you will see that the wildtype curves in panels A, B, and C are slightly different since each was performed a different day.
-But each panel is for a single antibody, and all curves with that antibody were performed on the same day---we repeated the wildtype in each batch of assays to avoid any confounding batch effects.
+If you look at Figure~6, you will see that the wildtype curves in panels A, B, and C are slightly different.
+But each panel is for a single antibody, and all curves with with each antibody were performed on the same day to avoid any confounding batch effects.
 Therefore, we have already done what the reviewer suggests, and the new Figure~S9 now clearly explains this.
@@
 As described in the response to question 5 by Reviewer \#1, the new statistical analysis finds that the effect for K(-8)T is significant, whereas that for K280A is not.
-We agree that the effect for K(-8)T is somewhat perplexing.
-In the original manuscript, we said, \textsl{``we are not confident that this effect exceeds the noise in our neutralization assays.''}.
-However, the new statistical analysis described in response to question 5 by Reviewer \#1 finds the effect is significant, but we continue to add strong caveats (see the response to question 5 by Reviewer \#1).
-So rest assured that we are not trying to exaggerate the effect of K(-8)T; we sincerely wish it would go away as it would simplify the reviewer responses!
-However, this mutation was identified by our mutational antigenic profiling (which was performed in triplicate), and then again by a totally independent approach (neutralization assays, again performed in triplicate).
-So in good conscience, we feel that we have to present the result with K(-8)T, however perplexing.
-
-The reviewer also points out that it might be interesting to look at the $F_{r,a}$ value for K280A, since it might have reduced neutralization resistance.
-First, we note that the statistical analysis in the new Figure~S9 shows that the difference between K280A and wildtype is \emph{not} statistically significant, so we do not comment on possible increased neutralization sensitivity or the $F_{r,a}$ for this mutation in the paper.
-However, we have added a new supplementary file (File~S6) that gives all the $F_{r,a}$ values in case anyone wants to examine them.
+The reviewer also suggests that it might be interesting to look at the $F_{r,a}$ value for K280A, since it might have reduced neutralization resistance.
+Since the statistical analysis in the new Figure~S9 shows that the difference between K280A and wildtype is \emph{not} statistically significant, we do not comment on possible increased neutralization sensitivity or the $F_{r,a}$ for this mutation in the paper.
+However, we have added a new supplementary file (File~S6) that gives all the $F_{r,a}$ values.
 The value of $F_{r,a}$ for this mutation is 0.00382, which is slightly below the library average of 0.00955.
@@
 {\color{black}
-For the quantitative analyses of IC50 values, we fit four-parameter logistic curves, although we fixed the top value to one, so there are really only three parameters.
-This is now explained in the legend to the new Figure~S9, which reports the actual IC50 values and shows the curves for individual replicates.
-The exact computer code used to fit the curves is at \url{https://jbloomlab.github.io/dms_tools2/dms_tools2.neutcurve.html}.
+For the quantitative analyses of IC50's, we fit four-parameter logistic curves, although we fixed the top value to one.
+This is explained in the legend to the new Figure~S9, which reports the IC50's and shows the curves for individual replicates.
+The computer code used to fit the curves is at \url{https://jbloomlab.github.io/dms_tools2/dms_tools2.neutcurve.html}.
 }
@@
 {\color{black}
-We are disappointed that the reviewer does not view the work of sufficient impact for \textit{Nature Microbiology}, but are appreciative of the statement that the study is ``performed to an extremely high technical standard.''
+We are disappointed that the reviewer does not view the work of sufficient impact for \textit{Nature Microbiology}, but appreciate the statement that the study is ``performed to an extremely high technical standard.''
 We feel that direct quantitative comparison of the escape from narrow and broad antibodies is valuable, since other comparisons have been more anecdotal.
-We also note that at least two of the four reviewers found the work to be of high impact.
-Therefore, we think it is appropriate for \textit{Nature Communications}.
+We also note that at least two of the reviewers found the work of high impact.
 
-Thank you for the references mentioned above. 
-Interestingly, references 3, 4, and 6 all describe selecting escape from \emph{group 2} HAs (e.g., H3 or H7), not a group 1 HA like the one we have used.
+Thank you for the references. 
+Interestingly, (3), (4), and (6) all describe selecting escape from \emph{group 2} HAs (e.g., H3 or H7), not a group 1 HA like the one we used.
 We now cite these references and mention this point as described in the response to question 7 by Reviewer \#1.
-It is true that reference 5 mentioned by the reviewer does mention an escape mutation from a group 1 HA from CR6261, but it never actually shows any neutralization data for that mutation, so we are unable to comment on it in detail.
+It is true that (5) does mention an escape mutation from a group 1 HA from CR6261, but it never actually shows any neutralization data for that mutation, so we are unable to comment on it.
 }
@@
 {\color{black}
-The original paper on C179 selected a strong escape mutant virus in an H1 background.
-That virus had three mutations, and although the authors argue in the text that only one of them is important, they never actually show that the others are not required as secondary changes.
-We cite this paper, and use it to inform the discussion of multiple mutants that is described in response to question 1 from Reviewer \#1.
+The original paper on C179 selected a strong escape mutant in an H1 background.
+That virus had three mutations, and although the paper argues that only one is important, it never shows that the others are not required as secondary changes.
+We cite this paper, and use it to inform the discussion of multiple mutants described in response to question 1 from Reviewer \#1.
 }
@@
 \color{black}
-This question was also asked by other reviewers.
-We now better describe the neutralization curves, show curves for all replicates, calculate IC50s and the statistical significance of their differences from wildtype, and describe the curve fitting procedures.
-These revisions are detailed in response to question 5 of Reviewer \#1 and questions 5 and 6 of Reviewer \#2.}
+This point was also mentioned by other reviewers.
+See our response to question 5 of Reviewer \#1 and questions 5 and 6 of Reviewer \#2.}
 
@@
 In addition, it is difficult to relate \textit{in vitro} neutralization to \textit{in vivo} activity due to Fc-mediated effector functions.
-This is a limitation not just of our study, but of virtually all work on antibody neutralization.
+This is a limitation of virtually all work on antibody neutralization.
 We mention this caveat in the Discussion:
@@
 {\color{black}
-This is an important question.
-Unfortunately, most of the antibodies have been characterized against only a very small number of the very large number of strains within HA subtypes.
+Unfortunately, most of the antibodies have been characterized against only a small number of the many strains within HA subtypes.
 We summarize the subtype-level information in Figure 2B, and cite all studies that look at breadth within subtypes (references 24, 25, 29, 45, 46).
@@
 {\color{black}
-This is a good reference, and we now cite it at the relevant point.
-Interestingly, the distribution of infection frequencies they derive from epidemiology is similar to the 4-7 years we had previously cited from a human challenge study.
-However, we have updated the text to say ``$\sim$5 years'' to better reflect the ambiguity.}
+This is a good reference, and we now cite it.
+The distribution of infection frequencies they derive from epidemiology is similar to the 4-7 years we had previously cited from a human challenge study.
+However, we have updated the text to ``$\sim$5 years'' to better reflect the ambiguity.}
 

### 2018-01-22 paper/ResponseToNatureMicrobiologyReviews.tex
version submitted to *Nature Communications*

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format

\usepackage[margin=1.1in]{geometry}

\usepackage{hyperref}
%\hypersetup{colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue}

\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
\usepackage{color}
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

\title{Response to reviews of ``Quantifying the effects of single mutations on viral escape from broad and narrow antibodies to an H1 influenza hemagglutinin'' from \textit{Nature Microbiology}}
\author{Michael B. Doud, Juhye M. Lee, and Jesse D. Bloom}

\begin{document}
\maketitle

\subsection*{Overview of reviewer response and changes to the manuscript.}

We received four reviews for the original submission to \textit{Nature Microbiology}.
All four reviewers found the work to be scientifically sound.
Two reviewers (\#1 and \#4) were also enthusiastic about impact.
The other two (\#2 and \#3) concurred about scientific soundness, but were less enthusiastic about impact.
The most critical was Reviewer \#3, who still wrote that ``[t]he study is performed to an extremely high technical standard''---a positive comment to receive from the most negative reviewer!

Because all reviewers found the work to be scientifically sound and at least half found it to be high impact, we request that the paper be considered for \textit{Nature Communications} with the current reviews.
Impact is subjective, and the fact that half the reviewers found the paper of sufficient impact for \textit{Nature Microbiology} indicates that it will interest many in the field. 
We emphasize the following impactful aspects of our study:
\begin{itemize}\parskip1pt
\item It introduces a rigorous method to compare the potential for immune-escape by single amino-acid mutations across antibodies.
\item It is the first paper to completely map how all single amino-acid mutations affect neutralization by broadly neutralizing anti-influenza antibodies.
\item It is the first direct head-to-head comparison of immune-escape from anti-head versus anti-stalk antibodies to influenza hemagglutinin.
\end{itemize}

Below we provide a detailed response to the reviews.
We have addressed all the questions / concerns.
The only major request was that we show all triplicate neutralization assays and perform statistical analyses, which we now do.
The rest of the response is rather long because the reviews were so thorough, but you will see that all points are addressed.

In conclusion, our paper has been evaluated by four reviewers who all found it to be sound, raising only minor points that we have clarified in our revision.
The only question was whether our work was of sufficient impact for \textit{Nature Microbiology}, with the reviewers splitting on this question.
The \textit{Nature Communications} policies read: ``While \textit{Nature} and Nature research journals strive to publish the highest impact research within these disciplines, \textit{Nature Communications} is committed to publishing important advances of significance to specialists within each field.''
Our study clearly meets these criteria according to the evaluations of all reviewers.

\subsection*{Response to reviews}
Below find our responses to the reviews received from \textit{Nature Microbiology}.
The original comments {\color{blue} are in blue}, and our responses are in black.

\color{blue}

\subsubsection*{Reviewer Expertise:}

\begin{itemize}\parskip1pt
\item Referee \#1: influenza antibody neutralization and escape
\item Referee \#2: viral evolution, drift, deep-sequencing, computation
\item Referee \#3: influenza antibody neutralization and escape
\item Referee \#4: computational analysis of viral evolution
\end{itemize}

{\color{black}
Thanks for selecting reviewers who cover the full spectrum of relevant expertise.}

\subsubsection*{Reviewer \#1 (Remarks to the Author):}

In their manuscript Doud and colleagues quantify the effects of single mutations on viral escape from both narrow and broadly binding anti-hemagglutinin antibodies. The methodology is based on a mutant library of an H1N1 strain which are analyzed by deep sequencing after a single infection cycle in the presence or absence of mAb. The data presented in the manuscript is of high significance for the field and the work is technically sound. There are some issues that the authors need to address but no major weaknesses were identified.

{\color{black}
We appreciate the fact that the reviewer finds our work to be both technically sound and of high impact.}

Specific points

1) Multicylce replication might have changed the data somewhat because it would have allowed the virus to acquire additional mutations.

{\color{black}
This is true.
Our approach systematically quantifies how all \emph{single} mutations affect viral neutralization.
This contrasts with an alternative approach of using multiple rounds of growth to select for multi-mutants.
There are strengths and weaknesses to both approaches.
Examining all single mutations makes the results complete and quantitative---but we can't make claims about the effects of multiple mutations.
In contrast, multi-cycle growth sometimes finds combinations of mutations, but is not complete and quantitative, since it is random which mutations arise in any given multi-cycle growth experiment.
Therefore, the strength of our approach (examining all single mutations in a way that is comparable across antibodies) comes with the limitation that we do not find multiple mutants.

We have taken the following pains to emphasize this limitation. 
First, the title of the paper reads ``Quantifying the effects of \emph{single mutations} on viral escape from broad and narrow antibodies to an H1 influenza hemagglutinin'' (emphasis added).

In addition, we have included the following paragraph in the Discussion:

\begin{quote}
\textsl{Another important caveat is that our experiments examine {\bf single} amino-acid mutations to the HA from one influenza virus strain.
The protein evolution literature is full of examples of epistatic interactions that enable multiple mutations to access phenotypes not accessible by single mutations.
Such epistasis is relevant to HA's evolution.  
For instance, work by Das et al suggests that the sequential accumulation of mutations can shift the spectrum of available antibody-escape mutations.
Wu et al have used deep mutational scanning to directly demonstrate that rampant epistasis enables HA's receptor-binding pocket to accommodate combinations of individually deleterious mutations, some of which affect sensitivity to antibodies.
Therefore, our work does not imply any absolute limits on the possibilities for antibody escape when evolution is given sufficient time to explore combinations of mutations.
However, single mutations are the most accessible form of genetic variation, and much of influenza virus's natural antigenic drift involves individual mutations that reduce sensitivity to immunodominant antibody specificities.
Quantifying the antigenic effects of all such mutations therefore provides a relevant measure of ease of viral antibody escape. }
\end{quote}
}

2) The virus used, WSN, is highly artificial and does not reflect actual human influenza viruses. It is trypsin independent (human viruses are not) and it is actually neurotropic (which is highly unusual). Also, it is a mouse adapted virus. While this is not necessarily a problem in the context of the presented data this caveat should be extensively discussed and avoided in the future.

{\color{black}
The reviewer is correct that A/WSN/1933 is a lab-adapted virus.
As the reviewer also notes, this is not actually a problem for our work since we are examining antibody neutralization rather than pathogenesis, fitness, or transmission. 
Nonetheless, the reviewer is correct that it is important to clearly mention this caveat. 
We have added the following text:

\begin{quote}
\textsl{We performed our experiments using the lab-adapted A/WSN/1933 (H1N1) strain of influenza.
This strain is derived from an early seasonal H1N1 that was extensively passaged in the lab, where it adapted to become neurotropic and trypsin independent.
But despite these unusual properties, the virus is neutralized by most broad antibodies that target other H1 viruses, including those used in this study (Figure~3).}
\end{quote}
}

3) Page 2, beginning of second last paragraph: It is unclear if such a vaccine would really enhance the pressure dramatically. This would depend on vaccine coverage and potency. Currently, the drift of influenza viruses is driven by immunity induced by natural infection, not by vaccines (global vaccination rates are extremely low, even in high income countries in Europe the coverage is low).

{\color{black}
This is a good point. We have changed ``will be under strong antigenic selection'' to ``could come under stronger antigenic selection.''
}

4) Data deposition at GitHub: How long will data/software stored on this platform be available? If versions are updated, is this tracked/recorded?

{\color{black}
GitHub is the leading platform for archiving and tracking computer code.
It is used widely in computational biology, and also by Google and Microsoft, among others. For instance, see: \url{https://www.wired.com/2015/03/github-conquered-google-microsoft-everyone-else/}

GitHub uses version tracking to record all updates to computer code, and makes it possible to access any prior version of the software.
For instance, while \url{https://github.com/jbloomlab/dms_tools2} provides the latest version of the code, the version used in this study (version 2.2.1) will always be available \url{https://github.com/jbloomlab/dms_tools2/tree/2.2.1}, and it is easy to examine modifications made to this version in any new updates.

While it is impossible to guarantee that GitHub will exist forever, the same could be said for \textit{Nature Press} (which could go out of business) or PubMed (which could be de-funded by the NIH or the Trump administration).
The best that we can say is that GitHub is considered the most stable and reliable way to archive and track software, as evidenced by its use not only in computational biology but also by major corporations such as Google and Microsoft.

In addition, all the Jupyter notebooks and major results are provided in supplementary files, which will be available on the \textit{Nature Press} website.
}

5) Figure 6: It might be a good idea to quantify the IC50 shifts and - since the neutralization assay was hopefully done at least in triplicates - check what is significant and what not. E.g. the effect of the -8 mutation for FI6 doesn't look real.

{\color{black}
This is a good comment that was also made by several other reviewers.
We therefore now show full data for triplicate measurements with significance testing.

Specifically, all neutralization assays with FI6v3 were performed in triplicate on the same day to eliminate batch effects, with each replicate involving independent serial dilutions of the antibody in a separate column of a 96-well plate (this is now explained in the Methods).
In the original manuscript, we only showed the mean and standard deviation in Figure~6A.
We have added Figure~S9, which shows all the replicates.
In addition, as explained in Figure~S9, we fit IC50 values to each replicate, and used the three IC50's to test the null hypothesis that each mutant has an IC50 indistinguishable from wildtype.

The statistical testing in Figure~S9B supports the conclusions in the text. 
The four mutations that are \emph{not} expected to differ from wildtype with respect to FI6v3 neutralization (K280A, V135T, P80D, and M17L(HA2)) do \emph{not} have IC50s that are significantly different than wildtype.
The four mutations expected to have the largest increases in neutralization resistance (K280S, G47R(HA2), K280T, and N291S) are all significantly more resistant to FI6v3 than wildtype, with corrected $P$-values ranging from 0.00072 to 0.023.
The mutation found by our mutational antigenic profiling that elicited the most skepticism both from us and the reviewers is K(-8)T.
However, the statistical testing suggests that it really is slightly but significantly more resistant to FI6v3, with a corrected $P$-value of 0.04. 
However the effect is very small, and in the revised text we clearly emphasize this fact and advise caution in over-interpreting the result:

\begin{quote}
\textsl{The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.
We tested one of these mutations, K(-8)T, and it did lead to a very slight increase in neutralization resistance (Figure~6A, Figure~S9)---although despite the significance testing in Figure~S9, we remain circumspect about the magnitude of this effect relative to the noise in our neutralization assays.
}
\end{quote}

These revisions address this important comment made by several reviewers.
}

6) It would be important to quantify differences in mAb binding to the different escape mutants shown in Figure 6, e.g. by SPR/BLI. Right now it is unclear which mutant produces a complete loss of binding, if affinity changes result in the loss of neutralization etc. This needs to be explored.

{\color{black}
The reviewer is correct that we do not include biophysical measurements of why the mutants have increased neutralization resistance.
However, biophysically characterizing antibody binding is beyond the scope of our study.
We do not make any claims about the mechanisms of the mutations---we simply say that they reduce neutralization.
This is in line with many other studies in the field.
While biophysically oriented studies certainly include such measurements, there are numerous papers on viral evolution and antibody breadth that characterize neutralization as the phenotype of interest.
Our study falls into this latter class.
}

7) Page 15/first page of discussion: The authors did not find a single-mutant virus with full loss of neutralization by anti-stalk mAbs. It has been noted that it is relatively difficult to make escape mutants against anti-stalk mAbs and the process usually involves serial passaging of virus in the presence of sub-neutralizing mAb concentrations - but usually full escape mutants are then recovered. In most cases these escape mutants have a) more than one mutation and b) have no completely loss of binding (but complete loss of neutralization). But there are also exceptions to this rule. Two papers have so far described escape from stalk-based mAbs in detail (\url{https://www.ncbi.nlm.nih.gov/pubmed/27351973} and \url{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4362269/}). It is strongly suggested that the authors consult these papers and include their findings in the discussion.

{\color{black}
Thank you for pointing out these important references, which we now discuss.
Two more references were pointed out by Reviewer \#3.
Interestingly, all these studies selected their escape mutants in group 2 HAs rather than a group 1 HA like the one in our study.
Anecdotally, it seems that there are more described broad anti-stalk antibodies against group 1 than group 2 HAs.
We obviously have no way to quantify this rigorously, but in the added text that mentions the studies suggested by the reviewer, we also point out the difference in HA groups:

\begin{quote}
\textsl{Interestingly, most previous studies (Chai, 2016; Ekiert, 2011; Friesen, 2014; Dunand, 2016) that have reported selecting single mutations with large effects ($\gg$10-fold) on neutralization by anti-stalk antibodies have used group 2 (e.g., H3 or H7) HAs rather than group 1 HAs like the one used in our work.}
\end{quote}
}

8) It would be good to have a negative control experiment with a non-binding human or mouse mAb. This is missing but should actually be the ``real'' negative control.

{\color{black}
We see the rationale behind this suggestion, but do we not believe it is essential for the current work. 
Our claim is that we specifically identify mutations that increase resistance to each of the individual antibodies.
With the three broad antibodies tested in this paper (FI6v3, C179, S139/1) and the four narrow antibodies tested by Doud et al (2017), we have examined seven different antibodies.
The only overlap in the identified resistance mutations comes when the antibodies target similar epitopes (e.g., FI6v3 and C179)---there is no overlap in the mutations identified by antibodies targeting different epitopes.
This fact indicates that we are identifying bona fide resistance mutations, rather than mutations with some general effect on all IgGs.

Figure~6 provides further data indicating that we find antibody-specific resistance mutations rather than mutations generally selected by any IgG.
For instance, the mutations that increase resistance to H17-L19 (V135T) or H17-L7 (P80D) do \emph{not} affect neutralization by FI6v3, showing they are specific to the antibodies that selected them.
Finally, note that there are numerous papers where a monoclonal antibody is tested against a range of viral variants to compare IC50s without including data showing that all viral variants are unaffected by an irrelevant antibody (e.g., Table~1 of \textit{Nature}, 489:526-532; Figure~3 of \textit{Nature Microbiology},; 2:16199).
}

\subsubsection*{Reviewer \#2 (Remarks to the Author):}

The authors use a previously described method of mutational antigenic profiling, a very clever and elegant approach exploited in previous papers in PLOS Pathogens and CHM. This approach involves generating viral libraries containing all mutations in the protein of interest, selecting these viruses with or without antibody, and deep-sequencing to determine the relative frequencies of each mutation. The full methodology (except for summary statistics described in Equations 10 and 11), as well as all the results for the narrow antibodies, were described in their previous paper (Doud et al., 2017, PLOS Pathogens). Figure 7 seems to be imported from another paper (Doud et al., 2016, Viruses).

{\color{black}
All the results for the broadly neutralizing antibodies are new to this study.
The data for the narrow antibodies in Figure 6B,C is from Doud et al (2017), and the mutational tolerance data in Figure 7 is from Doud and Bloom (2016).
This fact is clearly stated; for instance, the figure legends say ``Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 \emph{taken from Doud et al (2017)}'' and ``The logo plots at bottom show the preference of each HA site of each amino acid under selection for viral replication \emph{as measured by Doud and Bloom (2016)}.''
There is nothing wrong with including previous data for comparative purposes.
For instance, papers on crystal structures often compare the results to prior structures on related molecules.
In fact, a major strength of our approach is that it puts the results into units that can be meaningfully compared across studies.

The experimental methodology is the same as in Doud et al (2017), although the entire analysis framework (all equations, not just Equations 10 and 11) are new.
The strength of this framework is that it enables the results to be compared across antibodies, which was not possible with the approach in Doud et al (2017). 
}

In this work, they compare three broad and three narrow antibodies. The broad: FI6v3, binds stalk, neutralizes both group 1 and 2 HAs; C179 binds stalk, neutralizes only some group 1 HAs; S139/1, binds receptor-binding pocket, neutralizes both group 1 and group 2 HAs. The authors compare these in a re-analysis of data from narrow, strain-specific antibodies: H17-L19, H17-L10, and H17-L7, bind the Ca2, Ca1, and Cb antigenic regions on HA's globular head.

Using these three broad antibodies and WSN/33 Flu, they selected for escape mutations, and compared them to previous narrow antibody data. They find that escape from narrow or RBS-targeting Ab is relatively easy by single point mutation; while those targeting the stalk are difficult to escape by single point mutation. So breadth of antibody activity per se, is not a predictor of likelihood of escape.

They go on to show that the selected mutations are nearly all in or close to the antibody-binding footprint. While this may not be surprising, it is a nice set of data. Another nice aspect, is that the authors show that numerous different amino acid mutations at each site confer neutralization resistance - most studies tend to identify just one or two, rather than all of the possible aa.

They further show that the mutational antigenic profiling is highly predictive of the
results of the neutralization assays, even for small-effect mutations, as was previously shown for large-effect mutations from narrow antibody escape.

{\color{black}
The foregoing paragraphs are an accurate summary.
We think it is valuable to know that the mutational antigenic profiling also works well for small-effect mutations.
}

They compare the sites where mutations arise from broad antibody, to previous mutational scanning data that reveals whether these regions are tolerant or not to mutation. They find that this is only partially true, but cannot account for all lack of escape mutations; however, they do not uncover what the other reasons are.

Some questions or comments that come to mind:

1. If the rationale is to ask whether current circulating strains could escape immunity based on broad neutralising antibody elicited by vaccination, could the authors not use circulating strains for their experiments, or at least more lab strains that only WSN?

{\color{black}
We agree that it would be interesting to examine additional strains.
However, prior to our work, no study had completely mapped the effect of all single mutations to any viral strain with respect to neutralization by any broad antibodies---so doing this with one strain is a substantial step forward.
As mentioned in the response to question 2 by Reviewer \#1, we now more clearly explain that WSN is a lab-adapted strain.
}

2. What difference in terms of viral fitness is there between a mutation present in large proportion (for the narrow antibodies) vs present at low frequency (for the stalk-targeting antibodies). Figure 7 partly answers this question; however, it would be interesting to confirm the fitness of those mutants using more conventional growth curves, binding assays, etc.

{\color{black}
The neutralization assays in Figure~6 show that mutations for the stalk-targeting antibodies only have weak effects on neutralization, while Figure~7 shows that many of them don't have much effect on viral growth.
In addition, we were able to generate all the stalk-antibody escape mutants for the neutralization assays, so they are not too badly attenuated in cell culture.
However, we avoid making strong statements about the effects of mutations on viral growth beyond the data in Figure~7 because it is beyond the scope of our study to do growth curves for such a large number of mutations.}

3. Are all changes in amino-acid obtained after one nucleotide change? Are the authors able to detect multiple nucleotide or amino acid changes?

{\color{black}
Our study tested all amino-acid changes, including those that involved multiple-nucleotide changes.
The reason is that our libraries contained all codon mutations (e.g., the 63 non-wildtype codons at each site) rather than just all single nucleotide mutations.
We have added text that clarifies this point:
\begin{quote}
\textsl{Importantly, as described previously (Doud and Bloom 2016), these virus libraries were generated by mutagenizing HA at the \emph{codon} level rather than at the nucleotide level.
Performing codon mutagenesis is important, because single-nucleotide mutations access only about a third of the possible amino-acid mutations from a given codon, whereas codon mutations access all possible amino-acid mutations.}
\end{quote}
}

4. The major caveat of this study is that it only takes into account one mutation. Authors could check a combination of the mutations they obtain with their experiments (especially the ones escaping the stalk-targeting antibodies).

{\color{black}
We agree that this is an important caveat.
See our response to question 1 by Reviewer \#1.}

5. Figure 6 confirms some of the findings using neutralisation assays. As the authors state, it is unclear whether some of the observations actually exceed the noise of the assay. In particular, K(-8)T looks as far from WT as does K280A. It could be that (1) all these differences are due to noise or that (2) there is actually a biological difference between WT and K280A. Whether (1) occurs is unclear: ``three biological replicates'' should perhaps be more detailed in the Materials and Methods section as to what this means: are they three rows of the same plate performed on the same day, on different days, etc.? According to the authors' previous paper (Doud et al., 2017, PLOS Pathogens), they performed ``three replicate dilution columns''. The neutralisation curves seem therefore to have been performed at different days (besides, the WT curves are the same or similar to that of Figure 3, and the neutralisation curves from the narrow antibodies were imported from the authors' previous paper). This can induce a batch effect, which may not be linear, thus the normalisation performed (with no antibody and background noise) may not regress this out. From the standpoint of experimental design, these experiments should be repeated, with all viruses for a given antibody performed in the same batch; the triplicates will then give a more accurate measurement of noise. In general, all neutralisation experiments should be repeated, so as to make sure the findings are reproducible. If this has been done, it was not clear from the figure legend, and could be added for clarity. To address (2), authors should provide details of $F_{r,a}$ for K280A. If I understood correctly, because $F_{r,a}^{\rm{excess}}$ (defined in Equation 2) hides mutations with lower fitness than average, it is not possible for the reader to assess whether this mutation is ``neutral'' or it has lower fitness with the antibody under study.

{\color{black}
A similar point was made by several other reviewers.
We have provided a detailed response in answer to question 5 by Reviewer \#1.

Our revisions show the neutralization curves for all replicates and perform statistical testing.
All neutralization curves for antibody FI6v3 were performed on the same day.
This is important because as the reviewer notes, there can be batch effects.
If you look at Figure~6, you will see that the wildtype curves in panels A, B, and C are slightly different.
But each panel is for a single antibody, and all curves with each antibody were performed on the same day to avoid any confounding batch effects.
Therefore, we have already done what the reviewer suggests, and the new Figure~S9 now clearly explains this.

As described in the response to question 5 by Reviewer \#1, the new statistical analysis finds that the effect for K(-8)T is significant, whereas that for K280A is not.
The reviewer also suggests that it might be interesting to look at the $F_{r,a}$ value for K280A, since it might have reduced neutralization resistance.
Since the statistical analysis in the new Figure~S9 shows that the difference between K280A and wildtype is \emph{not} statistically significant, we do not comment on possible increased neutralization sensitivity or the $F_{r,a}$ for this mutation in the paper.
However, we have added a new supplementary file (File~S6) that gives all the $F_{r,a}$ values.
The value of $F_{r,a}$ for this mutation is 0.00382, which is slightly below the library average of 0.00955.
}

6. Could the authors detail, in the legend, how the curves of Figure 6 were fitted?

{\color{black}
For the quantitative analyses of IC50's, we fit four-parameter logistic curves, although we fixed the top value to one.
This is explained in the legend to the new Figure~S9, which reports the IC50's and shows the curves for individual replicates.
The computer code used to fit the curves is at \url{https://jbloomlab.github.io/dms_tools2/dms_tools2.neutcurve.html}.
}

\subsubsection*{Reviewer \#3 (Remarks to the Author):}

This paper identifies escape mutations against anti-influenza HA antibodies and quantifies their effects. The authors employed a high-throughput approach known as mutational antigenic profiling, which was developed in their previous study (1). The authors made a minor technical improvement in this study to enable a fairer comparison across profiling results from different antibodies. Two broadly neutralizing antibodies that target the HA stem (mouse C179 and human FI6v3), and one that targets the HA receptor-binding site (S139/1) were examined in this study. Additionally, the authors reanalyzed the results from their previous study (1) on three strain-specific antibodies (H17-L7, H17-L10, and H17-L19). The authors observed strong escape mutations against the strain-specific antibodies and broadly neutralizing antibody S139/1. In contrast, only weak escape mutations were identified against the two stem antibodies. The authors conclude that ``breadth is an imperfect indicator of the potential for viral escape''. The authors also conclude that broadly neutralizing antibodies targeting the H1 hemagglutinin stalk are ``quantifiably harder to escape than the other antibodies tested here''. The last two conclusions are not unexpected and are not particularly novel insights.

The study is, however, performed to an extremely high technical standard. Nevertheless, the impact is limited. The improvement in methodology over their previous study (1) is minor. Also, it is known that strong escape mutations against S139/1, one of the broadly neutralizing antibodies tested in this study (although not nearly as broad as FI6v3), can be generated rapidly through passaging the virus in the presence of the antibody (2). In fact, strong escape mutations against some of the stalk antibodies that were not tested in this study, such as CR8020 (3), CR8043 (4), CR6261 (5) and 39.29 (6), have also been generated through single point mutations. Overall, the impact and novelty of the study is not sufficient for publication in Nature Microbiology.

{\color{black}
We are disappointed that the reviewer does not view the work of sufficient impact for \textit{Nature Microbiology}, but appreciate the statement that the study is ``performed to an extremely high technical standard.''
We feel that direct quantitative comparison of the escape from narrow and broad antibodies is valuable, since other comparisons have been more anecdotal.
We also note that at least two of the reviewers found the work of high impact.

Thank you for the references. 
Interestingly, (3), (4), and (6) all describe selecting escape from \emph{group 2} HAs (e.g., H3 or H7), not a group 1 HA like the one we used.
We now cite these references and mention this point as described in the response to question 7 by Reviewer \#1.
It is true that (5) does mention an escape mutation from a group 1 HA from CR6261, but it never actually shows any neutralization data for that mutation, so we are unable to comment on it.
}

Specific comments:

1.  Strong escape mutations to C179 have been identified in the original paper that discovered the antibody (7). But it seems like those mutations were not identified here. Can the authors provide some explanations for such a discrepancy?

{\color{black}
The original paper on C179 selected a strong escape mutant in an H1 background.
That virus had three mutations, and although the paper argues that only one is important, it never shows that the others are not required as secondary changes.
We cite this paper, and use it to inform the discussion of multiple mutants described in response to question 1 from Reviewer \#1.
}

2.  How is the curve fitting performed for the neutralization assay results? I do not seem to see any description for the curve-fitting method.

{
\color{black}
This point was also mentioned by other reviewers.
See our response to question 5 of Reviewer \#1 and questions 5 and 6 of Reviewer \#2.}

References

1.  Doud MB, Hensley SE, Bloom JD. Complete mapping of viral escape from neutralizing antibodies. PLoS Pathog. 2017;13(3):e1006271.

2.  Yoshida R, Igarashi M, Ozaki H, Kishida N, Tomabechi D, Kida H, et al. Cross-protective potential of a novel monoclonal antibody directed against antigenic site B of the hemagglutinin of influenza A viruses. PLoS Pathog. 2009;5(3):e1000350.

3.  Ekiert DC, Friesen RH, Bhabha G, Kwaks T, Jongeneelen M, Yu W, et al. A highly conserved neutralizing epitope on group 2 influenza A viruses. Science. 2011;333(6044):843-50.

4.  Friesen RH, Lee PS, Stoop EJ, Hoffman RM, Ekiert DC, Bhabha G, et al. A common solution to group 2 influenza virus neutralization. Proc Natl Acad Sci USA. 2014;111(1):445-50.

5.  Throsby M, van den Brink E, Jongeneelen M, Poon LL, Alard P, Cornelissen L, et al. Heterosubtypic neutralizing monoclonal antibodies cross-protective against H5N1 and H1N1 recovered from human IgM+ memory B cells. PLoS One. 2008;3(12):e3942.

6.  Chai N, Swem LR, Reichelt M, Chen-Harris H, Luis E, Park S, et al. Two escape mechanisms of influenza A virus to a broadly neutralizing stalk-binding antibody. PLoS Pathog. 2016;12(6):e1005702.

7.  Okuno Y, Isegawa Y, Sasao F, Ueda S. A common neutralizing epitope conserved between the hemagglutinins of influenza A virus H1 and H2 strains. J Virol. 1993;67(5):2552-8.

{\color{black}
Thanks for these references, they are now all cited.
}

\subsubsection*{Reviewer \#4 (Remarks to the Author):}

This paper tackles a really nice question: Are some antibodies broad because they target sites that face little immune pressure, or are they broad because the epitopes they target cannot easily mutate to escape them? Put differently, are conserved epitopes that are targeted by some antibodies conserved because they've not had to change, or are they conserved because no beneficial changes are accessible? This question is obviously relevant for ``universal'' vaccine development, but it's also important for understanding strain structure and evolution more generally.

To investigate this question, the authors analyzed the ability of a H1 flu strain to escape six well characterized antibodies via single mutations. They showed that for most antibodies, including a broadly neutralizing antibody, single mutations allowed escape from antibody detection. In the other two antibodies, no mutations really rescued the virus. By comparing the mutations conferring antibody escape to those that are generally tolerated during replication, the authors further showed that mutational tolerance does not fully explain failure to escape: despite substitutions, some antibodies are harder to escape than others.

These are important observations. My only real criticism of the work is that I found the description of mutational effects confusing as early as the abstract. Does the measured frequency of antibody escape mutants reflect only antibody selection, or does it also include selection for replicative capacity? The distinction between antibody selection and general fitness is not made until p. 13, but I think readers (especially those familiar with other work from the lab) might benefit if these distinctions were made in the introduction and review of the methods. My take is that the fraction surviving antibody escape includes both, in that non-viable mutants were rare in the initial library (again implied on p. 13), but this could be clarified. It seems the contribution of each might be quantifiable, or at the relative mutational tolerances of the different epitopes compared.

{\color{black}
Thanks for the kind comments on the importance of the paper and the questions it addresses.

The point about clearly distinguishing between how mutations affect antigenicity and viral growth is well taken.
We have added text at several places to clarify this.
For instance, as soon as we define our measures of antigenic effects, we say:
\begin{quote}
\textsl{Importantly, Equations~1 and 2 correct for effects on viral growth due to normalization by the mock-selected control, and so measure only antigenicity and not viral growth provided that the virus at least grows well enough to be present in the library.}
\end{quote}}

Minor points:

1. How do the concentrations for selecting escape mutants in vitro compare to likely concentrations in vivo? Concentrations may be too high or low in vivo to promote adaptation.

{\color{black}
We are unable to answer this question, because the concentrations of any given antibody specificity can vary widely among people.
In addition, it is difficult to relate \textit{in vitro} neutralization to \textit{in vivo} activity due to Fc-mediated effector functions.
This is a limitation of virtually all work on antibody neutralization.
We mention this caveat in the Discussion:
\begin{quote}
\textsl{When interpreting the magnitude of these effects, it is important to note that our experiments only measure how mutations affect neutralization, and not how they affect Fc-mediated functions that are responsible for much of the in vivo protection afforded by anti-stalk antibodies.}
\end{quote}
}

2. Outside these experiments, how much data is there on the antibodies' neutralization capacities within HA subtypes? i.e., how broad or narrow do we really know them to be? Obviously such data might be complementary.

{\color{black}
Unfortunately, most of the antibodies have been characterized against only a small number of the many strains within HA subtypes.
We summarize the subtype-level information in Figure 2B, and cite all studies that look at breadth within subtypes (references 24, 25, 29, 45, 46).
However, because different studies use different strains, the data in these studies do not enable useful comparisons across antibodies.
Our study only looks at one strain---but at least it is complete for the mutations to this HA, and so can be compared across antibodies.}

3. Additional suggested references:

``the viruses's rapid antigenic evolution erodes the effectiveness of this immunity to that strain's descendants within four to seven years...'' -- Kucharski et al. PLOS Biol 2015 (although numbers might need revising)

{\color{black}
This is a good reference, and we now cite it.
The distribution of infection frequencies they derive from epidemiology is similar to the 4-7 years we had previously cited from a human challenge study.
However, we have updated the text to ``$\sim$5 years'' to better reflect the ambiguity.}

on immune status driving the rate of antigenic evolution -- Wen et al. Proc B 2016

{\color{black}
This is a good reference, and we now cite it at the relevant point.}

4. In ``...we expect a larger fraction of a library to survive an antibody with many escape mutations...'' (and similar phrases), the ``many escape mutations'' read as though they modify ``antibody.''

{\color{black}
This is a good point, we have clarified the wording in these phrases.}

\end{document}

### 2018-02-21 paper/ResponseToFinalNatureCommunicationsReviews.tex
added reviewer response: very brief, as reviews only asked for one minor thing

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format

\usepackage[margin=1.1in]{geometry}

\usepackage{hyperref}
%\hypersetup{colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue}

\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
\usepackage{color}
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

\title{Response to reviews of ``Quantifying the effects of single mutations on viral escape from broad and narrow antibodies to an H1 influenza hemagglutinin'' from \textit{Nature Communications}}
\author{Michael B. Doud, Juhye M. Lee, and Jesse D. Bloom}

\begin{document}
\maketitle

\subsection*{Overview of reviewer response and changes to the manuscript.}

The original manuscript was reviewed for \textit{Nature Microbiology} by four reviewers.
They made a number of helpful suggestions, and the revised manuscript was transferred to \textit{Nature Communications}.

Three of the four reviewers then provided additional feedback on that revised manuscript. 
Here we respond to their feedback and describe the changes we have made.

\subsection*{Response to reviews}
Below find our responses to the reviews received from \textit{Nature Communications}.
The original comments {\color{blue} are in blue}, and our responses are in black.

\color{blue}

\subsection*{Reviewer \#1 (Remarks to the Author):}

The authors addressed all point sufficiently.

{\color{black}
Thanks---your points in the original review were very helpful.}

\subsection*{Reviewer \#2 (Remarks to the Author):}

I had previously reviewed this article for Nature Microbiology. The authors have very justly considered the points raised by all the reviewers and addressed points in an exemplary manner. The key points that I consider have improved the work or its presentation:

The authors clarified and emphasized in the text the limitations of examining only single point mutations.

They agree on the caveat of using the WSN strain, and discuss it.

Most importantly, they provide more complete data to accompany one of the major figures, in figure S9.

It was my opinion and impression of other reviewer comments that the lack of testing using a large panel of viral strains was missing from the original submission. However, considering the transferred manuscript, I do not think that this should be counted against the current submission.

The paper remains technically strong, well written and statistically sound as the original, and improved by added data and discussion.

{\color{black}
Thanks---we appreciate your original suggestion to show all the neutralization data, and are glad that you like the new display in Figure S9.
}

\subsection*{Reviewer \#3 (Remarks to the Author):}

Comment:

The authors have responded well to my comments from the previous round of review. However, there is one concern that remains unaddressed.

In the response, the authors claim that ?It is true that [Throsby et al. 2008] does mention an escape mutation from a group 1 HA from CR6261, but it never actually shows any neutralization data for that mutation, so we are unable to comment on it.? In Throsby et al. 2008, a group 1 (H5) mutant H111L was confirmed to be resistant to CR6261 (IC50>100 µg/ml vs. 6.25 µg/ml for wild type). The author should still acknowledge this finding and cite Throsby et al. 2008.

{\color{black}
We apologize for the oversight about this mutation to an H5 HA in Throsby et al.
We have added a sentence discussing the Throsby et al result in the Discussion.
}

\end{document}

### 2018-02-21 paper/manuscript.tex
added reviewer response: very brief, as reviews only asked for one minor thing

@@
 
-The correlations among replicates of the mutational antigenic profiling in terms of the measured fraction-surviving above average for each possible amino-acid mutations are shown in Figure~\ref{suppfig:corr}.
+The correlations among replicates of the mutational antigenic profiling, in terms of the measured fraction-surviving above average for each possible amino-acid mutation, are shown in Figure~\ref{suppfig:corr}.
 For the remainder of this paper, we will refer to the median antigenic effect of each mutation across replicates.
@@
 However, our approach shows that there are mutations that have more modest ($<$10-fold) effects on neutralization by even the broadest antibody. 
-Interestingly, most previous studies~\cite{chai2016two,ekiert2011highly,friesen2014common,dunand2015preexisting} that have reported selecting single mutations with large effects ($\gg$10-fold) on neutralization by anti-stalk antibodies have used group 2 (e.g., H3 or H7) HAs rather than group 1 HAs like the one used in our work.
-When interpreting the magnitude of these effects, it is important to note that our experiments only measure how mutations affect neutralization, and not how they affect Fc-mediated functions that are responsible for much of the \textit{in vivo} protection afforded by anti-stalk antibodies\cite{dilillo2014broadly,dilillo2016broadly}.
+Interestingly, most previous studies~\cite{chai2016two,ekiert2011highly,friesen2014common,dunand2015preexisting} that have reported selecting single mutations with large effects ($\gg$10-fold) on neutralization by anti-stalk antibodies have used group 2 (e.g., H3 or H7) HAs rather than group 1 HAs like the one used in our work---although at least one study has selected large-effect escape mutations to broad anti-stalk antibody in an H5 group 1 HA~\cite{throsby2008heterosubtypic}.
+In addition, when interpreting the magnitude of the effects measured in our experiments, it is important to note that we are only assessing how mutations affect neutralization, and not how they affect Fc-mediated functions that are responsible for much of the \textit{in vivo} protection afforded by anti-stalk antibodies\cite{dilillo2014broadly,dilillo2016broadly}.
 

### 2018-02-22 paper/ResponseToFinalNatureCommunicationsReviews.tex
made all editor requested revisions except to supp material and moving figures to end

@@
 
-\subsection*{Overview of reviewer response and changes to the manuscript.}
+\subsection*{Overview of revisions to the manuscript made in response to comments by reviewers and the editor.}
 
@@
 Three of the four reviewers then provided additional feedback on that revised manuscript. 
-Here we respond to their feedback and describe the changes we have made.
+Here we respond to their feedback and describe the changes we have made in response to these reviewer comments.
+We then describe the changes that we have made in response to the editor's comments.
 
@@
 
-\subsection*{Reviewer \#1 (Remarks to the Author):}
+\subsubsection*{Reviewer \#1 (Remarks to the Author):}
 
@@
 
-\subsection*{Reviewer \#2 (Remarks to the Author):}
+\subsubsection*{Reviewer \#2 (Remarks to the Author):}
 
@@
 
-\subsection*{Reviewer \#3 (Remarks to the Author):}
+\subsubsection*{Reviewer \#3 (Remarks to the Author):}
 
@@
 
+\subsection*{Editor comments:}
+Comments by the editor were provided in a marked PDF.
+
+{\color{black}
+We have made the following changes in response to these editor comments:
+\begin{itemize}
+\item We have shortened the title from ``Quantifying the effects of single mutations on viral escape from broad and narrow antibodies to an H1 influenza hemagglutinin'' (19 words) to ``How single mutations affect viral escape from broad and narrow antibodies to H1 influenza hemagglutinin'' (15 words).
+\item We have provided the full address for each affiliation as requested.
+\item We have shortened the abstract from the original length of 160 words to the requested length of 150 words.
+\item We have shortened all subheadings in the RESULTS and METHODS to no more than 60 characters.
+\item {\color{red} We have provided publication-quality versions of all figures, and moved the legends to the end of the manuscript. STILL NEEDS TO BE DONE.}
+\item The naming of supplementary figures has been changed from ``Figure S1'' to ``Supplementary Fig. 1''. Similar changes have been made for supplementary tables and data as requested.
+\item We have clarified that the error bars in Figure 6 represent standard errors.
+\item We have added the source of the 293F cells.
+The 293T-CMV-PB1 and MDCK-SIAT1-CMV-PB1 cells were generated in the cited reference 70, the pHW* series plasmids were generated in the cited reference 71, and the pHH-PB1flank-eGFP plasmid was generated in the cited reference 70.
+\item We have expanded the Data Availability section of the METHODS as requested. 
+\item We have added a conflict of interest statement under the heading COMPETING INTERESTS. We have no conflicts.
+\item {\color{red} Still need to make adjustments to supporting material.}
+\end{itemize}
+
+}
+
 \end{document}

### 2018-02-22 paper/manuscript.tex
made all editor requested revisions except to supp material and moving figures to end

@@
 \usepackage{newfloat}
-\DeclareFloatingEnvironment[name={Figure}]{suppfigure}
-\renewcommand{\thesuppfigure}{S\arabic{suppfigure}}
-\DeclareFloatingEnvironment[name={Table}]{supptable}
-\renewcommand{\thesupptable}{S\arabic{supptable}}
-\DeclareFloatingEnvironment[name={File}]{suppfile}
-\renewcommand{\thesuppfile}{S\arabic{suppfile}}
+\DeclareFloatingEnvironment[name={Supplementary Fig.}]{suppfigure}
+\renewcommand{\thesuppfigure}{\arabic{suppfigure}}
+\DeclareFloatingEnvironment[name={Supplementary Table}]{supptable}
+\renewcommand{\thesupptable}{\arabic{supptable}}
+\DeclareFloatingEnvironment[name={Supplementary Data}]{suppdata}
+\renewcommand{\thesuppdata}{\arabic{suppdata}}
 
@@
 
-\title{Quantifying the effects of single mutations on viral escape from broad and narrow antibodies to an H1 influenza hemagglutinin}
+\title{How single mutations affect viral escape from broad and narrow antibodies to H1 influenza hemagglutinin}
 
@@
 \\
-\scriptsize{$^1$Basic Sciences and Computational Biology Program, Fred Hutchinson Cancer Research Center}\\
-\scriptsize{$^2$Department of Genome Sciences and $^3$Medical Scientist Training Program, University of Washington} \\
-\scriptsize{Seattle, WA, USA} \\
+\scriptsize{$^1$Basic Sciences and Computational Biology, Fred Hutchinson Cancer Research Center, 1100 Fairview Ave N, Seattle, WA  98109}\\
+\scriptsize{$^2$Department of Genome Sciences, University of Washington, 3720 15th Ave NE, Seattle, WA  98195} \\
+\scriptsize{$^3$Medical Scientist Training Program, University of Washington, 1959 NE Pacific Street, Seattle, WA  98195} \\
 \scriptsize{$^{\dagger}$These authors contributed equally} \\
@@
 \begin{abstract}
-Influenza virus can completely escape most antibodies with single mutations.
+Influenza virus can escape most antibodies with single mutations.
 However, rare antibodies broadly neutralize many viral strains.
-It is unclear how easily influenza virus might escape such antibodies if it was under strong pressure to do so.
-Here we map all single amino-acid mutations that increase resistance to broad antibodies targeting an H1 hemagglutinin.
-Crucially, our approach not only identifies antigenic mutations but also quantifies their effect sizes.
+It is unclear how easily influenza virus might escape such antibodies if there was strong pressure to do so.
+Here we map all single amino-acid mutations that increase resistance to broad antibodies to H1 hemagglutinin.
+Our approach not only identifies antigenic mutations but also quantifies their effect sizes.
 All antibodies select mutations, but the effect sizes vary widely. 
-The virus can escape a broad antibody that targets residues in hemagglutinin's receptor-binding site the same way it escapes narrow strain-specific antibodies: via single mutations with huge effects.
-In contrast, broad antibodies targeting hemagglutinin's stalk only select mutations with small effects. 
-Therefore, among the antibodies we have examined, breadth is an imperfect indicator of the potential for viral escape via single mutations.
-Broadly neutralizing antibodies targeting the H1 hemagglutinin stalk are quantifiably harder to escape than the other antibodies tested here.
+The virus can escape a broad antibody to hemagglutinin's receptor-binding site the same way it escapes narrow strain-specific antibodies: via single mutations with huge effects.   
+In contrast, broad antibodies to hemagglutinin's stalk only select mutations with small effects. 
+Therefore, among the antibodies we examine, breadth is an imperfect indicator of the potential for viral escape via single mutations.
+Antibodies targeting the H1 hemagglutinin stalk are quantifiably harder to escape than the other antibodies tested here.
 \end{abstract}
@@
 
-\subsection*{An approach to quantify the fraction of virions with each mutation that escape antibody neutralization}
+\subsection*{Fraction of each viral mutant that escapes neutralization}
 We can visualize the outcome of antibody selection on viral populations containing antigenic mutations as in Figure~\ref{fig:fracsurvive_example}.
@@
 There are multiple sites of large-effect mutations for H17L19, H17L10, H17L7, and S139/1---but none for FI6v3 and C179.
-Figure~\ref{suppfig:maxfracsurvive} shows the excess fraction surviving for the largest-effect mutation at each site.
-Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo} show all mutations using logo plots.
+Supplementary Fig.~\ref{suppfig:maxfracsurvive} shows the excess fraction surviving for the largest-effect mutation at each site.
+Supplementary Figs.~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo} show all mutations using logo plots.
 Sites are labeled in H3 numbering.
@@
 
-\subsection*{Quantifying the antigenic effects of all mutations selected by each antibody}
+\subsection*{The effects of all mutations on antibody neutralization}
 We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}. 
@@
 
-The correlations among replicates of the mutational antigenic profiling, in terms of the measured fraction-surviving above average for each possible amino-acid mutation, are shown in Figure~\ref{suppfig:corr}.
+The correlations among replicates of the mutational antigenic profiling, in terms of the measured fraction-surviving above average for each possible amino-acid mutation, are shown in Supplementary Fig.~\ref{suppfig:corr}.
 For the remainder of this paper, we will refer to the median antigenic effect of each mutation across replicates.
@@
 For all four of these antibodies, there are multiple sites in HA where mutations enable a substantial fraction of virions to survive high antibody concentrations (Figure~\ref{fig:avgfracsurvive}).
-Specifically, there are mutations that enable over a third of virions to survive at concentrations where virtually all wildtype virions are neutralized (Figure~\ref{suppfig:maxfracsurvive}).
+Specifically, there are mutations that enable over a third of virions to survive at concentrations where virtually all wildtype virions are neutralized (Supplementary Fig.~\ref{suppfig:maxfracsurvive}).
 Therefore, the virus can escape these four antibodies with the sort of large-effect single amino-acid mutations that characterize traditional influenza antigenic drift\cite{yewdell1979antigenic,webster1980determination,koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction}.
@@
 In contrast, the stalk-targeting antibodies C179 and FI6v3 select no strong escape mutants. 
-If we look at the results for these antibodies on the same scale as the other antibodies, we see only a few small bumps in the fraction of virions surviving (Figures~\ref{fig:avgfracsurvive} and \ref{suppfig:maxfracsurvive}).
-Only if we zoom in can we see that there are actually a few sites where mutations slightly increase the fraction of virions surviving C179 and FI6v3 (Figures~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo}).
+If we look at the results for these antibodies on the same scale as the other antibodies, we see only a few small bumps in the fraction of virions surviving (Figures~\ref{fig:avgfracsurvive} and Supplementary Fig.~\ref{suppfig:maxfracsurvive}).
+Only if we zoom in can we see that there are actually a few sites where mutations slightly increase the fraction of virions surviving C179 and FI6v3 (Supplementary Figs.~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo}).
 But the effect sizes of these antigenic mutations are tiny compared to the other antibodies---especially for FI6v3.
@@
 
-\subsection*{The selected mutations are near the binding footprints of the antibodies}
+\subsection*{Selected mutations are near antibody binding footprints}
 Antigenic mutations selected by narrow strain-specific antibodies against HA are thought to occur at residues in or near the physical binding footprint of the antibody\cite{yewdell1979antigenic,webster1980determination,caton1982antigenic}.
@@
 
-However, C179 selects several mutations that do not have any apparent effect on FI6v3 (Figure~\ref{fig:structures}A, Figure~\ref{suppfig:FI6v3logo}).
+However, C179 selects several mutations that do not have any apparent effect on FI6v3 (Figure~\ref{fig:structures}A, Supplementary Fig.~\ref{suppfig:FI6v3logo}).
 The most notable of these C179-specific mutations are at site 38.
@@
 
-\subsection*{Neutralization assays validate the mutational antigenic profiling}
+\subsection*{Validation by neutralization assays}
 Do the mutations identified in our mutational antigenic profiling actually have the expected effect on antibody neutralization?
@@
 (A) Neutralization curves of individual viral mutants with FI6v3.
-The mutations K280S, K280T, N291S, G47R (HA2), and K(-8)T are all expected to increase neutralization resistance based on the mutational antigenic profiling (Figure~\ref{fig:structures}A), whereas K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization (Figure~\ref{suppfig:FI6v3logo}).
+The mutations K280S, K280T, N291S, G47R (HA2), and K(-8)T are all expected to increase neutralization resistance based on the mutational antigenic profiling (Figure~\ref{fig:structures}A), whereas K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization (Supplementary Fig~\ref{suppfig:FI6v3logo}).
 All neutralization curves in this panel were performed in triplicate on the same day.
-This panel shows the average of the replicates; Figure~\ref{suppfig:FI6v3replicates} shows the curves for each replicate individually and performs statistical testing of whether the IC50s for mutants are significantly different than for wildtype.
+This panel shows the average of the replicates; Supplementary Fig.~\ref{suppfig:FI6v3replicates} shows the curves for each replicate individually and performs statistical testing of whether the IC50s for mutants are significantly different than for wildtype.
 (B), (C) In contrast to FI6v3, mutations selected by narrow antibodies have very large effects on neutralization.
 Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 taken from Doud et al\cite{doud2017complete}.
+Points indicate mean and standard error of three replicates.
 }
@@
 As discussed in the previous section, the three mutations most strongly selected by FI6v3 introduce glycosylation motifs at sites 278-280 or 289-291 (Figure~\ref{fig:structures}A,B).
-We created viruses carrying each of these mutations (K280S, K280T, and N291S) and validated that all three modestly but significantly increased resistance to FI6v3 (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates}).
-As a control, we also validated that a mutation at one of these sites (K280A) that does \emph{not} have an effect in our mutational antigenic profiling does not significantly shift the neutralization curve (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates}).
+We created viruses carrying each of these mutations (K280S, K280T, and N291S) and validated that all three modestly but significantly increased resistance to FI6v3 (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~\ref{suppfig:FI6v3replicates}).
+As a control, we also validated that a mutation at one of these sites (K280A) that does \emph{not} have an effect in our mutational antigenic profiling does not significantly shift the neutralization curve (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~\ref{suppfig:FI6v3replicates}).
 
 Our mutational antigenic profiling also identified several non-glycosylation-motif mutations that were selected by FI6v3.
-We validated that one of these mutations, G47R in the HA2 chain, significantly increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates})---although as predicted by the mutational antigenic profiling, the magnitude of the effect was small.
+We validated that one of these mutations, G47R in the HA2 chain, significantly increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~\ref{suppfig:FI6v3replicates})---although as predicted by the mutational antigenic profiling, the magnitude of the effect was small.
 The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.
-We tested one of these mutations, K(-8)T, and it did lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates})---although despite the significance testing in Figure~\ref{suppfig:FI6v3replicates}, we remain circumspect about the magnitude of this effect relative to the noise in our neutralization assays.
-As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them significantly affected neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Figure~\ref{suppfig:FI6v3replicates}).
+We tested one of these mutations, K(-8)T, and it did lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~\ref{suppfig:FI6v3replicates})---although despite the significance testing in Supplementary Fig.~\ref{suppfig:FI6v3replicates}, we remain circumspect about the magnitude of this effect relative to the noise in our neutralization assays.
+As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them significantly affected neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~\ref{suppfig:FI6v3replicates}).
 
@@
 
-\subsection*{Limited HA mutational tolerance does not fully explain the lack of strong escape mutations from the anti-stalk antibodies}
+\subsection*{HA mutational tolerance and antibody escape}
 Why are there no large-effect escape mutations from the anti-stalk antibodies?
@@
 C179 IgG was purchased from Takara Bio Inc (Catalog \#M145).
-FI6v3 was purified from 293F cells transduced with a lentiviral vector encoding a commercially synthesized gene for the IgG form of the antibody, with the heavy and light chains reverse-translated from the protein sequence in the PDB structure 3ZTN\cite{corti2011neutralizing} as described previously\cite{balazs2013broad}.
+FI6v3 was purified from 293F cells (ThermoFisher R79007) transduced with a lentiviral vector encoding a commercially synthesized gene for the IgG form of the antibody, with the heavy and light chains reverse-translated from the protein sequence in the PDB structure 3ZTN\cite{corti2011neutralizing} as described previously\cite{balazs2013broad}.
 Genes encoding S139/1 in IgG form were were reverse-translated from the protein sequence in PDB structure 4GMS\cite{lee2012heterosubtypic}, and used to express and purify protein by the Fred Hutchinson Cancer Research Center protein expression core.
@@
 We performed neutralization assays using influenza viruses that carried GFP in the PB1 segment.
-These PB1flank-eGFP were generated in co-cultures of 293T-CMV-PB1 and MDCK-SIAT1-CMV-PB1 cells as described previously\cite{bloom2010permissive}, using the standard bi-directional pHW181-PB2, ..., pHW188-NS reverse-genetics plasmids\cite{hoffmann2000dna} for all genes \emph{except} PB1, plus the pHH-PB1flank-eGFP plasmid\cite{bloom2010permissive}.
+These PB1flank-eGFP were generated in co-cultures of 293T-CMV-PB1 and MDCK-SIAT1-CMV-PB1 cells as described previously\cite{bloom2010permissive}, using the standard bi-directional pHW181-PB2, pHW182-PB1, pHW183-PA, pHW184-HA, pHW185-NP, pHW186-NA, pHW187-M and pHW188-NS reverse-genetics plasmids\cite{hoffmann2000dna} for all genes \emph{except} PB1, plus the pHH-PB1flank-eGFP plasmid\cite{bloom2010permissive}.
 Each mutant was generated by repeating this process using a version of the pHW184-HA plasmid that had been engineered by site-directed mutagenesis to carry the indicated mutation.
 The neutralization assays themselves were performed by using a plate reader to quantify the GFP signal produced by MDCK-SIAT1-CMV-PB1 cells infected by PB1flank-eGFP virus that had been incubated with the indicated antibody concentration as described previously\cite{hooper2013mutant}.
-All neutralization curves in Figure~\ref{fig:FI6v3neutcurves}A represent the mean and standard deviation of three measurements, with the individual replicates shown in Figure~\ref{suppfig:FI6v3replicates}.
+All neutralization curves in Figure~\ref{fig:FI6v3neutcurves}A represent the mean and standard deviation of three measurements, with the individual replicates shown in Supplementary Fig.~\ref{suppfig:FI6v3replicates}.
 All the neutralization assays for FI6v3 were performed on the same day to eliminate batch effects, with each replicate involving independent serial dilution of the antibody in a separate column of a 96-well plate.
@@
 The conversion between sequential numbering of the A/WSN/1933 HA and the H3 numbering scheme was performed using the Python script available at \url{https://github.com/jbloomlab/HA_numbering}.
-File~\ref{suppfile:HAnumbering} gives the numbering conversion.
+Supplementary Data~\ref{suppdata:HAnumbering} gives the numbering conversion.
 
@@
 To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database\cite{zhang2017influenza}, inferred the phylogenetic tree using RaxML\cite{stamatakis2014raxml} with a GTR model, and visualized the tree using FigTree (\url{http://tree.bio.ed.ac.uk/software/figtree/}).
-The HA sequences used are in File~\ref{suppfile:HAsubtypes}. 
+The HA sequences used are in Supplementary Data~\ref{suppdata:HAsubtypes}. 
 In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize\cite{yoshida2009cross, lee2012heterosubtypic, okuno1993common, dreyfus2013structure, corti2011neutralizing}.
@@
 For the cladogram in Figure~\ref{fig:structures}C, the amino-acid identities at site 38 are from the strains tested against C179 by by Dreyfus et al\cite{dreyfus2013structure}. 
-For subtypes not tested, the amino-acid identity reported is that in the strain for that subtype in File~\ref{suppfile:HAsubtypes}.
+For subtypes not tested, the amino-acid identity reported is that in the strain for that subtype in Supplementary Data~\ref{suppdata:HAsubtypes}.
 
@@
 We performed three fully independent replicates of each selection using the three replicate mutant virus libraries.
-In addition, we performed technical replicates (independent neutralization experiments on the \emph{same} virus library) in some cases as indicated in Figure~\ref{suppfig:corr}.
+In addition, we performed technical replicates (independent neutralization experiments on the \emph{same} virus library) in some cases as indicated in Supplementary Fig.~\ref{suppfig:corr}.
 The virus-antibody mixtures were used to infect cells, and viral RNA was extracted, reverse-transcribed, and PCR amplified as described previously\cite{doud2017complete}.
@@
 These fractions are denoted by $\gamma$ in this paper.
-The average of these fractions across libraries are reported in Figure~\ref{fig:neutcurves}, and the values for each individual replicate are in Table~\ref{supptab:fracsurvive}.
+The average of these fractions across libraries are reported in Figure~\ref{fig:neutcurves}, and the values for each individual replicate are in Supplementary Table~\ref{supptab:fracsurvive}.
 The fractions were estimated using qRT-PCR against the viral NP and canine GAPDH as described previously\cite{doud2017complete}.
@@
 The deep sequencing data were analyzed using version 2.2.1 of the \texttt{dms\_tools2} software package\cite{bloom2015software}, which is available at \url{http://jbloomlab.github.io/dms_tools2}.
-File~\ref{suppfile:analysis_code} contains a Jupyter notebook that performs all steps of the analysis beginning with downloading the FASTQ files from the Sequence Read Archive.
-Detailed statistics about the sequencing depth and error rates are shown in this Jupyter notebook and its HTML rendering in File~\ref{suppfile:analysis_html}.
+Supplementary Data~\ref{suppdata:analysis_code} contains a Jupyter notebook that performs all steps of the analysis beginning with downloading the FASTQ files from the Sequence Read Archive.
+Detailed statistics about the sequencing depth and error rates are shown in this Jupyter notebook and its HTML rendering in Supplementary Data~\ref{suppdata:analysis_html}.
 
-\subsection*{Calculating the fraction of virions with each mutation that escapes antibody neutralization}
+\subsection*{Calculating fraction of mutant that survive neutralization}
 In prior mutational antigenic profiling work\cite{doud2017complete,dingens2017comprehensive}, we calculated the differential selection on each mutation as the logarithm of its enrichment relative to wildtype in an antibody-selected sample versus a mock-selected control.
@@
 
-In this paper, Figures~\ref{fig:avgfracsurvive} and \ref{suppfig:maxfracsurvive} show the median of excess fraction surviving taken across all biological and technical replicates at a given antibody concentration (Equation~\ref{eq:fracsurvive_excess}).
+In this paper, Supplementary Figs.~\ref{fig:avgfracsurvive} and \ref{suppfig:maxfracsurvive} show the median of excess fraction surviving taken across all biological and technical replicates at a given antibody concentration (Equation~\ref{eq:fracsurvive_excess}).
 The subsequent logo plots show the medians of these values taken across all concentrations for each antibody.
-The numerical values plotted in these logo plots are in File~\ref{suppfile:fracsurvive_excess}.
-The fraction surviving values \emph{not} adjusted to be in excess of the library average (Equation~\ref{eq:fracsurvive}) are in File~\ref{suppfile:fracsurvive}.
+The numerical values plotted in these logo plots are in Supplementary Data~\ref{suppdata:fracsurvive_excess}.
+The fraction surviving values \emph{not} adjusted to be in excess of the library average (Equation~\ref{eq:fracsurvive}) are in Supplementary Data~\ref{suppdata:fracsurvive}.
 
@@
 \subsection*{Data availability and source code}
-Deep sequencing data are available from the Sequence Read Archive under BioSample accession SAMN05789126.
-Computer code that analyzes these data to generate all the results described in this paper is in File~\ref{suppfile:analysis_code}, and an HTML version of the analysis notebook is in File~\ref{suppfile:analysis_html}.
+Deep sequencing data are available from the Sequence Read Archive under BioSample accession SAMN05789126 at \url{https://www.ncbi.nlm.nih.gov/sra/?term=SAMN05789126}.
+Computer code that analyzes these data to generate all the results described in this paper is in Supplementary Data~\ref{suppdata:analysis_code}, and an HTML version of the analysis notebook is in Supplementary Data~\ref{suppdata:analysis_html}.
+In addition, all of this code as well as the manuscript itself and other data are available on GitHub at \url{https://github.com/jbloomlab/HA_antibody_ease_of_escape}.
+Finally, the \texttt{dms\_tools2} software\cite{bloom2015software} that performs most of the analysis is available at \url{https://jbloomlab.github.io/dms_tools2/}.
+The authors declare that all other data supporting the findings of this study are available within the article and its Supplementary Information files, or are available from the authors upon request
 
@@
 All three authors designed the project, contributed to the computer code, analyzed the data, and wrote the paper.
+
+\subsection*{COMPETING INTERESTS}
+The authors declare no competing financial interests.
 
@@
 
-\begin{suppfile}
-\caption{\label{suppfile:HAnumbering}
+\begin{suppdata}
+\caption{\label{suppdata:HAnumbering}
 {\bf Conversion from sequential numbering of the A/WSN/1933 HA to H3 numbering.}
@@
 }
-\end{suppfile}
+\end{suppdata}
 
-\begin{suppfile}
-\caption{\label{suppfile:HAsubtypes}
+\begin{suppdata}
+\caption{\label{suppdata:HAsubtypes}
 {\bf Sequences used to infer the tree for all HA subtypes.}
@@
 }
-\end{suppfile}
+\end{suppdata}
 
-\begin{suppfile}
-\caption{\label{suppfile:analysis_code}
+\begin{suppdata}
+\caption{\label{suppdata:analysis_code}
 {\bf Computer code and data for the analysis of the mutational antigenic profiling data.}
@@
 The actual analysis is performed by the Jupyter notebook \texttt{analysis\_notebook.ipynb}, which includes embedded plots summarizing key statistics and results.
-An HTML version of this notebook is also included as File~\ref{suppfile:analysis_code}.
+An HTML version of this notebook is also included as Supplementary Data~\ref{suppdata:analysis_code}.
 }
-\end{suppfile}
+\end{suppdata}
 
-\begin{suppfile}
-\caption{\label{suppfile:analysis_html}
+\begin{suppdata}
+\caption{\label{suppdata:analysis_html}
 {\bf HTML version of the analysis notebook.}
-This file is an HTML rendering of the Jupyter notebook in File~\ref{suppfile:analysis_code}.
+This file is an HTML rendering of the Jupyter notebook in Supplementary Data~\ref{suppdata:analysis_code}.
 It contains detailed plots for all aspects of the deep sequencing data and its analysis.
 }
-\end{suppfile}
+\end{suppdata}
 
-\begin{suppfile}
-\caption{\label{suppfile:fracsurvive_excess}
+\begin{suppdata}
+\caption{\label{suppdata:fracsurvive_excess}
 {\bf The excess fraction surviving for each mutation for each antibody.}
@@
 }
-\end{suppfile}
+\end{suppdata}
 
-\begin{suppfile}
-\caption{\label{suppfile:fracsurvive}
+\begin{suppdata}
+\caption{\label{suppdata:fracsurvive}
 {\bf The fraction surviving for each mutation for each antibody.}
-This file differs from File~\ref{suppfile:fracsurvive_excess} only in that the values are \emph{not} adjusted to be in excess of the library average (e.g., they are from Equation~\ref{eq:fracsurvive} rather than Equation~\ref{eq:fracsurvive_excess}).
+This file differs from File~\ref{suppdata:fracsurvive_excess} only in that the values are \emph{not} adjusted to be in excess of the library average (e.g., they are from Equation~\ref{eq:fracsurvive} rather than Equation~\ref{eq:fracsurvive_excess}).
 }
-\end{suppfile}
+\end{suppdata}
 

### 2018-02-22 paper/ResponseToFinalNatureCommunicationsReviews.tex
revised manuscript and response to reviews for _Nature Communications_

--- 2018-02-21 paper/ResponseToFinalNatureCommunicationsReviews.tex
@@
 \usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
-\usepackage{graphicx}				% Use pdf, png, jpg, or epsÂ§ with pdflatex; use eps in DVI mode
+\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
 								% TeX will automatically convert eps --> pdf in pdflatex		
@@
 
-\subsection*{Overview of reviewer response and changes to the manuscript.}
+\subsection*{Overview of revisions to the manuscript made in response to comments by reviewers.}
 
@@
 Three of the four reviewers then provided additional feedback on that revised manuscript. 
-Here we respond to their feedback and describe the changes we have made.
+Here we respond to their feedback and describe the changes we have made in response to these reviewer comments.
 
@@
 
-In the response, the authors claim that ?It is true that [Throsby et al.
-2008] does mention an escape mutation from a group 1 HA from CR6261, but it never actually shows any neutralization data for that mutation, so we are unable to comment on it.?
-In Throsby et al.
-2008, a group 1 (H5) mutant H111L was confirmed to be resistant to CR6261 (IC50>100 µg/ml vs.
-6.25 µg/ml for wild type).
+In the response, the authors claim that ``It is true that [Throsby et al.
+2008] does mention an escape mutation from a group 1 HA from CR6261, but it never actually shows any neutralization data for that mutation, so we are unable to comment on it.'' In Throsby et al.
+2008, a group 1 (H5) mutant H111L was confirmed to be resistant to CR6261 (IC50$>$100 $\mu$g/ml vs.
+6.25 $\mu$g/ml for wild type).
 The author should still acknowledge this finding and cite Throsby et al.

### 2018-02-22 paper/manuscript.tex
revised manuscript and response to reviews for _Nature Communications_

@@
 
-\subsection*{Calculating fraction of mutant that survive neutralization}
+\subsection*{Calculating fraction of mutants that survive neutralization}
 In prior mutational antigenic profiling work\cite{doud2017complete,dingens2017comprehensive}, we calculated the differential selection on each mutation as the logarithm of its enrichment relative to wildtype in an antibody-selected sample versus a mock-selected control.
@@
 
-\section*{Supplementary Material}
+\section*{Supplementary Information}
 \FloatBarrier

### 2018-02-22 paper/CoverLetter_FinalNatureCommunications.tex
added cover letter and a few more tweaks

\documentclass[a4paper,11pt]{letter}

% Some of the article customisations are relevant for this class

\name{Jesse D. Bloom} % To be used for the return address on the envelope
\signature{} % Goes after the closing (ie at the end of the letter, with space for a signature)
\address{Fred Hutch Cancer Research Center \\ jbloom@fredhutch.org}
% Alternatively, these may be set on an individual basis within each letter environment.

\usepackage{color}

\makelabels % this command prints envelope labels on the final page of the document

\begin{document}
\begin{letter}{Sonja Schmid, PhD \\ Associate Editor \\ \textit{Nature Communications} }

\opening{Dear Sonja,} % eg Hello.

Thank you for the rapid review of our manuscript at \textit{Nature Communications}. 
We have prepared a revised version that addresses the few remaining reviewer requests and your editorial requests.

A description of the changes made in response to the reviewer requests is including in a separate reviewer response letter.
Here we detail the changes made in response to your queries, and provide other requested information.
Also, note that we have included a track-changes PDF that shows all the textual changes in the revised version relative to the initial submission.

Regarding the two-sentence editor's summary, you provided the following sentences:
\begin{quote}
\textsl{
Influenza A virus can escape neutralization antibodies (Abs), but it is unclear how ease of escape differs between Ab types. Here, the authors show that neutralization breadth is an imperfect indicator of viral escape, and that Abs targeting the stalk of hemagglutinin are harder to escape from.}
\end{quote}

Could we instead suggest the following slightly revised version:
\begin{quote}
\textsl{
Influenza A virus can escape antibodies, but it is unclear how the ease of escape depends on the epitope targeted by an antibody. Here, the authors show that neutralization breadth is an imperfect indicator of the ease of viral escape by single mutations, and that antibodies targeting the stalk of hemagglutinin are harder to escape.
}
\end{quote}

We have made the following changes to the manuscript in response to these editor comments:
\begin{itemize}

\item We have shortened the title from ``Quantifying the effects of single mutations on viral escape from broad and narrow antibodies to an H1 influenza hemagglutinin'' (19 words) to ``How single mutations affect viral escape from broad and narrow antibodies to H1 influenza hemagglutinin'' (15 words).

\item We have provided the full address for each affiliation as requested.

\item We have shortened the abstract from the original length of 160 words to the requested length of 150 words.

\item We have shortened all subheadings in the RESULTS and METHODS to no more than 60 characters.

\item We have provided publication-quality versions of all figures as separate files, and moved the legends to the end of the manuscript. {\color{red}STILL NEEDS TO BE DONE.}

\item The naming of supplementary figures has been changed from ``Figure S1'' to ``Supplementary Fig. 1''. Similar changes have been made for supplementary tables and data as requested.

\item We have clarified that the error bars in Figure 6 represent standard errors.

\item We have added the source of the 293F cells. The 293T-CMV-PB1 and MDCK-SIAT1-CMV-PB1 cells were generated in the cited reference 70, the pHW* series plasmids were generated in the cited reference 71, and the pHH-PB1flank-eGFP plasmid was generated in the cited reference 70.

\item We have expanded the Data Availability section of the METHODS as requested.

\item We have added a conflict of interest statement under the heading COMPETING INTERESTS. We have no conflicts.

\item We have provided the Supplementary Information as a separate PDF, which the \textit{Nature Communications} author instructions indicate is an acceptable substitute for a Word document. The reason that we are providing it is a PDF rather than a Word document is that it is written in LaTex. {\color{red} Still need to be done.}

\item Here are the legends for the Supplementary Data files as requested:

\begin{itemize}

\item {\bf Supplementary Data 1: Conversion from sequential numbering of the A/WSN/1933 HA to H3 numbering.}
In this CSV file, the \emph{original} column gives the residue number in sequential (1, 2, ...) numbering of the A/WSN/1933 HA, and the \emph{new} column gives the residue number in H3 numbering.

\item {\bf Supplementary Data 2: Sequences used to infer the tree for all HA subtypes.}
This FASTA file gives the HA sequences used to infer the tree of subtypes in Figure 2.

\item {\bf Supplementary Data 3: Computer code and data for the analysis of the mutational antigenic profiling data.}
The code in this ZIP file performs the entire computational analysis beginning with downloading the FASTQ files from the Sequence Read Archive.
The ZIP file contains a \texttt{README} file that explains the contents in detail.
The actual analysis is performed by the Jupyter notebook \texttt{analysis\_notebook.ipynb}, which includes embedded plots summarizing key statistics and results.
An HTML version of this notebook is also included as Supplementary Data 4.

\item {\bf Supplementary Data 4: HTML version of the analysis notebook.}
This file is an HTML rendering of the Jupyter notebook in Supplementary Data 3.
It contains detailed plots for all aspects of the deep sequencing data and its analysis.

\item {\bf Supplementary Data 5: The excess fraction surviving for each mutation for each antibody.}
This file is a ZIP of CSV files giving the numerical values plotted in the logo plots.
These are median excess fraction surviving taken first across replicates and then across antibody concentrations.
See Equation 2.

\item {\bf Supplementary Data 6: The fraction surviving for each mutation for each antibody.}
This file differs from Supplementary Data 5 only in that the values are \emph{not} adjusted to be in excess of the library average (e.g., they are from Equation 1 rather than Equation 2).

\end{itemize}
\end{itemize}

We have also included all requested files in our submission.
Please let us know if you have any other comments or questions.

\closing{Sincerely} % eg Regards,

\end{letter}
\end{document}

### 2018-02-22 paper/manuscript.tex
added cover letter and a few more tweaks

@@
 The actual analysis is performed by the Jupyter notebook \texttt{analysis\_notebook.ipynb}, which includes embedded plots summarizing key statistics and results.
-An HTML version of this notebook is also included as Supplementary Data~\ref{suppdata:analysis_code}.
+An HTML version of this notebook is also included as Supplementary Data~\ref{suppdata:analysis_html}.
 }
@@
 {\bf The fraction surviving for each mutation for each antibody.}
-This file differs from File~\ref{suppdata:fracsurvive_excess} only in that the values are \emph{not} adjusted to be in excess of the library average (e.g., they are from Equation~\ref{eq:fracsurvive} rather than Equation~\ref{eq:fracsurvive_excess}).
+This file differs from Supplementary Data~\ref{suppdata:fracsurvive_excess} only in that the values are \emph{not} adjusted to be in excess of the library average (e.g., they are from Equation~\ref{eq:fracsurvive} rather than Equation~\ref{eq:fracsurvive_excess}).
 }

### 2018-02-22 paper/manuscript.tex
fix error caught by Juhye

@@
 However, our approach shows that there are mutations that have more modest ($<$10-fold) effects on neutralization by even the broadest antibody. 
-Interestingly, most previous studies~\cite{chai2016two,ekiert2011highly,friesen2014common,dunand2015preexisting} that have reported selecting single mutations with large effects ($\gg$10-fold) on neutralization by anti-stalk antibodies have used group 2 (e.g., H3 or H7) HAs rather than group 1 HAs like the one used in our work---although at least one study has selected large-effect escape mutations to broad anti-stalk antibody in an H5 group 1 HA~\cite{throsby2008heterosubtypic}.
+Interestingly, most previous studies~\cite{chai2016two,ekiert2011highly,friesen2014common,dunand2015preexisting} that have reported selecting single mutations with large effects ($\gg$10-fold) on neutralization by anti-stalk antibodies have used group 2 (e.g., H3 or H7) HAs rather than group 1 HAs like the one used in our work---although at least one study has selected a large-effect escape mutation to a broad anti-stalk antibody in an H5 group 1 HA~\cite{throsby2008heterosubtypic}.
 In addition, when interpreting the magnitude of the effects measured in our experiments, it is important to note that we are only assessing how mutations affect neutralization, and not how they affect Fc-mediated functions that are responsible for much of the \textit{in vivo} protection afforded by anti-stalk antibodies\cite{dilillo2014broadly,dilillo2016broadly}.

### 2018-02-22 paper/FinalNatureCommunicationsSubmissionFiles/manuscript_figsatend_nosupp.tex
near complete final version to submit, just need to add signature to Licence To Publish Form

--- 2018-02-22 paper/manuscript.tex
@@
 
-\begin{figure}
-\centerline{\includegraphics[width=\textwidth]{figs/fracsurvive_example/fracsurvive_fig.pdf}}
-\caption{\label{fig:fracsurvive_example}
-{\bf Quantifying the fraction of virions with each mutation that escape antibody neutralization.}
-This figure shows hypothetical data for four viral variants.
-(A) Virions with the V1K mutation (orange) completely survive an antibody concentration where most other virions are neutralized.
-(B) This resistance is manifested by a large shift in V1K's neutralization curve.
-(C) For each dotted vertical line drawn through the neutralization curves in (B), we calculate the fraction of virions with that mutation that survive the antibody, and indicate this fraction by the height of the letter corresponding to that amino acid at that site.
-(D-F) Similar data to the first three panels, but now V1K has only a small antigenic effect, and so only modestly increases the fraction of virions that survive antibody treatment.
-}
-\end{figure}
-
 Now consider the case where a mutation has just a small antigenic effect, and so only slightly increases the fraction of virions that survive neutralization (Figure~\ref{fig:fracsurvive_example}D).
@@
 
-\begin{figure}
-\centerline{\includegraphics[width=\textwidth]{figs/antibody_summary_fig/Ab_summary.pdf}}
-\caption{\label{fig:antibody_summary}
-{\bf Epitopes and breadth of broad and narrow antibodies targeting HA.}
-(A) Crystal structures of the broad antibodies and sites of escape mutations selected by the narrow ones superimposed on the structure of the HA trimer (PDB 1RVX\cite{gamblin2004structure}). 
-S139/1 (PDB 4GMS\cite{lee2012heterosubtypic}) targets residues in the receptor-binding pocket; C179 (PDB 4HLZ\cite{dreyfus2013structure}) and FI6v3 (PDB 3ZTN\cite{corti2011neutralizing}) target the stalk.
-The sites of escape mutations for H17-L19, H17-L10, and H17-L7 are those mapped by Doud et al\cite{doud2017complete}. 
-(B) A phylogenetic tree of HA subtypes.
-Circles (broad antibodies) and squares (narrow antibodies) denote reported antibody binding or neutralization activity against that subtype. 
-Not all antibodies have been tested against all subtypes. 
-}
-\end{figure}
-
 \subsection*{Broad and narrow antibodies that neutralize influenza virus}
@@
 
-\begin{figure}
-\centerline{\includegraphics[width=0.8\textwidth]{figs/neutralization_curves/WT_neutralization_curves.pdf}}
-\caption{\label{fig:neutcurves}
-{\bf Neutralization of wildtype virus by each antibody, and the fraction of mutant library virions surviving at each concentration used in our experiments.}
-The curves show neutralization of the wildtype A/WSN/1933 virus. 
-Each point represents the mean and standard deviation of three measurements. 
-The vertical dotted lines show the concentrations of antibody that were then used in the mutant virus library selections, and the tables give the overall fraction of the mutant virus libraries that survived at each concentration, determined by qRT-PCR.
-As described in the text, the antibody concentrations were chosen to give similar fractions of the mutant virus libraries that survive, rather than to fall at uniform positions on the neutralization curves of the wildtype virus.
-}
-\end{figure}
-
 The wildtype virus is neutralized by all the antibodies, with IC50s between 0.01 and 1 $\mu$g/ml (Figure~\ref{fig:neutcurves}).
@@
 
-\begin{figure}
-\centerline{\includegraphics[width=\textwidth]{figs/avgfracsurvive.pdf}}
-\caption{
-\label{fig:avgfracsurvive}
-{\bf Strain-specific and anti-receptor-binding-site antibodies select mutations with large antigenic effects, but anti-stalk antibodies only select small-effect mutations.}
-The excess fraction of virions with a mutation at each site that survive the antibody, averaging across all amino-acid mutations at each site (see Equation~\ref{eq:avgfracsurvive}).
-There are multiple sites of large-effect mutations for H17L19, H17L10, H17L7, and S139/1---but none for FI6v3 and C179.
-Supplementary Fig.~\ref{suppfig:maxfracsurvive} shows the excess fraction surviving for the largest-effect mutation at each site.
-Supplementary Figs.~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo} show all mutations using logo plots.
-Sites are labeled in H3 numbering.
-}
-\end{figure}
-
 \subsection*{The effects of all mutations on antibody neutralization}
@@
 
-The correlations among replicates of the mutational antigenic profiling, in terms of the measured fraction-surviving above average for each possible amino-acid mutation, are shown in Supplementary Fig.~\ref{suppfig:corr}.
+The correlations among replicates of the mutational antigenic profiling, in terms of the measured fraction-surviving above average for each possible amino-acid mutation, are shown in Supplementary Fig.~1.
 For the remainder of this paper, we will refer to the median antigenic effect of each mutation across replicates.
@@
 For all four of these antibodies, there are multiple sites in HA where mutations enable a substantial fraction of virions to survive high antibody concentrations (Figure~\ref{fig:avgfracsurvive}).
-Specifically, there are mutations that enable over a third of virions to survive at concentrations where virtually all wildtype virions are neutralized (Supplementary Fig.~\ref{suppfig:maxfracsurvive}).
+Specifically, there are mutations that enable over a third of virions to survive at concentrations where virtually all wildtype virions are neutralized (Supplementary Fig.~2).
 Therefore, the virus can escape these four antibodies with the sort of large-effect single amino-acid mutations that characterize traditional influenza antigenic drift\cite{yewdell1979antigenic,webster1980determination,koel2013substitutions,chambers2015identification,petrie2016antibodies,neher2016prediction}.
@@
 In contrast, the stalk-targeting antibodies C179 and FI6v3 select no strong escape mutants. 
-If we look at the results for these antibodies on the same scale as the other antibodies, we see only a few small bumps in the fraction of virions surviving (Figures~\ref{fig:avgfracsurvive} and Supplementary Fig.~\ref{suppfig:maxfracsurvive}).
-Only if we zoom in can we see that there are actually a few sites where mutations slightly increase the fraction of virions surviving C179 and FI6v3 (Supplementary Figs.~\ref{suppfig:H17L19logo}, \ref{suppfig:H17L10logo}, \ref{suppfig:H17L7logo}, \ref{suppfig:FI6v3logo}, \ref{suppfig:C179logo}, and \ref{suppfig:S139logo}).
+If we look at the results for these antibodies on the same scale as the other antibodies, we see only a few small bumps in the fraction of virions surviving (Figures~\ref{fig:avgfracsurvive} and Supplementary Fig.~2).
+Only if we zoom in can we see that there are actually a few sites where mutations slightly increase the fraction of virions surviving C179 and FI6v3 (Supplementary Figs.~3, 4, 5, 6, 7, and 8).
 But the effect sizes of these antigenic mutations are tiny compared to the other antibodies---especially for FI6v3.
@@
 It is immediately clear that the selected mutations are nearly all in or close to the antibody-binding footprint.
-
-\begin{figure}[h!]
-\centerline{\includegraphics[width=0.9\textwidth]{figs/logoplots_pymol/logoplots_pymol.pdf}}
-\caption{
-\label{fig:structures}
-{\bf Mutations selected by broad and narrow antibodies.}
-(A) Logo plots show sites where mutations have the largest effect.
-Letter heights are proportional to the excess fraction of virions with that mutation that survive antibody, as indicated by the scale bars.
-Structures are colored white to red by the excess fraction surviving for the largest-effect mutation at each site, with each antibody scaled separately.  
-(B) Sites of selection from anti-stalk antibodies, with the same coloring scale for both antibodies.
-Selection for serine or threonine at sites 280 and 291 introduces glycosylation sites at 278 and 289, respectively.
-(C) Cladogram of group 1 HA subtypes.
-The amino acid at site 38 is indicated. 
-Colors indicate whether a subtype has been reported in the literature to be bound or neutralized by C179.  
-}
-\end{figure}
 
@@
 
-However, C179 selects several mutations that do not have any apparent effect on FI6v3 (Figure~\ref{fig:structures}A, Supplementary Fig.~\ref{suppfig:FI6v3logo}).
+However, C179 selects several mutations that do not have any apparent effect on FI6v3 (Figure~\ref{fig:structures}A, Supplementary Fig.~6).
 The most notable of these C179-specific mutations are at site 38.
@@
 Interestingly, group 1 HA subtypes that are susceptible to C179 tend to possess a histidine at site 38, but subtypes that are not bound or neutralized by C179 often possess a serine (Figure~\ref{fig:structures}C).
-%The amino-acid identity at site 111 of HA2 can change the orientation of a conserved Trp21 in HA2, also resulting in group 1 and group 2 differences in the binding ability of stalk-targeting antibodies.
-An H111T mutation in the A/South Carolina/1/1918 (H1N1) has been reported to abrogate C179 binding\cite{dreyfus2013structure}.
-%However, we did not observe escape mutations at site 111, but this may be due to strain-specific differences.
 
@@
 
-\begin{figure}
-\centerline{\includegraphics[width=0.8\textwidth]{figs/FI6v3mutant_neutcurves/FI6v3_mutant_neutcurves.pdf}}
-\caption{
-\label{fig:FI6v3neutcurves}
-{\bf The mutations selected by FI6v3 increase neutralization resistance, but the effects are small.}
-(A) Neutralization curves of individual viral mutants with FI6v3.
-The mutations K280S, K280T, N291S, G47R (HA2), and K(-8)T are all expected to increase neutralization resistance based on the mutational antigenic profiling (Figure~\ref{fig:structures}A), whereas K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization (Supplementary Fig~\ref{suppfig:FI6v3logo}).
-All neutralization curves in this panel were performed in triplicate on the same day.
-This panel shows the average of the replicates; Supplementary Fig.~\ref{suppfig:FI6v3replicates} shows the curves for each replicate individually and performs statistical testing of whether the IC50s for mutants are significantly different than for wildtype.
-(B), (C) In contrast to FI6v3, mutations selected by narrow antibodies have very large effects on neutralization.
-Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 taken from Doud et al\cite{doud2017complete}.
-Points indicate mean and standard error of three replicates.
-}
-\end{figure}
-
 Figure~\ref{fig:FI6v3neutcurves} shows that the mutational antigenic profiling is highly predictive of the results of the neutralization assays, even for small-effect mutations.
 As discussed in the previous section, the three mutations most strongly selected by FI6v3 introduce glycosylation motifs at sites 278-280 or 289-291 (Figure~\ref{fig:structures}A,B).
-We created viruses carrying each of these mutations (K280S, K280T, and N291S) and validated that all three modestly but significantly increased resistance to FI6v3 (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~\ref{suppfig:FI6v3replicates}).
-As a control, we also validated that a mutation at one of these sites (K280A) that does \emph{not} have an effect in our mutational antigenic profiling does not significantly shift the neutralization curve (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~\ref{suppfig:FI6v3replicates}).
+We created viruses carrying each of these mutations (K280S, K280T, and N291S) and validated that all three modestly but significantly increased resistance to FI6v3 (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~9).
+As a control, we also validated that a mutation at one of these sites (K280A) that does \emph{not} have an effect in our mutational antigenic profiling does not significantly shift the neutralization curve (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~9).
 
 Our mutational antigenic profiling also identified several non-glycosylation-motif mutations that were selected by FI6v3.
-We validated that one of these mutations, G47R in the HA2 chain, significantly increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~\ref{suppfig:FI6v3replicates})---although as predicted by the mutational antigenic profiling, the magnitude of the effect was small.
+We validated that one of these mutations, G47R in the HA2 chain, significantly increased neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~9)---although as predicted by the mutational antigenic profiling, the magnitude of the effect was small.
 The most unexpected mutations identified in our mutational antigenic profiling were at site -8 in the signal peptide.
-We tested one of these mutations, K(-8)T, and it did lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~\ref{suppfig:FI6v3replicates})---although despite the significance testing in Supplementary Fig.~\ref{suppfig:FI6v3replicates}, we remain circumspect about the magnitude of this effect relative to the noise in our neutralization assays.
-As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them significantly affected neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~\ref{suppfig:FI6v3replicates}).
+We tested one of these mutations, K(-8)T, and it did lead to a very slight increase in neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~9)---although despite the significance testing in Supplementary Fig.~9, we remain circumspect about the magnitude of this effect relative to the noise in our neutralization assays.
+As controls, we also tested three mutations (P80D and V135T, which are escape mutations for H17-L7 and H17-L19, and M17L in HA2) that did \emph{not} have substantial effects in the mutational antigenic profiling, and confirmed that none of them significantly affected neutralization resistance (Figure~\ref{fig:FI6v3neutcurves}A, Supplementary Fig.~9).
 
@@
 Another possibility is that mutations are tolerated at some HA sites in the antibody footprint, but that the binding energetics are distributed across sites in such a way that none of these tolerated mutations strongly affect neutralization.
-
-\begin{figure}
-\centerline{\includegraphics[width=\textwidth]{figs/prefs_fracsurvive/prefs_fracsurvive_logoplots.pdf}}
-\caption{
-\label{fig:muttolerance}
-{\bf Mutational tolerance of HA sites in the antibody-binding footprints.}
-These plots show all HA sites within 4 angstroms of the antibody in the crystal structure, plus any additional sites (marked with a *) where we identified antigenic mutations. 
-The logo plots at bottom show the preference of each HA site for each amino acid under selection for viral replication as measured by Doud and Bloom\cite{doud2016accurate}.
-For instance, site 153 only tolerates tryptophan, so W occupies the entire height of the preference logo stack.
-In contrast, site 156 tolerates many amino acids, all of which contribute to the height of the preference logo stack. 
-Above the preference logo stacks are logo plots showing the excess fraction surviving antibody treatment as measured in the current study.
-Note that scale for these antigenic effects is 10$\times$ smaller for FI6v3 and C179 than for S139/1.
-}
-\end{figure}
 
@@
 The neutralization assays themselves were performed by using a plate reader to quantify the GFP signal produced by MDCK-SIAT1-CMV-PB1 cells infected by PB1flank-eGFP virus that had been incubated with the indicated antibody concentration as described previously\cite{hooper2013mutant}.
-All neutralization curves in Figure~\ref{fig:FI6v3neutcurves}A represent the mean and standard deviation of three measurements, with the individual replicates shown in Supplementary Fig.~\ref{suppfig:FI6v3replicates}.
+All neutralization curves in Figure~\ref{fig:FI6v3neutcurves}A represent the mean and standard deviation of three measurements, with the individual replicates shown in Supplementary Fig.~9.
 All the neutralization assays for FI6v3 were performed on the same day to eliminate batch effects, with each replicate involving independent serial dilution of the antibody in a separate column of a 96-well plate.
@@
 The conversion between sequential numbering of the A/WSN/1933 HA and the H3 numbering scheme was performed using the Python script available at \url{https://github.com/jbloomlab/HA_numbering}.
-Supplementary Data~\ref{suppdata:HAnumbering} gives the numbering conversion.
+Supplementary Data~1 gives the numbering conversion.
 
@@
 To infer the phylogenetic tree in Figure~\ref{fig:antibody_summary}, we downloaded one HA sequence per subtype from the Influenza Research Database\cite{zhang2017influenza}, inferred the phylogenetic tree using RaxML\cite{stamatakis2014raxml} with a GTR model, and visualized the tree using FigTree (\url{http://tree.bio.ed.ac.uk/software/figtree/}).
-The HA sequences used are in Supplementary Data~\ref{suppdata:HAsubtypes}. 
+The HA sequences used are in Supplementary Data~2. 
 In Figure~\ref{fig:antibody_summary}, we indicate which HAs each antibody has been reported to bind or neutralize\cite{yoshida2009cross, lee2012heterosubtypic, okuno1993common, dreyfus2013structure, corti2011neutralizing}.
@@
 For the cladogram in Figure~\ref{fig:structures}C, the amino-acid identities at site 38 are from the strains tested against C179 by by Dreyfus et al\cite{dreyfus2013structure}. 
-For subtypes not tested, the amino-acid identity reported is that in the strain for that subtype in Supplementary Data~\ref{suppdata:HAsubtypes}.
+For subtypes not tested, the amino-acid identity reported is that in the strain for that subtype in Supplementary Data~2.
 
@@
 We performed three fully independent replicates of each selection using the three replicate mutant virus libraries.
-In addition, we performed technical replicates (independent neutralization experiments on the \emph{same} virus library) in some cases as indicated in Supplementary Fig.~\ref{suppfig:corr}.
+In addition, we performed technical replicates (independent neutralization experiments on the \emph{same} virus library) in some cases as indicated in Supplementary Fig.~1.
 The virus-antibody mixtures were used to infect cells, and viral RNA was extracted, reverse-transcribed, and PCR amplified as described previously\cite{doud2017complete}.
@@
 These fractions are denoted by $\gamma$ in this paper.
-The average of these fractions across libraries are reported in Figure~\ref{fig:neutcurves}, and the values for each individual replicate are in Supplementary Table~\ref{supptab:fracsurvive}.
+The average of these fractions across libraries are reported in Figure~\ref{fig:neutcurves}, and the values for each individual replicate are in Supplementary Table~1.
 The fractions were estimated using qRT-PCR against the viral NP and canine GAPDH as described previously\cite{doud2017complete}.
@@
 The deep sequencing data were analyzed using version 2.2.1 of the \texttt{dms\_tools2} software package\cite{bloom2015software}, which is available at \url{http://jbloomlab.github.io/dms_tools2}.
-Supplementary Data~\ref{suppdata:analysis_code} contains a Jupyter notebook that performs all steps of the analysis beginning with downloading the FASTQ files from the Sequence Read Archive.
-Detailed statistics about the sequencing depth and error rates are shown in this Jupyter notebook and its HTML rendering in Supplementary Data~\ref{suppdata:analysis_html}.
+Supplementary Data~3 contains a Jupyter notebook that performs all steps of the analysis beginning with downloading the FASTQ files from the Sequence Read Archive.
+Detailed statistics about the sequencing depth and error rates are shown in this Jupyter notebook and its HTML rendering in Supplementary Data~4.
 
@@
 
-In this paper, Supplementary Figs.~\ref{fig:avgfracsurvive} and \ref{suppfig:maxfracsurvive} show the median of excess fraction surviving taken across all biological and technical replicates at a given antibody concentration (Equation~\ref{eq:fracsurvive_excess}).
+In this paper, Supplementary Figs.~\ref{fig:avgfracsurvive} and 2 show the median of excess fraction surviving taken across all biological and technical replicates at a given antibody concentration (Equation~\ref{eq:fracsurvive_excess}).
 The subsequent logo plots show the medians of these values taken across all concentrations for each antibody.
-The numerical values plotted in these logo plots are in Supplementary Data~\ref{suppdata:fracsurvive_excess}.
-The fraction surviving values \emph{not} adjusted to be in excess of the library average (Equation~\ref{eq:fracsurvive}) are in Supplementary Data~\ref{suppdata:fracsurvive}.
+The numerical values plotted in these logo plots are in Supplementary Data~5.
+The fraction surviving values \emph{not} adjusted to be in excess of the library average (Equation~\ref{eq:fracsurvive}) are in Supplementary Data~6.
 
@@
 Deep sequencing data are available from the Sequence Read Archive under BioSample accession SAMN05789126 at \url{https://www.ncbi.nlm.nih.gov/sra/?term=SAMN05789126}.
-Computer code that analyzes these data to generate all the results described in this paper is in Supplementary Data~\ref{suppdata:analysis_code}, and an HTML version of the analysis notebook is in Supplementary Data~\ref{suppdata:analysis_html}.
+Computer code that analyzes these data to generate all the results described in this paper is in Supplementary Data~3, and an HTML version of the analysis notebook is in Supplementary Data~4.
 In addition, all of this code as well as the manuscript itself and other data are available on GitHub at \url{https://github.com/jbloomlab/HA_antibody_ease_of_escape}.
@@
 \bibliographystyle{naturemag}
-\bibliography{references.bib}
+\begin{thebibliography}{10}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem{birrer1981antigenic}
+\bibinfo{author}{Birrer, M.~J.}, \bibinfo{author}{Udem, S.},
+  \bibinfo{author}{Nathenson, S.} \& \bibinfo{author}{Bloom, B.~R.}
+\newblock \bibinfo{title}{Antigenic variants of measles virus}.
+\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{293}},
+  \bibinfo{pages}{67--69} (\bibinfo{year}{1981}).
+
+\bibitem{ter1981antigenic}
+\bibinfo{author}{Ter~Meulen, V.}, \bibinfo{author}{L{\"o}ffler, S.},
+  \bibinfo{author}{Carter, M.~J.} \& \bibinfo{author}{Stephenson, J.~R.}
+\newblock \bibinfo{title}{Antigenic characterization of measles and {SSPE}
+  virus haemagglutinin by monoclonal antibodies}.
+\newblock \emph{\bibinfo{journal}{Journal of General Virology}}
+  \textbf{\bibinfo{volume}{57}}, \bibinfo{pages}{357--364}
+  (\bibinfo{year}{1981}).
+
+\bibitem{crainic1983natural}
+\bibinfo{author}{Crainic, R.} \emph{et~al.}
+\newblock \bibinfo{title}{Natural variation of poliovirus neutralization
+  epitopes.}
+\newblock \emph{\bibinfo{journal}{Infection and Immunity}}
+  \textbf{\bibinfo{volume}{41}}, \bibinfo{pages}{1217--1225}
+  (\bibinfo{year}{1983}).
+
+\bibitem{diamond1985antigenic}
+\bibinfo{author}{Diamond, D.~C.} \emph{et~al.}
+\newblock \bibinfo{title}{Antigenic variation and resistance to neutralization
+  in poliovirus type 1}.
+\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{229}},
+  \bibinfo{pages}{1090--1093} (\bibinfo{year}{1985}).
+
+\bibitem{drexler2014robustness}
+\bibinfo{author}{Drexler, J.~F.} \emph{et~al.}
+\newblock \bibinfo{title}{Robustness against serum neutralization of a
+  poliovirus type 1 from a lethal epidemic of poliomyelitis in the {Republic of
+  Congo} in 2010}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{111}}, \bibinfo{pages}{12889--12894}
+  (\bibinfo{year}{2014}).
+
+\bibitem{panum1847iagttagelser}
+\bibinfo{author}{Panum, P.~L.}
+\newblock \emph{\bibinfo{title}{Iagttagelser, anstillede under
+  Maesllinge-Epidemien paa Faeroerne i Aaret 1846:(Observations made during the
+  epidemic of measles on the Faroe Islands in the year 1846)}},
+  vol.~\bibinfo{volume}{3} (\bibinfo{publisher}{Bibliothek for Laeger},
+  \bibinfo{year}{1847}).
+
+\bibitem{salk1984one}
+\bibinfo{author}{Salk, J.}
+\newblock \bibinfo{title}{One-dose immunization against paralytic poliomyelitis
+  using a noninfectious vaccine}.
+\newblock \emph{\bibinfo{journal}{Reviews of Infectious Diseases}}
+  \textbf{\bibinfo{volume}{6}}, \bibinfo{pages}{S444--S450}
+  (\bibinfo{year}{1984}).
+
+\bibitem{fluinboardingschool1978}
+\bibinfo{author}{{British Medical Journal}}.
+\newblock \bibinfo{title}{Influenza in a boarding school}.
+\newblock \emph{\bibinfo{journal}{British Medical Journal}}
+  \textbf{\bibinfo{volume}{1}}, \bibinfo{pages}{587} (\bibinfo{year}{1978}).
+
+\bibitem{davies1982christ}
+\bibinfo{author}{Davies, J.}, \bibinfo{author}{Smith, A.},
+  \bibinfo{author}{Grilli, E.} \& \bibinfo{author}{Hoskins, T.}
+\newblock \bibinfo{title}{Christ's hospital 1978--79: An account of two
+  outbreaks of influenza a h1n1}.
+\newblock \emph{\bibinfo{journal}{Journal of Infection}}
+  \textbf{\bibinfo{volume}{5}}, \bibinfo{pages}{151--156}
+  (\bibinfo{year}{1982}).
+
+\bibitem{yu2008neutralizing}
+\bibinfo{author}{Yu, X.} \emph{et~al.}
+\newblock \bibinfo{title}{Neutralizing antibodies derived from the {B} cells of
+  1918 influenza pandemic survivors}.
+\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{455}},
+  \bibinfo{pages}{532} (\bibinfo{year}{2008}).
+
+\bibitem{couch1983immunity}
+\bibinfo{author}{Couch, R.~B.} \& \bibinfo{author}{Kasel, J.~A.}
+\newblock \bibinfo{title}{Immunity to influenza in man}.
+\newblock \emph{\bibinfo{journal}{Annual Reviews in Microbiology}}
+  \textbf{\bibinfo{volume}{37}}, \bibinfo{pages}{529--549}
+  (\bibinfo{year}{1983}).
+
+\bibitem{kucharski2015estimating}
+\bibinfo{author}{Kucharski, A.~J.} \emph{et~al.}
+\newblock \bibinfo{title}{Estimating the life course of influenza {A (H3N2)}
+  antibody responses from cross-sectional data}.
+\newblock \emph{\bibinfo{journal}{PLoS Biology}} \textbf{\bibinfo{volume}{13}},
+  \bibinfo{pages}{e1002082} (\bibinfo{year}{2015}).
+
+\bibitem{lipsitch2007patterns}
+\bibinfo{author}{Lipsitch, M.} \& \bibinfo{author}{O'Hagan, J.~J.}
+\newblock \bibinfo{title}{Patterns of antigenic diversity and the mechanisms
+  that maintain them}.
+\newblock \emph{\bibinfo{journal}{Journal of the Royal Society Interface}}
+  \textbf{\bibinfo{volume}{4}}, \bibinfo{pages}{787--802}
+  (\bibinfo{year}{2007}).
+
+\bibitem{cobey2014pathogen}
+\bibinfo{author}{Cobey, S.}
+\newblock \bibinfo{title}{Pathogen evolution and the immunological niche}.
+\newblock \emph{\bibinfo{journal}{Annals of the New York Academy of Sciences}}
+  \textbf{\bibinfo{volume}{1320}}, \bibinfo{pages}{1--15}
+  (\bibinfo{year}{2014}).
+
+\bibitem{fulton2015mutational}
+\bibinfo{author}{Fulton, B.~O.} \emph{et~al.}
+\newblock \bibinfo{title}{Mutational analysis of measles virus suggests
+  constraints on antigenic variation of the glycoproteins}.
+\newblock \emph{\bibinfo{journal}{Cell Reports}} \textbf{\bibinfo{volume}{11}},
+  \bibinfo{pages}{1331--1338} (\bibinfo{year}{2015}).
+
+\bibitem{yewdell1979antigenic}
+\bibinfo{author}{Yewdell, J.}, \bibinfo{author}{Webster, R.} \&
+  \bibinfo{author}{Gerhard, W.}
+\newblock \bibinfo{title}{Antigenic variation in three distinct determinants of
+  an influenza type {A} haemagglutinin molecule}.
+\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{279}},
+  \bibinfo{pages}{246--248} (\bibinfo{year}{1979}).
+
+\bibitem{webster1980determination}
+\bibinfo{author}{Webster, R.} \& \bibinfo{author}{Laver, W.}
+\newblock \bibinfo{title}{Determination of the number of nonoverlapping
+  antigenic areas on {Hong Kong (H3N2)} influenza virus hemagglutinin with
+  monoclonal antibodies and the selection of variants with potential
+  epidemiological significance}.
+\newblock \emph{\bibinfo{journal}{Virology}} \textbf{\bibinfo{volume}{104}},
+  \bibinfo{pages}{139--148} (\bibinfo{year}{1980}).
+
+\bibitem{koel2013substitutions}
+\bibinfo{author}{Koel, B.~F.} \emph{et~al.}
+\newblock \bibinfo{title}{Substitutions near the receptor binding site
+  determine major antigenic change during influenza virus evolution}.
+\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{342}},
+  \bibinfo{pages}{976--979} (\bibinfo{year}{2013}).
+
+\bibitem{chambers2015identification}
+\bibinfo{author}{Chambers, B.~S.}, \bibinfo{author}{Parkhouse, K.},
+  \bibinfo{author}{Ross, T.~M.}, \bibinfo{author}{Alby, K.} \&
+  \bibinfo{author}{Hensley, S.~E.}
+\newblock \bibinfo{title}{Identification of hemagglutinin residues responsible
+  for {H3N2} antigenic drift during the 2014--2015 influenza season}.
+\newblock \emph{\bibinfo{journal}{Cell Reports}} \textbf{\bibinfo{volume}{12}},
+  \bibinfo{pages}{1--6} (\bibinfo{year}{2015}).
+
+\bibitem{petrie2016antibodies}
+\bibinfo{author}{Petrie, J.~G.} \emph{et~al.}
+\newblock \bibinfo{title}{Antibodies against the current influenza {A (H1N1)}
+  vaccine strain do not protect some individuals from infection with
+  contemporary circulating influenza {A (H1N1)} virus strains}.
+\newblock \emph{\bibinfo{journal}{The Journal of infectious diseases}}
+  \textbf{\bibinfo{volume}{214}}, \bibinfo{pages}{1947--1951}
+  (\bibinfo{year}{2016}).
+
+\bibitem{neher2016prediction}
+\bibinfo{author}{Neher, R.~A.}, \bibinfo{author}{Bedford, T.},
+  \bibinfo{author}{Daniels, R.~S.}, \bibinfo{author}{Russell, C.~A.} \&
+  \bibinfo{author}{Shraiman, B.~I.}
+\newblock \bibinfo{title}{Prediction, dynamics, and visualization of antigenic
+  phenotypes of seasonal influenza viruses}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{113}}, \bibinfo{pages}{E1701--E1709}
+  (\bibinfo{year}{2016}).
+
+\bibitem{wen2016explaining}
+\bibinfo{author}{Wen, F.}, \bibinfo{author}{Bedford, T.} \&
+  \bibinfo{author}{Cobey, S.}
+\newblock \bibinfo{title}{Explaining the geographical origins of seasonal
+  influenza {A (H3N2)}}.
+\newblock In \emph{\bibinfo{booktitle}{Proc. R. Soc. B}}, vol.
+  \bibinfo{volume}{283}, \bibinfo{pages}{20161312} (\bibinfo{organization}{The
+  Royal Society}, \bibinfo{year}{2016}).
+
+\bibitem{xue2017parallel}
+\bibinfo{author}{Xue, K.~S.} \emph{et~al.}
+\newblock \bibinfo{title}{Parallel evolution of influenza across multiple
+  spatiotemporal scales}.
+\newblock \emph{\bibinfo{journal}{eLife}} \textbf{\bibinfo{volume}{6}},
+  \bibinfo{pages}{e26875} (\bibinfo{year}{2017}).
+
+\bibitem{sui2009structural}
+\bibinfo{author}{Sui, J.} \emph{et~al.}
+\newblock \bibinfo{title}{Structural and functional bases for broad-spectrum
+  neutralization of avian and human influenza {A} viruses}.
+\newblock \emph{\bibinfo{journal}{Nature Structural \& Molecular Biology}}
+  \textbf{\bibinfo{volume}{16}}, \bibinfo{pages}{265--273}
+  (\bibinfo{year}{2009}).
+
+\bibitem{ekiert2009antibody}
+\bibinfo{author}{Ekiert, D.~C.} \emph{et~al.}
+\newblock \bibinfo{title}{Antibody recognition of a highly conserved influenza
+  virus epitope}.
+\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{324}},
+  \bibinfo{pages}{246--251} (\bibinfo{year}{2009}).
+
+\bibitem{corti2011neutralizing}
+\bibinfo{author}{Corti, D.} \emph{et~al.}
+\newblock \bibinfo{title}{A neutralizing antibody selected from plasma cells
+  that binds to group 1 and group 2 influenza {A} hemagglutinins}.
+\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{333}},
+  \bibinfo{pages}{850--856} (\bibinfo{year}{2011}).
+
+\bibitem{lee2012heterosubtypic}
+\bibinfo{author}{Lee, P.~S.} \emph{et~al.}
+\newblock \bibinfo{title}{Heterosubtypic antibody recognition of the influenza
+  virus hemagglutinin receptor binding site enhanced by avidity}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{109}}, \bibinfo{pages}{17040--17045}
+  (\bibinfo{year}{2012}).
+
+\bibitem{ekiert2012cross}
+\bibinfo{author}{Ekiert, D.~C.} \emph{et~al.}
+\newblock \bibinfo{title}{Cross-neutralization of influenza a viruses mediated
+  by a single antibody loop}.
+\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{489}},
+  \bibinfo{pages}{526--532} (\bibinfo{year}{2012}).
+
+\bibitem{schmidt2015viral}
+\bibinfo{author}{Schmidt, A.~G.} \emph{et~al.}
+\newblock \bibinfo{title}{Viral receptor-binding site antibodies with diverse
+  germline origins}.
+\newblock \emph{\bibinfo{journal}{Cell}} \textbf{\bibinfo{volume}{161}},
+  \bibinfo{pages}{1026--1034} (\bibinfo{year}{2015}).
+
+\bibitem{ellebedy2014induction}
+\bibinfo{author}{Ellebedy, A.~H.} \emph{et~al.}
+\newblock \bibinfo{title}{Induction of broadly cross-reactive antibody
+  responses to the influenza ha stem region following {H5N1} vaccination in
+  humans}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{111}}, \bibinfo{pages}{13133--13138}
+  (\bibinfo{year}{2014}).
+
+\bibitem{andrews2015immune}
+\bibinfo{author}{Andrews, S.} \emph{et~al.}
+\newblock \bibinfo{title}{Immune history profoundly affects broadly protective
+  {B} cell responses to influenza.}
+\newblock \emph{\bibinfo{journal}{Science Translational Medicine}}
+  \textbf{\bibinfo{volume}{7}}, \bibinfo{pages}{316ra192}
+  (\bibinfo{year}{2015}).
+
+\bibitem{krammer2015advances}
+\bibinfo{author}{Krammer, F.} \& \bibinfo{author}{Palese, P.}
+\newblock \bibinfo{title}{Advances in the development of influenza virus
+  vaccines}.
+\newblock \emph{\bibinfo{journal}{Nature Reviews Drug Discovery}}
+  \textbf{\bibinfo{volume}{14}}, \bibinfo{pages}{167--182}
+  (\bibinfo{year}{2015}).
+
+\bibitem{corti2017tackling}
+\bibinfo{author}{Corti, D.} \emph{et~al.}
+\newblock \bibinfo{title}{Tackling influenza with broadly neutralizing
+  antibodies}.
+\newblock \emph{\bibinfo{journal}{Current Opinion in Virology}}
+  \textbf{\bibinfo{volume}{24}}, \bibinfo{pages}{60--69}
+  (\bibinfo{year}{2017}).
+
+\bibitem{sheerar1989antigenic}
+\bibinfo{author}{Sheerar, M.}, \bibinfo{author}{Easterday, B.} \&
+  \bibinfo{author}{Hinshaw, V.}
+\newblock \bibinfo{title}{Antigenic conservation of {H1N1} swine influenza
+  viruses}.
+\newblock \emph{\bibinfo{journal}{Journal of General Virology}}
+  \textbf{\bibinfo{volume}{70}}, \bibinfo{pages}{3297--3303}
+  (\bibinfo{year}{1989}).
+
+\bibitem{luoh1992hemagglutinin}
+\bibinfo{author}{Luoh, S.}, \bibinfo{author}{McGregor, M.} \&
+  \bibinfo{author}{Hinshaw, V.}
+\newblock \bibinfo{title}{Hemagglutinin mutations related to antigenic
+  variation in {H1} swine influenza viruses.}
+\newblock \emph{\bibinfo{journal}{Journal of Virology}}
+  \textbf{\bibinfo{volume}{66}}, \bibinfo{pages}{1066} (\bibinfo{year}{1992}).
+
+\bibitem{lee2004effect}
+\bibinfo{author}{Lee, C.-W.}, \bibinfo{author}{Senne, D.~A.} \&
+  \bibinfo{author}{Suarez, D.~L.}
+\newblock \bibinfo{title}{Effect of vaccine use in the evolution of mexican
+  lineage {H5N2} avian influenza virus}.
+\newblock \emph{\bibinfo{journal}{Journal of Virology}}
+  \textbf{\bibinfo{volume}{78}}, \bibinfo{pages}{8372--8381}
+  (\bibinfo{year}{2004}).
+
+\bibitem{cattoli2011antigenic}
+\bibinfo{author}{Cattoli, G.} \emph{et~al.}
+\newblock \bibinfo{title}{Antigenic drift in {H5N1} avian influenza virus in
+  poultry is driven by mutations in major antigenic sites of the hemagglutinin
+  molecule analogous to those for human influenza virus}.
+\newblock \emph{\bibinfo{journal}{Journal of Virology}}
+  \textbf{\bibinfo{volume}{85}}, \bibinfo{pages}{8718--8724}
+  (\bibinfo{year}{2011}).
+
+\bibitem{thyagarajan2014inherent}
+\bibinfo{author}{Thyagarajan, B.} \& \bibinfo{author}{Bloom, J.~D.}
+\newblock \bibinfo{title}{The inherent mutational tolerance and antigenic
+  evolvability of influenza hemagglutinin}.
+\newblock \emph{\bibinfo{journal}{eLife}} \textbf{\bibinfo{volume}{3}},
+  \bibinfo{pages}{e03300} (\bibinfo{year}{2014}).
+
+\bibitem{wu2014high}
+\bibinfo{author}{Wu, N.~C.} \emph{et~al.}
+\newblock \bibinfo{title}{High-throughput profiling of influenza {A} virus
+  hemagglutinin gene at single-nucleotide resolution}.
+\newblock \emph{\bibinfo{journal}{Scientific Reports}}
+  \textbf{\bibinfo{volume}{4}}, \bibinfo{pages}{4942} (\bibinfo{year}{2014}).
+
+\bibitem{heaton2013genome}
+\bibinfo{author}{Heaton, N.~S.}, \bibinfo{author}{Sachs, D.},
+  \bibinfo{author}{Chen, C.-J.}, \bibinfo{author}{Hai, R.} \&
+  \bibinfo{author}{Palese, P.}
+\newblock \bibinfo{title}{Genome-wide mutagenesis of influenza virus reveals
+  unique plasticity of the hemagglutinin and {NS1} proteins}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{110}}, \bibinfo{pages}{20248--20253}
+  (\bibinfo{year}{2013}).
+
+\bibitem{yoshida2009cross}
+\bibinfo{author}{Yoshida, R.} \emph{et~al.}
+\newblock \bibinfo{title}{Cross-protective potential of a novel monoclonal
+  antibody directed against antigenic site {B} of the hemagglutinin of
+  influenza {A} viruses}.
+\newblock \emph{\bibinfo{journal}{PLoS Pathogens}}
+  \textbf{\bibinfo{volume}{5}}, \bibinfo{pages}{e1000350}
+  (\bibinfo{year}{2009}).
+
+\bibitem{chai2016two}
+\bibinfo{author}{Chai, N.} \emph{et~al.}
+\newblock \bibinfo{title}{Two escape mechanisms of influenza {A} virus to a
+  broadly neutralizing stalk-binding antibody}.
+\newblock \emph{\bibinfo{journal}{PLoS Pathogens}}
+  \textbf{\bibinfo{volume}{12}}, \bibinfo{pages}{e1005702}
+  (\bibinfo{year}{2016}).
+
+\bibitem{ekiert2011highly}
+\bibinfo{author}{Ekiert, D.~C.} \emph{et~al.}
+\newblock \bibinfo{title}{A highly conserved neutralizing epitope on group 2
+  influenza {A} viruses}.
+\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{333}},
+  \bibinfo{pages}{843--850} (\bibinfo{year}{2011}).
+
+\bibitem{friesen2014common}
+\bibinfo{author}{Friesen, R.~H.} \emph{et~al.}
+\newblock \bibinfo{title}{A common solution to group 2 influenza virus
+  neutralization}.
+\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
+  Sciences}} \textbf{\bibinfo{volume}{111}}, \bibinfo{pages}{445--450}
+  (\bibinfo{year}{2014}).
+
+\bibitem{dunand2015preexisting}
+\bibinfo{author}{Dunand, C. J.~H.} \emph{et~al.}
+\newblock \bibinfo{title}{Preexisting human antibodies neutralize recently
+  emerged {H7N9} influenza strains}.
+\newblock \emph{\bibinfo{journal}{The Journal of Clinical Investigation}}
+  \textbf{\bibinfo{volume}{125}}, \bibinfo{pages}{1255} (\bibinfo{year}{2015}).
+
+\bibitem{anderson2017natural}
+\bibinfo{author}{Anderson, C.~S.} \emph{et~al.}
+\newblock \bibinfo{title}{Natural and directed antigenic drift of the h1
+  influenza virus hemagglutinin stalk domain}.
+\newblock \emph{\bibinfo{journal}{Scientific Reports}}
+  \textbf{\bibinfo{volume}{7}}, \bibinfo{pages}{14614} (\bibinfo{year}{2017}).
+
+\bibitem{doud2017complete}
+\bibinfo{author}{Doud, M.~B.}, \bibinfo{author}{Hensley, S.~E.} \&
+  \bibinfo{author}{Bloom, J.~D.}
+\newblock \bibinfo{title}{Complete mapping of viral escape from neutralizing
+  antibodies}.
+\newblock \emph{\bibinfo{journal}{PLoS Pathogens}}
+  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{e1006271}
+  (\bibinfo{year}{2017}).
+
+\bibitem{dingens2017comprehensive}
+\bibinfo{author}{Dingens, A.~S.}, \bibinfo{author}{Haddox, H.~K.},
+  \bibinfo{author}{Overbaugh, J.} \& \bibinfo{author}{Bloom, J.~D.}
+\newblock \bibinfo{title}{Comprehensive mapping of {HIV-1} escape from a
+  broadly neutralizing antibody}.
+\newblock \emph{\bibinfo{journal}{Cell Host \& Microbe}}
+  \textbf{\bibinfo{volume}{21}}, \bibinfo{pages}{777--787}
+  (\bibinfo{year}{2017}).
+
+\bibitem{gamblin2004structure}
+\bibinfo{author}{Gamblin, S.} \emph{et~al.}
+\newblock \bibinfo{title}{The structure and receptor binding properties of the
+  1918 influenza hemagglutinin}.
+\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{303}},
+  \bibinfo{pages}{1838--1842} (\bibinfo{year}{2004}).
+
+\bibitem{dreyfus2013structure}
+\bibinfo{author}{Dreyfus, C.}, \bibinfo{author}{Ekiert, D.~C.} \&
+  \bibinfo{author}{Wilson, I.~A.}
+\newblock \bibinfo{title}{Structure of a classical broadly neutralizing stem
+  antibody in complex with a pandemic {H2} influenza virus hemagglutinin}.
+\newblock \emph{\bibinfo{journal}{Journal of Virology}}
+  \textbf{\bibinfo{volume}{87}}, \bibinfo{pages}{7149--7154}
+  (\bibinfo{year}{2013}).
+
+\bibitem{okuno1993common}
+\bibinfo{author}{Okuno, Y.}, \bibinfo{author}{Isegawa, Y.},
+  \bibinfo{author}{Sasao, F.} \& \bibinfo{author}{Ueda, S.}
+\newblock \bibinfo{title}{A common neutralizing epitope conserved between the
+  hemagglutinins of influenza {A} virus {H1} and {H2} strains.}
+\newblock \emph{\bibinfo{journal}{Journal of Virology}}
+  \textbf{\bibinfo{volume}{67}}, \bibinfo{pages}{2552--2558}
+  (\bibinfo{year}{1993}).
+
+\bibitem{caton1982antigenic}
+\bibinfo{author}{Caton, A.~J.}, \bibinfo{author}{Brownlee, G.~G.},
+  \bibinfo{author}{Yewdell, J.~W.} \& \bibinfo{author}{Gerhard, W.}
+\newblock \bibinfo{title}{The antigenic structure of the influenza virus
+  {A/PR/8/34} hemagglutinin ({H1} subtype)}.
+\newblock \emph{\bibinfo{journal}{Cell}} \textbf{\bibinfo{volume}{31}},
+  \bibinfo{pages}{417--427} (\bibinfo{year}{1982}).
+
+\bibitem{sun2010modifications}
+\bibinfo{author}{Sun, X.}, \bibinfo{author}{Longping, V.~T.},
+  \bibinfo{author}{Ferguson, A.~D.} \& \bibinfo{author}{Whittaker, G.~R.}
+\newblock \bibinfo{title}{Modifications to the hemagglutinin cleavage site
+  control the virulence of a neurotropic {H1N1} influenza virus}.
+\newblock \emph{\bibinfo{journal}{Journal of virology}}
+  \textbf{\bibinfo{volume}{84}}, \bibinfo{pages}{8683--8690}
+  (\bibinfo{year}{2010}).
+
+\bibitem{joyce2016vaccine}
+\bibinfo{author}{Joyce, M.~G.} \emph{et~al.}
+\newblock \bibinfo{title}{Vaccine-induced antibodies that neutralize group 1
+  and group 2 influenza a viruses}.
+\newblock \emph{\bibinfo{journal}{Cell}} \textbf{\bibinfo{volume}{166}},
+  \bibinfo{pages}{609--623} (\bibinfo{year}{2016}).
+
+\bibitem{doud2016accurate}
+\bibinfo{author}{Doud, M.~B.} \& \bibinfo{author}{Bloom, J.~D.}
+\newblock \bibinfo{title}{Accurate measurement of the effects of all amino-acid
+  mutations to influenza hemagglutinin}.
+\newblock \emph{\bibinfo{journal}{Viruses}} \textbf{\bibinfo{volume}{8}},
+  \bibinfo{pages}{155} (\bibinfo{year}{2016}).
+
+\bibitem{daniels2003n}
+\bibinfo{author}{Daniels, R.}, \bibinfo{author}{Kurowski, B.},
+  \bibinfo{author}{Johnson, A.~E.} \& \bibinfo{author}{Hebert, D.~N.}
+\newblock \bibinfo{title}{N-linked glycans direct the cotranslational folding
+  pathway of influenza hemagglutinin}.
+\newblock \emph{\bibinfo{journal}{Molecular cell}}
+  \textbf{\bibinfo{volume}{11}}, \bibinfo{pages}{79--90}
+  (\bibinfo{year}{2003}).
+
+\bibitem{burke2014recommended}
+\bibinfo{author}{Burke, D.~F.} \& \bibinfo{author}{Smith, D.~J.}
+\newblock \bibinfo{title}{A recommended numbering scheme for influenza {A} {HA}
+  subtypes}.
+\newblock \emph{\bibinfo{journal}{PloS One}} \textbf{\bibinfo{volume}{9}},
+  \bibinfo{pages}{e112302} (\bibinfo{year}{2014}).
+
+\bibitem{nordholm2017translational}
+\bibinfo{author}{Nordholm, J.} \emph{et~al.}
+\newblock \bibinfo{title}{Translational regulation of viral secretory proteins
+  by the 5′ coding regions and a viral {RNA}-binding protein}.
+\newblock \emph{\bibinfo{journal}{J Cell Biol}} \textbf{\bibinfo{volume}{216}},
+  \bibinfo{pages}{2283--2293} (\bibinfo{year}{2017}).
+
+\bibitem{jin1992high}
+\bibinfo{author}{Jin, L.}, \bibinfo{author}{Fendly, B.~M.} \&
+  \bibinfo{author}{Wells, J.~A.}
+\newblock \bibinfo{title}{High resolution functional analysis of
+  antibody-antigen interactions}.
+\newblock \emph{\bibinfo{journal}{Journal of Molecular Biology}}
+  \textbf{\bibinfo{volume}{226}}, \bibinfo{pages}{851--865}
+  (\bibinfo{year}{1992}).
+
+\bibitem{cunningham1993comparison}
+\bibinfo{author}{Cunningham, B.~C.} \& \bibinfo{author}{Wells, J.~A.}
+\newblock \bibinfo{title}{Comparison of a structural and a functional epitope}.
+\newblock \emph{\bibinfo{journal}{Journal of Molecular Biology}}
+  \textbf{\bibinfo{volume}{234}}, \bibinfo{pages}{554--563}
+  (\bibinfo{year}{1993}).
+
+\bibitem{dall1998mutational}
+\bibinfo{author}{Dall'Acqua, W.} \emph{et~al.}
+\newblock \bibinfo{title}{A mutational analysis of binding interactions in an
+  antigen-antibody protein-protein complex}.
+\newblock \emph{\bibinfo{journal}{Biochemistry}} \textbf{\bibinfo{volume}{37}},
+  \bibinfo{pages}{7981--7991} (\bibinfo{year}{1998}).
+
+\bibitem{throsby2008heterosubtypic}
+\bibinfo{author}{Throsby, M.} \emph{et~al.}
+\newblock \bibinfo{title}{Heterosubtypic neutralizing monoclonal antibodies
+  cross-protective against {H5N1} and {H1N1} recovered from human {IgM+} memory
+  {B} cells}.
+\newblock \emph{\bibinfo{journal}{PloS One}} \textbf{\bibinfo{volume}{3}},
+  \bibinfo{pages}{e3942} (\bibinfo{year}{2008}).
+
+\bibitem{dilillo2014broadly}
+\bibinfo{author}{DiLillo, D.~J.}, \bibinfo{author}{Tan, G.~S.},
+  \bibinfo{author}{Palese, P.} \& \bibinfo{author}{Ravetch, J.~V.}
+\newblock \bibinfo{title}{Broadly neutralizing hemagglutinin stalk-specific
+  antibodies require {Fc$\gamma$R} interactions for protection against
+  influenza virus in vivo}.
+\newblock \emph{\bibinfo{journal}{Nature Medicine}}
+  \textbf{\bibinfo{volume}{20}}, \bibinfo{pages}{143--151}
+  (\bibinfo{year}{2014}).
+
+\bibitem{dilillo2016broadly}
+\bibinfo{author}{DiLillo, D.~J.}, \bibinfo{author}{Palese, P.},
+  \bibinfo{author}{Wilson, P.~C.} \& \bibinfo{author}{Ravetch, J.~V.}
+\newblock \bibinfo{title}{Broadly neutralizing anti-influenza antibodies
+  require fc receptor engagement for in vivo protection}.
+\newblock \emph{\bibinfo{journal}{The Journal of Clinical Investigation}}
+  \textbf{\bibinfo{volume}{126}}, \bibinfo{pages}{605} (\bibinfo{year}{2016}).
+
+\bibitem{gong2013stability}
+\bibinfo{author}{Gong, L.~I.}, \bibinfo{author}{Suchard, M.~A.} \&
+  \bibinfo{author}{Bloom, J.~D.}
+\newblock \bibinfo{title}{Stability-mediated epistasis constrains the evolution
+  of an influenza protein}.
+\newblock \emph{\bibinfo{journal}{eLife}} \textbf{\bibinfo{volume}{2}},
+  \bibinfo{pages}{e00631} (\bibinfo{year}{2013}).
+
+\bibitem{harms2014historical}
+\bibinfo{author}{Harms, M.~J.} \& \bibinfo{author}{Thornton, J.~W.}
+\newblock \bibinfo{title}{Historical contingency and its biophysical basis in
+  glucocorticoid receptor evolution}.
+\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{512}},
+  \bibinfo{pages}{203--207} (\bibinfo{year}{2014}).
+
+\bibitem{starr2017alternative}
+\bibinfo{author}{Starr, T.~N.}, \bibinfo{author}{Picton, L.~K.} \&
+  \bibinfo{author}{Thornton, J.~W.}
+\newblock \bibinfo{title}{Alternative evolutionary histories in the sequence
+  space of an ancient protein}.
+\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{549}},
+  \bibinfo{pages}{409--413} (\bibinfo{year}{2017}).
+
+\bibitem{das2013defining}
+\bibinfo{author}{Das, S.~R.} \emph{et~al.}
+\newblock \bibinfo{title}{Defining influenza a virus hemagglutinin antigenic
+  drift by sequential monoclonal antibody selection}.
+\newblock \emph{\bibinfo{journal}{Cell Host \& Microbe}}
+  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{314--323}
+  (\bibinfo{year}{2013}).
+
+\bibitem{wu2017diversity}
+\bibinfo{author}{Wu, N.~C.} \emph{et~al.}
+\newblock \bibinfo{title}{Diversity of functionally permissive sequences in the
+  receptor-binding site of influenza hemagglutinin}.
+\newblock \emph{\bibinfo{journal}{Cell Host \& Microbe}}
+  \textbf{\bibinfo{volume}{21}}, \bibinfo{pages}{742--753}
+  (\bibinfo{year}{2017}).
+
+\bibitem{balazs2013broad}
+\bibinfo{author}{Balazs, A.~B.}, \bibinfo{author}{Bloom, J.~D.},
+  \bibinfo{author}{Hong, C.~M.}, \bibinfo{author}{Rao, D.~S.} \&
+  \bibinfo{author}{Baltimore, D.}
+\newblock \bibinfo{title}{Broad protection against influenza infection by
+  vectored immunoprophylaxis in mice}.
+\newblock \emph{\bibinfo{journal}{Nature Biotechnology}}
+  \textbf{\bibinfo{volume}{31}}, \bibinfo{pages}{647--652}
+  (\bibinfo{year}{2013}).
+
+\bibitem{bloom2010permissive}
+\bibinfo{author}{Bloom, J.~D.}, \bibinfo{author}{Gong, L.~I.} \&
+  \bibinfo{author}{Baltimore, D.}
+\newblock \bibinfo{title}{Permissive secondary mutations enable the evolution
+  of influenza oseltamivir resistance}.
+\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{328}},
+  \bibinfo{pages}{1272--1275} (\bibinfo{year}{2010}).
+
+\bibitem{hoffmann2000dna}
+\bibinfo{author}{Hoffmann, E.}, \bibinfo{author}{Neumann, G.},
+  \bibinfo{author}{Kawaoka, Y.}, \bibinfo{author}{Hobom, G.} \&
+  \bibinfo{author}{Webster, R.~G.}
+\newblock \bibinfo{title}{A {DNA} transfection system for generation of
+  influenza {A} virus from eight plasmids}.
+\newblock \emph{\bibinfo{journal}{Proc. Natl. Acad. Sci. USA}}
+  \textbf{\bibinfo{volume}{97}}, \bibinfo{pages}{6108--6113}
+  (\bibinfo{year}{2000}).
+
+\bibitem{hooper2013mutant}
+\bibinfo{author}{Hooper, K.~A.} \& \bibinfo{author}{Bloom, J.~D.}
+\newblock \bibinfo{title}{A mutant influenza virus that uses an {N1}
+  neuraminidase as the receptor-binding protein}.
+\newblock \emph{\bibinfo{journal}{Journal of Virology}}
+  \textbf{\bibinfo{volume}{87}}, \bibinfo{pages}{12531--12540}
+  (\bibinfo{year}{2013}).
+
+\bibitem{zhang2017influenza}
+\bibinfo{author}{Zhang, Y.} \emph{et~al.}
+\newblock \bibinfo{title}{{Influenza Research Database}: An integrated
+  bioinformatics resource for influenza virus research.}
+\newblock \emph{\bibinfo{journal}{Nucleic Acids Research}}
+  \textbf{\bibinfo{volume}{45}}, \bibinfo{pages}{D466--D474}
+  (\bibinfo{year}{2017}).
+
+\bibitem{stamatakis2014raxml}
+\bibinfo{author}{Stamatakis, A.}
+\newblock \bibinfo{title}{{RAxML} version 8: a tool for phylogenetic analysis
+  and post-analysis of large phylogenies}.
+\newblock \emph{\bibinfo{journal}{Bioinformatics}}
+  \textbf{\bibinfo{volume}{30}}, \bibinfo{pages}{1312--1313}
+  (\bibinfo{year}{2014}).
+
+\bibitem{bloom2014experimentally}
+\bibinfo{author}{Bloom, J.~D.}
+\newblock \bibinfo{title}{An experimentally determined evolutionary model
+  dramatically improves phylogenetic fit}.
+\newblock \emph{\bibinfo{journal}{Molecular Biology and Evolution}}
+  \textbf{\bibinfo{volume}{31}}, \bibinfo{pages}{1956--1978}
+  (\bibinfo{year}{2014}).
+
+\bibitem{bloom2015software}
+\bibinfo{author}{Bloom, J.~D.}
+\newblock \bibinfo{title}{Software for the analysis and visualization of deep
+  mutational scanning data}.
+\newblock \emph{\bibinfo{journal}{BMC Bioinformatics}}
+  \textbf{\bibinfo{volume}{16}}, \bibinfo{pages}{1} (\bibinfo{year}{2015}).
+
+\end{thebibliography}
 
 \clearpage
-\normalsize
 
-\section*{Supplementary Information}
-\FloatBarrier
-\pagenumbering{arabic}% resets `page` counter to 1
-\renewcommand*{\thepage}{S\arabic{page}}
+\begin{figure}
+\centerline{\includegraphics[width=\textwidth]{Fig1.pdf}}
+\caption{\label{fig:fracsurvive_example}
+{\bf Quantifying the fraction of virions with each mutation that escape antibody neutralization.}
+This figure shows hypothetical data for four viral variants.
+(A) Virions with the V1K mutation (orange) completely survive an antibody concentration where most other virions are neutralized.
+(B) This resistance is manifested by a large shift in V1K's neutralization curve.
+(C) For each dotted vertical line drawn through the neutralization curves in (B), we calculate the fraction of virions with that mutation that survive the antibody, and indicate this fraction by the height of the letter corresponding to that amino acid at that site.
+(D-F) Similar data to the first three panels, but now V1K has only a small antigenic effect, and so only modestly increases the fraction of virions that survive antibody treatment.
+}
+\end{figure}
 
-\begin{suppfigure}
-\centerline{\includegraphics[width=0.84\textwidth]{figs/corrs/site_correlations.pdf}}
-\caption{\label{suppfig:corr}
-{\bf Correlations across experimental replicates.} 
-Each point represents one site in HA, and gives the fraction surviving above average across all amino-acid mutations at that site, as calculated using Equation~\ref{eq:avgfracsurvive}.
-The replicates are highly correlated for antibodies with strong escape mutations (S139/1, H17-L19, H17-L10, and H17-L7), and reasonably correlated for antibodies with only weak escape mutations (FI6v3 and C179).
-}
-\end{suppfigure}
-
-\begin{suppfigure}
-\centerline{\includegraphics[width=\textwidth]{figs/maxfracsurvive.pdf}}
-\caption{\label{suppfig:maxfracsurvive}
-{\bf The excess fraction surviving for the single strongest escape mutation at each site.}
-This plot differs from Figure~\ref{fig:avgfracsurvive} in that the height of the line indicates the excess fraction of virions that survive the antibody selection for the single strongest escape mutation at that site, rather than the average across all amino-acid mutations at that site.
-}
-\end{suppfigure}
-
-\begin{suppfigure}
-\centerline{\includegraphics[trim=0.1cm 0.02cm 0.1cm 0.03cm,clip=true,width=\textwidth]{figs/logoplots/H17L19_fracsurvive.pdf}}
-\caption{\label{suppfig:H17L19logo}
-{\bf The excess fraction surviving selection with antibody H17L19 for all amino-acid mutations.}
-The excess fraction surviving for each replicate was computed using Equation~\ref{eq:fracsurvive_excess}, then we took the median across all technical and biological replicates for each antibody concentration, and then took the medians of those values across concentrations.
-The height of each letter is proportional to the excess fraction surviving of virions with that mutation.
-The scale bar at the top of the plot relates the letter heights to the actual fractions.
-The sites are labeled using H3 numbering.
-}
-\end{suppfigure}
-
-\begin{suppfigure}
-\centerline{\includegraphics[trim=0.1cm 0.02cm 0.1cm 0.03cm,clip=true,width=\textwidth]{figs/logoplots/H17L10_fracsurvive.pdf}}
-\caption{\label{suppfig:H17L10logo}
-{\bf The excess fraction surviving selection with antibody H17L10 for all amino-acid mutations.}
-The excess fraction surviving for each replicate was computed using Equation~\ref{eq:fracsurvive_excess}, then we took the median across all technical and biological replicates for each antibody concentration, and then took the medians of those values across concentrations.
-The height of each letter is proportional to the excess fraction surviving of virions with that mutation.
-The scale bar at the top of the plot relates the letter heights to the actual fractions.
-The sites are labeled using H3 numbering.
-}
-\end{suppfigure}
-
-\begin{suppfigure}
-\centerline{\includegraphics[trim=0.1cm 0.02cm 0.1cm 0.03cm,clip=true,width=\textwidth]{figs/logoplots/H17L7_fracsurvive.pdf}}
-\caption{\label{suppfig:H17L7logo}
-{\bf The excess fraction surviving selection with antibody H17L7 for all amino-acid mutations.}
-The excess fraction surviving for each replicate was computed using Equation~\ref{eq:fracsurvive_excess}, then we took the median across all technical and biological replicates for each antibody concentration, and then took the medians of those values across concentrations.
-The height of each letter is proportional to the excess fraction surviving of virions with that mutation.
-The scale bar at the top of the plot relates the letter heights to the actual fractions.
-The sites are labeled using H3 numbering.
-}
-\end{suppfigure}
-
-\begin{suppfigure}
-\centerline{\includegraphics[trim=0.1cm 0.02cm 0.1cm 0.03cm,clip=true,width=\textwidth]{figs/logoplots/FI6v3_fracsurvive.pdf}}
-\caption{\label{suppfig:FI6v3logo}
-{\bf The excess fraction surviving selection with antibody FI6v3 for all amino-acid mutations.}
-The excess fraction surviving for each replicate was computed using Equation~\ref{eq:fracsurvive_excess}, then we took the median across all technical and biological replicates for each antibody concentration, and then took the medians of those values across concentrations.
-The height of each letter is proportional to the excess fraction surviving of virions with that mutation.
-The scale bar at the top of the plot relates the letter heights to the actual fractions.
-The sites are labeled using H3 numbering.
-}
-\end{suppfigure}
-
-\begin{suppfigure}
-\centerline{\includegraphics[trim=0.1cm 0.02cm 0.1cm 0.03cm,clip=true,width=\textwidth]{figs/logoplots/C179_fracsurvive.pdf}}
-\caption{\label{suppfig:C179logo}
-{\bf The excess fraction surviving selection with antibody C179 for all amino-acid mutations.}
-The excess fraction surviving for each replicate was computed using Equation~\ref{eq:fracsurvive_excess}, then we took the median across all technical and biological replicates for each antibody concentration, and then took the medians of those values across concentrations.
-The height of each letter is proportional to the excess fraction surviving of virions with that mutation.
-The scale bar at the top of the plot relates the letter heights to the actual fractions.
-The sites are labeled using H3 numbering.
-}
-\end{suppfigure}
-
-\begin{suppfigure}
-\centerline{\includegraphics[trim=0.1cm 0.02cm 0.1cm 0.03cm,clip=true,width=\textwidth]{figs/logoplots/S139_fracsurvive.pdf}}
-\caption{\label{suppfig:S139logo}
-{\bf The excess fraction surviving selection with antibody S139/1 for all amino-acid mutations.}
-The excess fraction surviving for each replicate was computed using Equation~\ref{eq:fracsurvive_excess}, then we took the median across all technical and biological replicates for each antibody concentration, and then took the medians of those values across concentrations.
-The height of each letter is proportional to the excess fraction surviving of virions with that mutation.
-The scale bar at the top of the plot relates the letter heights to the actual fractions.
-The sites are labeled using H3 numbering.
-}
-\end{suppfigure}
-
-\begin{suppfigure}
-{\bf \LARGE A} \\
-\centerline{\includegraphics[width=\textwidth]{figs/FI6v3_replicate_neutcurves.pdf}} \\
-\vspace{0.1in}
-{\bf \LARGE B}
-\begin{center}
-\input{figs/FI6v3_replicate_IC50s}
-\end{center}
-\caption{\label{suppfig:FI6v3replicates}
-{\bf Replicates of the FI6v3 neutralization curves in Figure~\ref{fig:FI6v3neutcurves}A.}
-The neutralization assays were performed in triplicate for all nine mutants and wildtype.
-Figure~\ref{fig:FI6v3neutcurves}A shows the \emph{average} of those replicates.
-(A) The neutralization data for each replicate shown individually, with IC50 values fit using a four-parameter logistic curve with the top value constrained to one (see \url{https://jbloomlab.github.io/dms_tools2/dms_tools2.neutcurve.html} for the code used for the fitting.)
-(B) Table of the IC50 values for each replicate.
-We used an unpaired Student's t-test with unequal variances to test the null hypothesis that each mutant had an IC50 indistinguishable from wildtype.
-We then used Bonferroni's method to correct the $P$-values for multiple testing, and report these corrected values.
-}
-\end{suppfigure}
-
-\begin{supptable}
-\centering
-\begin{tabular}{cccc}
-antibody &  concentration ($\mu$g/ml) &  replicate &  fraction surviving \\
-\hline
-FI6v3 &            0.1 &  1a &             0.01662 \\
- FI6v3 &            0.1 &  1b &             0.01390 \\
-FI6v3 &            0.2 &  1a &             0.00465 \\
-FI6v3 &            0.2 &  1b &             0.00345 \\
-FI6v3 &            0.1 &  2 &             0.02322 \\
-FI6v3 &            0.2 &  2 &             0.00278 \\
-FI6v3 &            0.1 &  3 &             0.00903 \\
-FI6v3 &            0.2 &  3 &             0.00144 \\
-S139/1 &          100.0 &  1 &             0.02490 \\
-S139/1 &          200.0 &  1 &             0.01470 \\
-S139/1 &          300.0 &  1 &             0.01270 \\
-S139/1 &          100.0 &  2 &             0.02190 \\
-S139/1 &          200.0 &  2 &             0.01720 \\
-S139/1 &          300.0 &  2 &             0.00854 \\
-S139/1 &          100.0 &  3 &             0.05180 \\
-S139/1 &          200.0 &  3 &             0.04060 \\
-S139/1 &          300.0 &  3 &             0.03750 \\
-C179 &            1.0 &    1a &             0.00941 \\
-C179 &            1.0 &    1b &             0.00890 \\
-C179 &            1.0 &    1c &             0.00960 \\
-C179 &            2.5 &    1 &             0.00450 \\
-C179 &            1.0 &    2 &             0.00554 \\
-C179 &            2.5 &    2 &             0.00198 \\
-C179 &            1.0 &    3 &             0.00256 \\
-C179 &            2.5 &    3 &             0.00100 \\
-\end{tabular}
-\caption{\label{supptab:fracsurvive}
-{\bf The total fraction of virions surviving each antibody treatment at each concentration as estimated by qPCR.}
-These are the quantities referred to as $\gamma$.
-This table shows the values for the broad antibodies; values for the narrow H17-L17, H17-L10, and H17-L7 antibodies have been reported previously\cite{doud2017complete}.
-}
-\end{supptable}
 \clearpage
 
-\begin{suppdata}
-\caption{\label{suppdata:HAnumbering}
-{\bf Conversion from sequential numbering of the A/WSN/1933 HA to H3 numbering.}
-In this CSV file, the \emph{original} column gives the residue number in sequential (1, 2, ...) numbering of the A/WSN/1933 HA, and the \emph{new} column gives the residue number in H3 numbering.
+\begin{figure}
+\centerline{\includegraphics[width=\textwidth]{Fig2.pdf}}
+\caption{\label{fig:antibody_summary}
+{\bf Epitopes and breadth of broad and narrow antibodies targeting HA.}
+(A) Crystal structures of the broad antibodies and sites of escape mutations selected by the narrow ones superimposed on the structure of the HA trimer (PDB 1RVX\cite{gamblin2004structure}). 
+S139/1 (PDB 4GMS\cite{lee2012heterosubtypic}) targets residues in the receptor-binding pocket; C179 (PDB 4HLZ\cite{dreyfus2013structure}) and FI6v3 (PDB 3ZTN\cite{corti2011neutralizing}) target the stalk.
+The sites of escape mutations for H17-L19, H17-L10, and H17-L7 are those mapped by Doud et al\cite{doud2017complete}. 
+(B) A phylogenetic tree of HA subtypes.
+Circles (broad antibodies) and squares (narrow antibodies) denote reported antibody binding or neutralization activity against that subtype. 
+Not all antibodies have been tested against all subtypes. 
 }
-\end{suppdata}
+\end{figure}
 
-\begin{suppdata}
-\caption{\label{suppdata:HAsubtypes}
-{\bf Sequences used to infer the tree for all HA subtypes.}
-This FASTA file gives the HA sequences used to infer the tree of subtypes in Figure~\ref{fig:antibody_summary}.
+\clearpage
+
+\begin{figure}
+\centerline{\includegraphics[width=0.8\textwidth]{Fig3.pdf}}
+\caption{\label{fig:neutcurves}
+{\bf Neutralization of wildtype virus by each antibody, and the fraction of mutant library virions surviving at each concentration used in our experiments.}
+The curves show neutralization of the wildtype A/WSN/1933 virus. 
+Each point represents the mean and standard deviation of three measurements. 
+The vertical dotted lines show the concentrations of antibody that were then used in the mutant virus library selections, and the tables give the overall fraction of the mutant virus libraries that survived at each concentration, determined by qRT-PCR.
+As described in the text, the antibody concentrations were chosen to give similar fractions of the mutant virus libraries that survive, rather than to fall at uniform positions on the neutralization curves of the wildtype virus.
 }
-\end{suppdata}
+\end{figure}
 
-\begin{suppdata}
-\caption{\label{suppdata:analysis_code}
-{\bf Computer code and data for the analysis of the mutational antigenic profiling data.}
-The code in this ZIP file performs the entire computational analysis beginning with downloading the FASTQ files from the Sequence Read Archive.
-The ZIP file contains a \texttt{README} file that explains the contents in detail.
-The actual analysis is performed by the Jupyter notebook \texttt{analysis\_notebook.ipynb}, which includes embedded plots summarizing key statistics and results.
-An HTML version of this notebook is also included as Supplementary Data~\ref{suppdata:analysis_html}.
+\clearpage
+
+\begin{figure}
+\centerline{\includegraphics[width=\textwidth]{Fig4.pdf}}
+\caption{
+\label{fig:avgfracsurvive}
+{\bf Strain-specific and anti-receptor-binding-site antibodies select mutations with large antigenic effects, but anti-stalk antibodies only select small-effect mutations.}
+The excess fraction of virions with a mutation at each site that survive the antibody, averaging across all amino-acid mutations at each site (see Equation~\ref{eq:avgfracsurvive}).
+There are multiple sites of large-effect mutations for H17L19, H17L10, H17L7, and S139/1---but none for FI6v3 and C179.
+Supplementary Fig.~2 shows the excess fraction surviving for the largest-effect mutation at each site.
+Supplementary Figs.~3, 4, 5, 6, 7, and 8 show all mutations using logo plots.
+Sites are labeled in H3 numbering.
 }
-\end{suppdata}
+\end{figure}
 
-\begin{suppdata}
-\caption{\label{suppdata:analysis_html}
-{\bf HTML version of the analysis notebook.}
-This file is an HTML rendering of the Jupyter notebook in Supplementary Data~\ref{suppdata:analysis_code}.
-It contains detailed plots for all aspects of the deep sequencing data and its analysis.
+\clearpage
+
+\begin{figure}[h!]
+\centerline{\includegraphics[width=0.9\textwidth]{Fig5.pdf}}
+\caption{
+\label{fig:structures}
+{\bf Mutations selected by broad and narrow antibodies.}
+(A) Logo plots show sites where mutations have the largest effect.
+Letter heights are proportional to the excess fraction of virions with that mutation that survive antibody, as indicated by the scale bars.
+Structures are colored white to red by the excess fraction surviving for the largest-effect mutation at each site, with each antibody scaled separately.  
+(B) Sites of selection from anti-stalk antibodies, with the same coloring scale for both antibodies.
+Selection for serine or threonine at sites 280 and 291 introduces glycosylation sites at 278 and 289, respectively.
+(C) Cladogram of group 1 HA subtypes.
+The amino acid at site 38 is indicated. 
+Colors indicate whether a subtype has been reported in the literature to be bound or neutralized by C179.  
 }
-\end{suppdata}
+\end{figure}
 
-\begin{suppdata}
-\caption{\label{suppdata:fracsurvive_excess}
-{\bf The excess fraction surviving for each mutation for each antibody.}
-This file is a ZIP of CSV files giving the numerical values plotted in the logo plots.
-These are median excess fraction surviving taken first across replicates and then across antibody concentrations.
-See Equation~\ref{eq:fracsurvive_excess}.
+\clearpage
+
+\begin{figure}
+\centerline{\includegraphics[width=0.8\textwidth]{Fig6.pdf}}
+\caption{
+\label{fig:FI6v3neutcurves}
+{\bf The mutations selected by FI6v3 increase neutralization resistance, but the effects are small.}
+(A) Neutralization curves of individual viral mutants with FI6v3.
+The mutations K280S, K280T, N291S, G47R (HA2), and K(-8)T are all expected to increase neutralization resistance based on the mutational antigenic profiling (Figure~\ref{fig:structures}A), whereas K280A, M17L (HA2), P80D, and V135T are \emph{not} expected to affect neutralization (Supplementary Fig~6).
+All neutralization curves in this panel were performed in triplicate on the same day.
+This panel shows the average of the replicates; Supplementary Fig.~9 shows the curves for each replicate individually and performs statistical testing of whether the IC50s for mutants are significantly different than for wildtype.
+(B), (C) In contrast to FI6v3, mutations selected by narrow antibodies have very large effects on neutralization.
+Shown are neutralization curves for representative escape mutants from H17-L19 and H17-L7 taken from Doud et al\cite{doud2017complete}.
+Points indicate mean and standard error of three replicates.
 }
-\end{suppdata}
+\end{figure}
 
-\begin{suppdata}
-\caption{\label{suppdata:fracsurvive}
-{\bf The fraction surviving for each mutation for each antibody.}
-This file differs from Supplementary Data~\ref{suppdata:fracsurvive_excess} only in that the values are \emph{not} adjusted to be in excess of the library average (e.g., they are from Equation~\ref{eq:fracsurvive} rather than Equation~\ref{eq:fracsurvive_excess}).
+\clearpage
+
+\begin{figure}
+\centerline{\includegraphics[width=\textwidth]{Fig7.pdf}}
+\caption{
+\label{fig:muttolerance}
+{\bf Mutational tolerance of HA sites in the antibody-binding footprints.}
+These plots show all HA sites within 4 angstroms of the antibody in the crystal structure, plus any additional sites (marked with a *) where we identified antigenic mutations. 
+The logo plots at bottom show the preference of each HA site for each amino acid under selection for viral replication as measured by Doud and Bloom\cite{doud2016accurate}.
+For instance, site 153 only tolerates tryptophan, so W occupies the entire height of the preference logo stack.
+In contrast, site 156 tolerates many amino acids, all of which contribute to the height of the preference logo stack. 
+Above the preference logo stacks are logo plots showing the excess fraction surviving antibody treatment as measured in the current study.
+Note that scale for these antigenic effects is 10$\times$ smaller for FI6v3 and C179 than for S139/1.
 }
-\end{suppdata}
+\end{figure}
 

### 2018-02-23 paper/CoverLetter_FinalNatureCommunications.tex
near complete final version to submit, just need to add signature to Licence To Publish Form
slight revisions requested by editorial staff at _Nature Communications_

@@
 
+Regarding the question of whether we want transparent peer review, the substantive reviewer comments were made in the initial round of reviews for \textit{Nature Microbiology} that were then transferred to \textit{Nature Communications}.
+If those reviewer comments can be used, then yes we would like to opt for publication of the reviewer comments.
+However, our understanding is that perhaps this option can only be used for the reviews for the second round with \textit{Nature Communications.}
+If that is the case, then we opt \emph{not} to include the reviewer comments, since these comments for this second round of review are cursory and not particularly informative in the absence of the initial reviews for \textit{Nature Microbiology}.
+
 We have made the following changes to the manuscript in response to these editor comments:
@@
 
-\item We have provided publication-quality versions of all figures as separate files, and moved the legends to the end of the manuscript. {\color{red}STILL NEEDS TO BE DONE.}
+\item We have provided publication-quality versions of all figures as separate files, and moved the legends to the end of the manuscript.
 
@@
 The reason that we are providing it is a PDF rather than a Word document is that it is written in LaTex.
-{\color{red} Still need to be done.}
 

### 2018-02-23 paper/FinalNatureCommunicationsSubmissionFiles/manuscript_figsatend_nosupp.tex
slight revisions requested by editorial staff at _Nature Communications_

@@
 \subsection*{The effects of all mutations on antibody neutralization}
-We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}. 
+We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves} (the fraction of each library neutralized at each of these concentrations is listed in Supplementary Table~1).
 All experiments were performed in full biological triplicate using three independently generated virus libraries carrying single amino-acid mutations to HA\cite{doud2016accurate}.
@@
 \subsection*{COMPETING INTERESTS}
-The authors declare no competing financial interests.
+The authors declare no competing interests, financial or otherwise.
 

### 2018-02-23 paper/manuscript.tex
slight revisions requested by editorial staff at _Nature Communications_

@@
 \subsection*{The effects of all mutations on antibody neutralization}
-We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves}. 
+We performed mutational antigenic profiling using the three broad antibodies at the concentrations indicated in Figure~\ref{fig:neutcurves} (the fraction of each library neutralized at each of these concentrations is listed in Supplementary Table~\ref{supptab:fracsurvive}).
 All experiments were performed in full biological triplicate using three independently generated virus libraries carrying single amino-acid mutations to HA\cite{doud2016accurate}.
